### 2014-11-05 paper/paper.tex
One pass through the semantics, reformatting and fixing bugs + adding todos towards the paper

\documentclass[twocolumn]{article}
\usepackage{verbatim}
\usepackage[utf8x]{inputenc}
\title{A semantics of core AUTOSAR}
\author{Johan Nordlander \and Patrik Jansson}

\begin{document}
\maketitle
\begin{abstract}

\end{abstract}

\section{Introduction}
\label{sec:Intro}

Enkla exempel redan tidigt (informellt)

Kort om begränsningar och contributions

\section{A calculus for AUTOSAR software components}
\label{sec:Calc}

TODO: referenser från AUTOSAR-syntax och koncept i artikeln till AUTSAR-dokumentation (standarden)

TODO: mer detaljer om begränsningar: vad vi inte täcker in i kalkylen

\section{Semantics}
\label{sec:Sem}

TODO: referenser från AUTOSAR-regler i artikeln till AUTSAR-dokumentation (standarden)

\section{Examples}
\label{sec:Examples}

\begin{itemize}
\item Mer detaljer om de enkla exemplen + ett större exempel.
\item några exempel, valda för att illustrera värdet av kalkylen och vår semantik
\end{itemize}

\section{Discussion / results}
\label{sec:Disc}

\begin{itemize}
\item osäkerheter: där det finns alternativa tolkningar
\item förslag på förtydliganden / förbättringar / exempel på oklarheter
\end{itemize}

\section{Conclusions and Future Work}
\label{sec:Conc}

Return to the limitation - what is the next step (which can be lifted more easily)

\onecolumn
\appendix
\section{Prolog code}
\label{sec:Prolog}

\verbatiminput{../prolog/semantics.pl}

\end{document}

### 2014-11-05 prolog/semantics.pl

@@ -5,147 +5,87 @@
 
 %%%%% Combining reductions
 
-[]  ---hear(_A,_L)--->  []
-    .
-[ P | Q ]  ---say(A,L)--->  R
-    :-
-    P  ---say(A,L)--->  P2
-    ,
-    Q  ---hear(A,L)--->  Q2
-    ,
-    flateval(P2, Q2, R)                 % Intuitively the same as R = [ P2 | Q2 ]
-    .
-[ P | Q ]  ---say(A,L)--->  [ P2 | Q2 ]
-    :-
-    P  ---hear(A,L)--->  P2
-    ,
-    Q  ---say(A,L)--->  Q2
-    .
-[ P | Q ]  ---hear(A,L)--->  [ P2 | Q2 ]
-    :-
-    P  ---hear(A,L)--->  P2
-    ,
-    Q  ---hear(A,L)--->  Q2
-    .
-[ P | Q ]  ---say(A,L)--->  [ P | Q2 ]
-    :-
-    Q  ---say(A,L)--->  Q2 
-    ,
-    tolerate(A,P)
-    .
-[ P | Q ]  ---hear(A,L)--->  [ P | Q2 ]
-    :-
-    Q  ---hear(A,L)--->  Q2
-    ,
-    tolerate(A,P)
-    .
-[]  ---delta(_T)--->  []
-    .
-[ P | Q ]  ---delta(T)--->  [ P2 | Q2 ]
-    :-
-    P  ---delta(T)--->  P2
-    ,
-    Q  ---delta(T)--->  Q2
-    .
+[]  ---hear(_A,_L)--->  []   .
+
+% [ P | Ps ]  ---say(A,L)--->   [ P2 | Ps2 ]  :-   P  ---say(A,L)--->   P2  ,  Ps  ---hear(A,L)--->  Ps2 .
+
+[ P | Ps ]  ---say(A,L)--->   Rs            :-   P  ---say(A,L)--->   P2  ,  Ps  ---hear(A,L)--->  Ps2 ,
+                                                 flateval(P2, Ps2, Rs).
+[ P | Ps ]  ---say(A,L)--->   [ P2 | Ps2 ]  :-   P  ---hear(A,L)--->  P2  ,  Ps  ---say(A,L)--->   Ps2 .
[…]

### 2014-11-05 prolog/test.pl

@@ -4,11 +4,31 @@
 
 :- consult('semantics.pl').
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Static info examples %%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+connected(i:p0:e, i:p3:e).
+connected(_, _) :- false.
+
+event(i:r1, timing(300)).
+event(i:r1, data_received(p3:e)).
+
+implementation(i:r1, return(void)).
+
+minimum_start_interval(i:r1, 10).
+
+can_be_invoked_concurrently(i:r2).
+
+server_call_point(i:r1, sync(p2:o)).
+
+%%%%%%%%%%%%%%%%
+
 example1(L,R) :- 
+%  Cont1 = fn(X,rte_exit(x, Cont2)),
+%  Cont2 = fn(X,rte_exit(x, fn(Y,return(void)))),
   Proc1 = rinst(i:r1,nil,[],rte_enter(x,Cont1)),
   Proc2 = rinst(i:r2,nil,[],rte_enter(x,Cont2)),
   Block = [excl(i:x,free), Proc1, Proc2],
-  [excl(i:z,free) | Block] ---L---> R.
+  [[excl(i:z,free), []] | Block] ---L---> R.
 
 
 badexample(L,R) :- 

### 2015-03-19 paper/paper.tex
Minor updates about prolog implementations
Intro draft.

@@
 \usepackage[utf8x]{inputenc}
+\usepackage[a4paper,margin=1.4cm]{geometry}
+% \usepackage{minted}
 \title{A semantics of core AUTOSAR}
@@
 
-Enkla exempel redan tidigt (informellt)
+The AUTOSAR standard is an open software component architecture for the automotive industry.
+Its main purpose is to enable interoperability of software modules among different vendors and on heterogeneous platforms, via an extensive set of standardized interfaces and libraries, and a common software development methodology.
 
-Kort om begränsningar och contributions
+The standard has a rather wide scope and covers many features normally associated with complex operating systems, like I/O abstraction, concurrency, communication, distribution and real-time predictability.
+Unlike existing operating systems, however, AUTOSAR is not de facto defined in terms of a particular implementation.
+Instead, an explicit goal of AUTOSAR is to constitute an abstract specification that allows multiple competing realizations, and even systems built as an assembly of fragments from many different (and competing) vendors.
+Such a goal naturally puts the focus on the standard specification itself.
 
-\section{A calculus for AUTOSAR software components}
+Unfortunately, the AUTOSAR specification is not very rigorous, despite a sheer size of more than one hundred pdf documents and over 12,500 pages of text and UML diagrams in total.
+It is also not very abstract, in that it makes frequent references to assumed implementation techniques for the purpose of defining its semantics.
+In practice, the AUTOSAR standard becomes blurred with the specific behavior of one's chosen platform and development tools.
+And because the standard is open to interpretation, the interoperability of software components across tools and platforms is often seriously hampered.
+What is more, a single software component cannot easily be studied and understood in isolation, since its interactive behavior is only indirectly defined in terms of the concrete C-code and OS tasks that realize it and every interacting component.
+
+This paper takes an important step towards a remedy to these problems, by contributing a formal specification of a substantial core of the AUTOSAR standard.
+The formalization covers most of the Software Component Template (SWC) and its accompanying Run-Time Environment (RTE) (Section~\ref{sec:Calc}), and is able to directly express every legal way a system of SWCs can evolve in at run-time on an arbitrarily fast platform (Section~\ref{sec:Sem}).
+It can thus serve as a basis for both a concrete AUTOSAR implementation on a specific platform (whose supported behaviors must be a subset of those defined by the formal semantics), and a platform-independent AUTOSAR simulator (where behaviors from the legal set can be picked at random).
+Examples of semantic derivations and their applications are found in Section~\ref{sec:Examples}.
+
+The semantics has been formalized with the intent to accurately capture the informal meaning of the AUTOSAR standard, although mistakes and misunderstandings are certainly both possible and plausible.
+A formal notation is nevertheless a good starting point for any discussion on the resolution of such issues.
+The semantics is furthermore written to be unambiguous, except where concurrent execution should allow for more than one observed behavior.
+At some points, the AUTOSAR documents have been found unclear; here a specific choice has been made but alternative interpretations will be discussed separately (Section~\ref{sec:DiscAmb}).
+At other points, the standard documents are clear but the resulting semantics is still dubious.
+These cases will also be emphasized and discussed, together with suggested ways forward (Section~\ref{sec:DiscImp}).
+
+TODO: Limitations
+
+TODO: Simple examples early (informally)
+
+\section{Syntax and prelliminaries}
 \label{sec:Calc}
 
-TODO: referenser från AUTOSAR-syntax och koncept i artikeln till AUTSAR-dokumentation (standarden)
+TODO: links from syntax and concepts introduced here to the AUTSAR spec
 
-TODO: mer detaljer om begränsningar: vad vi inte täcker in i kalkylen
+TODO: details on limitations: what's not covered by the calculus
 
@@
 
-TODO: referenser från AUTOSAR-regler i artikeln till AUTSAR-dokumentation (standarden)
+TODO: links from rules and choices made here to the AUTSAR spec
 
@@
 \begin{itemize}
-\item Mer detaljer om de enkla exemplen + ett större exempel.
-\item några exempel, valda för att illustrera värdet av kalkylen och vår semantik
+\item More details on the simple examples + a bigger example.
+\item Some examples chosen to illustrate the value of a formal calculus
 \end{itemize}
@@
 
-\begin{itemize}
-\item osäkerheter: där det finns alternativa tolkningar
-\item förslag på förtydliganden / förbättringar / exempel på oklarheter
-\end{itemize}
+\subsection{Ambiguities and alternative interpretations}
+\label{sec:DiscAmb}
+
+\subsection{Clarification proposals and improvements}
+\label{sec:DiscImp}
 
@@
 
+% \inputminted{prolog}{../prolog/semantics.pl}
+
 \verbatiminput{../prolog/semantics.pl}

### 2015-04-02 ABS.hs
Added license info.

@@ -1,3 +1,32 @@
+{-
+Copyright (c) 2014-2015, Johan Nordlander, Jonas Duregård, Michał Pałka,
+                         Patrik Jansson and Josef Svenningsson
+All rights reserved.
+
+Redistribution and use in source and binary forms, with or without
+modification, are permitted provided that the following conditions are met:
+
+   * Redistributions of source code must retain the above copyright notice,
+     this list of conditions and the following disclaimer.
+   * Redistributions in binary form must reproduce the above copyright
+     notice, this list of conditions and the following disclaimer in the
+     documentation and/or other materials provided with the distribution.
+   * Neither the name of the Chalmers University of Technology nor the names of its 
+     contributors may be used to endorse or promote products derived from this 
+     software without specific prior written permission.
+
+THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE
+DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE
+FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
+DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
+SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER
+CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY,
+OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
+OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
+-}
+
 module Main where
         
 import ARSim

### 2015-04-02 AR.hs

@@ -1,3 +1,32 @@
+{-
+Copyright (c) 2014-2015, Johan Nordlander, Jonas Duregård, Michał Pałka,
+                         Patrik Jansson and Josef Svenningsson
+All rights reserved.
+
+Redistribution and use in source and binary forms, with or without
+modification, are permitted provided that the following conditions are met:
+
+   * Redistributions of source code must retain the above copyright notice,
+     this list of conditions and the following disclaimer.
+   * Redistributions in binary form must reproduce the above copyright
+     notice, this list of conditions and the following disclaimer in the
+     documentation and/or other materials provided with the distribution.
+   * Neither the name of the Chalmers University of Technology nor the names of its 
+     contributors may be used to endorse or promote products derived from this 
+     software without specific prior written permission.
+
+THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE
+DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE
+FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
+DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
+SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER
+CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY,
+OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
+OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
+-}
+
 {-# LANGUAGE FlexibleInstances #-}
 module AR where
 import qualified Data.Map as Map

### 2015-04-02 ARComp.hs

@@ -1,3 +1,32 @@
+{-
+Copyright (c) 2014-2015, Johan Nordlander, Jonas Duregård, Michał Pałka,
+                         Patrik Jansson and Josef Svenningsson
+All rights reserved.
+
+Redistribution and use in source and binary forms, with or without
+modification, are permitted provided that the following conditions are met:
+
+   * Redistributions of source code must retain the above copyright notice,
+     this list of conditions and the following disclaimer.
+   * Redistributions in binary form must reproduce the above copyright
+     notice, this list of conditions and the following disclaimer in the
+     documentation and/or other materials provided with the distribution.
+   * Neither the name of the Chalmers University of Technology nor the names of its 
+     contributors may be used to endorse or promote products derived from this 
+     software without specific prior written permission.
+
+THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE
+DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE
+FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
+DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
+SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER
+CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY,
+OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
+OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
+-}
+
 {-# LANGUAGE RankNTypes       #-}
 {-# LANGUAGE GADTs            #-}
 {-# LANGUAGE TypeFamilies     #-}

### 2015-04-02 ARSim.hs

@@ -1,3 +1,32 @@
+{-
+Copyright (c) 2014-2015, Johan Nordlander, Jonas Duregård, Michał Pałka,
+                         Patrik Jansson and Josef Svenningsson
+All rights reserved.
+
+Redistribution and use in source and binary forms, with or without
+modification, are permitted provided that the following conditions are met:
+
+   * Redistributions of source code must retain the above copyright notice,
+     this list of conditions and the following disclaimer.
+   * Redistributions in binary form must reproduce the above copyright
+     notice, this list of conditions and the following disclaimer in the
+     documentation and/or other materials provided with the distribution.
+   * Neither the name of the Chalmers University of Technology nor the names of its 
+     contributors may be used to endorse or promote products derived from this 
+     software without specific prior written permission.
+
+THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE
+DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE
+FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
+DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
+SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER
+CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY,
+OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
+OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
+-}
+
 {-# LANGUAGE FlexibleInstances, MultiParamTypeClasses, FunctionalDependencies,
              RankNTypes, ExistentialQuantification #-}
 

### 2015-04-02 ARSim/Dynamics.hs

@@ -1,3 +1,32 @@
+{-
+Copyright (c) 2014-2015, Johan Nordlander, Jonas Duregård, Michał Pałka,
+                         Patrik Jansson and Josef Svenningsson
+All rights reserved.
+
+Redistribution and use in source and binary forms, with or without
+modification, are permitted provided that the following conditions are met:
+
+   * Redistributions of source code must retain the above copyright notice,
+     this list of conditions and the following disclaimer.
+   * Redistributions in binary form must reproduce the above copyright
+     notice, this list of conditions and the following disclaimer in the
+     documentation and/or other materials provided with the distribution.
+   * Neither the name of the Chalmers University of Technology nor the names of its 
+     contributors may be used to endorse or promote products derived from this 
+     software without specific prior written permission.
+
+THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE
+DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE
+FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
+DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
+SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER
+CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY,
+OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
+OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
+-}
+
 {-#LANGUAGE GADTs, DeriveDataTypeable, StandaloneDeriving #-}
 module Dynamics(Value, toValue, value, value', Data, Typeable) where
 

### 2015-04-02 ARSim/NewABS.hs

@@ -1,3 +1,32 @@
+{-
+Copyright (c) 2014-2015, Johan Nordlander, Jonas Duregård, Michał Pałka,
+                         Patrik Jansson and Josef Svenningsson
+All rights reserved.
+
+Redistribution and use in source and binary forms, with or without
+modification, are permitted provided that the following conditions are met:
+
+   * Redistributions of source code must retain the above copyright notice,
+     this list of conditions and the following disclaimer.
+   * Redistributions in binary form must reproduce the above copyright
+     notice, this list of conditions and the following disclaimer in the
+     documentation and/or other materials provided with the distribution.
+   * Neither the name of the Chalmers University of Technology nor the names of its 
+     contributors may be used to endorse or promote products derived from this 
+     software without specific prior written permission.
+
+THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE
+DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE
+FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
+DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
+SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER
+CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY,
+OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
+OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
+-}
+
 {-# LANGUAGE DeriveDataTypeable #-}
 module Main where
 

### 2015-04-02 ARSim/TicketDispenser.hs

@@ -1,3 +1,32 @@
+{-
+Copyright (c) 2014-2015, Johan Nordlander, Jonas Duregård, Michał Pałka,
+                         Patrik Jansson and Josef Svenningsson
+All rights reserved.
+
+Redistribution and use in source and binary forms, with or without
+modification, are permitted provided that the following conditions are met:
+
+   * Redistributions of source code must retain the above copyright notice,
+     this list of conditions and the following disclaimer.
+   * Redistributions in binary form must reproduce the above copyright
+     notice, this list of conditions and the following disclaimer in the
+     documentation and/or other materials provided with the distribution.
+   * Neither the name of the Chalmers University of Technology nor the names of its 
+     contributors may be used to endorse or promote products derived from this 
+     software without specific prior written permission.
+
+THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE
+DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE
+FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
+DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
+SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER
+CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY,
+OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
+OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
+-}
+
 {-# LANGUAGE RankNTypes #-}
 module Main where
 

### 2015-04-02 ARSim/Tricky.hs

@@ -1,3 +1,32 @@
+{-
+Copyright (c) 2014-2015, Johan Nordlander, Jonas Duregård, Michał Pałka,
+                         Patrik Jansson and Josef Svenningsson
+All rights reserved.
+
+Redistribution and use in source and binary forms, with or without
+modification, are permitted provided that the following conditions are met:
+
+   * Redistributions of source code must retain the above copyright notice,
+     this list of conditions and the following disclaimer.
+   * Redistributions in binary form must reproduce the above copyright
+     notice, this list of conditions and the following disclaimer in the
+     documentation and/or other materials provided with the distribution.
+   * Neither the name of the Chalmers University of Technology nor the names of its 
+     contributors may be used to endorse or promote products derived from this 
+     software without specific prior written permission.
+
+THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE
+DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE
+FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
+DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
+SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER
+CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY,
+OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
+OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
+-}
+
 module Tricky where
 
 import NewARSim

### 2015-04-02 ARText.hs

@@ -1,3 +1,32 @@
+{-
+Copyright (c) 2014-2015, Johan Nordlander, Jonas Duregård, Michał Pałka,
+                         Patrik Jansson and Josef Svenningsson
+All rights reserved.
+
+Redistribution and use in source and binary forms, with or without
+modification, are permitted provided that the following conditions are met:
+
+   * Redistributions of source code must retain the above copyright notice,
+     this list of conditions and the following disclaimer.
+   * Redistributions in binary form must reproduce the above copyright
+     notice, this list of conditions and the following disclaimer in the
+     documentation and/or other materials provided with the distribution.
+   * Neither the name of the Chalmers University of Technology nor the names of its 
+     contributors may be used to endorse or promote products derived from this 
+     software without specific prior written permission.
+
+THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE
+DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE
+FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
+DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
+SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER
+CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY,
+OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
+OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
+-}
+
 module ARText where
 
 import Data.Map

### 2015-04-02 ARXML.hs

@@ -1,3 +1,31 @@
+{-
+Copyright (c) 2014-2015, Active Group GmbH
+All rights reserved.
+
+Redistribution and use in source and binary forms, with or without
+modification, are permitted provided that the following conditions are met:
+
+   * Redistributions of source code must retain the above copyright notice,
+     this list of conditions and the following disclaimer.
+   * Redistributions in binary form must reproduce the above copyright
+     notice, this list of conditions and the following disclaimer in the
+     documentation and/or other materials provided with the distribution.
+   * Neither the name of the Chalmers University of Technology nor the names of its 
+     contributors may be used to endorse or promote products derived from this 
+     software without specific prior written permission.
+
+THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE
+DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE
+FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
+DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
+SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER
+CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY,
+OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
+OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
+-}
+
 {-# LANGUAGE OverloadedStrings #-}
 module ARXML where
 

### 2015-04-02 AUTOSAR.hs

@@ -1,3 +1,33 @@
+{-
+Copyright (c) 2014-2015, Johan Nordlander, Jonas Duregård, Michał Pałka,
+                         Patrik Jansson, Josef Svenningsson and
+                         Active Group GmbH
+All rights reserved.
+
+Redistribution and use in source and binary forms, with or without
+modification, are permitted provided that the following conditions are met:
+
+   * Redistributions of source code must retain the above copyright notice,
+     this list of conditions and the following disclaimer.
+   * Redistributions in binary form must reproduce the above copyright
+     notice, this list of conditions and the following disclaimer in the
+     documentation and/or other materials provided with the distribution.
+   * Neither the name of the Chalmers University of Technology nor the names of its 
+     contributors may be used to endorse or promote products derived from this 
+     software without specific prior written permission.
+
+THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE
+DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE
+FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
+DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
+SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER
+CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY,
+OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
+OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
+-}
+
 {-# LANGUAGE MultiParamTypeClasses, GADTs, StandaloneDeriving, KindSignatures, FlexibleInstances, FlexibleContexts #-}
 module AUTOSAR where
 

### 2015-04-02 Main.hs

@@ -1,3 +1,32 @@
+{-
+Copyright (c) 2014-2015, Johan Nordlander, Jonas Duregård, Michał Pałka,
+                         Patrik Jansson and Josef Svenningsson
+All rights reserved.
+
+Redistribution and use in source and binary forms, with or without
+modification, are permitted provided that the following conditions are met:
+
+   * Redistributions of source code must retain the above copyright notice,
+     this list of conditions and the following disclaimer.
+   * Redistributions in binary form must reproduce the above copyright
+     notice, this list of conditions and the following disclaimer in the
+     documentation and/or other materials provided with the distribution.
+   * Neither the name of the Chalmers University of Technology nor the names of its 
+     contributors may be used to endorse or promote products derived from this 
+     software without specific prior written permission.
+
+THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE
+DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE
+FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
+DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
+SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER
+CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY,
+OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
+OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
+-}
+
 module Main where
 
 import ARSim

### 2015-04-02 prolog/semantics.pl

@@ -1,3 +1,30 @@
+% Copyright (c) 2014-2015, Johan Nordlander, Jonas Duregård, Michał Pałka,
+%                          Patrik Jansson and Josef Svenningsson
+% All rights reserved.
+% 
+% Redistribution and use in source and binary forms, with or without
+% modification, are permitted provided that the following conditions are met:
+% 
+%    * Redistributions of source code must retain the above copyright notice,
+%      this list of conditions and the following disclaimer.
+%    * Redistributions in binary form must reproduce the above copyright
+%      notice, this list of conditions and the following disclaimer in the
+%      documentation and/or other materials provided with the distribution.
+%    * Neither the name of the Chalmers University of Technology nor the names of its 
+%      contributors may be used to endorse or promote products derived from this 
+%      software without specific prior written permission.
+% 
+% THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+% AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+% IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE
+% DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE
+% FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
+% DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
+% SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER
+% CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY,
+% OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
+% OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
+
 :- op(901, xfy, ':').
 :- op(902, xfx, '==>').
 :- op(905, xfx, '---').

### 2015-04-02 prolog/test.pl

@@ -1,3 +1,30 @@
+% Copyright (c) 2014-2015, Johan Nordlander, Jonas Duregård, Michał Pałka,
+%                          Patrik Jansson and Josef Svenningsson
+% All rights reserved.
+% 
+% Redistribution and use in source and binary forms, with or without
+% modification, are permitted provided that the following conditions are met:
+% 
+%    * Redistributions of source code must retain the above copyright notice,
+%      this list of conditions and the following disclaimer.
+%    * Redistributions in binary form must reproduce the above copyright
+%      notice, this list of conditions and the following disclaimer in the
+%      documentation and/or other materials provided with the distribution.
+%    * Neither the name of the Chalmers University of Technology nor the names of its 
+%      contributors may be used to endorse or promote products derived from this 
+%      software without specific prior written permission.
+% 
+% THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+% AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+% IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE
+% DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE
+% FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
+% DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
+% SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER
+% CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY,
+% OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
+% OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Static info examples %%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 e:p0:i ==> e:p3:i.

### 2015-04-02 prolog/semantics.pl
Added a Prolog parser for testing

@@ -6,87 +6,87 @@
 
 %%%%% Combining reductions
 
-[]  ---hear(_A,_L)--->  []   .
+[]  --- hear(_A,_L) --->  []   .
 
-% [ P | Ps ]  ---say(A,L)--->   [ P2 | Ps2 ]  :-   P  ---say(A,L)--->   P2  ,  Ps  ---hear(A,L)--->  Ps2 .
+% [ P | Ps ]  --- say(A,L) --->   [ P2 | Ps2 ]  :-   P  --- say(A,L) --->   P2  ,  Ps  --- hear(A,L) --->  Ps2 .
 
-[ P | Ps ]  ---say(A,L)--->   Rs            :-   P  ---say(A,L)--->   P2  ,  Ps  ---hear(A,L)--->  Ps2 ,
+[ P | Ps ]  --- say(A,L) --->   Rs            :-   P  --- say(A,L) --->   P2  ,  Ps  --- hear(A,L) --->  Ps2 ,
                                                  flateval(P2, Ps2, Rs).
-[ P | Ps ]  ---say(A,L)--->   [ P2 | Ps2 ]  :-   P  ---hear(A,L)--->  P2  ,  Ps  ---say(A,L)--->   Ps2 .
-[ P | Ps ]  ---hear(A,L)--->  [ P2 | Ps2 ]  :-   P  ---hear(A,L)--->  P2  ,  Ps  ---hear(A,L)--->  Ps2 .
+[ P | Ps ]  --- say(A,L) --->   [ P2 | Ps2 ]  :-   P  --- hear(A,L) --->  P2  ,  Ps  --- say(A,L) --->   Ps2 .
+[ P | Ps ]  --- hear(A,L) --->  [ P2 | Ps2 ]  :-   P  --- hear(A,L) --->  P2  ,  Ps  --- hear(A,L) --->  Ps2 .
 
-[ P | Ps ]  ---say(A,L)--->   [ P  | Ps2 ]  :-   Ps  ---say(A,L)--->   Ps2  ,  ignore((A,L),P)    .
-[ P | Ps ]  ---hear(A,L)--->  [ P  | Ps2 ]  :-   Ps  ---hear(A,L)--->  Ps2  ,  ignore((A,L),P)    .
-% Think of   ignore((A,L),P) :- P---hear2(A,L)--->P.  But implemented to avoid overlap.
[…]

### 2015-04-02 prolog/test.pl

@@ -1,5 +1,7 @@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Static info examples %%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+:- consult('semantics.pl').
+
 e:p0:i ==> e:p3:i.
 _ ==> _ :- false.
 
@@ -16,18 +18,10 @@ server_call_point(r1:i, sync(o:p2)).
 
 %%%%%%%%%%%%%%%%
 
-example1(L,R) :- 
+example(L,R) :-
 %  Cont1 = fn(X,rte_exit(x, Cont2)),
 %  Cont2 = fn(X,rte_exit(x, fn(Y,return(void)))),
   Proc1 = rinst(r1:i,nil,[],rte_enter(x,_Cont1)),
   Proc2 = rinst(r2:i,nil,[],rte_enter(x,_Cont2)),
-  Block = [excl(x:i,free), Proc1, Proc2],
-  [[excl(z:i,free), []] | Block] ---L---> R.
-
-
-badexample(L,R) :- 
-  Proc1 = rinst(r1:i,nil,[],rte_enter(x,_Cont1)),
-  Proc2 = rinst(r2:i,nil,[],rte_enter(x,_Cont2)),
-  Block = [excl(x:i,free), Proc1, Proc2],
-  [Block | excl(z:i,free)] ---L---> R.
-
+  Block = [excl(x:i,free) | [Proc1 | [Proc2 | []]]],
+  [excl(z:i,free) | Block] --- L ---> R.

### 2015-04-02 prolog/tool/example.pl

@@ -1,0 +1,23 @@
+
+eq(X,X).
+parent(patrik,julia).
+parent(patrik,endre).
+parent(tünde,julia).
+parent(tünde,endre).
+sibling(X,Y) :- parent(P,X), parent(P,Y).
+cousing(X,Y) :- parent(P,X), parent(Q,Y), sibling(P,Q).
+
+% Test file to check the parser on simple inputs
+
+X ==> Y.
+X:Y --- L ---> P.
+a:b:c ==> d:e:f.
+a:a0 ==> X.
+_ ==> _.
+e:p0:i ==> e:p3:i.
+_ ==> _ :- false.
+event(r1:i, timing(300)).
+event(r1:i, data_received(e:p3)).
+X = [A|List].
+example1(L,R) :- A ---L ---> R.
+X :- A ; B.

### 2015-08-21 paper/paper.tex
Added a Prolog parser for testing
Started hand-converting from the Prolog file to LaTeX.
Combining reductions transformed to LaTeX
More text.

@@
 \usepackage[a4paper,margin=1.4cm]{geometry}
+\usepackage[fleqn]{amsmath}
+\usepackage{mathpartir}
 % \usepackage{minted}
@@
 \author{Johan Nordlander \and Patrik Jansson}
+
+\newcommand{\V}[1]{\mbox{\it #1}}
+\newcommand{\Code}{\V{Code}}
+\newcommand{\Cont}{\V{Cont}}
+\newcommand{\State}{\V{State}}
+\newcommand{\Act}{\V{Act}}
+\newcommand{\Upd}{\V{Upd}}
+\newcommand{\Xs}{\V{Xs}}
+\newcommand{\Vs}{\V{Vs}}
+\newcommand{\Cs}{\V{Cs}}
+
+\newcommand{\void}{\mbox{\bf void}}
+\newcommand{\nil}{\mbox{\bf nil}}
+\newcommand{\true}{\mbox{\bf true}}
+\newcommand{\false}{\mbox{\bf false}}
+\newcommand{\idle}{\mbox{\bf idle}}
+\newcommand{\pending}{\mbox{\bf pending}}
+\newcommand{\serving}[2]{\mbox{\bf serving}(#1,#2)}
+
+\newcommand{\seq}[2]{#1\hspace{-0.7ex}:\hspace{-0.7ex}#2}
+\newcommand{\seqq}[3]{#1\hspace{-0.7ex}:\hspace{-0.7ex}#2\hspace{-0.7ex}:\hspace{-0.7ex}#3}
+\newcommand{\sep}{\hspace{2ex}|\hspace{2ex}}
+
+\newcommand{\runnable}[4]{\mbox{\bf runnable}(#1,#2,#3,#4)}
+\newcommand{\rinst}[4]{\mbox{\bf rinst}(#1,#2,#3,#4)}
+\newcommand{\excl}[2]{\mbox{\bf excl}(#1,#2)}
+\newcommand{\irv}[2]{\mbox{\bf irv}(#1,#2)}
+\newcommand{\qelem}[2]{\mbox{\bf qelem}(#1,#2)}
+\newcommand{\delem}[3]{\mbox{\bf delem}(#1,#2,#3)}
+\newcommand{\opres}[2]{\mbox{\bf opres}(#1,#2)}
+\newcommand{\timer}[3]{\mbox{\bf timer}(#1,#2, #3)}
+
+\newcommand{\Enter}[2]{\mbox{rte\_Enter}(#1,#2)}
+\newcommand{\Exit}[2]{\mbox{rte\_Exit}(#1,#2)}
+\newcommand{\IrvWrite}[3]{\mbox{rte\_IrvWrite}(#1,#2,#3)}
+\newcommand{\IrvRead}[2]{\mbox{rte\_IrvRead}(#1,#2)}
+\newcommand{\Send}[4]{\mbox{rte\_Send}(#1,#2,#3,#4)}
+\newcommand{\Receive}[3]{\mbox{rte\_Receive}(#1,#2,#3)}
+\newcommand{\Write}[4]{\mbox{rte\_Write}(#1,#2,#3,#4)}
+\newcommand{\Read}[3]{\mbox{rte\_Read}(#1,#2,#3)}
+\newcommand{\Invalidate}[3]{\mbox{rte\_Invalidate}(#1,#2,#3)}
+\newcommand{\IsUpdated}[3]{\mbox{rte\_IsUpdated}(#1,#2,#3)}
+\newcommand{\Call}[4]{\mbox{rte\_Call}(#1,#2,#3,#4)}
+\newcommand{\Result}[3]{\mbox{rte\_Result}(#1,#2,#3)}
+\newcommand{\Return}[1]{\mbox{return}(#1)}
+
+\newcommand{\red}[1]{\xrightarrow{#1}}
+\newcommand{\say}[2]{#1 ! #2}
+\newcommand{\hear}[2]{#1 ? #2}
+
+\newcommand{\ent}{\mbox{\small ent}}
+\newcommand{\ex}{\mbox{\small ex}}
 
@@
 Unlike existing operating systems, however, AUTOSAR is not de facto defined in terms of a particular implementation.
-Instead, an explicit goal of AUTOSAR is to constitute an abstract specification that allows multiple competing realizations, and even systems built as an assembly of fragments from many different (and competing) vendors.
+Instead, an explicit goal of AUTOSAR is to constitute an abstract specification that allows multiple and competing realizations, and even systems built as an assembly of fragments from many different (and competing) vendors.
 Such a goal naturally puts the focus on the standard specification itself.
@@
 And because the standard is open to interpretation, the interoperability of software components across tools and platforms is often seriously hampered.
-What is more, a single software component cannot easily be studied and understood in isolation, since its interactive behavior is only indirectly defined in terms of the concrete C-code and OS tasks that realize it and every interacting component.
+What is more, a single software component cannot easily be studied and understood in isolation, since its interactive behavior is only indirectly defined in terms of the concrete C-code and OS tasks that realize it and interaction environment.
 
 This paper takes an important step towards a remedy to these problems, by contributing a formal specification of a substantial core of the AUTOSAR standard.
-The formalization covers most of the Software Component Template (SWC) and its accompanying Run-Time Environment (RTE) (Section~\ref{sec:Calc}), and is able to directly express every legal way a system of SWCs can evolve in at run-time on an arbitrarily fast platform (Section~\ref{sec:Sem}).
+The formalization covers most of the Software Component Template and its accompanying Run-Time Environment (RTE) (Section~\ref{sec:Calc}), and is able to directly express every legal way a system of SWCs can evolve in at run-time on an arbitrarily fast platform (Section~\ref{sec:Sem}).
 It can thus serve as a basis for both a concrete AUTOSAR implementation on a specific platform (whose supported behaviors must be a subset of those defined by the formal semantics), and a platform-independent AUTOSAR simulator (where behaviors from the legal set can be picked at random).
@@
 
-\section{Syntax and prelliminaries}
+\section{AUTOSAR}
+\label{sec:autosar}
+
+An AUTOSAR model is primarily a structure of interconnected {\bf software components} (SWCs).
+Links between SWCs are called {\bf connectors}, which attach to {\bf ports} exposed by the SWCs.
+A port is either {\bf required} or {\bf provided}, and then classified according to what {\bf interface} it exposes: {\bf sender-receiver}, {\bf client-server}, {\bf trigger} and {\bf mode-switch} are common interface types.
+A sender-receiver interface is an aggregation of one or more {\bf data-elements}, with each element characterized by the type of data it carries and optional further communication specification details.
+A client-server interface likewise consists of one or more {\bf operations}, that each give the signature of a procedure call and the possible errors that may result.
+
+Some components merely act as containers of other components structures, these are called {\bf composition} SWCs.
+In contrast, {\bf atomic} SWCs define their own behavior in terms of {\bf runnables} and {\bf inter-runnable variables}, primarily.
+A runnable denotes a sequential and (normally) terminating piece of code, although its actual behavioral definition is typically relegated to external C or Matlab files rather than being part of the AUTOSAR model proper.
+Instead, an AUTOSAR runnable  only provides constraints on what ports and inter-runnable variables the actual implementation can access; as well as {\bf event} declarations that determine under what conditions it will get executed.
+
+A runnable under execution is called a {\bf runnable instance}, and these may execute concurrently even if they belong to the same SWC.
+Each runnable indicates whether it may be instantiated concurrently with itself.
+Runnable code can also use {\bf exclusive areas} (a form of mutex semaphores) to further control concurrent behavior.
+
+To interact with ports, inter-runnable variables and exclusive areas, runnables utilize what is abstractly known as the {\bf virtual function bus} (VFB).
+Concretely, the VFB takes the shape of a C programming interface to the {\bf run-time environment} (RTE), custom-generated for each runnable on basis of its access constraints.
+Behind the RTE, a collection of OS kernels, communication stacks and other basic software modules implement the VBF services for the platform at hand.
+Complete AUTOSAR designs also typically contain {\bf ECU} and {\bf task assignments}, which are manually built tables that control how SWCs and runnables map onto the available hardware and OS resources, respectively.
+
+AUTOSAR models are completely static, which means that all components, runnables, ports, connectors, etc, are created ahead of execution time.
+Runnable instances are an exception, but they are only identified by AUTOSAR for conceptual purposes and never part of any model syntax.
+This static nature renders AUTOSAR models very suitable to graphical notations, which by far is the most widespread representation format for AUTOSAR models.
+An graphical model example is shown in Figure~\ref{fig:model}.
+
+\section{Syntax and preliminaries}
 \label{sec:Calc}
+
+The formal approach we have chosen is that of a typical process calculus \cite{}, where a system of concurrent processes evolves in a sequence of atomic steps, as determined by a global set of transition rules.
+Our process terms correspond to the individual state-carrying parts of an AUTOSAR system, like the runnables, runnable instances, inter-runnable variables and port elements.
+Each such term is also assigned a constant address parameter for identification purposes.
+
+Components and ports do not carry any dynamic state besides the state of their constituent parts, so they need no explicit representation as process terms.
+However, in order to still be able to talk about model elements using their locally scoped names, we employ a hierarchical addressing scheme, such that address $\seq{R}{I} $ means runnable $R$ of component $I$ and $\seqq{E}{P}{I}$ means element $E$ of port $P$ within component $I$.
+The following meta-variable conventions will be used:
+$$
+\begin{array}{rcll}
+	I	& \in		& \mbox{Component names} \\
+	R	& \in		& \mbox{Runnable names} \\
+	U	& \in		& \mbox{Port names} \\
+	E	& \in		& \mbox{Sender-receiver element names} \\
+	O	& \in		& \mbox{Client-server operation names} \\
+	S	& \in		& \mbox{Inter-runnable variable names} \\
+	X	& \in		& \mbox{Exclusive area names} \\
+	\\
+	A	& \in		& \mbox{Addresses} \\
+	A	& ::=		& \seq{R}{I}   \sep   \seqq{E}{P}{I}   \sep   \seqq{O}{U}{I}   \sep   \seq{S}{I}   \sep   \seq{X}{I}   \\
+\end{array}
+$$
+Since AUTOSAR components can be arbitrarily nested, the component names $I$ can also be considered to have a similarly nested internal structure (component $I_1$ within component $I_2$ within component $I_3$, etc).
+We can fully ignore this detail here, though, as the hierarchical layout of components in an AUTOSAR system has no run-time implications \cite{}.
+
+The process terms of our calculus are as follows:
+\begin{itemize}
+\item $\runnable{\seq{R}{I}}{T}{\Act}{N}$  \newline
+Common state for runnable $R$ of component $I$, showing $N$ current instances, $T$ seconds left until next possible instantiation, and activation state $\Act$.
+\item $\rinst{\seq{R}{I}}{C}{\Xs}{\Code}$ \newline
+An instance of runnable $R$ of component $I$, currently executing $\Code$ and owning the exclusive areas $\Xs$, possibly running on behalf of client $C$.
+\item $\excl{\seq{X}{I}}{B}$ \newline
+Exclusive area $X$ of component $I$, with boolean occupation state $B$.
+\item $\irv{\seq{S}{I}}{V}$ \newline
+Inter-runnable variable $S$ of component $I$, currently holding value $V$.
+\item $\delem{\seqq{E}{U}{I}}{B}{V}$ \newline
+Non-queued sender-receiver data element $E$ in port $U$ of component $I$, currently holding value $V$ with boolean update status flag $B$.
+\item $\qelem{\seqq{E}{U}{I}}{\Vs}$ \newline
+Queued sender-receiver data element $E$ in port $U$ of component $I$, currently holding value sequence $\Vs$.
+\item $\opres{\seqq{O}{U}{I}}{\Vs}$ \newline
+Client-server operation $O$ in port $U$ of component $I$, currently holding result value sequence $\Vs$.
+\item $\timer{\seq{R}{I}}{T_p}{T}$ \newline
+Timer state for runnable $R$ of component $I$ and period $T_p$, with $T$ seconds left of its current cycle.
+\end{itemize}
+
+In the untyped setting of this report, a value $V$ can be any data item computed and communicated by an AUTOSAR system, including the unit value $\void$ and the error codes that might be returned from RTE calls.
+Meta-variables $N$, $B$ and $T$ range over natural numbers, boolean values and (floating point) time values, respectively.
+The client $C$ of a runnable instance is the address of the invoking port operation (if the runnable was invoked by a client-server call), or $\void$ otherwise.
+An activation state $\Act$ either toggles between $\idle$ and $\pending$, or it has the special shape $\serving{\Cs}{\Vs}$ (in the case of a runnable triggered by client-server calls).
+Formally,
+$$
+\begin{array}{rcll}
+	V	& \in		& \mbox{Values} \\
+	N	& \in		& \mbox{Natural numbers} \\
+	B	& \in		& \mbox{Boolean values} \\
+	T	& \in		& \mbox{Time values} \\
+	C	& ::=		& \seqq{O}{U}{I} \sep \void \\
+	\Act	& ::=		& \idle \sep \pending \sep \serving{\Cs}{\Vs} \\
+\end{array}
+$$
+We write $\Vs$ for a sequence of values, $\Cs$ for a sequence of clients, etc.
+Sequences may be empty, which is written $\epsilon$.
+Left (right) concatenation of an element with a sequence is written $\seq{V}{\Vs}$ ($\seq{\Vs\,}{V}$).
+
+The code part of a runnable instance should technically be a representation of the C or Matlab function that must accompany an AUTOSAR runnable definition once it is complete.
+However, since our task in this report is not to investigate the semantics of these languages in every detail, a more abstract notion of executable code will be required.
+To this end, we have chosen to ignore all details of internal computations, and just capture the observable effects of runnable execution as a pure sequence of RTE (VFB) calls.
+Because the result from an RTE call may in general affect subsequent runnable behavior, we use a continuation-passing style to allow any data-dependencies to be expressed.
+$$
+\begin{array}{lll}
+	\Code	& ::=	& \Enter{X}{\Cont}			\\
+			& |	& \Exit{X}{\Cont}			\\
+			& |	& \IrvWrite{S}{V}{\Cont}		\\
+			& |	& \IrvRead{S}{\Cont}			\\
+			& |	& \Send{E}{U}{V}{\Cont}		\\
+			& |	& \Receive{E}{U}{\Cont}		\\
+			& |	& \Write{E}{U}{V}{\Cont}		\\
+			& |	& \Read{E}{U}{\Cont}		\\
+			& |	& \Invalidate{E}{U}{\Cont}		\\
+			& |	& \IsUpdated{E}{U}{\Cont}		\\
+			& |	& \Call{O}{U}{V}{\Cont}		\\
+			& |	& \Result{O}{U}{\Cont}		\\
+			& |	& \Return{V}				\\
+	\\
+	\Cont	& \in	& V \rightarrow \Code		\\
+\end{array}
+$$
+
+Together with an operator for composing process terms in parallel, the process grammar for our AUTOSAR calculus becomes
+$$
+\begin{array}{rcll}
+	P,Q 	& ::= &
+		\runnable{A}{T}{\Act}{N}	\\
+		& | &
+		\rinst{A}{C}{\Xs}{\Code}	\\
+		& | &
+		\excl{A}{B}			\\
+		& | &
+		\irv{A}{V}				\\
+		& | &
+		\delem{A}{B}{V}			\\
+		& | &
+		\qelem{A}{\Vs}			\\
+		& | &
+		\opres{A}{\Vs}			\\
+		& | &
+		\timer{A}{T_p}{T}		\\
+		& | &
+		P|Q					\\
+\end{array}
+$$
 
@@
 
-TODO: links from rules and choices made here to the AUTSAR spec
+Our semantic approach defines the meaning of an AUTOSAR system as the possible chains of state changes that can be applied to the system's initial state; or equivalently, as the set of {\em traces} that can be generated from the corresponding process term $P_0$.
+A trace is a possibly infinite sequence of transitions
+$$
+\begin{array}{c}
+	P_0 \red{L_1} P_1 \red{L_2} P_2 \red{L_3} \cdots
+\end{array}
+$$
+where each step
+$$
+\begin{array}{c}
+	P_{i-1} \red{L_i} P_i
+\end{array}
+$$
+states that the system state captured as process term $P_{i-1}$ can evolve into state $P_i$ by means of a single transition labelled $L_i$.
+
+The label $L$ of a transition can be ow two kinds, indicating that the affected process term either "says" or "hears" something during the transition.
+A label also always consists of an address $A$ paired with some additional detail $D$, as captured in the following label grammar:
+$$
+\begin{array}{rcll}
+	L	& ::=		& A!D		& \hspace{4em}\mbox{Say $A$ and $D$} \\
+		&		& A?D		& \hspace{4em}\mbox{Hear $A$ and $D$} \\
+\end{array}
+$$
+The possible forms of the extra label information $D$ will be introduced as the different transition rules are defined.
+
+Two processes $P$ and $Q$ can make transitions in parallel if they agree on what is being said or heard.
+Just as the saying/hearing intuition suggests, at most one of the processes can have the saying role in such an agreement.
+The following set of inference rules capture this intuition formally.
+\begin{mathpar}
+	\inferrule
+		{P \red{A!D} P' \\ Q \red{A?D} Q'}
+		{P | Q \red{A!D} P' | Q'}
+	\and
+	\inferrule
+		{P \red{A?D} P' \\ Q \red{A!D} Q'}
+		{P | Q \red{A!D} P' | Q'}
+	\and
+	\inferrule
+		{P \red{A?D} P' \\ Q \red{A?D} Q'}
+		{P | Q \red{A?D} P' | Q'}	
+\end{mathpar}
+
+The effect of these agreement rules closely resembles the idea of a {\em broadcast}: what one process says may be heard by every other process in a large parallel composition.
+This behavior serves us well, because in general, what one AUTOSAR runnable does may have impact on many -- if not all -- parts of the executing system.
+It is however important not to confuse this broadcasting notion with any particular form of AUTOSAR communication.
+The way transition labels distribute over a parallel system is merely a technical aspect of our process calculus, and will be used to express several different AUTOSAR communication semantics, among other things.
+
+The initial process $P_0$ is determined fully by the AUTOSAR model under study, as the parallel composition of the following terms:
+\begin{enumerate}
+
+\item For each runnable $R$ of each component $I$, a term
+$$
+\begin{array}{c}
+	\runnable{\seq{R}{I}}{0.0}{\Act}{0}
+\end{array}
+$$
+where $\Act$ is $\serving{\epsilon}{\epsilon}$ if $R$ is triggered by an {\em OperationInvokedEvent}; else $\Act$ is $\pending$ if $R$ is triggered by an {\em InitEvent}; otherwise $\Act$ is $\idle$.
+% SWS_Rte_03524 (SWS_Rte_03526?) forbids mixing OperationInvokedEvents with other events for the same runnable
+
+\item For each exclusive area $X$ of each atomic component $I$, a term
+$$
+\begin{array}{c}
+	\excl{\seq{X}{I}}{\false}
+\end{array}
+$$
+
+\item For each inter-runnable variable $S$ of each atomic component $I$, a term
+$$
+\begin{array}{c}
+	\irv{\seq{S}{I}}{V}
+\end{array}
+$$
+where $V$ is the initial value attributed to $S$ if it exists, otherwise $V$ is $\void$.
+
+\item For each data element $E$ of each required port $U$ of each atomic component $I$, a term
+$$
+\begin{array}{c}
+	\qelem{\seqq{E}{U}{I}}{\epsilon}
+\end{array}
+$$
+if the {\em swImplPolicy} attribute of element $E$ is {\em queued}; otherwise, a term
+$$
+\begin{array}{c}
+	\delem{\seqq{E}{U}{I}}{\false}{\mbox{NO\_DATA}}
+\end{array}
+$$
+
+\item For each operation $O$ of each required port $U$ of each atomic component $I$, a term
+$$
+\begin{array}{c}
+	\opres{\seqq{O}{U}{I}}{\epsilon}
+\end{array}
+$$
+
+\item For each {\em TimingEvent} triggering each runnable $R$ of each atomic component $I$, a term
+$$
+\begin{array}{c}
+	\timer{\seq{R}{I}}{T}{0.0}
+\end{array}
+$$
+where $T$ is the period of the event.
+
+\end{enumerate}
+
+Some further static model information will also be referenced by the transition rules. We refer to the {\em connectors} of a model using a binary relation $\Rightarrow$ on addresses, defined as
+$$
+	\seqq{E}{U_1}{I_1} \Rightarrow \seqq{E}{U_2}{I_2} \;\;\mbox{for all elements $E$ of port $U_2$}
+$$
+if there is a connector in the model from provided sender-receiver port $U_1$ of component $I_1$ to required port $U_2$ of component $I_2$; and
+$$
+	\seqq{O}{U_1}{I_1} \Rightarrow \seqq{O}{U_2}{I_2} \;\;\mbox{for all operations $O$ of port $U_1$}
+$$
+if there is a connector in the model from required client-server port $U_1$ of component $I_1$ to provided port $U_2$ of component $I_2$.
+
+We expect the {\em implementation} of a runnable to be available as a continuation $\Cont$, even though the real AUTOSAR model will be refering to concrete functions in external C/Matlab files.
+We will further abstract away from actual numbers of input and output parameters to these function by assuming that records (structs) are used as the single continuation argument and result whenever the arity so requires.
+Absent arguments and results will be replaced by the unit value $\void$.
+
+\subsection{Exclusive areas}
+
+\begin{gather*}
+	\rinst{\seq{R}{I}}{C}{\Xs}{\Enter{X}{\Cont}}  \red{\say{X:I}{\ent}}  \\
+		\rinst{\seq{R}{I}}{C}{\seq X \Xs}{\Cont(\void)} \\
+	\\
+	\rinst{\seq{R}{I}}{C}{\seq X \Xs}{\Exit{X}{\Cont}}  \red{\say{X:I}{\ex}}  \\
+		\rinst{\seq{R}{I}}{C}{\Xs}{\Cont(\void)} \\
+	\\
+	\excl{\seq X I}{\false} \red{\hear{X:I}{\ent}} \excl{\seq X I}{\true} \\
+	\\
+	\excl{\seq X I}{\true} \red{\hear{X:I}{\ex}} \excl{\seq X I}{\false} \\
+	\\
+\end{gather*}
+
+\subsection{Inter-runnable variables}
+
+\begin{gather*}
+	\rinst{\seq{R}{I}}{C}{\Xs}{\IrvWrite{X}{\Cont}}  \red{\say{X:I}{\ent}}  \\
+		\rinst{\seq{R}{I}}{C}{\seq X \Xs}{\Cont(\void)} \\
+	\\
+	\rinst{\seq{R}{I}}{C}{\seq X \Xs}{\Exit{X}{\Cont}}  \red{\say{X:I}{\ex}}  \\
+		\rinst{\seq{R}{I}}{C}{\Xs}{\Cont(\void)} \\
+	\\
+	\excl{\seq X I}{\false} \red{\hear{X:I}{\ent}} \excl{\seq X I}{\true} \\
+	\\
+	\excl{\seq X I}{\true} \red{\hear{X:I}{\ex}} \excl{\seq X I}{\false} \\
+	\\
+\end{gather*}
+
+TODO: links from rules and choices made here to the AUTOSAR spec
 
@@
 
+irv without initValue vs.\ delem without initValue -- "undefined" vs.\ NO\_DATA at initial read.
+
+Mixing OperationInvokedEvents with others for same runnable...
+
+Multiple timingEvent for same runnable, possibly with same period?
+
+Exclusive area nesting
+
+Return value of qelem write with fanout > 1 and *some* LIMIT failures
+
+Timer initial offsets
+
 \section{Conclusions and Future Work}

### 2015-08-21 paper/paper.tex

@@
 
-\newcommand{\seq}[2]{#1\hspace{-0.7ex}:\hspace{-0.7ex}#2}
-\newcommand{\seqq}[3]{#1\hspace{-0.7ex}:\hspace{-0.7ex}#2\hspace{-0.7ex}:\hspace{-0.7ex}#3}
+\newcommand{\adr}[1]{\mbox{\em #1}}
 \newcommand{\sep}{\hspace{2ex}|\hspace{2ex}}
@@
 \newcommand{\ex}{\mbox{\small ex}}
+
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{4pt plus 2pt minus 1pt}
 
@@
 Components and ports do not carry any dynamic state besides the state of their constituent parts, so they need no explicit representation as process terms.
-However, in order to still be able to talk about model elements using their locally scoped names, we employ a hierarchical addressing scheme, such that address $\seq{R}{I} $ means runnable $R$ of component $I$ and $\seqq{E}{P}{I}$ means element $E$ of port $P$ within component $I$.
+However, in order to still be able to talk about model elements using their locally scoped names, we employ a hierarchical addressing scheme, such that address $\adr{R:I} $ means runnable $R$ of component $I$ and $\adr{E:P:I}$ means element $E$ of port $P$ within component $I$.
 The following meta-variable conventions will be used:
@@
 	A	& \in		& \mbox{Addresses} \\
-	A	& ::=		& \seq{R}{I}   \sep   \seqq{E}{P}{I}   \sep   \seqq{O}{U}{I}   \sep   \seq{S}{I}   \sep   \seq{X}{I}   \\
+	A	& ::=		& \adr{R:I}   \sep   \adr{E:P:I}   \sep   \adr{O:U:I}   \sep   \adr{S:I}   \sep   \adr{X:I}   \\
 \end{array}
@@
 \begin{itemize}
-\item $\runnable{\seq{R}{I}}{T}{\Act}{N}$  \newline
+\item $\runnable{\adr{R:I}}{T}{\Act}{N}$  \newline
 Common state for runnable $R$ of component $I$, showing $N$ current instances, $T$ seconds left until next possible instantiation, and activation state $\Act$.
-\item $\rinst{\seq{R}{I}}{C}{\Xs}{\Code}$ \newline
+\item $\rinst{\adr{R:I}}{C}{\Xs}{\Code}$ \newline
 An instance of runnable $R$ of component $I$, currently executing $\Code$ and owning the exclusive areas $\Xs$, possibly running on behalf of client $C$.
-\item $\excl{\seq{X}{I}}{B}$ \newline
+\item $\excl{\adr{X:I}}{B}$ \newline
 Exclusive area $X$ of component $I$, with boolean occupation state $B$.
-\item $\irv{\seq{S}{I}}{V}$ \newline
+\item $\irv{\adr{S:I}}{V}$ \newline
 Inter-runnable variable $S$ of component $I$, currently holding value $V$.
-\item $\delem{\seqq{E}{U}{I}}{B}{V}$ \newline
+\item $\delem{\adr{E:U:I}}{B}{V}$ \newline
 Non-queued sender-receiver data element $E$ in port $U$ of component $I$, currently holding value $V$ with boolean update status flag $B$.
-\item $\qelem{\seqq{E}{U}{I}}{\Vs}$ \newline
+\item $\qelem{\adr{E:U:I}}{\Vs}$ \newline
 Queued sender-receiver data element $E$ in port $U$ of component $I$, currently holding value sequence $\Vs$.
-\item $\opres{\seqq{O}{U}{I}}{\Vs}$ \newline
+\item $\opres{\adr{O:U:I}}{\Vs}$ \newline
 Client-server operation $O$ in port $U$ of component $I$, currently holding result value sequence $\Vs$.
-\item $\timer{\seq{R}{I}}{T_p}{T}$ \newline
+\item $\timer{\adr{R:I}}{T_p}{T}$ \newline
 Timer state for runnable $R$ of component $I$ and period $T_p$, with $T$ seconds left of its current cycle.
@@
 	T	& \in		& \mbox{Time values} \\
-	C	& ::=		& \seqq{O}{U}{I} \sep \void \\
+	C	& ::=		& \adr{O:U:I} \sep \void \\
 	\Act	& ::=		& \idle \sep \pending \sep \serving{\Cs}{\Vs} \\
@@
 Sequences may be empty, which is written $\epsilon$.
-Left (right) concatenation of an element with a sequence is written $\seq{V}{\Vs}$ ($\seq{\Vs\,}{V}$).
+Left (right) concatenation of an element with a sequence is written $\adr{V:Vs}$ ($\adr{Vs:V}$).
 
@@
 \begin{array}{c}
-	\runnable{\seq{R}{I}}{0.0}{\Act}{0}
+	\runnable{\adr{R:I}}{0.0}{\Act}{0}
 \end{array}
@@
 \begin{array}{c}
-	\excl{\seq{X}{I}}{\false}
+	\excl{\adr{X:I}}{\false}
 \end{array}
@@
 \begin{array}{c}
-	\irv{\seq{S}{I}}{V}
+	\irv{\adr{S:I}}{V}
 \end{array}
@@
 \begin{array}{c}
-	\qelem{\seqq{E}{U}{I}}{\epsilon}
+	\qelem{\adr{E:U:I}}{\epsilon}
 \end{array}
@@
 \begin{array}{c}
-	\delem{\seqq{E}{U}{I}}{\false}{\mbox{NO\_DATA}}
+	\delem{\adr{E:U:I}}{\false}{\mbox{NO\_DATA}}
 \end{array}
@@
 \begin{array}{c}
-	\opres{\seqq{O}{U}{I}}{\epsilon}
+	\opres{\adr{O:U:I}}{\epsilon}
 \end{array}
@@
 \begin{array}{c}
-	\timer{\seq{R}{I}}{T}{0.0}
+	\timer{\adr{R:I}}{T}{0.0}
 \end{array}
@@
 $$
-	\seqq{E}{U_1}{I_1} \Rightarrow \seqq{E}{U_2}{I_2} \;\;\mbox{for all elements $E$ of port $U_2$}
+	\adr{E:$U_1$:$I_1$} \Rightarrow \adr{E:$U_2$:$I_2$} \;\;\mbox{for all elements $E$ of port $U_2$}
 $$
@@
 $$
-	\seqq{O}{U_1}{I_1} \Rightarrow \seqq{O}{U_2}{I_2} \;\;\mbox{for all operations $O$ of port $U_1$}
+	\adr{O:$U_1$:$I_1$} \Rightarrow \adr{O:$U_2$:$I_2$} \;\;\mbox{for all operations $O$ of port $U_1$}
 $$
@@
 \begin{gather*}
-	\rinst{\seq{R}{I}}{C}{\Xs}{\Enter{X}{\Cont}}  \red{\say{X:I}{\ent}}  \\
-		\rinst{\seq{R}{I}}{C}{\seq X \Xs}{\Cont(\void)} \\
+	\rinst{\adr{R:I}}{C}{\Xs}{\Enter{X}{\Cont}}  \red{\say{X:I}{\ent}}  \\
+		\rinst{\adr{R:I}}{C}{\adr{X:Xs}}{\Cont(\void)} \\
 	\\
-	\rinst{\seq{R}{I}}{C}{\seq X \Xs}{\Exit{X}{\Cont}}  \red{\say{X:I}{\ex}}  \\
-		\rinst{\seq{R}{I}}{C}{\Xs}{\Cont(\void)} \\
+	\rinst{\adr{R:I}}{C}{\adr{X:Xs}}{\Exit{X}{\Cont}}  \red{\say{X:I}{\ex}}  \\
+		\rinst{\adr{R:I}}{C}{\Xs}{\Cont(\void)} \\
 	\\
-	\excl{\seq X I}{\false} \red{\hear{X:I}{\ent}} \excl{\seq X I}{\true} \\
+	\excl{\adr{X:I}}{\false} \red{\hear{X:I}{\ent}} \excl{\adr{X:I}}{\true} \\
 	\\
-	\excl{\seq X I}{\true} \red{\hear{X:I}{\ex}} \excl{\seq X I}{\false} \\
+	\excl{\adr{X:I}}{\true} \red{\hear{X:I}{\ex}} \excl{\adr{X:I}}{\false} \\
 	\\
@@
 \begin{gather*}
-	\rinst{\seq{R}{I}}{C}{\Xs}{\IrvWrite{X}{\Cont}}  \red{\say{X:I}{\ent}}  \\
-		\rinst{\seq{R}{I}}{C}{\seq X \Xs}{\Cont(\void)} \\
+	\rinst{\adr{R:I}}{C}{\Xs}{\IrvWrite{X}{\Cont}}  \red{\say{X:I}{\ent}}  \\
+		\rinst{\adr{R:I}}{C}{\adr{X:Xs}}{\Cont(\void)} \\
 	\\
-	\rinst{\seq{R}{I}}{C}{\seq X \Xs}{\Exit{X}{\Cont}}  \red{\say{X:I}{\ex}}  \\
-		\rinst{\seq{R}{I}}{C}{\Xs}{\Cont(\void)} \\
+	\rinst{\adr{R:I}}{C}{\adr{X:Xs}}{\Exit{X}{\Cont}}  \red{\say{X:I}{\ex}}  \\
+		\rinst{\adr{R:I}}{C}{\Xs}{\Cont(\void)} \\
 	\\
-	\excl{\seq X I}{\false} \red{\hear{X:I}{\ent}} \excl{\seq X I}{\true} \\
+	\excl{\adr{X:I}}{\false} \red{\hear{X:I}{\ent}} \excl{\adr{X:I}}{\true} \\
 	\\
-	\excl{\seq X I}{\true} \red{\hear{X:I}{\ex}} \excl{\seq X I}{\false} \\
+	\excl{\adr{X:I}}{\true} \red{\hear{X:I}{\ex}} \excl{\adr{X:I}}{\false} \\
 	\\
@@
 
-\verbatiminput{../prolog/semantics.pl}
+\input{semantics.tex}
+
+% \verbatiminput{../prolog/semantics.pl}
 

### 2015-08-22 paper/paper.tex
Some LaTeX formatting

@@
 
-\newcommand{\V}[1]{\mbox{\it #1}}
-\newcommand{\Code}{\V{Code}}
-\newcommand{\Cont}{\V{Cont}}
+\newcommand{\V}[1]{\text{\it #1}}
+\newcommand{\Code} {\V{Code}}
+\newcommand{\Cont} {\V{Cont}}
 \newcommand{\State}{\V{State}}
-\newcommand{\Act}{\V{Act}}
-\newcommand{\Upd}{\V{Upd}}
-\newcommand{\Xs}{\V{Xs}}
-\newcommand{\Vs}{\V{Vs}}
-\newcommand{\Cs}{\V{Cs}}
+\newcommand{\Act}  {\V{Act}}
+\newcommand{\Upd}  {\V{Upd}}
+\newcommand{\Xs}   {\V{Xs}}
+\newcommand{\Vs}   {\V{Vs}}
+\newcommand{\Cs}   {\V{Cs}}
 
-\newcommand{\void}{\mbox{\bf void}}
-\newcommand{\nil}{\mbox{\bf nil}}
-\newcommand{\true}{\mbox{\bf true}}
-\newcommand{\false}{\mbox{\bf false}}
-\newcommand{\idle}{\mbox{\bf idle}}
-\newcommand{\pending}{\mbox{\bf pending}}
-\newcommand{\serving}[2]{\mbox{\bf serving}(#1,#2)}
+\newcommand{\kw}[1]{\textbf{#1}}
+\newcommand{\void}    {\kw{void}}
+\newcommand{\nil}     {\kw{nil}}
+\newcommand{\true}    {\kw{true}}
+\newcommand{\false}   {\kw{false}}
+\newcommand{\idle}    {\kw{idle}}
+\newcommand{\pending} {\kw{pending}}
+\newcommand{\serving}[2]{\kw{serving}(#1,#2)}
 
-\newcommand{\adr}[1]{\mbox{\em #1}}
+\newcommand{\adr}[1]{\emph{#1}}
 \newcommand{\sep}{\hspace{2ex}|\hspace{2ex}}
 
-\newcommand{\runnable}[4]{\mbox{\bf runnable}(#1,#2,#3,#4)}
-\newcommand{\rinst}[4]{\mbox{\bf rinst}(#1,#2,#3,#4)}
-\newcommand{\excl}[2]{\mbox{\bf excl}(#1,#2)}
-\newcommand{\irv}[2]{\mbox{\bf irv}(#1,#2)}
-\newcommand{\qelem}[2]{\mbox{\bf qelem}(#1,#2)}
-\newcommand{\delem}[3]{\mbox{\bf delem}(#1,#2,#3)}
-\newcommand{\opres}[2]{\mbox{\bf opres}(#1,#2)}
-\newcommand{\timer}[3]{\mbox{\bf timer}(#1,#2, #3)}
+\newcommand{\runnable}[4]{\kw{runnable}(#1,#2,#3,#4)}
+\newcommand{\rinst}   [4]{\kw{rinst}(#1,#2,#3,#4)}
+\newcommand{\excl}    [2]{\kw{excl}(#1,#2)}
+\newcommand{\irv}     [2]{\kw{irv}(#1,#2)}
+\newcommand{\qelem}   [2]{\kw{qelem}(#1,#2)}
+\newcommand{\delem}   [3]{\kw{delem}(#1,#2,#3)}
+\newcommand{\opres}   [2]{\kw{opres}(#1,#2)}
+\newcommand{\timer}   [3]{\kw{timer}(#1,#2, #3)}
 
-\newcommand{\Enter}[2]{\mbox{rte\_Enter}(#1,#2)}
-\newcommand{\Exit}[2]{\mbox{rte\_Exit}(#1,#2)}
-\newcommand{\IrvWrite}[3]{\mbox{rte\_IrvWrite}(#1,#2,#3)}
-\newcommand{\IrvRead}[2]{\mbox{rte\_IrvRead}(#1,#2)}
-\newcommand{\Send}[4]{\mbox{rte\_Send}(#1,#2,#3,#4)}
-\newcommand{\Receive}[3]{\mbox{rte\_Receive}(#1,#2,#3)}
-\newcommand{\Write}[4]{\mbox{rte\_Write}(#1,#2,#3,#4)}
-\newcommand{\Read}[3]{\mbox{rte\_Read}(#1,#2,#3)}
-\newcommand{\Invalidate}[3]{\mbox{rte\_Invalidate}(#1,#2,#3)}
-\newcommand{\IsUpdated}[3]{\mbox{rte\_IsUpdated}(#1,#2,#3)}
-\newcommand{\Call}[4]{\mbox{rte\_Call}(#1,#2,#3,#4)}
-\newcommand{\Result}[3]{\mbox{rte\_Result}(#1,#2,#3)}
-\newcommand{\Return}[1]{\mbox{return}(#1)}
+\newcommand{\rte}[1]{\text{rte\_#1}}
+\newcommand{\Enter}     [2]{\rte{Enter}(#1,#2)}
+\newcommand{\Exit}      [2]{\rte{Exit}(#1,#2)}
+\newcommand{\IrvWrite}  [3]{\rte{IrvWrite}(#1,#2,#3)}
+\newcommand{\IrvRead}   [2]{\rte{IrvRead}(#1,#2)}
+\newcommand{\Send}      [4]{\rte{Send}(#1,#2,#3,#4)}
+\newcommand{\Receive}   [3]{\rte{Receive}(#1,#2,#3)}
+\newcommand{\Write}     [4]{\rte{Write}(#1,#2,#3,#4)}
+\newcommand{\Read}      [3]{\rte{Read}(#1,#2,#3)}
+\newcommand{\Invalidate}[3]{\rte{Invalidate}(#1,#2,#3)}
+\newcommand{\IsUpdated} [3]{\rte{IsUpdated}(#1,#2,#3)}
+\newcommand{\Call}      [4]{\rte{Call}(#1,#2,#3,#4)}
+\newcommand{\Result}    [3]{\rte{Result}(#1,#2,#3)}
+\newcommand{\Return}    [1]{\text{return}(#1)}
 
@@
 
-\newcommand{\ent}{\mbox{\small ent}}
-\newcommand{\ex}{\mbox{\small ex}}
+\newcommand{\ent}{\text{\small ent}}
+\newcommand{\ex}{\text{\small ex}}
 
@@
 
-An AUTOSAR model is primarily a structure of interconnected {\bf software components} (SWCs).
-Links between SWCs are called {\bf connectors}, which attach to {\bf ports} exposed by the SWCs.
-A port is either {\bf required} or {\bf provided}, and then classified according to what {\bf interface} it exposes: {\bf sender-receiver}, {\bf client-server}, {\bf trigger} and {\bf mode-switch} are common interface types.
-A sender-receiver interface is an aggregation of one or more {\bf data-elements}, with each element characterized by the type of data it carries and optional further communication specification details.
-A client-server interface likewise consists of one or more {\bf operations}, that each give the signature of a procedure call and the possible errors that may result.
+An AUTOSAR model is primarily a structure of interconnected \textbf{software components} (SWCs).
+Links between SWCs are called \textbf{connectors}, which attach to \textbf{ports} exposed by the SWCs.
+A port is either \textbf{required} or \textbf{provided}, and then classified according to what \textbf{interface} it exposes: \textbf{sender-receiver}, \textbf{client-server}, \textbf{trigger} and \textbf{mode-switch} are common interface types.
+A sender-receiver interface is an aggregation of one or more \textbf{data-elements}, with each element characterized by the type of data it carries and optional further communication specification details.
+A client-server interface likewise consists of one or more \textbf{operations}, that each give the signature of a procedure call and the possible errors that may result.
 
-Some components merely act as containers of other components structures, these are called {\bf composition} SWCs.
-In contrast, {\bf atomic} SWCs define their own behavior in terms of {\bf runnables} and {\bf inter-runnable variables}, primarily.
+Some components merely act as containers of other components structures, these are called \textbf{composition} SWCs.
+In contrast, \textbf{atomic} SWCs define their own behavior in terms of \textbf{runnables} and \textbf{inter-runnable variables}, primarily.
 A runnable denotes a sequential and (normally) terminating piece of code, although its actual behavioral definition is typically relegated to external C or Matlab files rather than being part of the AUTOSAR model proper.
-Instead, an AUTOSAR runnable  only provides constraints on what ports and inter-runnable variables the actual implementation can access; as well as {\bf event} declarations that determine under what conditions it will get executed.
+Instead, an AUTOSAR runnable  only provides constraints on what ports and inter-runnable variables the actual implementation can access; as well as \textbf{event} declarations that determine under what conditions it will get executed.
 
-A runnable under execution is called a {\bf runnable instance}, and these may execute concurrently even if they belong to the same SWC.
+A runnable under execution is called a \textbf{runnable instance}, and these may execute concurrently even if they belong to the same SWC.
 Each runnable indicates whether it may be instantiated concurrently with itself.
-Runnable code can also use {\bf exclusive areas} (a form of mutex semaphores) to further control concurrent behavior.
+Runnable code can also use \textbf{exclusive areas} (a form of mutex semaphores) to further control concurrent behavior.
 
-To interact with ports, inter-runnable variables and exclusive areas, runnables utilize what is abstractly known as the {\bf virtual function bus} (VFB).
-Concretely, the VFB takes the shape of a C programming interface to the {\bf run-time environment} (RTE), custom-generated for each runnable on basis of its access constraints.
+To interact with ports, inter-runnable variables and exclusive areas, runnables utilize what is abstractly known as the \textbf{virtual function bus} (VFB).
+Concretely, the VFB takes the shape of a C programming interface to the \textbf{run-time environment} (RTE), custom-generated for each runnable on basis of its access constraints.
 Behind the RTE, a collection of OS kernels, communication stacks and other basic software modules implement the VBF services for the platform at hand.
-Complete AUTOSAR designs also typically contain {\bf ECU} and {\bf task assignments}, which are manually built tables that control how SWCs and runnables map onto the available hardware and OS resources, respectively.
+Complete AUTOSAR designs also typically contain \textbf{ECU} and \textbf{task assignments}, which are manually built tables that control how SWCs and runnables map onto the available hardware and OS resources, respectively.
 
@@
 Components and ports do not carry any dynamic state besides the state of their constituent parts, so they need no explicit representation as process terms.
-However, in order to still be able to talk about model elements using their locally scoped names, we employ a hierarchical addressing scheme, such that address $\adr{R:I} $ means runnable $R$ of component $I$ and $\adr{E:P:I}$ means element $E$ of port $P$ within component $I$.
+However, in order to still be able to talk about model elements using their locally scoped names, we employ a hierarchical addressing scheme, such that address $\adr{R:I}$ means runnable $R$ of component $I$ and $\adr{E:P:I}$ means element $E$ of port $P$ within component $I$.
 The following meta-variable conventions will be used:
@@
 \begin{array}{rcll}
-	I	& \in		& \mbox{Component names} \\
-	R	& \in		& \mbox{Runnable names} \\
-	U	& \in		& \mbox{Port names} \\
-	E	& \in		& \mbox{Sender-receiver element names} \\
-	O	& \in		& \mbox{Client-server operation names} \\
-	S	& \in		& \mbox{Inter-runnable variable names} \\
-	X	& \in		& \mbox{Exclusive area names} \\
+	I	& \in		& \text{Component names} \\
+	R	& \in		& \text{Runnable names} \\
+	U	& \in		& \text{Port names} \\
+	E	& \in		& \text{Sender-receiver element names} \\
+	O	& \in		& \text{Client-server operation names} \\
+	S	& \in		& \text{Inter-runnable variable names} \\
+	X	& \in		& \text{Exclusive area names} \\
 	\\
-	A	& \in		& \mbox{Addresses} \\
+	A	& \in		& \text{Addresses} \\
 	A	& ::=		& \adr{R:I}   \sep   \adr{E:P:I}   \sep   \adr{O:U:I}   \sep   \adr{S:I}   \sep   \adr{X:I}   \\
@@
 \begin{array}{rcll}
-	V	& \in		& \mbox{Values} \\
-	N	& \in		& \mbox{Natural numbers} \\
-	B	& \in		& \mbox{Boolean values} \\
-	T	& \in		& \mbox{Time values} \\
+	V	& \in		& \text{Values} \\
+	N	& \in		& \text{Natural numbers} \\
+	B	& \in		& \text{Boolean values} \\
+	T	& \in		& \text{Time values} \\
 	C	& ::=		& \adr{O:U:I} \sep \void \\
@@
 
-Our semantic approach defines the meaning of an AUTOSAR system as the possible chains of state changes that can be applied to the system's initial state; or equivalently, as the set of {\em traces} that can be generated from the corresponding process term $P_0$.
+Our semantic approach defines the meaning of an AUTOSAR system as the possible chains of state changes that can be applied to the system's initial state; or equivalently, as the set of \emph{traces} that can be generated from the corresponding process term $P_0$.
 A trace is a possibly infinite sequence of transitions
@@
 \begin{array}{rcll}
-	L	& ::=		& A!D		& \hspace{4em}\mbox{Say $A$ and $D$} \\
-		&		& A?D		& \hspace{4em}\mbox{Hear $A$ and $D$} \\
+	L	& ::=		& A!D		& \hspace{4em}\text{Say $A$ and $D$} \\
+		&		& A?D		& \hspace{4em}\text{Hear $A$ and $D$} \\
 \end{array}
@@
 
-The effect of these agreement rules closely resembles the idea of a {\em broadcast}: what one process says may be heard by every other process in a large parallel composition.
+The effect of these agreement rules closely resembles the idea of a \emph{broadcast}: what one process says may be heard by every other process in a large parallel composition.
 This behavior serves us well, because in general, what one AUTOSAR runnable does may have impact on many -- if not all -- parts of the executing system.
@@
 $$
-where $\Act$ is $\serving{\epsilon}{\epsilon}$ if $R$ is triggered by an {\em OperationInvokedEvent}; else $\Act$ is $\pending$ if $R$ is triggered by an {\em InitEvent}; otherwise $\Act$ is $\idle$.
+where $\Act$ is $\serving{\epsilon}{\epsilon}$ if $R$ is triggered by an \emph{OperationInvokedEvent}; else $\Act$ is $\pending$ if $R$ is triggered by an \emph{InitEvent}; otherwise $\Act$ is $\idle$.
 % SWS_Rte_03524 (SWS_Rte_03526?) forbids mixing OperationInvokedEvents with other events for the same runnable
@@
 $$
-if the {\em swImplPolicy} attribute of element $E$ is {\em queued}; otherwise, a term
+if the \emph{swImplPolicy} attribute of element $E$ is \emph{queued}; otherwise, a term
 $$
 \begin{array}{c}
-	\delem{\adr{E:U:I}}{\false}{\mbox{NO\_DATA}}
+	\delem{\adr{E:U:I}}{\false}{\text{NO\_DATA}}
 \end{array}
@@
 
-\item For each {\em TimingEvent} triggering each runnable $R$ of each atomic component $I$, a term
+\item For each \emph{TimingEvent} triggering each runnable $R$ of each atomic component $I$, a term
 $$
@@
 
-Some further static model information will also be referenced by the transition rules. We refer to the {\em connectors} of a model using a binary relation $\Rightarrow$ on addresses, defined as
+Some further static model information will also be referenced by the transition rules. We refer to the \emph{connectors} of a model using a binary relation $\Rightarrow$ on addresses, defined as
 $$
-	\adr{E:$U_1$:$I_1$} \Rightarrow \adr{E:$U_2$:$I_2$} \;\;\mbox{for all elements $E$ of port $U_2$}
+	\adr{E:$U_1$:$I_1$} \Rightarrow \adr{E:$U_2$:$I_2$} \;\;\text{for all elements $E$ of port $U_2$}
 $$
@@
 $$
-	\adr{O:$U_1$:$I_1$} \Rightarrow \adr{O:$U_2$:$I_2$} \;\;\mbox{for all operations $O$ of port $U_1$}
+	\adr{O:$U_1$:$I_1$} \Rightarrow \adr{O:$U_2$:$I_2$} \;\;\text{for all operations $O$ of port $U_1$}
 $$
@@
 
-We expect the {\em implementation} of a runnable to be available as a continuation $\Cont$, even though the real AUTOSAR model will be refering to concrete functions in external C/Matlab files.
+We expect the \emph{implementation} of a runnable to be available as a continuation $\Cont$, even though the real AUTOSAR model will be refering to concrete functions in external C/Matlab files.
 We will further abstract away from actual numbers of input and output parameters to these function by assuming that records (structs) are used as the single continuation argument and result whenever the arity so requires.

### 2015-08-22 paper/paper.tex
Details

@@
 
-\newcommand{\red}[1]{\xrightarrow{#1}}
-\newcommand{\say}[2]{#1 ! #2}
+\newcommand{\red} [1]{\xrightarrow{#1}}
+\newcommand{\say} [2]{#1 ! #2}
 \newcommand{\hear}[2]{#1 ? #2}
@@
 \newcommand{\ent}{\text{\small ent}}
-\newcommand{\ex}{\text{\small ex}}
+\newcommand{\ex} {\text{\small ex}}
 
@@
 
-The formal approach we have chosen is that of a typical process calculus \cite{}, where a system of concurrent processes evolves in a sequence of atomic steps, as determined by a global set of transition rules.
+The formal approach we have chosen is that of a typical process calculus \cite{TODO}, where a system of concurrent processes evolves in a sequence of atomic steps, as determined by a global set of transition rules.
 Our process terms correspond to the individual state-carrying parts of an AUTOSAR system, like the runnables, runnable instances, inter-runnable variables and port elements.
@@
 The following meta-variable conventions will be used:
-$$
+%
+\[
 \begin{array}{rcll}
@@
 \end{array}
-$$
+\]
+%
 Since AUTOSAR components can be arbitrarily nested, the component names $I$ can also be considered to have a similarly nested internal structure (component $I_1$ within component $I_2$ within component $I_3$, etc).
-We can fully ignore this detail here, though, as the hierarchical layout of components in an AUTOSAR system has no run-time implications \cite{}.
+We can fully ignore this detail here, though, as the hierarchical layout of components in an AUTOSAR system has no run-time implications \cite{TODO2}.
 
@@
 A trace is a possibly infinite sequence of transitions
-$$
+\[
 \begin{array}{c}
@@
 \end{array}
-$$
+\]
 where each step
-$$
+\[
 \begin{array}{c}
@@
 \end{array}
-$$
+\]
 states that the system state captured as process term $P_{i-1}$ can evolve into state $P_i$ by means of a single transition labelled $L_i$.
@@
 A label also always consists of an address $A$ paired with some additional detail $D$, as captured in the following label grammar:
-$$
+\[
 \begin{array}{rcll}
@@
 \end{array}
-$$
+\]
 The possible forms of the extra label information $D$ will be introduced as the different transition rules are defined.

### 2015-12-16 ARSim/NewARSim.hs
New deriving Interface based on unsafeCoerce.

@@ -36,6 +36,7 @@ OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
 {-# LANGUAGE ExistentialQuantification #-}
 {-# LANGUAGE DeriveFunctor #-}
 {-# LANGUAGE TypeFamilies #-}
+{-# LANGUAGE DeriveAnyClass #-}
 
 module NewARSim (module NewARSim, Typeable, Data, mkStdGen, StdGen) where
 
@@ -47,6 +48,7 @@ import qualified Data.Map as Map
 import Data.Maybe
 import Dynamics
 import System.Random
+import qualified Unsafe.Coerce
 
 
 -- The RTE monad -------------------------------------------------------------
@@ -103,6 +105,7 @@ rteResult      rop     = singleton $ Result     rop
 printlog id val             = singleton $ Printlog id val
 
 data StdRet a               = Ok a
+                            | Error Int
                             | NO_DATA
                             | NEVER_RECEIVED
                             | LIMIT
@@ -112,7 +115,11 @@ data StdRet a               = Ok a
                             deriving Show
 
 newtype DataElement q a r c             = DE Address      -- Async channel of "a" data
+    deriving Interface
+    
 newtype ClientServerOperation a b r c   = OP Address      -- Sync channel of an "a->b" service
+    deriving Interface
+    
 
 data Queued         -- Parameter q above
 data Unqueued
@@ -338,12 +345,13 @@ instance Addressed (ClientServerOperation a b r c) where
 
 class Interface m where
         seal                :: m c -> m ()
+        seal                = Unsafe.Coerce.unsafeCoerce
 
-instance Interface (DataElement q a r) where
-        seal (DE a)         = DE a
+--instance Interface (DataElement q a r) where
+--        seal (DE a)         = DE a
 
-instance Interface (ClientServerOperation a b r) where
-        seal (OP a)         = OP a
+--instance Interface (ClientServerOperation a b r) where
+--        seal (OP a)         = OP a
 
 
 connectEach :: Port p => [p Provided ()] -> [p Required ()] -> AR c ()

### 2016-04-23 prolog/tool/LaTeX.hs
Simple converter from eprover to LaTeX added

@@ -1,0 +1,159 @@
+module LaTeX where
+-- pretty-printer generated by the BNF converter + tweaked
+import Eprover.Abs
+import Data.Char
+
+import Sem
+
+-- the top-level printing method
+printTree :: Print a => a -> String
+printTree = render . prt 0
+
+type Doc = [ShowS] -> [ShowS]
+
+doc :: ShowS -> Doc
+doc = (:)
+
+render :: Doc -> String
+render d = rend 0 (map ($ "") $ d []) "" where
+  rend i ss = case ss of
+    "["      :ts -> showChar '[' . rend i ts
+    "("      :ts -> showChar '(' . rend i ts
+    "{"      :ts -> showChar '{' . new (i+1) . rend (i+1) ts
+    "}" : ";":ts -> new (i-1) . space "}" . showChar ';' . new (i-1) . rend (i-1) ts
+    "}"      :ts -> new (i-1) . showChar '}' . new (i-1) . rend (i-1) ts
+    ";"      :ts -> showChar ';' . new i . rend i ts
+    t  : "," :ts -> showString t . space "," . rend i ts
+    t  : ")" :ts -> showString t . showChar ')' . rend i ts
+    t  : "]" :ts -> showString t . showChar ']' . rend i ts
+    t        :ts -> space t . rend i ts
+    _            -> id
+  new i   = id             -- spacing handled by LaTeX
+  space t = showString t   -- spacing handled by LaTeX
+--  new i   = showChar '\n' . replicateS (2*i) (showChar ' ') . dropWhile isSpace
+--  space t = showString t . (\s -> if null s then "" else (' ':s))
+
+parenth :: Doc -> Doc
+parenth ss = doc (showChar '(') . ss . doc (showChar ')')
+
+concatS :: [ShowS] -> ShowS
+concatS = foldr (.) id
+
+concatD :: [Doc] -> Doc
+concatD = foldr (.) id
+
+replicateS :: Int -> ShowS -> ShowS
+replicateS n f = concatS (replicate n f)
+
+-- the printer class does the job
+class Print a where
+  prt :: Int -> a -> Doc
+  prtList :: Int -> [a] -> Doc
+  prtList i = concatD . map (prt i)
+
+instance Print a => Print [a] where
+  prt = prtList
+
+instance Print Char where
+  prt _ s = doc (showChar '\'' . mkEsc '\'' s . showChar '\'')
[…]

### 2016-04-23 prolog/tool/Sem.hs

@@ -1,0 +1,4 @@
+module Sem where
+import Eprover.Abs
+sem :: Database
+sem = Db [Rule (TPred (Complex (Atm (LIdent "step")) [Complex (Atm (LIdent "par")) [VarT (V (UIdent "P1")),VarT (V (UIdent "Q1"))],Complex (Atm (LIdent "say")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],Complex (Atm (LIdent "par")) [VarT (V (UIdent "P2")),VarT (V (UIdent "Q2"))]])) [TPred (Complex (Atm (LIdent "step")) [VarT (V (UIdent "P1")),Complex (Atm (LIdent "say")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],VarT (V (UIdent "P2"))]),TPred (Complex (Atm (LIdent "step")) [VarT (V (UIdent "Q1")),Complex (Atm (LIdent "hear")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],VarT (V (UIdent "Q2"))])],Rule (TPred (Complex (Atm (LIdent "step")) [Complex (Atm (LIdent "par")) [VarT (V (UIdent "P1")),VarT (V (UIdent "Q1"))],Complex (Atm (LIdent "say")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],Complex (Atm (LIdent "par")) [VarT (V (UIdent "P2")),VarT (V (UIdent "Q2"))]])) [TPred (Complex (Atm (LIdent "step")) [VarT (V (UIdent "P1")),Complex (Atm (LIdent "hear")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],VarT (V (UIdent "P2"))]),TPred (Complex (Atm (LIdent "step")) [VarT (V (UIdent "Q1")),Complex (Atm (LIdent "say")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],VarT (V (UIdent "Q2"))])],Rule (TPred (Complex (Atm (LIdent "step")) [Complex (Atm (LIdent "par")) [VarT (V (UIdent "P1")),VarT (V (UIdent "Q1"))],Complex (Atm (LIdent "hear")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],Complex (Atm (LIdent "par")) [VarT (V (UIdent "P2")),VarT (V (UIdent "Q2")
[…]

### 2016-04-24 prolog/tool/LaTeX.hs
More work on the eprover to LaTeX conversion

@@ -82,14 +82,14 @@ instance Print LitString where
 
 
 instance Print UIdent where
-  prt _ (UIdent i) = doc (showString i)
+  prt _ (UIdent i) = docS ("\\Tvar{"++i++"}")
 
 instance Print LIdent where
   prt _ (LIdent i) = doc (showString i)
 
 
 instance Print Wild where
-  prt _ (Wild i) = doc (showString i)
+  prt _ (Wild i) = doc (showString ('\\':i))
 
 instance Print Database where
   prt i e = case e of
@@ -141,7 +141,7 @@ instance Print Term where
                           , prt 0 terms
                           , docS "}"
                           ])
-    TAtom   atom  -> prPrec i 0 (concatD [prt 0 atom])
+    TAtom   atom  -> prPrec i 0 (concatD [docS "\\Tatom{", prt 0 atom, docS "}"])
     VarT    var   -> prPrec i 0 (concatD [prt 0 var])
     TInt    n     -> prPrec i 0 (concatD [prt 0 n])
     TParen  term  -> prPrec i 0 (concatD [doc (showString "("), prt 0 term, doc (showString ")")])
@@ -149,11 +149,12 @@ instance Print Term where
   prtList _ (x:xs)  = concatD [prt 0 x, doc (showString "}{"), prt 0 xs]
 instance Print Atom where
   prt i e = case e of
-    Atm lident -> prPrec i 0 (concatD [prt 0 lident])
+    Atm (LIdent s) | elem '_' s -> prPrec i 0 (concatD [prt 0 (LIdent (filter ('_'/=) s))])
+                   | otherwise  -> prPrec i 0 (concatD [prt 0 (LIdent s)])
 
 instance Print Var where
   prt i e = case e of
-    V uident -> prPrec i 0 (concatD [prt 0 uident])
-    A wild -> prPrec i 0 (concatD [prt 0 wild])
+    V uident  -> prPrec i 0 (concatD [prt 0 uident])
+    A wild    -> prPrec i 0 (concatD [prt 0 wild])
 
 main = putStrLn (printTree sem)

### 2016-04-24 paper/paper.tex
Some TODOs.
More work on the eprover to LaTeX conversion
Preparing macro merge by moving them to a separate file
LaTeX macros merged from paper and semantics. Next step i formatting.

@@
 \author{Johan Nordlander \and Patrik Jansson}
-
-\newcommand{\V}[1]{\text{\it #1}}
-\newcommand{\Code} {\V{Code}}
-\newcommand{\Cont} {\V{Cont}}
-\newcommand{\State}{\V{State}}
-\newcommand{\Act}  {\V{Act}}
-\newcommand{\Upd}  {\V{Upd}}
-\newcommand{\Xs}   {\V{Xs}}
-\newcommand{\Vs}   {\V{Vs}}
-\newcommand{\Cs}   {\V{Cs}}
-
-\newcommand{\kw}[1]{\textbf{#1}}
-\newcommand{\void}    {\kw{void}}
-\newcommand{\nil}     {\kw{nil}}
-\newcommand{\true}    {\kw{true}}
-\newcommand{\false}   {\kw{false}}
-\newcommand{\idle}    {\kw{idle}}
-\newcommand{\pending} {\kw{pending}}
-\newcommand{\serving}[2]{\kw{serving}(#1,#2)}
-
-\newcommand{\adr}[1]{\emph{#1}}
-\newcommand{\sep}{\hspace{2ex}|\hspace{2ex}}
-
-\newcommand{\runnable}[4]{\kw{runnable}(#1,#2,#3,#4)}
-\newcommand{\rinst}   [4]{\kw{rinst}(#1,#2,#3,#4)}
-\newcommand{\excl}    [2]{\kw{excl}(#1,#2)}
-\newcommand{\irv}     [2]{\kw{irv}(#1,#2)}
-\newcommand{\qelem}   [2]{\kw{qelem}(#1,#2)}
-\newcommand{\delem}   [3]{\kw{delem}(#1,#2,#3)}
-\newcommand{\opres}   [2]{\kw{opres}(#1,#2)}
-\newcommand{\timer}   [3]{\kw{timer}(#1,#2, #3)}
-
-\newcommand{\rte}[1]{\text{rte\_#1}}
-\newcommand{\Enter}     [2]{\rte{Enter}(#1,#2)}
-\newcommand{\Exit}      [2]{\rte{Exit}(#1,#2)}
-\newcommand{\IrvWrite}  [3]{\rte{IrvWrite}(#1,#2,#3)}
-\newcommand{\IrvRead}   [2]{\rte{IrvRead}(#1,#2)}
-\newcommand{\Send}      [4]{\rte{Send}(#1,#2,#3,#4)}
-\newcommand{\Receive}   [3]{\rte{Receive}(#1,#2,#3)}
-\newcommand{\Write}     [4]{\rte{Write}(#1,#2,#3,#4)}
-\newcommand{\Read}      [3]{\rte{Read}(#1,#2,#3)}
-\newcommand{\Invalidate}[3]{\rte{Invalidate}(#1,#2,#3)}
-\newcommand{\IsUpdated} [3]{\rte{IsUpdated}(#1,#2,#3)}
-\newcommand{\Call}      [4]{\rte{Call}(#1,#2,#3,#4)}
-\newcommand{\Result}    [3]{\rte{Result}(#1,#2,#3)}
-\newcommand{\Return}    [1]{\text{return}(#1)}
-
-\newcommand{\red} [1]{\xrightarrow{#1}}
-\newcommand{\say} [2]{#1 ! #2}
-\newcommand{\hear}[2]{#1 ? #2}
-
-\newcommand{\ent}{\text{\small ent}}
-\newcommand{\ex} {\text{\small ex}}
-
+%%%%%%%
+\input{macros.tex}
 \setlength{\parindent}{0pt}
@@
 Non-queued sender-receiver data element $E$ in port $U$ of component $I$, currently holding value $V$ with boolean update status flag $B$.
-\item $\qelem{\adr{E:U:I}}{\Vs}$ \newline
+\item $\qelem{\adr{E:U:I}}{\Vs}{TODO}$ \newline
 Queued sender-receiver data element $E$ in port $U$ of component $I$, currently holding value sequence $\Vs$.
@@
 Client-server operation $O$ in port $U$ of component $I$, currently holding result value sequence $\Vs$.
-\item $\timer{\adr{R:I}}{T_p}{T}$ \newline
+\item $\timer{\adr{R:I}}{T_p}{T}$ \newline  %TODO: timer has only 2 args in the semantics
 Timer state for runnable $R$ of component $I$ and period $T_p$, with $T$ seconds left of its current cycle.
@@
 Because the result from an RTE call may in general affect subsequent runnable behavior, we use a continuation-passing style to allow any data-dependencies to be expressed.
+%TODO: several LaTeX commands here are used with more arguments than in the semantics below
 $$
@@
 	P,Q 	& ::= &
-		\runnable{A}{T}{\Act}{N}	\\
+		\runnable{A}{K}{T}{\Act}{N}	\\
 		& | &
@@
 		& | &
-		\qelem{A}{\Vs}			\\
+		\qelem{A}{\Vs}{TODO}			\\
 		& | &
@@
 \begin{array}{c}
-	\runnable{\adr{R:I}}{0.0}{\Act}{0}
+	\runnable{\adr{R:I}}{TODO}{0.0}{\Act}{0}
 \end{array}
@@
 \begin{array}{c}
-	\qelem{\adr{E:U:I}}{\epsilon}
+	\qelem{\adr{E:U:I}}{\epsilon}{TODO}
 \end{array}
@@
 
+Rule and examples of provided delegation port compatibility don't match
+
+Multiple inconsistent CompSpecs
+
 \section{Conclusions and Future Work}
@@
 
-\input{semantics.tex}
+\input{../prolog/tool/semantics.tex}
 

### 2016-04-25 paper/paper.tex
Starting to merge the updated semantics into paper/semantics.tex

@@
 $$
-	\adr{E:$U_1$:$I_1$} \Rightarrow \adr{E:$U_2$:$I_2$} \;\;\text{for all elements $E$ of port $U_2$}
+	\Tconnect{\adr{E:$U_1$:$I_1$}}{\adr{E:$U_2$:$I_2$}} \;\;\text{for all elements $E$ of port $U_2$}
 $$
@@
 $$
-	\adr{O:$U_1$:$I_1$} \Rightarrow \adr{O:$U_2$:$I_2$} \;\;\text{for all operations $O$ of port $U_1$}
+	\Tconnect{\adr{O:$U_1$:$I_1$}}{\adr{O:$U_2$:$I_2$}} \;\;\text{for all operations $O$ of port $U_1$}
 $$

### 2016-04-26 prolog/tool/LaTeX.hs
Edits during PaJa + JoNo meeting

@@ -122,8 +122,8 @@ instance Print Clause where
                            endRule])
     Directive predicates -> prPrec i 0 (concatD [doc (showString ":-"), prt 0 predicates])
   prtList _ [] = (concatD [])
-  prtList _ (x:xs) = (concatD [prt 0 x, doc (showString ".\n\\\\"), prt 0 xs])
-    -- TODO perhaps include the . and newline in the \Prule and \Pfact commands?
+  prtList _ (x:xs) = (concatD [prt 0 x, doc (showString "\n\\\\"), prt 0 xs])
+  -- TODO: perhaps generate a macro for the "cons" in the list of clauses
 
 instance Print Predicate where
   prt i e = case e of

### 2016-04-27 paper/paper.tex
LaTeX translation of semantics done.
Edits during PaJa + JoNo meeting
Paper reading + updates pages 1--3

@@
 And because the standard is open to interpretation, the interoperability of software components across tools and platforms is often seriously hampered.
-What is more, a single software component cannot easily be studied and understood in isolation, since its interactive behavior is only indirectly defined in terms of the concrete C-code and OS tasks that realize it and interaction environment.
+What is more, a single software component cannot easily be studied and understood in isolation, since its interactive behavior is only indirectly defined in terms of the concrete C-code and OS tasks that realize it and its interaction environment.
 
 This paper takes an important step towards a remedy to these problems, by contributing a formal specification of a substantial core of the AUTOSAR standard.
-The formalization covers most of the Software Component Template and its accompanying Run-Time Environment (RTE) (Section~\ref{sec:Calc}), and is able to directly express every legal way a system of SWCs can evolve in at run-time on an arbitrarily fast platform (Section~\ref{sec:Sem}).
+The formalization covers most of the Software Component Template and its accompanying Run-Time Environment (RTE) (Section~\ref{sec:Calc}), and is able to directly express every legal way a system of SWCs can evolve at run-time on an arbitrarily fast platform (Section~\ref{sec:Sem}).
 It can thus serve as a basis for both a concrete AUTOSAR implementation on a specific platform (whose supported behaviors must be a subset of those defined by the formal semantics), and a platform-independent AUTOSAR simulator (where behaviors from the legal set can be picked at random).
@@
 A runnable under execution is called a \textbf{runnable instance}, and these may execute concurrently even if they belong to the same SWC.
-Each runnable indicates whether it may be instantiated concurrently with itself.
+Each runnable indicates whether or not it may be instantiated concurrently with itself.
 Runnable code can also use \textbf{exclusive areas} (a form of mutex semaphores) to further control concurrent behavior.
 
-To interact with ports, inter-runnable variables and exclusive areas, runnables utilize what is abstractly known as the \textbf{virtual function bus} (VFB).
-Concretely, the VFB takes the shape of a C programming interface to the \textbf{run-time environment} (RTE), custom-generated for each runnable on basis of its access constraints.
-Behind the RTE, a collection of OS kernels, communication stacks and other basic software modules implement the VBF services for the platform at hand.
+To interact with ports, inter-runnable variables and exclusive areas, runnables use what is abstractly known as the \textbf{virtual function bus} (VFB).
+Concretely, the VFB takes the shape of a C programming interface to the \textbf{run-time environment} (RTE), custom-generated for each runnable on the basis of its access constraints.
+Behind the RTE, a collection of OS kernels, communication stacks and other basic software modules implement the VFB services for the platform at hand.
 Complete AUTOSAR designs also typically contain \textbf{ECU} and \textbf{task assignments}, which are manually built tables that control how SWCs and runnables map onto the available hardware and OS resources, respectively.
@@
 This static nature renders AUTOSAR models very suitable to graphical notations, which by far is the most widespread representation format for AUTOSAR models.
-An graphical model example is shown in Figure~\ref{fig:model}.
+A graphical model example is shown in Figure~\ref{fig:model}.
+
+TODO: add fig:model
 
@@
 Components and ports do not carry any dynamic state besides the state of their constituent parts, so they need no explicit representation as process terms.
-However, in order to still be able to talk about model elements using their locally scoped names, we employ a hierarchical addressing scheme, such that address $\adr{R:I}$ means runnable $R$ of component $I$ and $\adr{E:P:I}$ means element $E$ of port $P$ within component $I$.
-The following meta-variable conventions will be used:
+However, in order to still be able to talk about model elements using their locally scoped names, we employ a hierarchical addressing scheme, such that address $\adrr{R}{I}$ means runnable $R$ of component $I$ and $\adrep{E}{P}{I}$ means element $E$ of port $P$ within component $I$.
+We use the following meta-variable conventions:
+
+%TODO: port name convention change from U to P? (see ../prolog/semantics.eprover)
 %
@@
 	A	& \in		& \text{Addresses} \\
-	A	& ::=		& \adr{R:I}   \sep   \adr{E:P:I}   \sep   \adr{O:U:I}   \sep   \adr{S:I}   \sep   \adr{X:I}   \\
+	A	& ::=		& \adrr{R}{I}   \sep   \adrep{E}{P}{I}   \sep \adrop{O}{P}{I}   \sep   \adrs{S}{I}   \sep   \adrx{X}{I}   \\
 \end{array}
@@
 \begin{itemize}
-\item $\runnable{\adr{R:I}}{T}{\Act}{N}$  \newline
+\item $\runnable{\adrr{R}{I}}{TODOK}{T}{\Act}{N}$  \newline
 Common state for runnable $R$ of component $I$, showing $N$ current instances, $T$ seconds left until next possible instantiation, and activation state $\Act$.
-\item $\rinst{\adr{R:I}}{C}{\Xs}{\Code}$ \newline
+\item $\rinst{\adrr{R}{I}}{C}{\Xs}{\Code}$ \newline
 An instance of runnable $R$ of component $I$, currently executing $\Code$ and owning the exclusive areas $\Xs$, possibly running on behalf of client $C$.
-\item $\excl{\adr{X:I}}{B}$ \newline
+\item $\excl{\adrx{X}{I}}{B}$ \newline
 Exclusive area $X$ of component $I$, with boolean occupation state $B$.
-\item $\irv{\adr{S:I}}{V}$ \newline
+\item $\irv{\adrs{S}{I}}{V}$ \newline
 Inter-runnable variable $S$ of component $I$, currently holding value $V$.
-\item $\delem{\adr{E:U:I}}{B}{V}$ \newline
+\item $\delem{\adrep{E}{U}{I}}{B}{V}$ \newline
 Non-queued sender-receiver data element $E$ in port $U$ of component $I$, currently holding value $V$ with boolean update status flag $B$.
-\item $\qelem{\adr{E:U:I}}{\Vs}{TODO}$ \newline
+\item $\qelem{\adrep{E}{U}{I}}{\Vs}{TODO}$ \newline
 Queued sender-receiver data element $E$ in port $U$ of component $I$, currently holding value sequence $\Vs$.
-\item $\opres{\adr{O:U:I}}{\Vs}$ \newline
+\item $\opres{\adrop{O}{U}{I}}{\Vs}$ \newline
 Client-server operation $O$ in port $U$ of component $I$, currently holding result value sequence $\Vs$.
-\item $\timer{\adr{R:I}}{T_p}{T}$ \newline  %TODO: timer has only 2 args in the semantics
-Timer state for runnable $R$ of component $I$ and period $T_p$, with $T$ seconds left of its current cycle.
+\item $\timer{\adrr{R}{I}}{T}$ \newline  %TODO: timer has only 2 args in the semantics so {T_p} is not described
+Countdown timer state for runnable $R$ of component $I$ with $T$ seconds left of its current cycle.
 \end{itemize}
@@
 	T	& \in		& \text{Time values} \\
-	C	& ::=		& \adr{O:U:I} \sep \void \\
+	C	& ::=		& \adrop{O}{U}{I} \sep \void \\
 	\Act	& ::=		& \idle \sep \pending \sep \serving{\Cs}{\Vs} \\
@@
 Sequences may be empty, which is written $\epsilon$.
-Left (right) concatenation of an element with a sequence is written $\adr{V:Vs}$ ($\adr{Vs:V}$).
+Left (right) concatenation of an element with a sequence is written $\Tcons{\V{V}}{\V{Vs}}$ ($\Tsnoc{\V{Vs}}{\V{V}}$).
 
@@
 		& | &
-		\timer{A}{T_p}{T}		\\
+		\timer{A}{T}		\\
 		& | &
@@
 
-The label $L$ of a transition can be ow two kinds, indicating that the affected process term either "says" or "hears" something during the transition.
+The label $L$ of a transition can be of two kinds, indicating that the affected process term either "says" or "hears" something during the transition.
 A label also always consists of an address $A$ paired with some additional detail $D$, as captured in the following label grammar:
@@
 	L	& ::=		& A!D		& \hspace{4em}\text{Say $A$ and $D$} \\
-		&		& A?D		& \hspace{4em}\text{Hear $A$ and $D$} \\
+		&	|	& A?D		& \hspace{4em}\text{Hear $A$ and $D$} \\
 \end{array}
@@
 \begin{array}{c}
-	\runnable{\adr{R:I}}{TODO}{0.0}{\Act}{0}
+	\runnable{\adrr{R}{I}}{TODO}{0.0}{\Act}{0}
 \end{array}
@@
 \begin{array}{c}
-	\excl{\adr{X:I}}{\false}
+	\excl{\adrx{X}{I}}{\false}
 \end{array}
@@
 \begin{array}{c}
-	\irv{\adr{S:I}}{V}
+	\irv{\adrs{S}{I}}{V}
 \end{array}
@@
 \begin{array}{c}
-	\qelem{\adr{E:U:I}}{\epsilon}{TODO}
+	\qelem{\adrep{E}{U}{I}}{\epsilon}{TODO}
 \end{array}
@@
 \begin{array}{c}
-	\delem{\adr{E:U:I}}{\false}{\text{NO\_DATA}}
+	\delem{\adrep{E}{U}{I}}{\false}{\text{nodata}}
 \end{array}
@@
 \begin{array}{c}
-	\opres{\adr{O:U:I}}{\epsilon}
+	\opres{\adrep{O}{U}{I}}{\epsilon}
 \end{array}
@@
 \begin{array}{c}
-	\timer{\adr{R:I}}{T}{0.0}
+	\timer{\adrr{R}{I}}{0.0}
 \end{array}
 $$
-where $T$ is the period of the event.
+% where $T$ is the period of the event.
+%TODO: Does this mean all timed runnables will want to act _immediately_ (at the same time)? No init delay?
 
@@
 $$
-	\Tconnect{\adr{E:$U_1$:$I_1$}}{\adr{E:$U_2$:$I_2$}} \;\;\text{for all elements $E$ of port $U_2$}
+	\Tconnect{\adrep{E}{$U_1$}{$I_1$}}{\adrep{E}{$U_2$}{$I_2$}} \;\;\text{for all elements $E$ of port $U_2$}
 $$
@@
 $$
-	\Tconnect{\adr{O:$U_1$:$I_1$}}{\adr{O:$U_2$:$I_2$}} \;\;\text{for all operations $O$ of port $U_1$}
+	\Tconnect{\adrop{O}{$U_1$}{$I_1$}}{\adrop{O}{$U_2$}{$I_2$}} \;\;\text{for all operations $O$ of port $U_1$}
 $$
@@
 
+%TODO: Patrik got here 2016-04-27 in a reading + macro conversion round
 \subsection{Exclusive areas}
@@
 
-% \inputminted{prolog}{../prolog/semantics.pl}
-
-\input{../prolog/tool/semantics.tex}
-
-% \verbatiminput{../prolog/semantics.pl}
+\input{semantics.tex}
+%\input{../prolog/tool/semantics.tex}
 

### 2016-04-29 paper/paper.tex
Tweaks...

@@
 Components and ports do not carry any dynamic state besides the state of their constituent parts, so they need no explicit representation as process terms.
-However, in order to still be able to talk about model elements using their locally scoped names, we employ a hierarchical addressing scheme, such that address $\adrr{R}{I}$ means runnable $R$ of component $I$ and $\adrep{E}{P}{I}$ means element $E$ of port $P$ within component $I$.
+However, in order to still be able to talk about model elements using their locally scoped names, we employ a hierarchical addressing scheme, such that address $\Tr{R}{I}$ means runnable $R$ of component $I$ and $\Te{E}{P}{I}$ means element $E$ of port $P$ within component $I$.
 We use the following meta-variable conventions:
@@
 	A	& \in		& \text{Addresses} \\
-	A	& ::=		& \adrr{R}{I}   \sep   \adrep{E}{P}{I}   \sep \adrop{O}{P}{I}   \sep   \adrs{S}{I}   \sep   \adrx{X}{I}   \\
+	A	& ::=		& \Tr{R}{I}   \sep   \Te{E}{P}{I}   \sep \To{O}{P}{I}   \sep   \Ts{S}{I}   \sep   \Tx{X}{I}   \\
 \end{array}
@@
 \begin{itemize}
-\item $\runnable{\adrr{R}{I}}{TODOK}{T}{\Act}{N}$  \newline
+\item $\runnable{\Tr{R}{I}}{TODOK}{T}{\Act}{N}$  \newline
 Common state for runnable $R$ of component $I$, showing $N$ current instances, $T$ seconds left until next possible instantiation, and activation state $\Act$.
-\item $\rinst{\adrr{R}{I}}{C}{\Xs}{\Code}$ \newline
+\item $\rinst{\Tr{R}{I}}{C}{\Xs}{\Code}$ \newline
 An instance of runnable $R$ of component $I$, currently executing $\Code$ and owning the exclusive areas $\Xs$, possibly running on behalf of client $C$.
-\item $\excl{\adrx{X}{I}}{B}$ \newline
+\item $\excl{\Tx{X}{I}}{B}$ \newline
 Exclusive area $X$ of component $I$, with boolean occupation state $B$.
-\item $\irv{\adrs{S}{I}}{V}$ \newline
+\item $\irv{\Ts{S}{I}}{V}$ \newline
 Inter-runnable variable $S$ of component $I$, currently holding value $V$.
-\item $\delem{\adrep{E}{U}{I}}{B}{V}$ \newline
+\item $\delem{\Te{E}{U}{I}}{B}{V}$ \newline
 Non-queued sender-receiver data element $E$ in port $U$ of component $I$, currently holding value $V$ with boolean update status flag $B$.
-\item $\qelem{\adrep{E}{U}{I}}{\Vs}{TODO}$ \newline
+\item $\qelem{\Te{E}{U}{I}}{\Vs}{TODO}$ \newline
 Queued sender-receiver data element $E$ in port $U$ of component $I$, currently holding value sequence $\Vs$.
-\item $\opres{\adrop{O}{U}{I}}{\Vs}$ \newline
+\item $\opres{\To{O}{U}{I}}{\Vs}$ \newline
 Client-server operation $O$ in port $U$ of component $I$, currently holding result value sequence $\Vs$.
-\item $\timer{\adrr{R}{I}}{T}$ \newline  %TODO: timer has only 2 args in the semantics so {T_p} is not described
+\item $\timer{\Tr{R}{I}}{T}$ \newline  %TODO: timer has only 2 args in the semantics so {T_p} is not described
 Countdown timer state for runnable $R$ of component $I$ with $T$ seconds left of its current cycle.
@@
 	T	& \in		& \text{Time values} \\
-	C	& ::=		& \adrop{O}{U}{I} \sep \void \\
+	C	& ::=		& \To{O}{U}{I} \sep \void \\
 	\Act	& ::=		& \idle \sep \pending \sep \serving{\Cs}{\Vs} \\
@@
 \begin{array}{c}
-	\runnable{\adrr{R}{I}}{TODO}{0.0}{\Act}{0}
+	\runnable{\Tr{R}{I}}{TODO}{0.0}{\Act}{0}
 \end{array}
@@
 \begin{array}{c}
-	\excl{\adrx{X}{I}}{\false}
+	\excl{\Tx{X}{I}}{\false}
 \end{array}
@@
 \begin{array}{c}
-	\irv{\adrs{S}{I}}{V}
+	\irv{\Ts{S}{I}}{V}
 \end{array}
@@
 \begin{array}{c}
-	\qelem{\adrep{E}{U}{I}}{\epsilon}{TODO}
+	\qelem{\Te{E}{U}{I}}{\epsilon}{TODO}
 \end{array}
@@
 \begin{array}{c}
-	\delem{\adrep{E}{U}{I}}{\false}{\text{nodata}}
+	\delem{\Te{E}{U}{I}}{\false}{\text{nodata}}
 \end{array}
@@
 \begin{array}{c}
-	\opres{\adrep{O}{U}{I}}{\epsilon}
+	\opres{\Te{O}{U}{I}}{\epsilon}
 \end{array}
@@
 \begin{array}{c}
-	\timer{\adrr{R}{I}}{0.0}
+	\timer{\Tr{R}{I}}{0.0}
 \end{array}
@@
 $$
-	\Tconnect{\adrep{E}{$U_1$}{$I_1$}}{\adrep{E}{$U_2$}{$I_2$}} \;\;\text{for all elements $E$ of port $U_2$}
+	\Tconnect{\Tep{E}{$U_1$}{$I_1$}}{\Tep{E}{$U_2$}{$I_2$}} \;\;\text{for all elements $E$ of port $U_2$}
 $$
@@
 $$
-	\Tconnect{\adrop{O}{$U_1$}{$I_1$}}{\adrop{O}{$U_2$}{$I_2$}} \;\;\text{for all operations $O$ of port $U_1$}
+	\Tconnect{\Top{O}{$U_1$}{$I_1$}}{\Top{O}{$U_2$}{$I_2$}} \;\;\text{for all operations $O$ of port $U_1$}
 $$

### 2016-04-29 prolog/tool/LaTeX.hs

@@ -99,7 +99,7 @@ docS :: String -> Doc
 docS = doc . showString
 
 beginSemantics, endSemantics :: Doc
-beginSemantics = docS "\\[\n\\begin{array}{r@{}c@{}l}\n"
+beginSemantics = docS "\\[\n\\begin{array}{r@{}c@{}llllll}\n"
 endSemantics   = docS "\\end{array}\n\\]"
 
 beginFact, endFact :: Doc
@@ -141,7 +141,7 @@ instance Print Term where
                           , prt 0 terms
                           , docS "}"
                           ])
-    TAtom   atom  -> prPrec i 0 (concatD [docS "\\Tatom{", prt 0 atom, docS "}"])
+    TAtom   atom  -> prPrec i 0 (concatD [docS "\\T", prt 0 atom])
     VarT    var   -> prPrec i 0 (concatD [prt 0 var])
     TInt    n     -> prPrec i 0 (concatD [prt 0 n])
     TParen  term  -> prPrec i 0 (concatD [doc (showString "("), prt 0 term, doc (showString ")")])

### 2016-04-29 prolog/tool/Sem.hs

@@ -1,4 +1,4 @@
 module Sem where
 import Eprover.Abs
 sem :: Database
-sem = Db [Rule (TPred (Complex (Atm (LIdent "step")) [Complex (Atm (LIdent "par")) [VarT (V (UIdent "P1")),VarT (V (UIdent "Q1"))],Complex (Atm (LIdent "say")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],Complex (Atm (LIdent "par")) [VarT (V (UIdent "P2")),VarT (V (UIdent "Q2"))]])) [TPred (Complex (Atm (LIdent "step")) [VarT (V (UIdent "P1")),Complex (Atm (LIdent "say")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],VarT (V (UIdent "P2"))]),TPred (Complex (Atm (LIdent "step")) [VarT (V (UIdent "Q1")),Complex (Atm (LIdent "hear")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],VarT (V (UIdent "Q2"))])],Rule (TPred (Complex (Atm (LIdent "step")) [Complex (Atm (LIdent "par")) [VarT (V (UIdent "P1")),VarT (V (UIdent "Q1"))],Complex (Atm (LIdent "say")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],Complex (Atm (LIdent "par")) [VarT (V (UIdent "P2")),VarT (V (UIdent "Q2"))]])) [TPred (Complex (Atm (LIdent "step")) [VarT (V (UIdent "P1")),Complex (Atm (LIdent "hear")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],VarT (V (UIdent "P2"))]),TPred (Complex (Atm (LIdent "step")) [VarT (V (UIdent "Q1")),Complex (Atm (LIdent "say")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],VarT (V (UIdent "Q2"))])],Rule (TPred (Complex (Atm (LIdent "step")) [Complex (Atm (LIdent "par")) [VarT (V (UIdent "P1")),VarT (V (UIdent "Q1"))],Complex (Atm (LIdent "hear")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],Complex (Atm (LIdent "par")) [VarT (V (UIdent "P2")),VarT (V (UIdent "Q2")
[…]

### 2016-04-29 paper/paper.tex
Ongoing tweaks...

@@
 Components and ports do not carry any dynamic state besides the state of their constituent parts, so they need no explicit representation as process terms.
-However, in order to still be able to talk about model elements using their locally scoped names, we employ a hierarchical addressing scheme, such that address $\Tr{R}{I}$ means runnable $R$ of component $I$ and $\Te{E}{P}{I}$ means element $E$ of port $P$ within component $I$.
+However, in order to still be able to talk about model elements using their locally scoped names, we employ a hierarchical addressing scheme, such that address $\Tr{R}{I}$ means runnable $R$ of component $I$ and $\Tep{E}{P}{I}$ means element $E$ of port $P$ within component $I$.
 We use the following meta-variable conventions:
@@
 	A	& \in		& \text{Addresses} \\
-	A	& ::=		& \Tr{R}{I}   \sep   \Te{E}{P}{I}   \sep \To{O}{P}{I}   \sep   \Ts{S}{I}   \sep   \Tx{X}{I}   \\
+	A	& ::=		& \Tr{R}{I}   \sep   \Tep{E}{P}{I}   \sep \Top{O}{P}{I}   \sep   \Ts{S}{I}   \sep   \Tx{X}{I}   \\
 \end{array}
@@
 Inter-runnable variable $S$ of component $I$, currently holding value $V$.
-\item $\delem{\Te{E}{U}{I}}{B}{V}$ \newline
+\item $\delem{\Tep{E}{U}{I}}{B}{V}$ \newline
 Non-queued sender-receiver data element $E$ in port $U$ of component $I$, currently holding value $V$ with boolean update status flag $B$.
-\item $\qelem{\Te{E}{U}{I}}{\Vs}{TODO}$ \newline
+\item $\qelem{\Tep{E}{U}{I}}{\Vs}{TODO}$ \newline
 Queued sender-receiver data element $E$ in port $U$ of component $I$, currently holding value sequence $\Vs$.
-\item $\opres{\To{O}{U}{I}}{\Vs}$ \newline
+\item $\opres{\Top{O}{U}{I}}{\Vs}$ \newline
 Client-server operation $O$ in port $U$ of component $I$, currently holding result value sequence $\Vs$.
@@
 \begin{array}{c}
-	\qelem{\Te{E}{U}{I}}{\epsilon}{TODO}
+	\qelem{\Tep{E}{U}{I}}{\epsilon}{TODO}
 \end{array}
@@
 \begin{array}{c}
-	\delem{\Te{E}{U}{I}}{\false}{\text{nodata}}
+	\delem{\Tep{E}{U}{I}}{\false}{\text{nodata}}
 \end{array}
@@
 \begin{array}{c}
-	\opres{\Te{O}{U}{I}}{\epsilon}
+	\opres{\Tep{O}{U}{I}}{\epsilon}
 \end{array}
@@
 
-\begin{gather*}
-	\rinst{\adr{R:I}}{C}{\Xs}{\Enter{X}{\Cont}}  \red{\say{X:I}{\ent}}  \\
-		\rinst{\adr{R:I}}{C}{\adr{X:Xs}}{\Cont(\void)} \\
-	\\
-	\rinst{\adr{R:I}}{C}{\adr{X:Xs}}{\Exit{X}{\Cont}}  \red{\say{X:I}{\ex}}  \\
-		\rinst{\adr{R:I}}{C}{\Xs}{\Cont(\void)} \\
-	\\
-	\excl{\adr{X:I}}{\false} \red{\hear{X:I}{\ent}} \excl{\adr{X:I}}{\true} \\
-	\\
-	\excl{\adr{X:I}}{\true} \red{\hear{X:I}{\ex}} \excl{\adr{X:I}}{\false} \\
-	\\
-\end{gather*}
-
 \subsection{Inter-runnable variables}
-
-\begin{gather*}
-	\rinst{\adr{R:I}}{C}{\Xs}{\IrvWrite{X}{\Cont}}  \red{\say{X:I}{\ent}}  \\
-		\rinst{\adr{R:I}}{C}{\adr{X:Xs}}{\Cont(\void)} \\
-	\\
-	\rinst{\adr{R:I}}{C}{\adr{X:Xs}}{\Exit{X}{\Cont}}  \red{\say{X:I}{\ex}}  \\
-		\rinst{\adr{R:I}}{C}{\Xs}{\Cont(\void)} \\
-	\\
-	\excl{\adr{X:I}}{\false} \red{\hear{X:I}{\ent}} \excl{\adr{X:I}}{\true} \\
-	\\
-	\excl{\adr{X:I}}{\true} \red{\hear{X:I}{\ex}} \excl{\adr{X:I}}{\false} \\
-	\\
-\end{gather*}
 

### 2016-04-29 prolog/tool/Sem.hs

@@ -1,4 +1,4 @@
 module Sem where
 import Eprover.Abs
 sem :: Database
-sem = Db [Rule (TPred (Complex (Atm (LIdent "step")) [Complex (Atm (LIdent "par")) [VarT (V (UIdent "P1")),VarT (V (UIdent "Q1"))],Complex (Atm (LIdent "say")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],Complex (Atm (LIdent "par")) [VarT (V (UIdent "P2")),VarT (V (UIdent "Q2"))]])) [TPred (Complex (Atm (LIdent "step")) [VarT (V (UIdent "P1")),Complex (Atm (LIdent "say")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],VarT (V (UIdent "P2"))]),TPred (Complex (Atm (LIdent "step")) [VarT (V (UIdent "Q1")),Complex (Atm (LIdent "hear")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],VarT (V (UIdent "Q2"))])],Rule (TPred (Complex (Atm (LIdent "step")) [Complex (Atm (LIdent "par")) [VarT (V (UIdent "P1")),VarT (V (UIdent "Q1"))],Complex (Atm (LIdent "say")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],Complex (Atm (LIdent "par")) [VarT (V (UIdent "P2")),VarT (V (UIdent "Q2"))]])) [TPred (Complex (Atm (LIdent "step")) [VarT (V (UIdent "P1")),Complex (Atm (LIdent "hear")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],VarT (V (UIdent "P2"))]),TPred (Complex (Atm (LIdent "step")) [VarT (V (UIdent "Q1")),Complex (Atm (LIdent "say")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],VarT (V (UIdent "Q2"))])],Rule (TPred (Complex (Atm (LIdent "step")) [Complex (Atm (LIdent "par")) [VarT (V (UIdent "P1")),VarT (V (UIdent "Q1"))],Complex (Atm (LIdent "hear")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],Complex (Atm (LIdent "par")) [VarT (V (UIdent "P2")),VarT (V (UIdent "Q2")
[…]

### 2016-04-29 prolog/tool/Sem.hs
More of the same.

@@ -1,4 +1,4 @@
 module Sem where
 import Eprover.Abs
 sem :: Database
-sem = Db [Rule (TPred (Complex (Atm (LIdent "step")) [Complex (Atm (LIdent "par")) [VarT (V (UIdent "P1")),VarT (V (UIdent "Q1"))],Complex (Atm (LIdent "say")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],Complex (Atm (LIdent "par")) [VarT (V (UIdent "P2")),VarT (V (UIdent "Q2"))]])) [TPred (Complex (Atm (LIdent "step")) [VarT (V (UIdent "P1")),Complex (Atm (LIdent "say")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],VarT (V (UIdent "P2"))]),TPred (Complex (Atm (LIdent "step")) [VarT (V (UIdent "Q1")),Complex (Atm (LIdent "hear")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],VarT (V (UIdent "Q2"))])],Rule (TPred (Complex (Atm (LIdent "step")) [Complex (Atm (LIdent "par")) [VarT (V (UIdent "P1")),VarT (V (UIdent "Q1"))],Complex (Atm (LIdent "say")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],Complex (Atm (LIdent "par")) [VarT (V (UIdent "P2")),VarT (V (UIdent "Q2"))]])) [TPred (Complex (Atm (LIdent "step")) [VarT (V (UIdent "P1")),Complex (Atm (LIdent "hear")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],VarT (V (UIdent "P2"))]),TPred (Complex (Atm (LIdent "step")) [VarT (V (UIdent "Q1")),Complex (Atm (LIdent "say")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],VarT (V (UIdent "Q2"))])],Rule (TPred (Complex (Atm (LIdent "step")) [Complex (Atm (LIdent "par")) [VarT (V (UIdent "P1")),VarT (V (UIdent "Q1"))],Complex (Atm (LIdent "hear")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],Complex (Atm (LIdent "par")) [VarT (V (UIdent "P2")),VarT (V (UIdent "Q2")
[…]

### 2016-04-30 prolog/tool/LaTeX.hs
Started generating semantic group commands

@@ -111,8 +111,13 @@ beginRule = docS "\\Prule{"
 midRule   = docS "}{"
 endRule   = docS "}"
 
+beginGroup a = concatD [docS "\\newcommand{\\", docS a, docS "}{%"]
+endGroup   a = docS "}"
+
 instance Print Clause where
   prt i e = case e of
+    Fact (TPred (Complex (Atm (LIdent "begingroup")) [VarT (V (UIdent a))])) -> prPrec i 0 (beginGroup a)
+    Fact (TPred (Complex (Atm (LIdent "endgroup"))   [VarT (V (UIdent a))])) -> prPrec i 0 (endGroup a)
     Fact predicate -> prPrec i 0 (concatD [beginFact, prt 0 predicate, endFact])
     Rule predicate predicates ->
       prPrec i 0 (concatD [beginRule,

### 2016-04-30 prolog/tool/Sem.hs

@@ -1,4 +1,4 @@
 module Sem where
 import Eprover.Abs
 sem :: Database
-sem = Db [Rule (TPred (Complex (Atm (LIdent "step")) [Complex (Atm (LIdent "par")) [VarT (V (UIdent "P_1")),VarT (V (UIdent "Q_1"))],Complex (Atm (LIdent "say")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],Complex (Atm (LIdent "par")) [VarT (V (UIdent "P_2")),VarT (V (UIdent "Q_2"))]])) [TPred (Complex (Atm (LIdent "step")) [VarT (V (UIdent "P_1")),Complex (Atm (LIdent "say")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],VarT (V (UIdent "P_2"))]),TPred (Complex (Atm (LIdent "step")) [VarT (V (UIdent "Q_1")),Complex (Atm (LIdent "hear")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],VarT (V (UIdent "Q_2"))])],Rule (TPred (Complex (Atm (LIdent "step")) [Complex (Atm (LIdent "par")) [VarT (V (UIdent "P_1")),VarT (V (UIdent "Q_1"))],Complex (Atm (LIdent "say")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],Complex (Atm (LIdent "par")) [VarT (V (UIdent "P_2")),VarT (V (UIdent "Q_2"))]])) [TPred (Complex (Atm (LIdent "step")) [VarT (V (UIdent "P_1")),Complex (Atm (LIdent "hear")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],VarT (V (UIdent "P_2"))]),TPred (Complex (Atm (LIdent "step")) [VarT (V (UIdent "Q_1")),Complex (Atm (LIdent "say")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],VarT (V (UIdent "Q_2"))])],Rule (TPred (Complex (Atm (LIdent "step")) [Complex (Atm (LIdent "par")) [VarT (V (UIdent "P_1")),VarT (V (UIdent "Q_1"))],Complex (Atm (LIdent "hear")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],Complex (Atm (LIdent "par")) [VarT (V (UIdent "P_2"))
[…]

### 2016-04-30 prolog/tool/LaTeX.hs
Appendix now consists of macro-definitions of groups of semantics rules

@@ -99,8 +99,8 @@ docS :: String -> Doc
 docS = doc . showString
 
 beginSemantics, endSemantics :: Doc
-beginSemantics = docS "\\[\n\\begin{array}{r@{}c@{}llllll}\n"
-endSemantics   = docS "\\end{array}\n\\]"
+beginSemantics = docS "" -- "\\[\n\\begin{array}{r@{}c@{}llllll}\n"
+endSemantics   = docS "" -- "\\end{array}\n\\]"
 
 beginFact, endFact :: Doc
 beginFact = docS "\\Pfact{"
@@ -127,7 +127,7 @@ instance Print Clause where
                            endRule])
     Directive predicates -> prPrec i 0 (concatD [doc (showString ":-"), prt 0 predicates])
   prtList _ [] = (concatD [])
-  prtList _ (x:xs) = (concatD [prt 0 x, doc (showString "\n\\\\"), prt 0 xs])
+  prtList _ (x:xs) = (concatD [prt 0 x, doc (showString "\n"), prt 0 xs])
   -- TODO: perhaps generate a macro for the "cons" in the list of clauses
 
 instance Print Predicate where

### 2016-04-30 prolog/tool/Sem.hs

@@ -1,4 +1,4 @@
 module Sem where
 import Eprover.Abs
 sem :: Database
-sem = Db [Fact (TPred (Complex (Atm (LIdent "startgroup")) [VarT (V (UIdent "CombRed"))])),Rule (TPred (Complex (Atm (LIdent "step")) [Complex (Atm (LIdent "par")) [VarT (V (UIdent "P_1")),VarT (V (UIdent "Q_1"))],Complex (Atm (LIdent "say")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],Complex (Atm (LIdent "par")) [VarT (V (UIdent "P_2")),VarT (V (UIdent "Q_2"))]])) [TPred (Complex (Atm (LIdent "step")) [VarT (V (UIdent "P_1")),Complex (Atm (LIdent "say")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],VarT (V (UIdent "P_2"))]),TPred (Complex (Atm (LIdent "step")) [VarT (V (UIdent "Q_1")),Complex (Atm (LIdent "hear")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],VarT (V (UIdent "Q_2"))])],Rule (TPred (Complex (Atm (LIdent "step")) [Complex (Atm (LIdent "par")) [VarT (V (UIdent "P_1")),VarT (V (UIdent "Q_1"))],Complex (Atm (LIdent "say")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],Complex (Atm (LIdent "par")) [VarT (V (UIdent "P_2")),VarT (V (UIdent "Q_2"))]])) [TPred (Complex (Atm (LIdent "step")) [VarT (V (UIdent "P_1")),Complex (Atm (LIdent "hear")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],VarT (V (UIdent "P_2"))]),TPred (Complex (Atm (LIdent "step")) [VarT (V (UIdent "Q_1")),Complex (Atm (LIdent "say")) [VarT (V (UIdent "A")),VarT (V (UIdent "L"))],VarT (V (UIdent "Q_2"))])],Rule (TPred (Complex (Atm (LIdent "step")) [Complex (Atm (LIdent "par")) [VarT (V (UIdent "P_1")),VarT (V (UIdent "Q_1"))],Complex (Atm (LIdent "hear")) [VarT (V (UIdent "A")
[…]

### 2016-04-30 prolog/tool/LaTeX.hs
The semantics appendix .tex is now auto-generated from semantics.eprover

@@ -1,9 +1,15 @@
 module LaTeX where
 -- pretty-printer generated by the BNF converter + tweaked
+import qualified System.IO           as Sys ( stdin, hGetContents )
+import qualified System.Environment  as Sys ( getArgs, getProgName )
+import qualified System.Exit         as Sys ( exitFailure, exitSuccess )
+
+import qualified Eprover.Par   as P
+import qualified Eprover.Print as P
+import qualified Eprover.Lex   as L
+import qualified Eprover.ErrM  as E
 import Eprover.Abs
 import Data.Char
-
-import Sem
 
 -- the top-level printing method
 printTree :: Print a => a -> String
@@ -162,4 +168,48 @@ instance Print Var where
     V uident  -> prPrec i 0 (concatD [prt 0 uident])
     A wild    -> prPrec i 0 (concatD [prt 0 wild])
 
-main = putStrLn (printTree sem)
+----------------------------------------------------------------
+type ParseFun a = [L.Token] -> E.Err a
+type Verbosity = Int
+
+run :: (Print a, Show a) => Verbosity -> ParseFun a -> String -> IO ()
+run v p s = let ts = P.myLexer s in case p ts of
+           E.Bad s    -> do putStrLn "\nParse              Failed...\n"
+                            putStrLnV v "Tokens:"
+                            putStrLnV v $ show ts
+                            putStrLn s
+                            Sys.exitFailure
+           E.Ok  tree -> do putStrLnV v "\nParse Successful!"
+                            putStrLn (printTree tree)
+                            Sys.exitSuccess
+
+putStrLnV :: Verbosity -> String -> IO ()
+putStrLnV v s = if v > 1 then putStrLn s else return ()
+
+runFile :: (Print a, Show a) => Verbosity -> ParseFun a -> FilePath -> IO ()
+runFile v p f = putStrLnV v f >>
+                readFile f >>= run v p
+
+usage :: IO ()
+usage = do
+  putStrLn $ unlines
+    [ "usage: Call with one of the following argument combinations:"
+    , "  --help          Display this help message."
+    , "  (no arguments)  Parse stdin verbosely."
+    , "  (files)         Parse content of files verbosely."
+    , "  -s (files)      Silent mode (default). Parse content of files silently."
+    , "  -v (files)      Verbose mode."
+    ]
+  Sys.exitFailure
+
[…]

### 2016-05-09 NewARSim/NewARSim.hs
Re-exporting Dynamics

@@ -36,7 +36,7 @@ OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
 {-# LANGUAGE TypeFamilies #-}
 {-# LANGUAGE DeriveAnyClass #-}
 
-module NewARSim (module NewARSim, Typeable, Data, mkStdGen, StdGen) where
+module NewARSim (module NewARSim, Typeable, Data, mkStdGen, StdGen, module Dynamics) where
 
 import Control.Monad.Operational
 import Control.Monad.Identity hiding (void)

### 2016-05-09 paper/paper.tex
Appendix now consists of macro-definitions of groups of semantics rules

@@
 
-In the untyped setting of this report, a value $V$ can be any data item computed and communicated by an AUTOSAR system, including the unit value $\void$ and the error codes that might be returned from RTE calls.
+In the untyped setting of this report, a value $V$ can be any data item computed and communicated by an AUTOSAR system, including the unit value $\Tvoid$ and the error codes that might be returned from RTE calls.
 Meta-variables $N$, $B$ and $T$ range over natural numbers, boolean values and (floating point) time values, respectively.
-The client $C$ of a runnable instance is the address of the invoking port operation (if the runnable was invoked by a client-server call), or $\void$ otherwise.
-An activation state $\Act$ either toggles between $\idle$ and $\pending$, or it has the special shape $\serving{\Cs}{\Vs}$ (in the case of a runnable triggered by client-server calls).
+The client $C$ of a runnable instance is the address of the invoking port operation (if the runnable was invoked by a client-server call), or $\Tvoid$ otherwise.
+An activation state $\Act$ either toggles between $\Tidle$ and $\Tpending$, or it has the special shape $\serving{\Cs}{\Vs}$ (in the case of a runnable triggered by client-server calls).
 Formally,
@@
 	T	& \in		& \text{Time values} \\
-	C	& ::=		& \To{O}{U}{I} \sep \void \\
-	\Act	& ::=		& \idle \sep \pending \sep \serving{\Cs}{\Vs} \\
+	C	& ::=		& \To{O}{U}{I} \sep \Tvoid \\
+	\Act	& ::=		& \Tidle \sep \Tpending \sep \serving{\Cs}{\Vs} \\
 \end{array}
@@
 $$
-where $\Act$ is $\serving{\epsilon}{\epsilon}$ if $R$ is triggered by an \emph{OperationInvokedEvent}; else $\Act$ is $\pending$ if $R$ is triggered by an \emph{InitEvent}; otherwise $\Act$ is $\idle$.
+where $\Act$ is $\serving{\epsilon}{\epsilon}$ if $R$ is triggered by an \emph{OperationInvokedEvent}; else $\Act$ is $\Tpending$ if $R$ is triggered by an \emph{InitEvent}; otherwise $\Act$ is $\Tidle$.
 % SWS_Rte_03524 (SWS_Rte_03526?) forbids mixing OperationInvokedEvents with other events for the same runnable
@@
 \begin{array}{c}
-	\excl{\Tx{X}{I}}{\false}
+	\excl{\Tx{X}{I}}{\Tfalse}
 \end{array}
@@
 $$
-where $V$ is the initial value attributed to $S$ if it exists, otherwise $V$ is $\void$.
+where $V$ is the initial value attributed to $S$ if it exists, otherwise $V$ is $\Tvoid$.
 
@@
 \begin{array}{c}
-	\delem{\Tep{E}{U}{I}}{\false}{\text{nodata}}
+	\delem{\Tep{E}{U}{I}}{\Tfalse}{\text{nodata}}
 \end{array}
@@
 We will further abstract away from actual numbers of input and output parameters to these function by assuming that records (structs) are used as the single continuation argument and result whenever the arity so requires.
-Absent arguments and results will be replaced by the unit value $\void$.
+Absent arguments and results will be replaced by the unit value $\Tvoid$.
 
@@
 \input{semantics.tex}
-%\input{../prolog/tool/semantics.tex}
+{\renewcommand{\Prule}[2]{#1 & \Pif & #2\\}
+
+\[ \begin{array}{r@{}c@{}lcr@{}c@{}lr@{}c@{}l}  \CombRed  \end{array} \]
+}
+\subsection{Exclusive areas}
+\[ \begin{array}{r@{}c@{}l}      \ExclusiveAreas          \end{array} \]
+\subsection{Inter-runnable variables}
+\[ \begin{array}{r@{}c@{}l}      \InterrunnableVariables  \end{array} \]
+\subsection{Sending/receiving}
+\[ \begin{array}{r@{}c@{}l}      \SendingReceiving        \end{array} \]
+%TODO: RuleLong?
+\subsection{Reading/writing (unbuffered versions of rcv and snd)}
+\[ \begin{array}{r@{}c@{}lcl}    \ReadingWriting          \end{array} \]
+\subsection{Calling a server}
+\[ \begin{array}{r@{}c@{}lcl}    \CallServer              \end{array} \]
+\subsection{Obtaining a server result}
+\[ \begin{array}{r@{}c@{}lcl}    \ServerResult            \end{array} \]
+\subsection{Spawning and terminating}
+\[ \begin{array}{r@{}c@{}lcl}    \SpawnTerminate          \end{array} \]
+\subsection{Passing time}
+\[ \begin{array}{r@{}c@{}lcl}    \PassingTime             \end{array} \]
+\subsection{Ignoring broadcasts}
+\[ \begin{array}{r@{}c@{}lcl}    \IgnoreBroadcast         \end{array} \]
 
 \end{document}
+
+%\[
+%\begin{array}{r@{}c@{}llllll}
+
+%\input{../prolog/tool/semantics.tex}

### 2016-05-15 paper/paper.tex
Rename U to P + small fixes

@@
 
-%TODO: port name convention change from U to P? (see ../prolog/semantics.eprover)
 %
@@
   R & \in & \text{Runnable names} \\
-	U	& \in		& \text{Port names} \\
+  P & \in & \text{Port names} \\
   E & \in & \text{Sender-receiver element names} \\
@@
 Inter-runnable variable $S$ of component $I$, currently holding value $V$.
-\item $\delem{\Tep{E}{U}{I}}{B}{V}$ \newline
-Non-queued sender-receiver data element $E$ in port $U$ of component $I$, currently holding value $V$ with boolean update status flag $B$.
-\item $\qelem{\Tep{E}{U}{I}}{\Vs}{TODO}$ \newline
-Queued sender-receiver data element $E$ in port $U$ of component $I$, currently holding value sequence $\Vs$.
-\item $\opres{\Top{O}{U}{I}}{\Vs}$ \newline
-Client-server operation $O$ in port $U$ of component $I$, currently holding result value sequence $\Vs$.
+\item $\delem{\Tep{E}{P}{I}}{B}{V}$ \newline
+Non-queued sender-receiver data element $E$ in port $P$ of component $I$, currently holding value $V$ with boolean update status flag $B$.
+\item $\qelem{\Tep{E}{P}{I}}{\Vs}{TODO}$ \newline
+Queued sender-receiver data element $E$ in port $P$ of component $I$, currently holding value sequence $\Vs$.
+\item $\opres{\Top{O}{P}{I}}{\Vs}$ \newline
+Client-server operation $O$ in port $P$ of component $I$, currently holding result value sequence $\Vs$.
 \item $\timer{\Tr{R}{I}}{T}$ \newline  %TODO: timer has only 2 args in the semantics so {T_p} is not described
@@
 Formally,
-$$
+\[
 \begin{array}{rcll}
@@
 	T	& \in		& \text{Time values} \\
-	C	& ::=		& \To{O}{U}{I} \sep \Tvoid \\
+	C	& ::=		& \Top{O}{P}{I} \sep \Tvoid \\
 	\Act	& ::=		& \Tidle \sep \Tpending \sep \serving{\Cs}{\Vs} \\
 \end{array}
-$$
+\]
 We write $\Vs$ for a sequence of values, $\Cs$ for a sequence of clients, etc.
@@
 %TODO: several LaTeX commands here are used with more arguments than in the semantics below
-$$
+\[
 \begin{array}{lll}
@@
 			& |	& \IrvRead{S}{\Cont}			\\
-			& |	& \Send{E}{U}{V}{\Cont}		\\
-			& |	& \Receive{E}{U}{\Cont}		\\
-			& |	& \Write{E}{U}{V}{\Cont}		\\
-			& |	& \Read{E}{U}{\Cont}		\\
-			& |	& \Invalidate{E}{U}{\Cont}		\\
-			& |	& \IsUpdated{E}{U}{\Cont}		\\
-			& |	& \Call{O}{U}{V}{\Cont}		\\
-			& |	& \Result{O}{U}{\Cont}		\\
+			& |	& \Send{E}{P}{V}{\Cont}		\\
+			& |	& \Receive{E}{P}{\Cont}		\\
+			& |	& \Write{E}{P}{V}{\Cont}		\\
+			& |	& \Read{E}{P}{\Cont}		\\
+			& |	& \Invalidate{E}{P}{\Cont}		\\
+			& |	& \IsUpdated{E}{P}{\Cont}		\\
+			& |	& \Call{O}{P}{V}{\Cont}		\\
+			& |	& \Result{O}{P}{\Cont}		\\
 			& |	& \Return{V}				\\
@@
 \end{array}
-$$
+\]
 
 Together with an operator for composing process terms in parallel, the process grammar for our AUTOSAR calculus becomes
-$$
+\[
 \begin{array}{rcll}
@@
 \end{array}
-$$
+\]
 
-TODO: links from syntax and concepts introduced here to the AUTSAR spec
+TODO: links from syntax and concepts introduced here to the AUTOSAR spec
 
@@
 \item For each runnable $R$ of each component $I$, a term
-$$
+\[
 \begin{array}{c}
@@
 \end{array}
-$$
+\]
 where $\Act$ is $\serving{\epsilon}{\epsilon}$ if $R$ is triggered by an \emph{OperationInvokedEvent}; else $\Act$ is $\Tpending$ if $R$ is triggered by an \emph{InitEvent}; otherwise $\Act$ is $\Tidle$.
@@
 \item For each exclusive area $X$ of each atomic component $I$, a term
-$$
+\[
 \begin{array}{c}
@@
 \end{array}
-$$
+\]
 
 \item For each inter-runnable variable $S$ of each atomic component $I$, a term
-$$
+\[
 \begin{array}{c}
@@
 \end{array}
-$$
+\]
 where $V$ is the initial value attributed to $S$ if it exists, otherwise $V$ is $\Tvoid$.
 
-\item For each data element $E$ of each required port $U$ of each atomic component $I$, a term
-$$
+\item For each data element $E$ of each required port $P$ of each atomic component $I$, a term
+\[
 \begin{array}{c}
-	\qelem{\Tep{E}{U}{I}}{\epsilon}{TODO}
+	\qelem{\Tep{E}{P}{I}}{\epsilon}{TODO}
 \end{array}
-$$
+\]
 if the \emph{swImplPolicy} attribute of element $E$ is \emph{queued}; otherwise, a term
-$$
+\[
 \begin{array}{c}
-	\delem{\Tep{E}{U}{I}}{\Tfalse}{\text{nodata}}
+	\delem{\Tep{E}{P}{I}}{\Tfalse}{\text{nodata}}
 \end{array}
-$$
+\]
 
-\item For each operation $O$ of each required port $U$ of each atomic component $I$, a term
-$$
+\item For each operation $O$ of each required port $P$ of each atomic component $I$, a term
+\[
 \begin{array}{c}
-	\opres{\Tep{O}{U}{I}}{\epsilon}
+	\opres{\Tep{O}{P}{I}}{\epsilon}
 \end{array}
-$$
+\]
 
 \item For each \emph{TimingEvent} triggering each runnable $R$ of each atomic component $I$, a term
-$$
+\[
 \begin{array}{c}
@@
 \end{array}
-$$
+\]
 % where $T$ is the period of the event.
@@
 Some further static model information will also be referenced by the transition rules. We refer to the \emph{connectors} of a model using a binary relation $\Rightarrow$ on addresses, defined as
-$$
-	\Tconnect{\Tep{E}{$U_1$}{$I_1$}}{\Tep{E}{$U_2$}{$I_2$}} \;\;\text{for all elements $E$ of port $U_2$}
-$$
-if there is a connector in the model from provided sender-receiver port $U_1$ of component $I_1$ to required port $U_2$ of component $I_2$; and
-$$
-	\Tconnect{\Top{O}{$U_1$}{$I_1$}}{\Top{O}{$U_2$}{$I_2$}} \;\;\text{for all operations $O$ of port $U_1$}
-$$
-if there is a connector in the model from required client-server port $U_1$ of component $I_1$ to provided port $U_2$ of component $I_2$.
+\[
+	\Tconnect{\Tep{E}{$P_1$}{$I_1$}}{\Tep{E}{$P_2$}{$I_2$}} \;\;\text{for all elements $E$ of port $P_2$}
+\]
+if there is a connector in the model from provided sender-receiver port $P_1$ of component $I_1$ to required port $P_2$ of component $I_2$; and
+\[
+	\Tconnect{\Top{O}{$P_1$}{$I_1$}}{\Top{O}{$P_2$}{$I_2$}} \;\;\text{for all operations $O$ of port $P_1$}
+\]
+if there is a connector in the model from required client-server port $P_1$ of component $I_1$ to provided port $P_2$ of component $I_2$.
 
@@
 
-%TODO: Patrik got here 2016-04-27 in a reading + macro conversion round
 \subsection{Exclusive areas}
@@
 \end{document}
-
-%\[
-%\begin{array}{r@{}c@{}llllll}
-
-%\input{../prolog/tool/semantics.tex}

### 2016-05-16 paper/paper.tex
Paper macro fixes

@@
 \begin{itemize}
-\item $\runnable{\Tr{R}{I}}{TODOK}{T}{\Act}{N}$  \newline
-Common state for runnable $R$ of component $I$, showing $N$ current instances, $T$ seconds left until next possible instantiation, and activation state $\Act$.
-\item $\rinst{\Tr{R}{I}}{C}{\Xs}{\Code}$ \newline
+\item $\Trunnable{\Tr{R}{I}}{K}{T}{\Act}{N}$  \newline
+Common state for runnable $R$ of component $I$ with continuation $K$, showing $N$ current instances, $T$ seconds left until next possible instantiation, and activation state $\Act$.
+\item $\Trinst{\Tr{R}{I}}{C}{\Xs}{\Code}$ \newline
 An instance of runnable $R$ of component $I$, currently executing $\Code$ and owning the exclusive areas $\Xs$, possibly running on behalf of client $C$.
-\item $\excl{\Tx{X}{I}}{B}$ \newline
+\item $\Texcl{\Tx{X}{I}}{B}$ \newline
 Exclusive area $X$ of component $I$, with boolean occupation state $B$.
-\item $\irv{\Ts{S}{I}}{V}$ \newline
+\item $\Tirv{\Ts{S}{I}}{V}$ \newline
 Inter-runnable variable $S$ of component $I$, currently holding value $V$.
-\item $\delem{\Tep{E}{P}{I}}{B}{V}$ \newline
+\item $\Tdelem{\Tep{E}{P}{I}}{B}{V}$ \newline
 Non-queued sender-receiver data element $E$ in port $P$ of component $I$, currently holding value $V$ with boolean update status flag $B$.
-\item $\qelem{\Tep{E}{P}{I}}{\Vs}{TODO}$ \newline
-Queued sender-receiver data element $E$ in port $P$ of component $I$, currently holding value sequence $\Vs$.
-\item $\opres{\Top{O}{P}{I}}{\Vs}$ \newline
+\item $\Tqelem{\Tep{E}{P}{I}}{N}{\Vs}$ \newline
+Queued sender-receiver data element $E$ in port $P$ of component $I$, currently holding value sequence $\Vs$. TODO Maximum queue capacity? $N$.
+\item $\Topres{\Top{O}{P}{I}}{\Vs}$ \newline
 Client-server operation $O$ in port $P$ of component $I$, currently holding result value sequence $\Vs$.
-\item $\timer{\Tr{R}{I}}{T}$ \newline  %TODO: timer has only 2 args in the semantics so {T_p} is not described
+\item $\Ttimer{\Tr{R}{I}}{T}$ \newline  %TODO: timer has only 2 args in the semantics so {T_p} is not described
 Countdown timer state for runnable $R$ of component $I$ with $T$ seconds left of its current cycle.
@@
 The client $C$ of a runnable instance is the address of the invoking port operation (if the runnable was invoked by a client-server call), or $\Tvoid$ otherwise.
-An activation state $\Act$ either toggles between $\Tidle$ and $\Tpending$, or it has the special shape $\serving{\Cs}{\Vs}$ (in the case of a runnable triggered by client-server calls).
+An activation state $\Act$ either toggles between $\Tidle$ and $\Tpending$, or it has the special shape $\Tserving{\Cs}{\Vs}$ (in the case of a runnable triggered by client-server calls).
 Formally,
@@
 	C	& ::=		& \Top{O}{P}{I} \sep \Tvoid \\
-	\Act	& ::=		& \Tidle \sep \Tpending \sep \serving{\Cs}{\Vs} \\
+	\Act	& ::=		& \Tidle \sep \Tpending \sep \Tserving{\Cs}{\Vs} \\
 \end{array}
@@
 	P,Q 	& ::= &
-		\runnable{A}{K}{T}{\Act}{N}	\\
+		\Trunnable{A}{K}{T}{\Act}{N}	\\
 		& | &
-		\rinst{A}{C}{\Xs}{\Code}	\\
+		\Trinst{A}{C}{\Xs}{\Code}	\\
 		& | &
-		\excl{A}{B}			\\
+		\Texcl{A}{B}			\\
 		& | &
-		\irv{A}{V}				\\
+		\Tirv{A}{V}				\\
 		& | &
-		\delem{A}{B}{V}			\\
+		\Tdelem{A}{B}{V}			\\
 		& | &
-		\qelem{A}{\Vs}{TODO}			\\
+		\Tqelem{A}{N}{\Vs}			\\
 		& | &
-		\opres{A}{\Vs}			\\
+		\Topres{A}{\Vs}			\\
 		& | &
-		\timer{A}{T}		\\
+		\Ttimer{A}{T}		\\
 		& | &
@@
 
-\item For each runnable $R$ of each component $I$, a term
+\item For each runnable $R$ of each component $I$ with continuation (code) $K$, a term
 \[
 \begin{array}{c}
-	\runnable{\Tr{R}{I}}{TODO}{0.0}{\Act}{0}
+	\Trunnable{\Tr{R}{I}}{K}{0.0}{\Act}{0}
 \end{array}
 \]
-where $\Act$ is $\serving{\epsilon}{\epsilon}$ if $R$ is triggered by an \emph{OperationInvokedEvent}; else $\Act$ is $\Tpending$ if $R$ is triggered by an \emph{InitEvent}; otherwise $\Act$ is $\Tidle$.
+where $\Act$ is $\Tserving{\epsilon}{\epsilon}$ if $R$ is triggered by an \emph{OperationInvokedEvent}; else $\Act$ is $\Tpending$ if $R$ is triggered by an \emph{InitEvent}; otherwise $\Act$ is $\Tidle$.
 % SWS_Rte_03524 (SWS_Rte_03526?) forbids mixing OperationInvokedEvents with other events for the same runnable
@@
 \begin{array}{c}
-	\excl{\Tx{X}{I}}{\Tfalse}
+	\Texcl{\Tx{X}{I}}{\Tfalse}
 \end{array}
@@
 \begin{array}{c}
-	\irv{\Ts{S}{I}}{V}
+	\Tirv{\Ts{S}{I}}{V}
 \end{array}
@@
 \begin{array}{c}
-	\qelem{\Tep{E}{P}{I}}{\epsilon}{TODO}
+	\Tqelem{\Tep{E}{P}{I}}{N}{\epsilon}
 \end{array}
 \]
-if the \emph{swImplPolicy} attribute of element $E$ is \emph{queued}; otherwise, a term
+if the \emph{swImplPolicy} attribute of element $E$ is \emph{queued} (with capacity $N$); otherwise, a term
 \[
 \begin{array}{c}
-	\delem{\Tep{E}{P}{I}}{\Tfalse}{\text{nodata}}
+	\Tdelem{\Tep{E}{P}{I}}{\Tfalse}{\text{nodata}}
 \end{array}
@@
 \begin{array}{c}
-	\opres{\Tep{O}{P}{I}}{\epsilon}
+	\Topres{\Tep{O}{P}{I}}{\epsilon}
 \end{array}
@@
 \begin{array}{c}
-	\timer{\Tr{R}{I}}{0.0}
+	\Ttimer{\Tr{R}{I}}{0.0}
 \end{array}

### 2016-05-26 paper/paper.tex
Paper: TABs eliminated
Paper: Fixed rte command arities
PaJa+JoNo meeting

@@
 \usepackage{mathpartir}
+\usepackage{ amssymb }
 % \usepackage{minted}
@@
 
+%TODO: replace varibles with \V{Variable} (to make sure the style matches the semantic rules in the appendix - lower case variables, for example)
+
 The formal approach we have chosen is that of a typical process calculus \cite{TODO}, where a system of concurrent processes evolves in a sequence of atomic steps, as determined by a global set of transition rules.
@@
 \item $\Tqelem{\Tep{E}{P}{I}}{N}{\Vs}$ \newline
-Queued sender-receiver data element $E$ in port $P$ of component $I$, currently holding value sequence $\Vs$. TODO Maximum queue capacity? $N$.
+Queued sender-receiver data element $E$ in port $P$ of component $I$ with maximum capacity $N$, currently holding value sequence $\Vs$.
 \item $\Topres{\Top{O}{P}{I}}{\Vs}$ \newline
 Client-server operation $O$ in port $P$ of component $I$, currently holding result value sequence $\Vs$.
-\item $\Ttimer{\Tr{R}{I}}{T}$ \newline  %TODO: timer has only 2 args in the semantics so {T_p} is not described
-Countdown timer state for runnable $R$ of component $I$ with $T$ seconds left of its current cycle.
+\item $\Ttimer{\Tr{R}{I}}{\V{T_p}}{T}$ \newline
+Countdown timer with period \V{T_p} for runnable $R$ of component $I$ with $T$ seconds left of its current cycle.
 \end{itemize}
@@
 Because the result from an RTE call may in general affect subsequent runnable behavior, we use a continuation-passing style to allow any data-dependencies to be expressed.
-%TODO: several LaTeX commands here are used with more arguments than in the semantics below
+%
 \[
@@
         & |   & \IrvRead{S}{\Cont}              \\
-			& |	& \Send{E}{P}{V}{\Cont}		\\
-			& |	& \Receive{E}{P}{\Cont}		\\
-			& |	& \Write{E}{P}{V}{\Cont}		\\
-			& |	& \Read{E}{P}{\Cont}		\\
-			& |	& \Invalidate{E}{P}{\Cont}		\\
-			& |	& \IsUpdated{E}{P}{\Cont}		\\
-			& |	& \Call{O}{P}{V}{\Cont}		\\
-			& |	& \Result{O}{P}{\Cont}		\\
+        & |   & \Send{\Te{E}{P}}{V}{\Cont}      \\
+        & |   & \Receive{\Te{E}{P}}{\Cont}      \\
+        & |   & \Write{\Te{E}{P}}{V}{\Cont}     \\
+        & |   & \Read{\Te{E}{P}}{\Cont}         \\
+        & |   & \Invalidate{\Te{E}{P}}{\Cont}   \\
+        & |   & \IsUpdated{\Te{E}{P}}{\Cont}    \\
+        & |   & \Call{\To{O}{P}}{V}{\Cont}      \\
+        & |   & \Result{\To{O}{P}}{\Cont}       \\
         & |   & \Return{V}                      \\
@@
 \begin{array}{rcll}
-	P,Q 	& ::= &
-		\Trunnable{A}{K}{T}{\Act}{N}	\\
-		& | &
-		\Trinst{A}{C}{\Xs}{\Code}	\\
-		& | &
-		\Texcl{A}{B}			\\
-		& | &
-		\Tirv{A}{V}				\\
-		& | &
-		\Tdelem{A}{B}{V}			\\
-		& | &
-		\Tqelem{A}{N}{\Vs}			\\
-		& | &
-		\Topres{A}{\Vs}			\\
-		& | &
-		\Ttimer{A}{T}		\\
-		& | &
-		P|Q					\\
+  P,Q   & ::= & \Trunnable{A}{K}{T}{\Act}{N}    \\
+        & |   & \Trinst{A}{C}{\Xs}{\Code}       \\
+        & |   & \Texcl{A}{B}                    \\
+        & |   & \Tirv{A}{V}                     \\
+        & |   & \Tdelem{A}{B}{V}                \\
+        & |   & \Tqelem{A}{N}{\Vs}              \\
+        & |   & \Topres{A}{\Vs}                 \\
+        & |   & \Ttimer{A}{T}                   \\
+        & |   & P|Q                             \\
 \end{array}
@@
 TODO: details on limitations: what's not covered by the calculus
+
+TODO: replace | with \Opar
 
@@
     {P \red{A!D} P' \\ Q \red{A?D} Q'}
-		{P | Q \red{A!D} P' | Q'}
+    {\Tpar{P}{Q} \red{A!D} \Tpar{P'}{Q'}}
 \and
@@
 \begin{array}{c}
-	\Ttimer{\Tr{R}{I}}{0.0}
+  \Ttimer{\Tr{R}{I}}{\V{T_p}}{0.0}
 \end{array}
 \]
-% where $T$ is the period of the event.
-%TODO: Does this mean all timed runnables will want to act _immediately_ (at the same time)? No init delay?
+where \V{T_p} is the period of the event.
+%TODO: This means that all timed runnables want to act _immediately_ (at the same time - with no init delay).
 
@@
 
-TODO: links from rules and choices made here to the AUTOSAR spec
+TODO: some links from rules and choices made here to the AUTOSAR spec
 
@@
 \appendix
-\section{Prolog code}
+\section{Semantic rules}
 \label{sec:Prolog}
@@
 \input{semantics.tex}
-{\renewcommand{\Prule}[2]{#1 & \Pif & #2\\}
+{
+\renewcommand{\Prule}[2]{#1 & \Pif & #2\\}
+%\renewcommand{\Pcomma}{,\quad}
+\renewcommand{\Pcomma}{\quad\text{and}\quad}
 
-\[ \begin{array}{r@{}c@{}lcr@{}c@{}lr@{}c@{}l}  \CombRed  \end{array} \]
+%TODO: fix the alignment of p2, q1
+\[ \begin{array}{r@{}c@{}lcr@{}c@{}l@{}r@{}c@{}l}  \CombRed  \end{array} \]
 }
@@
 \subsection{Spawning and terminating}
+%TODO: make the parallell composition more visible
 \[ \begin{array}{r@{}c@{}lcl}    \SpawnTerminate          \end{array} \]

### 2016-05-26 prolog/tool/LaTeX.hs

@@ -141,7 +141,7 @@ instance Print Predicate where
     NPred term -> prPrec i 0 (concatD [doc (showString "\\Pnegate{"), prt 0 term, doc (showString "}")])
     TPred term -> prPrec i 0 (concatD [prt 0 term])
   prtList _ [x] = (concatD [prt 0 x])
-  prtList _ (x:xs) = (concatD [prt 0 x, doc (showString ","), prt 0 xs])
+  prtList _ (x:xs) = (concatD [prt 0 x, doc (showString "\\Pcomma{}"), prt 0 xs])
 
 instance Print Term where
   prt i e = case e of

### 2016-05-28 paper/paper.tex
Papers: made par bar bold, made variable uniformly formatted with \V{} and a few minor fixes

@@
 
-%TODO: replace varibles with \V{Variable} (to make sure the style matches the semantic rules in the appendix - lower case variables, for example)
-
 The formal approach we have chosen is that of a typical process calculus \cite{TODO}, where a system of concurrent processes evolves in a sequence of atomic steps, as determined by a global set of transition rules.
@@
 Components and ports do not carry any dynamic state besides the state of their constituent parts, so they need no explicit representation as process terms.
-However, in order to still be able to talk about model elements using their locally scoped names, we employ a hierarchical addressing scheme, such that address $\Tr{R}{I}$ means runnable $R$ of component $I$ and $\Tep{E}{P}{I}$ means element $E$ of port $P$ within component $I$.
+However, in order to still be able to talk about model elements using their locally scoped names, we employ a hierarchical addressing scheme, such that address $\Tr{\V{R}}{\V{I}}$ means runnable \V{R} of component \V{I} and $\Tep{\V{E}}{\V{P}}{\V{I}}$ means element \V{E} of port \V{P} within component \V{I}.
 We use the following meta-variable conventions:
@@
 \begin{array}{rcll}
-  I & \in & \text{Component names} \\
-  R & \in & \text{Runnable names} \\
-  P & \in & \text{Port names} \\
-  E & \in & \text{Sender-receiver element names} \\
-  O & \in & \text{Client-server operation names} \\
-  S & \in & \text{Inter-runnable variable names} \\
-  X & \in & \text{Exclusive area names} \\
+  \V{I} & \in & \text{Component names} \\
+  \V{R} & \in & \text{Runnable names} \\
+  \V{P} & \in & \text{Port names} \\
+  \V{E} & \in & \text{Sender-receiver element names} \\
+  \V{O} & \in & \text{Client-server operation names} \\
+  \V{S} & \in & \text{Inter-runnable variable names} \\
+  \V{X} & \in & \text{Exclusive area names} \\
 \\
-  A & \in & \text{Addresses} \\
-  A & ::= & \Tr{R}{I}   \sep   \Tep{E}{P}{I}   \sep \Top{O}{P}{I}   \sep   \Ts{S}{I}   \sep   \Tx{X}{I}   \\
+  \V{A} & \in & \text{Addresses} \\
+  \V{A} & ::= & \Tr{\V{R}}{\V{I}}   \sep   \Tep{\V{E}}{\V{P}}{\V{I}}   \sep \Top{\V{O}}{\V{P}}{\V{I}}   \sep   \Ts{\V{S}}{\V{I}}   \sep   \Tx{\V{X}}{\V{I}}   \\
 \end{array}
@@
 %
-Since AUTOSAR components can be arbitrarily nested, the component names $I$ can also be considered to have a similarly nested internal structure (component $I_1$ within component $I_2$ within component $I_3$, etc).
+Since AUTOSAR components can be arbitrarily nested, the component names \V{I} can also be considered to have a similarly nested internal structure (component \V{I_1} within component \V{I_2} within component \V{I_3}, etc).
 We can fully ignore this detail here, though, as the hierarchical layout of components in an AUTOSAR system has no run-time implications \cite{TODO2}.
@@
 \begin{itemize}
-\item $\Trunnable{\Tr{R}{I}}{K}{T}{\Act}{N}$  \newline
-Common state for runnable $R$ of component $I$ with continuation $K$, showing $N$ current instances, $T$ seconds left until next possible instantiation, and activation state $\Act$.
-\item $\Trinst{\Tr{R}{I}}{C}{\Xs}{\Code}$ \newline
-An instance of runnable $R$ of component $I$, currently executing $\Code$ and owning the exclusive areas $\Xs$, possibly running on behalf of client $C$.
-\item $\Texcl{\Tx{X}{I}}{B}$ \newline
-Exclusive area $X$ of component $I$, with boolean occupation state $B$.
-\item $\Tirv{\Ts{S}{I}}{V}$ \newline
-Inter-runnable variable $S$ of component $I$, currently holding value $V$.
-\item $\Tdelem{\Tep{E}{P}{I}}{B}{V}$ \newline
-Non-queued sender-receiver data element $E$ in port $P$ of component $I$, currently holding value $V$ with boolean update status flag $B$.
-\item $\Tqelem{\Tep{E}{P}{I}}{N}{\Vs}$ \newline
-Queued sender-receiver data element $E$ in port $P$ of component $I$ with maximum capacity $N$, currently holding value sequence $\Vs$.
-\item $\Topres{\Top{O}{P}{I}}{\Vs}$ \newline
-Client-server operation $O$ in port $P$ of component $I$, currently holding result value sequence $\Vs$.
-\item $\Ttimer{\Tr{R}{I}}{\V{T_p}}{T}$ \newline
-Countdown timer with period \V{T_p} for runnable $R$ of component $I$ with $T$ seconds left of its current cycle.
+\item $\Trunnable{\Tr{\V{R}}{\V{I}}}{\V{K}}{\V{T}}{\Act}{\V{N}}$  \newline
+Common state for runnable \V{R} of component \V{I} with continuation \V{K}, showing \V{N} current instances, \V{T} seconds left until next possible instantiation, and activation state $\Act$.
+\item $\Trinst{\Tr{\V{R}}{\V{I}}}{\V{C}}{\Xs}{\Code}$ \newline
+An instance of runnable \V{R} of component \V{I}, currently executing $\Code$ and owning the exclusive areas $\Xs$, possibly running on behalf of client \V{C}.
+\item $\Texcl{\Tx{\V{X}}{\V{I}}}{\V{B}}$ \newline
+Exclusive area \V{X} of component \V{I}, with boolean occupation state \V{B}.
+\item $\Tirv{\Ts{\V{S}}{\V{I}}}{\V{V}}$ \newline
+Inter-runnable variable \V{S} of component \V{I}, currently holding value \V{V}.
+\item $\Tdelem{\Tep{\V{E}}{\V{P}}{\V{I}}}{\V{B}}{\V{V}}$ \newline
+Non-queued sender-receiver data element \V{E} in port \V{P} of component \V{I}, currently holding value \V{V} with boolean update status flag \V{B}.
+\item $\Tqelem{\Tep{\V{E}}{\V{P}}{\V{I}}}{\V{N}}{\Vs}$ \newline
+Queued sender-receiver data element \V{E} in port \V{P} of component \V{I} with maximum capacity \V{N}, currently holding value sequence \Vs.
+\item $\Topres{\Top{\V{O}}{\V{P}}{\V{I}}}{\Vs}$ \newline
+Client-server operation \V{O} in port \V{P} of component \V{I}, currently holding result value sequence \Vs.
+\item $\Ttimer{\Tr{\V{R}}{\V{I}}}{\V{T_p}}{\V{T}}$ \newline
+Countdown timer with period \V{T_p} for runnable \V{R} of component \V{I} with \V{T} seconds left of its current cycle.
 \end{itemize}
 
-In the untyped setting of this report, a value $V$ can be any data item computed and communicated by an AUTOSAR system, including the unit value $\Tvoid$ and the error codes that might be returned from RTE calls.
-Meta-variables $N$, $B$ and $T$ range over natural numbers, boolean values and (floating point) time values, respectively.
-The client $C$ of a runnable instance is the address of the invoking port operation (if the runnable was invoked by a client-server call), or $\Tvoid$ otherwise.
+In the untyped setting of this report, a value \V{V} can be any data item computed and communicated by an AUTOSAR system, including the unit value $\Tvoid$ and the error codes that might be returned from RTE calls.
+Meta-variables \V{N}, \V{B} and \V{T} range over natural numbers, boolean values and (floating point) time values, respectively.
+The client \V{C} of a runnable instance is the address of the invoking port operation (if the runnable was invoked by a client-server call), or $\Tvoid$ otherwise.
 An activation state $\Act$ either toggles between $\Tidle$ and $\Tpending$, or it has the special shape $\Tserving{\Cs}{\Vs}$ (in the case of a runnable triggered by client-server calls).
@@
 \begin{array}{rcll}
-  V    & \in & \text{Values} \\
-  N    & \in & \text{Natural numbers} \\
-  B    & \in & \text{Boolean values} \\
-  T    & \in & \text{Time values} \\
-  C    & ::= & \Top{O}{P}{I} \sep \Tvoid \\
+  \V{V}    & \in & \text{Values} \\
+  \V{N}    & \in & \text{Natural numbers} \\
+  \V{B}    & \in & \text{Boolean values} \\
+  \V{T}    & \in & \text{Time values} \\
+  \V{C}    & ::= & \Top{\V{O}}{\V{P}}{\V{I}} \sep \Tvoid \\
   \Act & ::= & \Tidle \sep \Tpending \sep \Tserving{\Cs}{\Vs} \\
@@
 \]
-We write $\Vs$ for a sequence of values, $\Cs$ for a sequence of clients, etc.
+We write \Vs for a sequence of values, $\Cs$ for a sequence of clients, etc.
 Sequences may be empty, which is written $\epsilon$.
@@
 \begin{array}{lll}
-  \Code & ::= & \Enter{X}{\Cont}                \\
-        & |   & \Exit{X}{\Cont}                 \\
-        & |   & \IrvWrite{S}{V}{\Cont}          \\
-        & |   & \IrvRead{S}{\Cont}              \\
-        & |   & \Send{\Te{E}{P}}{V}{\Cont}      \\
-        & |   & \Receive{\Te{E}{P}}{\Cont}      \\
-        & |   & \Write{\Te{E}{P}}{V}{\Cont}     \\
-        & |   & \Read{\Te{E}{P}}{\Cont}         \\
-        & |   & \Invalidate{\Te{E}{P}}{\Cont}   \\
-        & |   & \IsUpdated{\Te{E}{P}}{\Cont}    \\
-        & |   & \Call{\To{O}{P}}{V}{\Cont}      \\
-        & |   & \Result{\To{O}{P}}{\Cont}       \\
-        & |   & \Return{V}                      \\
+  \Code & ::= & \Enter{\V{X}}{\Cont}                \\
+        & |   & \Exit{\V{X}}{\Cont}                 \\
+        & |   & \IrvWrite{\V{S}}{\V{V}}{\Cont}          \\
+        & |   & \IrvRead{\V{S}}{\Cont}              \\
+        & |   & \Send{\Te{\V{E}}{\V{P}}}{\V{V}}{\Cont}      \\
+        & |   & \Receive{\Te{\V{E}}{\V{P}}}{\Cont}      \\
+        & |   & \Write{\Te{\V{E}}{\V{P}}}{\V{V}}{\Cont}     \\
+        & |   & \Read{\Te{\V{E}}{\V{P}}}{\Cont}         \\
+        & |   & \Invalidate{\Te{\V{E}}{\V{P}}}{\Cont}   \\
+        & |   & \IsUpdated{\Te{\V{E}}{\V{P}}}{\Cont}    \\
+        & |   & \Call{\To{\V{O}}{\V{P}}}{\V{V}}{\Cont}      \\
+        & |   & \Result{\To{\V{O}}{\V{P}}}{\Cont}       \\
+        & |   & \Return{\V{V}}                      \\
 \\
-  \Cont & \in & V \rightarrow \Code             \\
+  \Cont & \in & \V{V} \rightarrow \Code             \\
 \end{array}
@@
 \begin{array}{rcll}
-  P,Q   & ::= & \Trunnable{A}{K}{T}{\Act}{N}    \\
-        & |   & \Trinst{A}{C}{\Xs}{\Code}       \\
-        & |   & \Texcl{A}{B}                    \\
-        & |   & \Tirv{A}{V}                     \\
-        & |   & \Tdelem{A}{B}{V}                \\
-        & |   & \Tqelem{A}{N}{\Vs}              \\
-        & |   & \Topres{A}{\Vs}                 \\
-        & |   & \Ttimer{A}{T}                   \\
-        & |   & P|Q                             \\
+  \V{P},\V{Q}   & ::= & \Trunnable{\V{A}}{\V{K}}{\V{T}}{\Act}{\V{N}}    \\
+        & |   & \Trinst{\V{A}}{\V{C}}{\Xs}{\Code}       \\
+        & |   & \Texcl{\V{A}}{\V{B}}                    \\
+        & |   & \Tirv{\V{A}}{\V{V}}                     \\
+        & |   & \Tdelem{\V{A}}{\V{B}}{\V{V}}            \\
+        & |   & \Tqelem{\V{A}}{\V{N}}{\Vs}              \\
+        & |   & \Topres{\V{A}}{\Vs}                     \\
+        & |   & \Ttimer{\V{A}}{\V{T_p}}{\V{T}}          \\
+        & |   & \Tpar{\V{P}}{\V{Q}}                     \\
 \end{array}
@@
 
-TODO: replace | with \Opar
-
 \section{Semantics}
@@
 
-Our semantic approach defines the meaning of an AUTOSAR system as the possible chains of state changes that can be applied to the system's initial state; or equivalently, as the set of \emph{traces} that can be generated from the corresponding process term $P_0$.
+Our semantic approach defines the meaning of an AUTOSAR system as the possible chains of state changes that can be applied to the system's initial state; or equivalently, as the set of \emph{traces} that can be generated from the corresponding process term \V{P_0}.
 A trace is a possibly infinite sequence of transitions
@@
 \begin{array}{c}
-  P_0 \red{L_1} P_1 \red{L_2} P_2 \red{L_3} \cdots
+  \V{P_0} \red{\V{L_1}} \V{P_1} \red{\V{L_2}} \V{P_2} \red{\V{L_3}} \cdots
 \end{array}
@@
 \begin{array}{c}
-  P_{i-1} \red{L_i} P_i
+  \V{P_{i-1}} \red{\V{L_i}} \V{P_i}
 \end{array}
 \]
-states that the system state captured as process term $P_{i-1}$ can evolve into state $P_i$ by means of a single transition labelled $L_i$.
+states that the system state captured as process term \V{P_{i-1}} can evolve into state \V{P_i} by means of a single transition labelled \V{L_i}.
 
-The label $L$ of a transition can be of two kinds, indicating that the affected process term either "says" or "hears" something during the transition.
-A label also always consists of an address $A$ paired with some additional detail $D$, as captured in the following label grammar:
+The label \V{L} of a transition can be of two kinds, indicating that the affected process term either "says" or "hears" something during the transition.
+A label also always consists of an address \V{A} paired with some additional detail \V{D}, as captured in the following label grammar:
 \[
 \begin{array}{rcll}
-  L     & ::= & A!D & \hspace{4em}\text{Say  $A$ and $D$} \\
-        & |   & A?D & \hspace{4em}\text{Hear $A$ and $D$} \\
+  \V{L}     & ::= & \V{A}!\V{D} & \hspace{4em}\text{Say  \V{A} and \V{D}} \\
+            & |   & \V{A}?\V{D} & \hspace{4em}\text{Hear \V{A} and \V{D}} \\
 \end{array}
 \]
-The possible forms of the extra label information $D$ will be introduced as the different transition rules are defined.
+The possible forms of the extra label information \V{D} will be introduced as the different transition rules are defined.
 
-Two processes $P$ and $Q$ can make transitions in parallel if they agree on what is being said or heard.
+Two processes \V{P} and \V{Q} can make transitions in parallel if they agree on what is being said or heard.
 Just as the saying/hearing intuition suggests, at most one of the processes can have the saying role in such an agreement.
@@
   \inferrule
-    {P \red{A!D} P' \\ Q \red{A?D} Q'}
-    {\Tpar{P}{Q} \red{A!D} \Tpar{P'}{Q'}}
+    {\V{P} \red{\V{A}!\V{D}} \V{P'} \\ \V{Q} \red{\V{A}?\V{D}} \V{Q'}}
+    {\Tpar{\V{P}}{\V{Q}} \red{\V{A}!\V{D}} \Tpar{\V{P'}}{\V{Q'}}}
 \and
   \inferrule
-    {P \red{A?D} P' \\ Q \red{A!D} Q'}
-    {P | Q \red{A!D} P' | Q'}
+    {\V{P} \red{\V{A}?\V{D}} \V{P'} \\ \V{Q} \red{\V{A}!\V{D}} \V{Q'}}
+    {\Tpar{\V{P}}{\V{Q}} \red{\V{A}!\V{D}} \Tpar{\V{P'}}{\V{Q'}}}
 \and
   \inferrule
-    {P \red{A?D} P' \\ Q \red{A?D} Q'}
-    {P | Q \red{A?D} P' | Q'}
+    {\V{P} \red{\V{A}?\V{D}} \V{P'} \\ \V{Q} \red{\V{A}?\V{D}} \V{Q'}}
+    {\Tpar{\V{P}}{\V{Q}} \red{\V{A}?\V{D}} \Tpar{\V{P'}}{\V{Q'}}}
 \end{mathpar}
@@
 
-The initial process $P_0$ is determined fully by the AUTOSAR model under study, as the parallel composition of the following terms:
+The initial process \V{P_0} is determined fully by the AUTOSAR model under study, as the parallel composition of the following terms:
 \begin{enumerate}
 
-\item For each runnable $R$ of each component $I$ with continuation (code) $K$, a term
+\item For each runnable \V{R} of each component \V{I} with continuation (code) \V{K}, a term
 \[
 \begin{array}{c}
-  \Trunnable{\Tr{R}{I}}{K}{0.0}{\Act}{0}
+  \Trunnable{\Tr{\V{R}}{\V{I}}}{\V{K}}{0.0}{\Act}{0}
 \end{array}
 \]
-where $\Act$ is $\Tserving{\epsilon}{\epsilon}$ if $R$ is triggered by an \emph{OperationInvokedEvent}; else $\Act$ is $\Tpending$ if $R$ is triggered by an \emph{InitEvent}; otherwise $\Act$ is $\Tidle$.
+where $\Act$ is $\Tserving{\epsilon}{\epsilon}$ if \V{R} is triggered by an \emph{OperationInvokedEvent}; else $\Act$ is $\Tpending$ if \V{R} is triggered by an \emph{InitEvent}; otherwise $\Act$ is $\Tidle$.
 % SWS_Rte_03524 (SWS_Rte_03526?) forbids mixing OperationInvokedEvents with other events for the same runnable
 
-\item For each exclusive area $X$ of each atomic component $I$, a term
+\item For each exclusive area \V{X} of each atomic component \V{I}, a term
 \[
 \begin{array}{c}
-  \Texcl{\Tx{X}{I}}{\Tfalse}
+  \Texcl{\Tx{\V{X}}{\V{I}}}{\Tfalse}
 \end{array}
@@
 
-\item For each inter-runnable variable $S$ of each atomic component $I$, a term
+\item For each inter-runnable variable \V{S} of each atomic component \V{I}, a term
 \[
 \begin{array}{c}
-  \Tirv{\Ts{S}{I}}{V}
+  \Tirv{\Ts{\V{S}}{\V{I}}}{\V{V}}
 \end{array}
 \]
-where $V$ is the initial value attributed to $S$ if it exists, otherwise $V$ is $\Tvoid$.
+where \V{V} is the initial value attributed to \V{S} if it exists, otherwise \V{V} is $\Tvoid$.
 
-\item For each data element $E$ of each required port $P$ of each atomic component $I$, a term
+\item For each data element \V{E} of each required port \V{P} of each atomic component \V{I}, a term
 \[
 \begin{array}{c}
-  \Tqelem{\Tep{E}{P}{I}}{N}{\epsilon}
+  \Tqelem{\Tep{\V{E}}{\V{P}}{\V{I}}}{\V{N}}{\epsilon}
 \end{array}
 \]
-if the \emph{swImplPolicy} attribute of element $E$ is \emph{queued} (with capacity $N$); otherwise, a term
+if the \emph{swImplPolicy} attribute of element \V{E} is \emph{queued} (with capacity \V{N}); otherwise, a term
 \[
 \begin{array}{c}
-  \Tdelem{\Tep{E}{P}{I}}{\Tfalse}{\text{nodata}}
+  \Tdelem{\Tep{\V{E}}{\V{P}}{\V{I}}}{\Tfalse}{\text{nodata}}
 \end{array}
@@
 
-\item For each operation $O$ of each required port $P$ of each atomic component $I$, a term
+\item For each operation \V{O} of each required port \V{P} of each atomic component \V{I}, a term
 \[
 \begin{array}{c}
-  \Topres{\Tep{O}{P}{I}}{\epsilon}
+  \Topres{\Tep{\V{O}}{\V{P}}{\V{I}}}{\epsilon}
 \end{array}
@@
 
-\item For each \emph{TimingEvent} triggering each runnable $R$ of each atomic component $I$, a term
+\item For each \emph{TimingEvent} triggering each runnable \V{R} of each atomic component \V{I}, a term
 \[
 \begin{array}{c}
-  \Ttimer{\Tr{R}{I}}{\V{T_p}}{0.0}
+  \Ttimer{\Tr{\V{R}}{\V{I}}}{\V{T_p}}{0.0}
 \end{array}
@@
 \[
-  \Tconnect{\Tep{E}{$P_1$}{$I_1$}}{\Tep{E}{$P_2$}{$I_2$}} \;\;\text{for all elements $E$ of port $P_2$}
+  \Tconnect{\Tep{\V{E}}{\V{P_1}}{\V{I_1}}}{\Tep{\V{E}}{\V{P_2}}{\V{I_2}}} \;\;\text{for all elements \V{E} of port \V{P_2}}
 \]
-if there is a connector in the model from provided sender-receiver port $P_1$ of component $I_1$ to required port $P_2$ of component $I_2$; and
+if there is a connector in the model from provided sender-receiver port \V{P_1} of component \V{I_1} to required port \V{P_2} of component \V{I_2}; and
 \[
-  \Tconnect{\Top{O}{$P_1$}{$I_1$}}{\Top{O}{$P_2$}{$I_2$}} \;\;\text{for all operations $O$ of port $P_1$}
+  \Tconnect{\Top{\V{O}}{\V{P_1}}{\V{I_1}}}{\Top{\V{O}}{\V{P_2}}{\V{I_2}}} \;\;\text{for all operations \V{O} of port \V{P_1}}
 \]
-if there is a connector in the model from required client-server port $P_1$ of component $I_1$ to provided port $P_2$ of component $I_2$.
+if there is a connector in the model from required client-server port \V{P_1} of component \V{I_1} to provided port \V{P_2} of component \V{I_2}.
 
@@
 
-%TODO: fix the alignment of p2, q1
 \[ \begin{array}{r@{}c@{}lcr@{}c@{}l@{}r@{}c@{}l}  \CombRed  \end{array} \]
@@
 \subsection{Spawning and terminating}
-%TODO: make the parallell composition more visible
 \[ \begin{array}{r@{}c@{}lcl}    \SpawnTerminate          \end{array} \]

### 2016-06-02 NewARSim/NewARSim.hs
Generalized the probe signatures.

@@ -31,6 +31,7 @@ OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
 {-# LANGUAGE DeriveFunctor              #-}
 {-# LANGUAGE ExistentialQuantification  #-}
 {-# LANGUAGE FlexibleInstances          #-}
+{-# LANGUAGE FlexibleContexts           #-}
 {-# LANGUAGE GADTs                      #-}
 {-# LANGUAGE MultiParamTypeClasses      #-}
 {-# LANGUAGE RankNTypes                 #-}
@@ -376,16 +377,20 @@ connectEach prov req = forM_ (prov `zip` req) $ \(p,r) -> connect p r
 
 
 class Addressed a where
-        address                     :: a -> Address
+    type Payload a              :: *
+    address                     :: a -> Address
 
 instance Addressed (InterRunnableVariable a c) where
-        address (IV n)              = n
+    type Payload (InterRunnableVariable a c) = a
+    address (IV n)              = n
 
 instance Addressed (DataElement q a r c) where
-        address (DE n)              = n
+    type Payload (DataElement q a r c) = a
+    address (DE n)              = n
 
 instance Addressed (ClientServerOperation a b r c) where
-        address (OP n)              = n
+    type Payload (ClientServerOperation a b r c) = a
+    address (OP n)              = n
 
 
 type family Seal a where
@@ -462,7 +467,7 @@ fromDyn                     :: Data a => Value -> a
 fromDyn                     = value'
 
 -- TODO: add Reading/Writing classes instead of Addressed?
-probeRead                   :: (Data a, Addressed (e a r c)) => String -> e a r c -> AR c' ()
+probeRead                   :: (Addressed t, Data (Payload t)) => String -> t -> AR c ()
 probeRead s x              = singleton $ NewProbe s g
   where
     g (RD b (Ok v))    | a==b    = Just v
@@ -473,13 +478,13 @@ probeRead s x              = singleton $ NewProbe s g
     a = address x
 
 
-probeWrite                  :: (Data a, Addressed (e a r c)) => String -> e a r c -> AR c' ()
+probeWrite                  :: (Addressed t, Data (Payload t)) => String -> t -> AR c ()
 probeWrite s x            = singleton $ NewProbe s g
   where
     g (IRVW b v)     | a==b     = Just v
     g (WR b v)       | a==b     = Just v
[…]

### 2016-06-02 NewARSim/TicketDispenser.hs
Updated TicketDispenser to new API (bar QuickCheck)

@@ -31,57 +31,59 @@ OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
 module Main where
 
 import NewARSim
-import Test.QuickCheck
-import Test.QuickCheck.Property
+--import Test.QuickCheck
+--import Test.QuickCheck.Property
 import Data.List(nub)
 import System.Random
 import System.CPUTime
 
 
 -- Simple server returning unique tickets to clients.
-ticketDispenser :: AR c (ProvidedOperation () Int ())
-ticketDispenser = component $ do
-                     requestTicketP <- providedOperation
+--ticketDispenser :: AR c (ProvidedOperation () Int ())
+ticketDispenser = atomic $ do
+                     requestTicketP <- providedPort
                      cur <- interRunnableVariable (0 :: Int)
                      excl <- exclusiveArea
                      probeRead "Read" cur
                      probeWrite "Write" cur
-                     let rtBody = do --rteEnter excl
-                                     Ok v <- rteIrvRead cur
-                                     rteIrvWrite cur (v+1)
-                                     --rteExit excl
-                                     return v
-                     serverRunnable Concurrent [requestTicketP] (\() -> rtBody)
+                     serverRunnable Concurrent [OperationInvokedEvent requestTicketP] $ \() -> do
+                                --rteEnter excl
+                                Ok v <- rteIrvRead cur
+                                rteIrvWrite cur (v+1)
+                                --rteExit excl
+                                return v
                      return (seal requestTicketP)
 
-client :: Int -> AR c (RequiredOperation () Int ())
-client n        = component $ do
-                     requestTicketR <- requiredOperation
-                     let clientLoop = do Ok v <- rteCall requestTicketR ()
-                                         printlog "Ticket" v
-                                         clientLoop
-                     runnable Concurrent [Init] clientLoop
+--client :: Int -> AR c (RequiredOperation () Int ())
+client n        = atomic $ do
+                     requestTicketR <- requiredPort
+                     let clientLoop = do 
+                                Ok v <- rteCall requestTicketR ()
+                                printlog "Ticket" v
+                                clientLoop
+                     runnable Concurrent [InitEvent] clientLoop
                      return (seal requestTicketR)
 
 -- Create a server instance and 3 client instances and connect them
-system :: AR c ()
+--system :: AR c ()
 system          = do srv <- ticketDispenser
[…]

### 2016-07-07 paper/paper.tex
Paper overhaul.

@@
 \input{macros.tex}
+\input{semantics.tex}
+
 \setlength{\parindent}{0pt}
@@
 Instead, an explicit goal of AUTOSAR is to constitute an abstract specification that allows multiple and competing realizations, and even systems built as an assembly of fragments from many different (and competing) vendors.
-Such a goal naturally puts the focus on the standard specification itself.
+Such a goal naturally puts a heavy focus on the standard specification itself.
 
-Unfortunately, the AUTOSAR specification is not very rigorous, despite a sheer size of more than one hundred pdf documents and over 12,500 pages of text and UML diagrams in total.
+Unfortunately, the AUTOSAR specification is not very rigorous, despite a sheer size of more than 12,500 pages of text and UML diagrams in total.
 It is also not very abstract, in that it makes frequent references to assumed implementation techniques for the purpose of defining its semantics.
@@
 This paper takes an important step towards a remedy to these problems, by contributing a formal specification of a substantial core of the AUTOSAR standard.
-The formalization covers most of the Software Component Template and its accompanying Run-Time Environment (RTE) (Section~\ref{sec:Calc}), and is able to directly express every legal way a system of SWCs can evolve at run-time on an arbitrarily fast platform (Section~\ref{sec:Sem}).
+The formalization covers most of the Software Component Template and its accompanying Run-Time Environment (RTE) (Section~\ref{sec:Calc}), and is able to directly express every legal way a system of software components can evolve at run-time on an arbitrarily fast platform (Section~\ref{sec:Sem}).
 It can thus serve as a basis for both a concrete AUTOSAR implementation on a specific platform (whose supported behaviors must be a subset of those defined by the formal semantics), and a platform-independent AUTOSAR simulator (where behaviors from the legal set can be picked at random).
@@
 AUTOSAR models are completely static, which means that all components, runnables, ports, connectors, etc, are created ahead of execution time.
-Runnable instances are an exception, but they are only identified by AUTOSAR for conceptual purposes and never part of any model syntax.
+Runnable instances are an exception, but they are only identified by AUTOSAR for conceptual purposes and are never part of any model syntax.
 This static nature renders AUTOSAR models very suitable to graphical notations, which by far is the most widespread representation format for AUTOSAR models.
@@
 \\
-  \V{A} & \in & \text{Addresses} \\
-  \V{A} & ::= & \Tr{\V{R}}{\V{I}}   \sep   \Tep{\V{E}}{\V{P}}{\V{I}}   \sep \Top{\V{O}}{\V{P}}{\V{I}}   \sep   \Ts{\V{S}}{\V{I}}   \sep   \Tx{\V{X}}{\V{I}}   \\
+  \V{A},\V{B} & \in & \text{Addresses} \\
+  \V{A},\V{B} & ::= & \Tr{\V{R}}{\V{I}}   \sep   i.p  \sep  \Tep{\V{E}}{\V{P}}{\V{I}}   \sep \Top{\V{O}}{\V{P}}{\V{I}}   \sep   \Ts{\V{S}}{\V{I}}   \sep   \Tx{\V{X}}{\V{I}}   \\
 \end{array}
@@
 \begin{itemize}
-\item $\Trunnable{\Tr{\V{R}}{\V{I}}}{\V{K}}{\V{T}}{\Act}{\V{N}}$  \newline
-Common state for runnable \V{R} of component \V{I} with continuation \V{K}, showing \V{N} current instances, \V{T} seconds left until next possible instantiation, and activation state $\Act$.
+\item $\Trunnable{\Tr{\V{R}}{\V{I}}}{\V{T}}{\Act}{\V{N}}$  \newline
+Dynamic state for runnable \V{R} of component \V{I}, indicating \V{N} current instances, \V{T} seconds left until next possible instantiation, and activation state $\Act$.
 \item $\Trinst{\Tr{\V{R}}{\V{I}}}{\V{C}}{\Xs}{\Code}$ \newline
 An instance of runnable \V{R} of component \V{I}, currently executing $\Code$ and owning the exclusive areas $\Xs$, possibly running on behalf of client \V{C}.
-\item $\Texcl{\Tx{\V{X}}{\V{I}}}{\V{B}}$ \newline
-Exclusive area \V{X} of component \V{I}, with boolean occupation state \V{B}.
+\item $\Texcl{\Tx{\V{X}}{\V{I}}}{\V{U}}$ \newline
+Exclusive area \V{X} of component \V{I}, with boolean occupation state \V{U}.
 \item $\Tirv{\Ts{\V{S}}{\V{I}}}{\V{V}}$ \newline
 Inter-runnable variable \V{S} of component \V{I}, currently holding value \V{V}.
-\item $\Tdelem{\Tep{\V{E}}{\V{P}}{\V{I}}}{\V{B}}{\V{V}}$ \newline
-Non-queued sender-receiver data element \V{E} in port \V{P} of component \V{I}, currently holding value \V{V} with boolean update status flag \V{B}.
+\item $\Tdelem{\Tep{\V{E}}{\V{P}}{\V{I}}}{\V{U}}{\V{V}}$ \newline
+Non-queued sender-receiver data element \V{E} in port \V{P} of component \V{I}, currently holding value \V{V} with boolean update status flag \V{U}.
 \item $\Tqelem{\Tep{\V{E}}{\V{P}}{\V{I}}}{\V{N}}{\Vs}$ \newline
@@
 \item $\Ttimer{\Tr{\V{R}}{\V{I}}}{\V{T_p}}{\V{T}}$ \newline
-Countdown timer with period \V{T_p} for runnable \V{R} of component \V{I} with \V{T} seconds left of its current cycle.
+Periodic timer for runnable \V{R} of component \V{I}, with a period of \V{T_p} and \V{T} seconds left of its current cycle.
 \end{itemize}
@@
 In the untyped setting of this report, a value \V{V} can be any data item computed and communicated by an AUTOSAR system, including the unit value $\Tvoid$ and the error codes that might be returned from RTE calls.
-Meta-variables \V{N}, \V{B} and \V{T} range over natural numbers, boolean values and (floating point) time values, respectively.
+Meta-variables \V{N}, \V{U} and \V{T} range over natural numbers, boolean values and (floating point) time values, respectively.
 The client \V{C} of a runnable instance is the address of the invoking port operation (if the runnable was invoked by a client-server call), or $\Tvoid$ otherwise.
@@
   \V{N}    & \in & \text{Natural numbers} \\
-  \V{B}    & \in & \text{Boolean values} \\
+  \V{U}    & \in & \text{Boolean values} \\
   \V{T}    & \in & \text{Time values} \\
+  \V{C}    & \in & \text{Clients} \\
+  \Act     & \in & \text{Activation states} \\ \\
   \V{C}    & ::= & \Top{\V{O}}{\V{P}}{\V{I}} \sep \Tvoid \\
@@
 \]
-We write \Vs for a sequence of values, $\Cs$ for a sequence of clients, etc.
+We write $\Vs$ for a sequence of values \V{V}, $\Cs$ for a sequence of clients \V{C}, etc.
 Sequences may be empty, which is written $\epsilon$.
+By $\Tlength{\Vs}$ we mean the length of sequence $\Vs$.
 Left (right) concatenation of an element with a sequence is written $\Tcons{\V{V}}{\V{Vs}}$ ($\Tsnoc{\V{Vs}}{\V{V}}$).
 
-The code part of a runnable instance should technically be a representation of the C or Matlab function that must accompany an AUTOSAR runnable definition once it is complete.
-However, since our task in this report is not to investigate the semantics of these languages in every detail, a more abstract notion of executable code will be required.
-To this end, we have chosen to ignore all details of internal computations, and just capture the observable effects of runnable execution as a pure sequence of RTE (VFB) calls.
-Because the result from an RTE call may in general affect subsequent runnable behavior, we use a continuation-passing style to allow any data-dependencies to be expressed.
+The $\Code$ part of a runnable instance should technically be a representation of the C or Matlab implementation that must accompany an AUTOSAR runnable definition once it is complete.
+However, since our task in this report is not to investigate the semantics of these languages in any detail, a more abstract notion of executable code will be required.
+
+To this end, we have chosen to ignore all internal computations and just capture the observable effects of runnable execution as a pure sequence of RTE (VFB) calls.
+And because the result from an RTE call may in general affect subsequent runnable behavior, we use a continuation-passing style where every $\Code$ term except the final function return actually contains the sequence that follows it as a continuation function $\Cont$.
 %
@@
 \begin{array}{lll}
+  \Code & \in & \text{Code}								\\
+  \Cont & \in & \text{Values} \rightarrow \text{Code}   \\ \\
   \Code & ::= & \Enter{\V{X}}{\Cont}                \\
@@
         & |   & \Return{\V{V}}                      \\
-\\
-  \Cont & \in & \V{V} \rightarrow \Code             \\
 \end{array}
@@
 
-Together with an operator for composing process terms in parallel, the process grammar for our AUTOSAR calculus becomes
+In contrast to $\Code$ sequences, process terms are completely unordered.
+This is expressed in the standard process calculus style, using an associative operator $\Tpar{}{}$ that allows arbitrary sets of primitive processes to be composed in parallel.
+This operator also has a left and right unit in the form of process $\Tzero$, which stands for the empty, or terminated, process.
+The complete grammar for our process terms thus looks as follows:
+
 \[
 \begin{array}{rcll}
-  \V{P},\V{Q}   & ::= & \Trunnable{\V{A}}{\V{K}}{\V{T}}{\Act}{\V{N}}    \\
+  \V{P},\V{Q}   
+  		& ::= & \Trunnable{\V{A}}{\V{T}}{\Act}{\V{N}}    \\
         & |   & \Trinst{\V{A}}{\V{C}}{\Xs}{\Code}       \\
-        & |   & \Texcl{\V{A}}{\V{B}}                    \\
+        & |   & \Texcl{\V{A}}{\V{U}}                    \\
         & |   & \Tirv{\V{A}}{\V{V}}                     \\
-        & |   & \Tdelem{\V{A}}{\V{B}}{\V{V}}            \\
+        & |   & \Tdelem{\V{A}}{\V{U}}{\V{V}}            \\
         & |   & \Tqelem{\V{A}}{\V{N}}{\Vs}              \\
@@
         & |   & \Tpar{\V{P}}{\V{Q}}                     \\
+        & |	  & \Tzero									\\
 \end{array}
@@
 
-Our semantic approach defines the meaning of an AUTOSAR system as the possible chains of state changes that can be applied to the system's initial state; or equivalently, as the set of \emph{traces} that can be generated from the corresponding process term \V{P_0}.
+Our semantic approach defines the meaning of an AUTOSAR system as the possible chains of state changes that can be applied to the system's initial state; or equivalently, as the set of \emph{traces} that can be generated from the initial process term \V{P_0}.
 A trace is a possibly infinite sequence of transitions
@@
 
-The label \V{L} of a transition can be of two kinds, indicating that the affected process term either "says" or "hears" something during the transition.
-A label also always consists of an address \V{A} paired with some additional detail \V{D}, as captured in the following label grammar:
+The label \V{L} of a transition can essentially be of two kinds, indicating that the affected process either "says" or "hears" something during the transition.
+Such a label always consists of an address \V{A} paired with some additional detail \V{D}.
+There is also a special form of "hearing" that denotes letting time pass and which does not carry any address.
+The different labels are captured in the following label grammar.
 \[
@@
             & |   & \V{A}?\V{D} 	& \hspace{4em}\text{Hear \V{A} and \V{D}} \\
+            & |   & \Tdelta{\V{T}}	& \hspace{4em}\text{Let \V{T} seconds pass} \\
 \end{array}
@@
 The following set of inference rules capture this intuition formally.
-\begin{mathpar}
-  \inferrule
-    {\V{P} \red{\V{A}!\V{D}} \V{P'} \\ \V{Q} \red{\V{A}?\V{D}} \V{Q'}}
-    {\Tpar{\V{P}}{\V{Q}} \red{\V{A}!\V{D}} \Tpar{\V{P'}}{\V{Q'}}}
-\and
-  \inferrule
-    {\V{P} \red{\V{A}?\V{D}} \V{P'} \\ \V{Q} \red{\V{A}!\V{D}} \V{Q'}}
-    {\Tpar{\V{P}}{\V{Q}} \red{\V{A}!\V{D}} \Tpar{\V{P'}}{\V{Q'}}}
-\and
-  \inferrule
-    {\V{P} \red{\V{A}?\V{D}} \V{P'} \\ \V{Q} \red{\V{A}?\V{D}} \V{Q'}}
-    {\Tpar{\V{P}}{\V{Q}} \red{\V{A}?\V{D}} \Tpar{\V{P'}}{\V{Q'}}}
-\end{mathpar}
+{
+\renewcommand{\Prule}[2]{#1 & \Pif & #2\\}
+
+\[ \begin{array}{@{}r@{}c@{}l@{\quad\;}c@{\;\,}r@{}c@{}l@{}r@{}c@{}l}  \CombRed  \end{array} \]
+}
 
 The effect of these agreement rules closely resembles the idea of a \emph{broadcast}: what one process says may be heard by every other process in a large parallel composition.
-This behavior serves us well, because in general, what one AUTOSAR runnable does may have impact on many -- if not all -- parts of the executing system.
+This behavior serves us well, because in general, what one AUTOSAR runnable does may have impact on many (if not all) parts of the executing system.
 It is however important not to confuse this broadcasting notion with any particular form of AUTOSAR communication.
-The way transition labels distribute over a parallel system is merely a technical aspect of our process calculus, and will be used to express several different AUTOSAR communication semantics, among other things.
+The way transition labels distribute over the parallel composition operator is merely a technical aspect of our process calculus, and will be used to express several different AUTOSAR communication semantics, among other things.
 
@@
 
-\item For each runnable \V{R} of each component \V{I} with continuation (code) \V{K}, a term
+\item For each runnable \V{R} of each component \V{I}, a term
 \[
 \begin{array}{c}
-  \Trunnable{\Tr{\V{R}}{\V{I}}}{\V{K}}{0.0}{\Act}{0}
+  \Trunnable{\Tr{\V{R}}{\V{I}}}{0}{\Act}{0}
 \end{array}
@@
 \]
-where \V{V} is the initial value attributed to \V{S} if it exists, otherwise \V{V} is $\Tvoid$.
+where \V{V} is the \emph{initValue} attributed to \V{S} if it exists, otherwise \V{V} is $\Tvoid$.
 
@@
 \begin{array}{c}
-  \Tdelem{\Tep{\V{E}}{\V{P}}{\V{I}}}{\Tfalse}{\text{nodata}}
+  \Tdelem{\Tep{\V{E}}{\V{P}}{\V{I}}}{\Tfalse}{\Tnodata}
 \end{array}
@@
 \begin{array}{c}
-  \Ttimer{\Tr{\V{R}}{\V{I}}}{\V{T_p}}{0.0}
+  \Ttimer{\Tr{\V{R}}{\V{I}}}{\V{T_p}}{0}
 \end{array}
@@
 
-Some further static model information will also be referenced by the transition rules. We refer to the \emph{connectors} of a model using a binary relation $\Rightarrow$ on addresses, defined as
+Some further static model information will also be referenced by the transition rules.
+We refer to the connectors of a model using a binary relation $\Rightarrow$ on port addresses, where the arrow points in the \emph{provided-to-required} direction.
+We also assume that $\Rightarrow$ is transitively closed (i.e., it expresses end-to-end connectivity across \emph{Delegation-} as well as \emph{AssemblySwConnector}s), and that every connection $a \Rightarrow b$ is lifted to the \emph{DataElement}s and \emph{ClientServerOperation}s of the connected ports such that
 \[
-  \Tconnect{\Tep{\V{E}}{\V{P_1}}{\V{I_1}}}{\Tep{\V{E}}{\V{P_2}}{\V{I_2}}} \;\;\text{for all elements \V{E} of port \V{P_2}}
+\begin{array}{lll}
+  \Tconnect{\Tp{e}{a}}{\Tp{e}{b}} & \text{for all elements $e$ of port $b$} \\
+  \Tconnect{\Tp{o}{a}}{\Tp{o}{b}} & \text{for all operations $o$ of port $b$} \\
+\end{array}
 \]
-if there is a connector in the model from provided sender-receiver port \V{P_1} of component \V{I_1} to required port \V{P_2} of component \V{I_2}; and
-\[
-  \Tconnect{\Top{\V{O}}{\V{P_1}}{\V{I_1}}}{\Top{\V{O}}{\V{P_2}}{\V{I_2}}} \;\;\text{for all operations \V{O} of port \V{P_1}}
-\]
-if there is a connector in the model from required client-server port \V{P_1} of component \V{I_1} to provided port \V{P_2} of component \V{I_2}.
 
-We expect the \emph{implementation} of a runnable to be available as a continuation $\Cont$, even though the real AUTOSAR model will be refering to concrete functions in external C/Matlab files.
-We will further abstract away from actual numbers of input and output parameters to these function by assuming that records (structs) are used as the single continuation argument and result whenever the arity so requires.
-Absent arguments and results will be replaced by the unit value $\Tvoid$.
+We expect the implementation of a runnable to be available as a continuation function $\Cont$, even though the real AUTOSAR model will be refering to concrete functions in external C/Matlab files.
+We will further abstract away from actual numbers of input and output parameters to these functions by assuming that records (structs) are used as the single continuation argument and result whenever the arity so requires.
+Absent arguments or results will be replaced by the unit value $\Tvoid$.
+
+We write $\Timplementation{i.r}{k}$ to state the assumption that the static AUTOSAR model under study assigns implementation function $k$ to runnable $i.r$.
+Similar notations will be used to express other references to the underlying static model; for example $\emph{DataReceivedEvent}(i.r, p.e)$ to state that the model allows runnable $i.r$ to be triggered by data reception events on port $p.e$ (in component $i$).
+Although the chosen notation is sometimes significantly shorter than the UML/XML-based syntax used in the AUTOSAR specification, the intended meaning should nevertheless be clear.
 
@@
 
-\input{semantics.tex}
-{
-\renewcommand{\Prule}[2]{#1 & \Pif & #2\\}
-%\renewcommand{\Pcomma}{,\quad}
-\renewcommand{\Pcomma}{\quad\text{and}\quad}
-
-\[ \begin{array}{r@{}c@{}lcr@{}c@{}l@{}r@{}c@{}l}  \CombRed  \end{array} \]
-}
 \subsection{Exclusive areas}
@@
 \subsection{Calling a server}
+
+\renewcommand\Pcomma{\\&&\text{and}\;\;}
+
 \[ \begin{array}{r@{}c@{}lcl}    \CallServer              \end{array} \]
-\subsection{Obtaining a server result}
+\subsection{Passing back a server result}
 \[ \begin{array}{r@{}c@{}lcl}    \ServerResult            \end{array} \]
@@
 \subsection{Ignoring broadcasts}
-\[ \begin{array}{r@{}c@{}lcl}    \IgnoreBroadcast         \end{array} \]
+{
+\renewcommand{\Prule}[2]{#1 & \Pif \;\; #2\\}
+\[ \begin{array}{r@{}c@{}lcll}    \IgnoreBroadcast         \end{array} \]
+}
 

### 2016-07-10 ACC/Gearbox.hs
Updates to NewABS2. Hooked up NewABS2 to ACC demo, but it's too slow to run at reasonable speed.

@@ -5,10 +5,8 @@
 -- signal. This module is intended to control the transmission by providing
 -- gear selection signals based on the engine rpm differences over time.
 --
--- TODO: * Re-use Johan's sequencer for gear-switch timer.
---       * Shifting is slightly odd at the moment, but works under "normal"
---         conditions; a 1 second delay is enforced between each consecutive
---         shift, to let the engine revs adjust.
+-- TODO: The state machine could also read an acceleration estimate and provide
+--       more natural shifting that is based on acceleration as well.
 module Gearbox 
   ( GearCtrl(..)
   , gearController
@@ -46,7 +44,6 @@ gearController resolution =
      gears <- interRunnableVariable initialGear
      timer <- interRunnableVariable (False, 0.0)
 
-     -- TODO: Can re-use sequencer for the gear-switch timer!
      runnable (MinInterval 0) [TimingEvent resolution] $
        do res <- rteRead engineRPM
           case res of 
@@ -99,7 +96,7 @@ setGear rpm1 rpm0 gear
 -- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 -- Simple rev-limiting logic.
 --
--- TODO: Re-use Johan's sequencer for timer.
+-- TODO: Finish.
 
 data RevLimit = RevLimit
   { rpmIn          :: DataElem Unqueued Double Required
@@ -115,3 +112,4 @@ revLimit :: Time
          -- ^ RPM limit
          -> AUTOSAR RevLimit
 revLimit resolution limit = undefined
+

### 2016-07-10 ACC/PID.hs

@@ -1,0 +1,83 @@
+-- | AUTOSAR PID controller.
+--
+-- TODO: Apply digital low-pass filter to input; somewhat sensitive to noise/
+--       rapid changes.
+--
+module PID 
+  ( -- * Full PID controller
+    PIDCtrl(..)
+  , pidController
+    -- * PD controller
+  , pdController
+    -- * PI controller
+  , piController
+  ) where
+
+import NewARSim
+
+-- * PID controller
+-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+-- The PID controller produces a linear combination of Propotional-, 
+-- Integrating and Differentiating gain. A few useful special cases (PI-, PD-)
+-- are exported as well.
+
+-- | The PID controller requires a @(Set, Input)@ tuple and provides a control
+-- signal.
+data PIDCtrl = PIDCtrl
+  { pidInput  :: DataElem Unqueued (Double, Double) Required
+  , pidOutput :: DataElem Unqueued Double           Provided
+  }
+
+-- | Create a PID controller.
+pidController :: Time 
+              -- ^ Sample time
+              -> Time
+              -- ^ Derivative time 
+              -> Time
+              -- ^ Integral time
+              -> Double
+              -- ^ Proportional scale 
+              -> AUTOSAR PIDCtrl
+pidController deltaT td ti k = atomic $ do
+     state <- interRunnableVariable (0.0, 0.0)
+
+     pidInput  <- requiredPort
+     pidOutput <- providedPort 
+     comSpec pidOutput (InitValue 0.0)
+
+     runnable (MinInterval 0) [DataReceivedEvent pidInput] $
+       do Ok (ctrl, feedback)   <- rteRead pidInput
+          Ok (prevErr, prevInt) <- rteIrvRead state
+         
+          let err        = ctrl - feedback
+              derivative = (err - prevErr) / deltaT
+              integral   = prevInt + deltaT / ti * err
+              output     = k * (err + integral + td * derivative) 
+          rteIrvWrite state (err, integral) 
+          rteWrite pidOutput output
+     return $ sealBy PIDCtrl pidInput pidOutput
+
+-- | Create a PD controller (proportional and derivative control).
+pdController :: Time 
+             -- ^ Sample time
+             -> Time
[…]

### 2016-07-10 ACC/Simulink.hs

@@ -1,7 +1,7 @@
 -- | External simulation of the ACC component using Simulink (or similar). 
 module Main where
 
-import ACC 
+import Vehicle
 import Control.Monad
 import NewARSim 
 

### 2016-07-10 ACC/Vehicle.hs

@@ -1,0 +1,322 @@
+-- | This module contains all vehicle-related things, such as IO/control of
+-- vehicle subsystems, velocity and slip estimation, et cetera. Most of these
+-- things were previously located in "ACC" or "NewABS2", but were required by
+-- both modules.
+
+{-# LANGUAGE DeriveDataTypeable #-}
+{-# LANGUAGE RecordWildCards    #-}
+
+module Vehicle where
+
+import ACC
+import Control.Monad
+import Gearbox
+import Generic
+import NewARSim 
+import NewABS2
+import Target
+
+-- * Target vehicle sensor
+-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+-- A basic simulated radar. The system is intended to simulate the distance to
+-- some moving object ahead of the vehicle, and report both relative speed as
+-- well as distance to the object ahead. 
+
+-- | Radar controller exports.
+data RadarCtrl = RadarCtrl
+  { distance :: DataElem Unqueued (Maybe Distance) Required
+  , relative :: DataElem Unqueued (Maybe Velo)     Provided
+  }
+
+-- | The radar controller simulates a target vehicle sensor. By reading the
+-- distance to the target vehicle from the environment @radarCtrl@ computes the 
+-- relative velocity between the vehicle and the target vehicle.
+--
+-- TODO: Reads the /absolute/ position of the vehicle, thus computes absolute
+--       speed. Either clarify this or "fix" the MovingTarget code.
+radarCtrl :: Time 
+          -- ^ Sample time
+          -> AUTOSAR RadarCtrl
+radarCtrl deltaT = atomic $ do
+     state    <- interRunnableVariable 0.0
+     distance <- requiredPort
+     relative <- providedPort
+     comSpec relative (InitValue Nothing)
+
+     runnable (MinInterval 0) [TimingEvent deltaT] $ 
+       do Ok dist <- rteRead distance
+          case dist of 
+            Nothing -> rteWrite relative Nothing
+            Just d1 ->
+              do Ok d0 <- rteIrvRead state
+                 let deltaS = (d1 - d0) / deltaT
+                 rteWrite relative (Just deltaS)
+                 rteIrvWrite state d1
[…]

### 2016-07-10 NewARSim/NewARSim.hs
Removed processes from trace. Greatly reduced memory usage, and reduced GC time.

@@ -53,7 +53,7 @@ import           Control.Monad.State.Lazy   hiding (void)
 import           Data.Char                         (ord, chr)
 import           Data.List
 import           Data.Map                          (Map)
-import qualified Data.Map                        as Map
+import qualified Data.Map.Strict                 as Map
 import           Data.Set                          (Set)
 import qualified Data.Set                        as Set
 import           Data.Maybe
@@ -691,10 +691,6 @@ minstart s      = case invocation s of
 trig :: ConnRel -> Address -> Static c -> Bool
 trig conn a s   = or [ a `conn` b | b <- triggers s ]
 
--- Based on a connection relation @conn@, a label and a process, @mayHear@
--- produces a label. Through (mayHear conn) we create a binary operation which
--- is folded over a list of processes. Note that @VETO@ annihilates all elements
--- in the structure.
 mayHear :: ConnRel -> Label -> Proc -> Label
 mayHear conn VETO _                                             = VETO
 mayHear conn (ENTER a)       (Excl b Free)     | a==b           = ENTER a --
@@ -775,16 +771,11 @@ respond _    []     label = label
 respond conn (p:ps) label = respond conn ps acc 
   where acc = mayHear conn label p
 
--- | @'step' conn procs@ produces a list of scheduling options, i.e. progresses
--- the simulation by one step.
 step :: ConnRel -> [Proc] -> [SchedulerOption]
 step conn procs = explore conn [] labels procs  
   where 
     labels  = map (respond conn procs . maySay) procs
 
--- | @'explore' conn _ ls ps@ explores all scheduling options given the
--- connection relation @conn@, a list of labels @ls@ and a list of processes
--- @ps@.
 explore :: ConnRel -> [Proc] -> [Label] -> [Proc] -> [SchedulerOption]
 explore conn pre (VETO:labels) (p:post) = explore conn (p:pre) labels post
 explore conn pre (l:labels)    (p:post) = commit : explore conn (p:pre) labels post
@@ -794,20 +785,27 @@ explore conn pre (l:labels)    (p:post) = commit : explore conn (p:pre) labels p
 explore conn _ _ _                      = []
 
 
--- The simulator proper ---------------------------------------------------------------------------------------
+-- * The simulator proper 
+-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 
-type Logs                   = [(ProbeID,Value)]
[…]

### 2016-08-01 paper/paper.tex
Added equation numbering.
Dump.

@@
 
+The AUTOSAR naming scheme guarantees that all processes terms in such an initial composition will carry unique addresses, with exception of timer terms which share addresses with the runnables they belong to.
+We will later see how runnable instances reuse the address of their runnables in the same manner.
+
 Some further static model information will also be referenced by the transition rules.
@@
 \subsection{Exclusive areas}
+
+An exclusive area is the analogue of a binary semaphore, that can be acquired and released in an atomic fashion by competing runnable instances.
+Rule (1) states that a runnable instance executing $\Enter x k$ may successfully proceed to executing its continuation $k$ by broadcasting $\say{\Tx x i}{\Tenter}$ to its process environment.
+Rule (3) expresses that exclusive area $\Tx x i$ accepts hearing $\hear{\Tx x i}{\Tenter}$ if it is currently not taken; i.e., carries the boolean state $\Tfalse$.
+This makes it possible for a runnable instance and an exclusive area term to engage in a parallel transition:
+\begin{eqnarray}
+	\Tstep{\TparLong{\Trinst{\Tr{\Tvar{R}}{\Tvar{I}}}{\Tvar{C}}{\Tvar{Xs}}{\TrteEnter{\Tvar{X}}{\Tvar{K}}}}
+				{\Texcl{\Tx{\Tvar{X}}{\Tvar{I}}}{\Tfalse}}}
+	  	  {\say{\Tx x i}{\Tenter}}
+		  {\TparLong{\Trinst{\Tr{\Tvar{R}}{\Tvar{I}}}{\Tvar{C}}{\Tcons{\Tvar{X}}{\Tvar{Xs}}}{\Tap{\Tvar{K}}{\Tok}}}
+		  {\Texcl{\Tx{\Tvar{X}}{\Tvar{I}}}{\Ttrue}}}
+\end{eqnarray}
 
@@
 \subsection{Exclusive areas}
-\[ \begin{array}{r@{}c@{}l}      \ExclusiveAreas          \end{array} \]
+\begin{eqnarray}      \ExclusiveAreas          \notag \end{eqnarray}
+
 \subsection{Inter-runnable variables}
-\[ \begin{array}{r@{}c@{}l}      \InterrunnableVariables  \end{array} \]
+\begin{eqnarray}      \InterrunnableVariables  \notag \end{eqnarray}
+
 \subsection{Sending/receiving}
-\[ \begin{array}{r@{}c@{}l}      \SendingReceiving        \end{array} \]
+\begin{eqnarray}      \SendingReceiving        \notag \end{eqnarray}
+
 %TODO: RuleLong?
 \subsection{Reading/writing (unbuffered versions of rcv and snd)}
-\[ \begin{array}{r@{}c@{}lcl}    \ReadingWriting          \end{array} \]
+\begin{eqnarray}      \ReadingWriting          \notag \end{eqnarray}
+
 \subsection{Calling a server}
 
-\renewcommand\Pcomma{\\&&\text{and}\;\;}
+\renewcommand\Pcomma{\nonumber \\&&\text{and}\;\;}
 
-\[ \begin{array}{r@{}c@{}lcl}    \CallServer              \end{array} \]
+\begin{eqnarray}      \CallServer              \notag \end{eqnarray}
+
 \subsection{Passing back a server result}
-\[ \begin{array}{r@{}c@{}lcl}    \ServerResult            \end{array} \]
+\begin{eqnarray}      \ServerResult            \notag \end{eqnarray}
+
 \subsection{Spawning and terminating}
-\[ \begin{array}{r@{}c@{}lcl}    \SpawnTerminate          \end{array} \]
+\begin{eqnarray}      \SpawnTerminate          \notag \end{eqnarray}
+
 \subsection{Passing time}
-\[ \begin{array}{r@{}c@{}lcl}    \PassingTime             \end{array} \]
+\begin{eqnarray}      \PassingTime             \notag \end{eqnarray}
+
+%\renewcommand{\Prule}[2]{#1 & \Pif \;\; #2\\}
 \subsection{Ignoring broadcasts}
-{
-\renewcommand{\Prule}[2]{#1 & \Pif \;\; #2\\}
-\[ \begin{array}{r@{}c@{}lcll}    \IgnoreBroadcast         \end{array} \]
-}
+\begin{eqnarray}      \IgnoreBroadcast         \notag \end{eqnarray}
 

### 2016-08-03 paper/paper.tex
More paper text. Fixed the "limit" problem in the semantics.

@@
 \usepackage{ amssymb }
+\usepackage[makeroom]{cancel}
 % \usepackage{minted}
@@
 There is also a special form of "hearing" that denotes letting time pass and which does not carry any address.
-The different labels are captured in the following label grammar.
+The different labels are captured in the following grammar.
 \[
@@
 {
-\renewcommand{\Prule}[2]{#1 & \Pif & #2\\}
+\renewcommand{\Prule}[2]{#1 \quad \Pif\; #2\\}
+\renewcommand{\Tstep}[3]{#1 \red{#2} #3}
+\renewcommand\Pcomma{\;\text{and}\;}
 
-\[ \begin{array}{@{}r@{}c@{}l@{\quad\;}c@{\;\,}r@{}c@{}l@{}r@{}c@{}l}  \CombRed  \end{array} \]
+\begin{eqnarray}  \CombRed  \notag \end{eqnarray}
 }
@@
 \begin{array}{c}
-  \Tdelem{\Tep{\V{E}}{\V{P}}{\V{I}}}{\Tfalse}{\Tnodata}
+  \Tdelem{\Tep{\V{E}}{\V{P}}{\V{I}}}{\Tfalse}{\TneverReceived}
 \end{array}
@@
 \begin{array}{lll}
-  \Tconnect{\Tp{e}{a}}{\Tp{e}{b}} & \text{for all elements $e$ of port $b$} \\
-  \Tconnect{\Tp{o}{a}}{\Tp{o}{b}} & \text{for all operations $o$ of port $b$} \\
+  \Tconnect{a.e}{b.e} & \text{for all elements $e$ common to $a$ and $b$} \\
+  \Tconnect{a.o}{b.o} & \text{for all operations $o$ common to $a$ and $b$} \\
 \end{array}
@@
 We write $\Timplementation{i.r}{k}$ to state the assumption that the static AUTOSAR model under study assigns implementation function $k$ to runnable $i.r$.
-Similar notations will be used to express other references to the underlying static model; for example $\emph{DataReceivedEvent}(i.r, p.e)$ to state that the model allows runnable $i.r$ to be triggered by data reception events on port $p.e$ (in component $i$).
+Similar notations will be used to express other references to the underlying static model; for example $\TdataReceivedEvent {i.r} {p.e}$ to state that the model allows runnable $i.r$ to be triggered by data reception events on port $p.e$ (in component $i$).
 Although the chosen notation is sometimes significantly shorter than the UML/XML-based syntax used in the AUTOSAR specification, the intended meaning should nevertheless be clear.
+
+The semantic transition axioms are written in full in Appendix~\ref{sec:Rules}.
+We will discuss each group of axioms in turn, starting with the relatively simple behaviour of inter-runnable variables and exclusive areas.
+In most cases, the driving force behind a transition is some runnable instance wishing to say something which other processes either react to or ignore.
+How the runnable instances themselves come into existence will be detailed in Section~\ref{sect:SpawnTerm}.
+
+\subsection{Inter-runnable variables}
+
+An inter-runnable variable (irv) represents state that is local to a particular component and accessible by all runnables of that component.
+Axioms (5) and (6) show the \emph{say} transitions taken by a runnable instance wishing to read and write an irv $s$.
+In both cases the embedded continuation $k$ is applied to the command result.
+The corresponding \emph{hear} transitions of an irv appear as rules (7) and (8).
+Thanks to inference rule (1), a runnable instance and an irv can now engage in a parallel transition, for example expressing the writing of shared data.
+$$
+\begin{array}{c}
+	\Tpar {\Trinst {i.r} c \Xs {\TrteIrvWrite s v k}}
+	 	  {\Tirv {i.s} \_} \vspace{0.6ex} \\ 
+	\red {\say {i.s} {\Tirvw v}} \\
+	\Tpar {\Trinst {i.r} c \Xs {\Tap k \Tvoid}}
+		  {\Tirv {i.s} v}
+\end{array}
+$$
+The corresponding read transition looks as follows:
+$$
+\begin{array}{c}
+	\Tpar {\Trinst {i.r} c \Xs {\TrteIrvRead s k}}
+	 	  {\Tirv {i.s} v} \vspace{0.6ex} \\ 
+	\red{\say {i.s} {\Tirvr v}} \\
+	\Tpar{\Trinst {i.r} c \Xs {\Tap k v}}
+		 {\Tirv {i.s} v}
+\end{array}
+$$
+Section~\ref{sect:Ignore} will show how examples like these can be extended to the case when both participating processes are embedded in arbitrarily large parallel process compositions.
+
+Note that in the last example above, the read value $v$ flows from the irv process to the runnable instance, even though axiom (7) states that an irv actually \emph{hears} the transition payload $\Tirvr{v}$.
+This apparent paradox is simply a reminder that the saying/hearing distinction of our calculus, and the dataflow patterns it may define, are entirely separate notions.
 
@@
 
-An exclusive area is the analogue of a binary semaphore, that can be acquired and released in an atomic fashion by competing runnable instances.
-Rule (1) states that a runnable instance executing $\Enter x k$ may successfully proceed to executing its continuation $k$ by broadcasting $\say{\Tx x i}{\Tenter}$ to its process environment.
-Rule (3) expresses that exclusive area $\Tx x i$ accepts hearing $\hear{\Tx x i}{\Tenter}$ if it is currently not taken; i.e., carries the boolean state $\Tfalse$.
-This makes it possible for a runnable instance and an exclusive area term to engage in a parallel transition:
-\begin{eqnarray}
-	\Tstep{\TparLong{\Trinst{\Tr{\Tvar{R}}{\Tvar{I}}}{\Tvar{C}}{\Tvar{Xs}}{\TrteEnter{\Tvar{X}}{\Tvar{K}}}}
-				{\Texcl{\Tx{\Tvar{X}}{\Tvar{I}}}{\Tfalse}}}
-	  	  {\say{\Tx x i}{\Tenter}}
-		  {\TparLong{\Trinst{\Tr{\Tvar{R}}{\Tvar{I}}}{\Tvar{C}}{\Tcons{\Tvar{X}}{\Tvar{Xs}}}{\Tap{\Tvar{K}}{\Tok}}}
-		  {\Texcl{\Tx{\Tvar{X}}{\Tvar{I}}}{\Ttrue}}}
-\end{eqnarray}
+An exclusive area is the AUTOSAR equivalent of a binary semaphore, that can be acquired and released in an atomic fashion by competing runnable instances.
+Axiom (9) states that a runnable instance wishing to enter exclusive area $x$ may successfully proceed to its continuation $k$ by broadcasting $\say{\Tx x i}{\Tent}$.
+Axiom (11) expresses that exclusive area $\Tx x i$ accepts hearing an ${\Tent}$ payload if it is currently not taken; i.e., carries the boolean state $\Tfalse$.
+This makes the following parallel transition possible:
+$$
+\begin{array}{c}
+	\Trinst {i.r} c \Xs {\TrteEnter x k}
+	\Opar
+	\Texcl {i.x} \Tfalse \vspace{0.6ex} \\ 
+	\red{\say {i.x} \Tent} \\
+	\Trinst{i.r} c {\Tcons x \Xs} {\Tap k \Tvoid}
+	\Opar
+	\Texcl {i.x} \Ttrue
+\end{array}
+$$
+Exiting from an exclusive area is simply the reverse, as defined in axioms (10) and (12):
+$$
+\begin{array}{c}
+	\Trinst {i.r} c {\Tcons x \Xs} {\TrteExit x k}
+	\Opar
+	\Texcl {i.x} \Ttrue \vspace{0.6ex} \\ 
+	\red{\say {i.x} \Tex} \\
+	\Trinst{i.r} c \Xs {\Tap k \Tvoid}
+	\Opar
+	\Texcl {i.x} \Tfalse
+\end{array}
+$$
+However, because axiom (11) is only defined for exclusive areas in the not taken state, there is no way to derive any parallel transition of the following form:
+$$
+\begin{array}{c}
+	p \Opar {\Texcl {i.x} \Ttrue} \red{\say{i.x}{\Tent}} \ldots \\
+\end{array}
+$$
+This means that any runnable instance in $p$ wishing to enter $x$ is effectively blocked from making progress until some other process in $p$ choses to exit $x$.
+In the same manner, a process wishing to exit an exclusive area not taken is also effectively blocked.
+The conditions on exiting are actually stronger than those that govern entering, since axiom (10) is only applicable to a runnable instance for which the exited area $x$ is at the top of its stack of acquired exclusive areas (as expressed by the sequence pattern $\Tcons x \Xs$).
+An exclusive area can thus only be exited by the process that entered it, and only in the reverse order of entering.
+Section~\ref{sect:DiscExcl} will bring up some alternatives to this semantics, and also discuss the interpretation of blocked processes in more detail.
 
-\subsection{Inter-runnable variables}
+\subsection{Unbuffered sending/receiving}
+
+Axioms (13) and (14) define unbuffered interprocess communication from a runnable instance point of view.
+These transitions are just syntactic variants of shared variable reading and writing (axioms (5) and (6)).
+A data element ({\bf delem}) process term contains two pieces of state in addition to its address: a communicated data value and a boolean flag for keeping track of unread writes.
+Axioms (15) and (16) capture reads and writes to the data value and also set the update flag accordingly.
+The side-confition in axiom (16) makes the rule apply only if there is a static connection from the data element being written and the address of the matched term (recall that data element storage is associated with the receiving side of a connection).
+Because AUTOSAR allows sender-receiver communication patterns such that $\Tconnect {i.p.e} a$ and $\Tconnect {i.p.e} b$ are simultaneously true, parallel data element updates like the following are possible:
+$$
+\begin{array}{c}
+	\Trinst {i.r} c \Xs {\TrteWrite {p.e} v k} \,\Opar  \\
+	{\Tdelem a \_ \_} \Opar {\Tdelem b \_ \_} \vspace{0.6ex} \\
+	\red{\say{i.p.e}{\Twr v}} \\
+	\Trinst {i.r} c \Xs {\Tap k \Tvoid} \,\Opar \\
+	{\Tdelem a \Ttrue v} \Opar {\Tdelem b \Ttrue v}
+\end{array}
+$$
+As an aside, note that the associativity of the parallel composition operator is made entirely irrelevant by the symmetric shape of transition rules (1) and (2).
+
+Axiom (17) provides another means to intepret a data element write.
+It expresses that a runnable set up to trigger on data reception on required data element $\Tep e p i$ should be marked as {\bf pending} whenever a write occurs on an element connected to $\Tep e p i$.
+An example may be as follows (assuming $\Tconnect {i.p.e} a$ and $\TdataReceivedEvent b a$:
+$$
+\begin{array}{c}
+	\Trinst {i.r} c \Xs {\TrteWrite {p.e} v k} \,\Opar  \\
+	{\Tdelem a \_ \_} \Opar {\Trunnable b t \_ n} \vspace{0.6ex} \\
+	\red{\say{i.p.e}{\Twr v}} \\
+	\Trinst {i.r} c \Xs {\Tap k \Tvoid} \,\Opar \\
+	{\Tdelem a \Ttrue v} \Opar {\Trunnable b t \Tpending n}
+\end{array}
+$$
+Transitions that actually create an instance of a pending runnable will be introduced in Section~\ref{sect:SpawnTerm}.
+
+A data element can also be queried for its update status flag, and be marked as carrying no valid value. Axioms (18)-(21) define the corresponding transitions.
+
+\subsection{Buffered sending/receiving}
+
+Buffered interprocess communication (axioms (22) - (30)) is essentially a minor variant of the unbuffered mechanism, where the single-value store of a {\bf delem} term is replaced by a {\bf qelem} process holding a sequence of values.
+The buffer is consumed from the left (axiom (25)) and extended to the right (axiom (27)).
+When the buffer is empty, the special value (i.e., the error code) {\bf nodata} is returned to receiving runnable instances (axiom (26)).
+When the buffer is at its maximum capacity and cannot hold more values, the provided value is simply dropped (axiom (28)).
+
+However, the AUTOSAR specification requires
+
+\subsection{Calling a server}
+
+\subsection{Passing back a server result}
+
+\subsection{Spawning and terminating} \label{sect:SpawnTerm}
+
+\subsection{Passing time}
+
+\subsection{Ignoring broadcasts} \label{sect:Ignore}
 
@@
 
+DataReceivedEvent at Invalidate?
+
+DataReceivedEvent at limit response?
+
 Exclusive area nesting
@@
 
+Crashed runnables
+
 \section{Conclusions and Future Work}
@@
 \section{Semantic rules}
-\label{sec:Prolog}
+\label{sec:Rules}
+
+\setcounter{equation}{4}
+
+\subsection{Inter-runnable variables}
+\begin{eqnarray}      \InterrunnableVariables  \notag \end{eqnarray}
 
@@
 
-\subsection{Inter-runnable variables}
-\begin{eqnarray}      \InterrunnableVariables  \notag \end{eqnarray}
+\subsection{Unbuffered sending/receiving}
+\begin{eqnarray}      \ReadingWriting          \notag \end{eqnarray}
 
-\subsection{Sending/receiving}
+\subsection{Buffered sending/receiving}
 \begin{eqnarray}      \SendingReceiving        \notag \end{eqnarray}
@@
 %TODO: RuleLong?
-\subsection{Reading/writing (unbuffered versions of rcv and snd)}
-\begin{eqnarray}      \ReadingWriting          \notag \end{eqnarray}
-
 \subsection{Calling a server}

### 2016-08-10 paper/paper.tex
Major changes to server invocation and return semantics.

@@
 \begin{itemize}
-\item $\Trunnable{\Tr{\V{R}}{\V{I}}}{\V{T}}{\Act}{\V{N}}$  \newline
-Dynamic state for runnable \V{R} of component \V{I}, indicating \V{N} current instances, \V{T} seconds left until next possible instantiation, and activation state $\Act$.
-\item $\Trinst{\Tr{\V{R}}{\V{I}}}{\V{C}}{\Xs}{\Code}$ \newline
-An instance of runnable \V{R} of component \V{I}, currently executing $\Code$ and owning the exclusive areas $\Xs$, possibly running on behalf of client \V{C}.
-\item $\Texcl{\Tx{\V{X}}{\V{I}}}{\V{U}}$ \newline
-Exclusive area \V{X} of component \V{I}, with boolean occupation state \V{U}.
-\item $\Tirv{\Ts{\V{S}}{\V{I}}}{\V{V}}$ \newline
-Inter-runnable variable \V{S} of component \V{I}, currently holding value \V{V}.
-\item $\Tdelem{\Tep{\V{E}}{\V{P}}{\V{I}}}{\V{U}}{\V{V}}$ \newline
-Non-queued sender-receiver data element \V{E} in port \V{P} of component \V{I}, currently holding value \V{V} with boolean update status flag \V{U}.
-\item $\Tqelem{\Tep{\V{E}}{\V{P}}{\V{I}}}{\V{N}}{\Vs}$ \newline
-Queued sender-receiver data element \V{E} in port \V{P} of component \V{I} with maximum capacity \V{N}, currently holding value sequence \Vs.
-\item $\Topres{\Top{\V{O}}{\V{P}}{\V{I}}}{\Vs}$ \newline
-Client-server operation \V{O} in port \V{P} of component \V{I}, currently holding result value sequence \Vs.
-\item $\Ttimer{\Tr{\V{R}}{\V{I}}}{\V{T_p}}{\V{T}}$ \newline
-Periodic timer for runnable \V{R} of component \V{I}, with a period of \V{T_p} and \V{T} seconds left of its current cycle.
+\item $\Trunnable {\Tr r i} t \Act n$  \newline
+Dynamic state for runnable $r$ of component $i$, indicating $n$ current instances, $t$ seconds left until next possible instantiation, and activation state $\Act$.
+\item $\Trinst {\Tr r i} c \Xs \Code$ \newline
+An instance of runnable $r$ of component $i$, currently executing $\Code$ and owning the exclusive areas $\Xs$, possibly running on behalf of client $c$.
+\item $\Texcl {\Tx x i} u$ \newline
+Exclusive area $x$ of component $i$, with boolean occupation state $u$.
+\item $\Tirv {\Ts s i} v$ \newline
+Inter-runnable variable $s$ of component $i$, currently holding value $v$.
+\item $\Tdelem {\Tep e p i} u v$ \newline
+Non-queued sender-receiver data element $e$ in port $p$ of component $i$, currently holding value \V{V} with boolean update status flag $u$.
+\item $\Tqelem {\Tep e p i} n \Vs$ \newline
+Queued sender-receiver data element $e$ in port $p$ of component $i$ with maximum capacity $n$, currently holding value sequence $\Vs$.
+\item $\Toper {\Top o p i} m \Srv$ \newline
+Client-server operation $o$ in port $p$ of component $i$, currently holding sequence number $m$ and call state $\Srv$.
+\item $\Ttimer {\Tr r i} {t_p} t$ \newline
+Periodic timer for runnable $r$ of component $i$, with a period of $t_p$ and $t$ seconds left of its current cycle.
 \end{itemize}
 
-In the untyped setting of this report, a value \V{V} can be any data item computed and communicated by an AUTOSAR system, including the unit value $\Tvoid$ and the error codes that might be returned from RTE calls.
-Meta-variables \V{N}, \V{U} and \V{T} range over natural numbers, boolean values and (floating point) time values, respectively.
-The client \V{C} of a runnable instance is the address of the invoking port operation (if the runnable was invoked by a client-server call), or $\Tvoid$ otherwise.
-An activation state $\Act$ either toggles between $\Tidle$ and $\Tpending$, or it has the special shape $\Tserving{\Cs}{\Vs}$ (in the case of a runnable triggered by client-server calls).
+In the untyped setting of this report, a value $v$ can be any data item computed and communicated by an AUTOSAR system, including the unit value $\Tvoid$ and the error codes that might be returned from RTE calls.
+Meta-variables $n$ and $m$ stand for natural numbers, while $u$ and $t$ range over boolean values and (floating point) time values, respectively.
+The client $c$ of a runnable instance is a tuple holding the address, sequence number and argument value of the current invocation (if the runnable was invoked by a client-server call), or is $\Tvoid$ otherwise.
+An activation state $\Act$ either toggles between $\Tidle$ and $\Tpending$, or is a sequence of client tuples (in the case of a runnable triggered by client-server calls).
+A call state $\Srv$ alternates between indicating an ongoing call with a timeout and a finished call with a result value.
 Formally,
@@
 \begin{array}{rcll}
-  \V{V}    & \in & \text{Values} \\
-  \V{N}    & \in & \text{Natural numbers} \\
-  \V{U}    & \in & \text{Boolean values} \\
-  \V{T}    & \in & \text{Time values} \\
-  \V{C}    & \in & \text{Clients} \\
-  \Act     & \in & \text{Activation states} \\ \\
-  \V{C}    & ::= & \Top{\V{O}}{\V{P}}{\V{I}} \sep \Tvoid \\
-  \Act     & ::= & \Tidle \sep \Tpending \sep \Tserving{\Cs}{\Vs} \\
+  v        & \in & \text{Values} \\
+  n,m      & \in & \text{Natural numbers} \\
+  u        & \in & \text{Boolean values} \\
+  t        & \in & \text{Time values} \\
+  c        & \in & \text{Clients} \\
+  \Act     & \in & \text{Activation states} \\
+  \Srv	   & \in & \text{Call states} \\ \\
+  c        & ::= & \Tclient {\Top o p i} m v \sep \Tvoid \\
+  \Act     & ::= & \Tidle \sep \Tpending \sep \Cs \\
+  \Srv	   & ::= & \Tcalling t \sep \Tdone v
 \end{array}
@@
 To this end, we have chosen to ignore all internal computations and just capture the observable effects of runnable execution as a pure sequence of RTE (VFB) calls.
-And because the result from an RTE call may in general affect subsequent runnable behavior, we use a continuation-passing style where every $\Code$ term except the final function return actually contains the sequence that follows it as a continuation function $\Cont$.
+And because the result from an RTE call may in general affect subsequent runnable behavior, we use a continuation-passing style where every $\Code$ term except the final function return literally the sequence that follows it, as a continuation function $\Cont$.
 %
@@
   \Cont & \in & \text{Values} \rightarrow \text{Code}   \\ \\
-  \Code & ::= & \Enter{\V{X}}{\Cont}                \\
-        & |   & \Exit{\V{X}}{\Cont}                 \\
-        & |   & \IrvWrite{\V{S}}{\V{V}}{\Cont}          \\
-        & |   & \IrvRead{\V{S}}{\Cont}              \\
-        & |   & \Send{\Te{\V{E}}{\V{P}}}{\V{V}}{\Cont}      \\
-        & |   & \Receive{\Te{\V{E}}{\V{P}}}{\Cont}      \\
-        & |   & \Write{\Te{\V{E}}{\V{P}}}{\V{V}}{\Cont}     \\
-        & |   & \Read{\Te{\V{E}}{\V{P}}}{\Cont}         \\
-        & |   & \Invalidate{\Te{\V{E}}{\V{P}}}{\Cont}   \\
-        & |   & \IsUpdated{\Te{\V{E}}{\V{P}}}{\Cont}    \\
-        & |   & \Call{\To{\V{O}}{\V{P}}}{\V{V}}{\Cont}      \\
-        & |   & \Result{\To{\V{O}}{\V{P}}}{\Cont}       \\
-        & |   & \Return{\V{V}}                      \\
+  \Code & ::= & \Enter x \Cont               \\
+        & |   & \Exit x \Cont                \\
+        & |   & \IrvWrite s v \Cont          \\
+        & |   & \IrvRead s \Cont             \\
+        & |   & \Send {\Te e p} v \Cont      \\
+        & |   & \Receive {\Te e p} \Cont     \\
+        & |   & \Write {\Te e p} v \Cont     \\
+        & |   & \Read {\Te e p} \Cont        \\
+        & |   & \Invalidate {\Te e p} \Cont  \\
+        & |   & \IsUpdated {\Te e p} \Cont   \\
+        & |   & \Call {\To o p} v \Cont      \\
+        & |   & \Result {\To o p} \Cont      \\
+        & |   & \Return v                    \\
 \end{array}
@@
 In contrast to $\Code$ sequences, process terms are completely unordered.
-This is expressed in the standard process calculus style, using an associative operator $\Tpar{}{}$ that allows arbitrary sets of primitive processes to be composed in parallel.
+This is expressed in the standard process calculus style using an associative operator $\Tpar{}{}$ that allows arbitrary sets of primitive processes to be composed in parallel.
 This operator also has a left and right unit in the form of process $\Tzero$, which stands for the empty, or terminated, process.
@@
   \V{P},\V{Q}   
-  		& ::= & \Trunnable{\V{A}}{\V{T}}{\Act}{\V{N}}    \\
-        & |   & \Trinst{\V{A}}{\V{C}}{\Xs}{\Code}       \\
-        & |   & \Texcl{\V{A}}{\V{U}}                    \\
-        & |   & \Tirv{\V{A}}{\V{V}}                     \\
-        & |   & \Tdelem{\V{A}}{\V{U}}{\V{V}}            \\
-        & |   & \Tqelem{\V{A}}{\V{N}}{\Vs}              \\
-        & |   & \Topres{\V{A}}{\Vs}                     \\
-        & |   & \Ttimer{\V{A}}{\V{T_p}}{\V{T}}          \\
-        & |   & \Tpar{\V{P}}{\V{Q}}                     \\
+  		& ::= & \Trunnable a t {\Act} n	                \\
+        & |   & \Trinst a c {\Xs} {\Code}               \\
+        & |   & \Texcl a u                              \\
+        & |   & \Tirv a v                               \\
+        & |   & \Tdelem a u v                           \\
+        & |   & \Tqelem a n {\Vs}                       \\
+        & |   & \Toper a m \Srv                         \\
+        & |   & \Ttimer a {\V{T_p}} t                   \\
+        & |   & \Tpar p q                               \\
         & |	  & \Tzero									\\
@@
 \begin{array}{c}
-  \V{P_0} \red{\V{L_1}} \V{P_1} \red{\V{L_2}} \V{P_2} \red{\V{L_3}} \cdots
+  {p_0} \red{l_1} {p_1} \red{l_2} {p_2} \red{l_3} \cdots
 \end{array}
@@
 \begin{array}{c}
-  \V{P_{i-1}} \red{\V{L_i}} \V{P_i}
+  {p_{i-1}} \red{l_i} {p_i}
 \end{array}
 \]
-states that the system state captured as process term \V{P_{i-1}} can evolve into state \V{P_i} by means of a single transition labelled \V{L_i}.
+states that the system state captured as process term $p_{i-1}$ can evolve into state $p_i$ by means of a single transition labelled $l_i$.
 
-The label \V{L} of a transition can essentially be of two kinds, indicating that the affected process either "says" or "hears" something during the transition.
-Such a label always consists of an address \V{A} paired with some additional detail \V{D}.
+The label $l$ of a transition can essentially be of two kinds, indicating that the affected process either "says" or "hears" something during the transition.
+Such a label always consists of an address $a$ paired with some additional detail $d$.
 There is also a special form of "hearing" that denotes letting time pass and which does not carry any address.
@@
 \begin{array}{rcll}
-  \V{L}     & ::= & \V{A}!\V{D} 	& \hspace{4em}\text{Say  \V{A} and \V{D}} \\
-            & |   & \V{A}?\V{D} 	& \hspace{4em}\text{Hear \V{A} and \V{D}} \\
-            & |   & \Tdelta{\V{T}}	& \hspace{4em}\text{Let \V{T} seconds pass} \\
+  l         & ::= & a!d 	    & \hspace{4em}\text{Say  \V{A} and \V{D}} \\
+            & |   & a?d 	    & \hspace{4em}\text{Hear \V{A} and \V{D}} \\
+            & |   & \Tdelta t	& \hspace{4em}\text{Let \V{T} seconds pass} \\
 \end{array}
@@
 
-Two processes \V{P} and \V{Q} can make transitions in parallel if they agree on what is being said or heard.
+Two processes $p$ and $q$ can make transitions in parallel if they agree on what is being said or heard.
 Just as the saying/hearing intuition suggests, at most one of the processes can have the saying role in such an agreement.
@@
 
-The initial process \V{P_0} is determined fully by the AUTOSAR model under study, as the parallel composition of the following terms:
+The initial process $p_0$ is determined fully by the AUTOSAR model under study, as the parallel composition of the following terms:
 \begin{enumerate}
 
-\item For each runnable \V{R} of each component \V{I}, a term
+\item For each runnable $r$ of each component $i$, a term
 \[
 \begin{array}{c}
-  \Trunnable{\Tr{\V{R}}{\V{I}}}{0}{\Act}{0}
+  \Trunnable {\Tr r i} 0 \Act 0
 \end{array}
 \]
-where $\Act$ is $\Tserving{\epsilon}{\epsilon}$ if \V{R} is triggered by an \emph{OperationInvokedEvent}; else $\Act$ is $\Tpending$ if \V{R} is triggered by an \emph{InitEvent}; otherwise $\Act$ is $\Tidle$.
+where $\Act$ is $\epsilon$ if $r$ is triggered by an \emph{OperationInvokedEvent}; else $\Act$ is $\Tpending$ if $r$ is triggered by an \emph{InitEvent}; otherwise $\Act$ is $\Tidle$.
 % SWS_Rte_03524 (SWS_Rte_03526?) forbids mixing OperationInvokedEvents with other events for the same runnable
 
-\item For each exclusive area \V{X} of each atomic component \V{I}, a term
+\item For each exclusive area $x$ of each atomic component $i$, a term
 \[
 \begin{array}{c}
-  \Texcl{\Tx{\V{X}}{\V{I}}}{\Tfalse}
+  \Texcl {\Tx x i} \Tfalse
 \end{array}
@@
 
-\item For each inter-runnable variable \V{S} of each atomic component \V{I}, a term
+\item For each inter-runnable variable $s$ of each atomic component $i$, a term
 \[
 \begin{array}{c}
-  \Tirv{\Ts{\V{S}}{\V{I}}}{\V{V}}
+  \Tirv {\Ts s i} v
 \end{array}
 \]
-where \V{V} is the \emph{initValue} attributed to \V{S} if it exists, otherwise \V{V} is $\Tvoid$.
+where $v$ is the \emph{initValue} attributed to $s$ if it exists, otherwise $v$ is $\Tvoid$.
 
-\item For each data element \V{E} of each required port \V{P} of each atomic component \V{I}, a term
+\item For each data element $e$ of each required port $p$ of each atomic component $i$, a term
 \[
 \begin{array}{c}
-  \Tqelem{\Tep{\V{E}}{\V{P}}{\V{I}}}{\V{N}}{\epsilon}
+  \Tqelem {\Tep e p i} n \epsilon
 \end{array}
 \]
-if the \emph{swImplPolicy} attribute of element \V{E} is \emph{queued} (with capacity \V{N}); otherwise, a term
+if the \emph{swImplPolicy} attribute of element $e$ is \emph{queued} (with capacity $n$); otherwise, a term
 \[
 \begin{array}{c}
-  \Tdelem{\Tep{\V{E}}{\V{P}}{\V{I}}}{\Tfalse}{\TneverReceived}
+  \Tdelem {\Tep e p i} \Tfalse \TneverReceived
 \end{array}
@@
 
-\item For each operation \V{O} of each required port \V{P} of each atomic component \V{I}, a term
+\item For each operation $o$ of each required port $p$ of each atomic component $i$, a term
 \[
 \begin{array}{c}
-  \Topres{\Tep{\V{O}}{\V{P}}{\V{I}}}{\epsilon}
+  \Toper{\Top o p i} 0 {\Tdone \Tvoid}
 \end{array}
@@
 
-\item For each \emph{TimingEvent} triggering each runnable \V{R} of each atomic component \V{I}, a term
+\item For each \emph{TimingEvent} of each runnable $r$ of each atomic component $i$, a term
 \[
 \begin{array}{c}
-  \Ttimer{\Tr{\V{R}}{\V{I}}}{\V{T_p}}{0}
+  \Ttimer {\Tr r i } {t_p} 0
 \end{array}
 \]
-where \V{T_p} is the period of the event.
+where $\V{T_p}$ is the period of the event.
 %TODO: This means that all timed runnables want to act _immediately_ (at the same time - with no init delay).
@@
 An inter-runnable variable (irv) represents state that is local to a particular component and accessible by all runnables of that component.
-Axioms (5) and (6) show the \emph{say} transitions taken by a runnable instance wishing to read and write an irv $s$.
+Axioms (5) and (7) show the \emph{say} transitions taken by a runnable instance wishing to read and write an irv $s$.
 In both cases the embedded continuation $k$ is applied to the command result.
-The corresponding \emph{hear} transitions of an irv appear as rules (7) and (8).
+The corresponding \emph{hear} transitions of an irv appear as rules (6) and (8).
 Thanks to inference rule (1), a runnable instance and an irv can now engage in a parallel transition, for example expressing the writing of shared data.
@@
 	 	  {\Tirv {i.s} v} \vspace{0.6ex} \\ 
-	\red{\say {i.s} {\Tirvr v}} \\
-	\Tpar{\Trinst {i.r} c \Xs {\Tap k v}}
+	\red {\say {i.s} {\Tirvr v}} \\
+	\Tpar {\Trinst {i.r} c \Xs {\Tap k v}}
 		  {\Tirv {i.s} v}
@@
 
-Note that in the last example above, the read value $v$ flows from the irv process to the runnable instance, even though axiom (7) states that an irv actually \emph{hears} the transition payload $\Tirvr{v}$.
-This apparent paradox is simply a reminder that the saying/hearing distinction of our calculus, and the dataflow patterns it may define, are entirely separate notions.
+Note that in the last example above, the read value $v$ flows from the irv process to the runnable instance, even though axiom (6) states that an irv actually \emph{hears} the transition payload $\Tirvr{v}$.
+This apparent paradox is simply a reminder that the saying/hearing distinction of our calculus is entirely separate from the dataflow patterns it defines.
 
@@
 An exclusive area is the AUTOSAR equivalent of a binary semaphore, that can be acquired and released in an atomic fashion by competing runnable instances.
-Axiom (9) states that a runnable instance wishing to enter exclusive area $x$ may successfully proceed to its continuation $k$ by broadcasting $\say{\Tx x i}{\Tent}$.
-Axiom (11) expresses that exclusive area $\Tx x i$ accepts hearing an ${\Tent}$ payload if it is currently not taken; i.e., carries the boolean state $\Tfalse$.
+Axiom (9) states that a runnable instance wishing to enter exclusive area $x$ may successfully proceed to its continuation $k$ by broadcasting $\say {i.x}{\Tent}$.
+Axiom (10) expresses that exclusive area $i.x$ accepts hearing an ${\Tent}$ payload if it is currently not taken; i.e., carries the boolean state $\Tfalse$.
 This makes the following parallel transition possible:
@@
 	\Texcl {i.x} \Tfalse \vspace{0.6ex} \\ 
-	\red{\say {i.x} \Tent} \\
-	\Trinst{i.r} c {\Tcons x \Xs} {\Tap k \Tvoid}
+	\red {\say {i.x} \Tent} \\
+	\Trinst {i.r} c {\Tcons x \Xs} {\Tap k \Tvoid}
 	\Opar
@@
 $$
-Exiting from an exclusive area is simply the reverse, as defined in axioms (10) and (12):
+Exiting from an exclusive area is simply the reverse, as defined in axioms (11) and (12):
 $$
@@
 	\Texcl {i.x} \Ttrue \vspace{0.6ex} \\ 
-	\red{\say {i.x} \Tex} \\
-	\Trinst{i.r} c \Xs {\Tap k \Tvoid}
+	\red {\say {i.x} \Tex} \\
+	\Trinst {i.r} c \Xs {\Tap k \Tvoid}
 	\Opar
@@
 $$
-However, because axiom (11) is only defined for exclusive areas in the not taken state, there is no way to derive any parallel transition of the following form:
+
+Because axiom (10) is only defined for exclusive areas in the not taken state, there is no way to derive any parallel transition of the following form:
 $$
@@
 In the same manner, a process wishing to exit an exclusive area not taken is also effectively blocked.
-The conditions on exiting are actually stronger than those that govern entering, since axiom (10) is only applicable to a runnable instance for which the exited area $x$ is at the top of its stack of acquired exclusive areas (as expressed by the sequence pattern $\Tcons x \Xs$).
+The conditions on exiting are actually stronger than those that govern entering, since axiom (11) is only applicable to a runnable instance for which the exited area $x$ is at the top of its stack of acquired exclusive areas (as expressed by the sequence pattern $\Tcons x \Xs$).
 An exclusive area can thus only be exited by the process that entered it, and only in the reverse order of entering.
@@
 
-Axioms (13) and (14) define unbuffered interprocess communication from a runnable instance point of view.
-These transitions are just syntactic variants of shared variable reading and writing (axioms (5) and (6)).
+Axioms (13) and (15) define unbuffered interprocess communication from a runnable instance point of view.
+These transitions are just syntactic variants of shared variable reading and writing (axioms (5) and (7)).
 A data element ({\bf delem}) process term contains two pieces of state in addition to its address: a communicated data value and a boolean flag for keeping track of unread writes.
-Axioms (15) and (16) capture reads and writes to the data value and also set the update flag accordingly.
+Axioms (14) and (16) capture reads and writes to the data value and also set the update flag accordingly.
 The side-confition in axiom (16) makes the rule apply only if there is a static connection from the data element being written and the address of the matched term (recall that data element storage is associated with the receiving side of a connection).
@@
 
-Axiom (17) provides another means to intepret a data element write.
-It expresses that a runnable set up to trigger on data reception on required data element $\Tep e p i$ should be marked as {\bf pending} whenever a write occurs on an element connected to $\Tep e p i$.
+Axiom (17) provides another means of intepreting a data element write.
+It expresses that a runnable set up to trigger on data reception on required data element $i.p.e$ should be marked as {\bf pending} whenever a write occurs on an element connected to $i.p.e$.
 An example may be as follows (assuming $\Tconnect {i.p.e} a$ and $\TdataReceivedEvent b a$:
@@
 Buffered interprocess communication (axioms (22) - (30)) is essentially a minor variant of the unbuffered mechanism, where the single-value store of a {\bf delem} term is replaced by a {\bf qelem} process holding a sequence of values.
-The buffer is consumed from the left (axiom (25)) and extended to the right (axiom (27)).
-When the buffer is empty, the special value (i.e., the error code) {\bf nodata} is returned to receiving runnable instances (axiom (26)).
-When the buffer is at its maximum capacity and cannot hold more values, the provided value is simply dropped (axiom (28)).
+The buffer is consumed from the left (axiom (23)) and extended to the right (axiom (27)).
+When the buffer is empty, the special value (error code) {\bf nodata} is returned to receiving runnable instances (axiom (24)).
+When the buffer is at its maximum capacity, the communicated value is simply dropped (axiom (28)).
 
-However, the AUTOSAR specification requires
+However, the AUTOSAR specification requires the send command to inform its caller whether all connected buffers were able to successfully store the communicated value or whether some buffers had it dropped.
+Such information is naturally expressed as an additional parameter $\Tvar{as}$ in the {\em snd} transition, listing the addresses of connected {\bf qelem} processes that have no spare capacity.
+The preconditions of axioms (27) and (28) ensure that $\Tvar{as}$ reflects the truth, and axioms (25) and (26) feed different error codes to the sender's continuation depending on whether $\Tvar{as}$ is empty or not.
+Precondition $\Tconnect {i.p.e} \Tvar{as}$ in axiom (26) should be read $\Tconnect {i.p.e} a$ for all $a \in \Tvar{as}$, and its purpose is to outrule irrelevant addresses in $\Tvar{as}$ causing bogus {\bf limit} results.
+
+The following example illustrates a scenario where a sent value is only stored in a subset of the connected buffers. Assuming $\Tconnect {\Tep e p i} {a_1}$ and $\Tconnect {\Tep e p i} {a_2}$:
+$$
+\begin{array}{c}
+	\Trinst {i.r} c \Xs {\TrteSend {p.e} v k} \,\Opar  \\
+	{\Tqelem {a_1} 2 \Teps} \Opar {\Tqelem {a_2} 2 {\Tcons {v_1} {v_2}}} \vspace{0.6ex} \\
+	\red{\say{i.p.e}{\Tsnd v {a_2}}} \\
+	\Trinst {i.r} c \Xs {\Tap k \Tlimit} \,\Opar \\
+	{\Tqelem {a_1} 2 v} \Opar {\Tqelem {a_2} 2 {\Tcons {v_1} {v_2}}}
+\end{array}
+$$
 
@@
 
+The behavior of client-server communication in AUTOSAR depends on whether a client lists the required server port operation as a synchronous or an asynchronous callpoint (which are mutually exclusive model attributes).
+In the former case, the call command is treated as if it were immediately followed by a command for retrieving the server result (axiom (31)), whereas in the latter case, the call succeeds immediately and the result has to be retireved explicitly by the client's continuation code (axiom (32)).
+In both cases, however, an attempt to call the server before it is done processing a previous asynchronous call leads to immediate abortion with the error code {\bf limit} (axiom (35)).
+
+The AUTOSAR specification restricts client-server connections to the many-to-one pattern only, assigning a dedicated runnable resposible for implementing the offered service.
+It furthermore forbids runnables to be triggered by any events other than operation invocations of compatible signatures, making it natural to associate the state needed to buffer up server invocations with the server runnables themselves.
+On the other hand, the specification also introduces timeouts and sequence counters that are private to each connection, which is why our calculus needs process terms that correspond to each required (i.e., client-side) operation as well.
+Axiom (33) shows how the current sequence counter $m$ blazes a call transition if the client-side operation is not already busy, while the busy case is detected in axiom (36).
+In axiom (34), a successful call transition causes the client identifier $a$ to be buffered up in the server runnable together with sequence number $m$ and call parameter $v$.
+Axiom (37) lets a server runnable ignore an aborted call attempt.
+
+Here is an example of how a client runnable instance, a client port operation, and a server runnable interact during a call transition (assuming $b = {i_2.r_2}$, $\ToperationInvokedEvent b {p_2.o}$, $\Tconnect {i_2.p_2.o} {i.p.o}$, and $\TserverCallPointTimeout{i.p.o} t $).
+$$
+\begin{array}{c}
+	{\Trinst {i.r} c \Xs {\TrteCall {p.o} v k}} \,\Opar  \\
+	{\Toper {i.p.o} m {\Tdone \Tnodata}} \Opar \\
+	{\Trunnable b 0 \Cs 0} \vspace{0.6ex} \\
+	\red{\say{i.p.o}{\Tcall m v}} \\
+	{\Trinst {i.r} c \Xs {\Tap k \Tok}} \,\Opar \\
+	{\Toper {i.p.o} m {\Tcalling t}} \Opar \\
+	{\Trunnable b 0 {\Tcons \Cs {\Tclient {i.p.o} m v}}  0} 
+\end{array}
+$$
+
 \subsection{Passing back a server result}
+
+The {\em result} command can appear as an implicit effect of a previous call command, if the operation invoked is marked as an {\em synchronousServerCallPoint} (c.f.\ axiom (31)).
+It can also be referenced explicitly in a client runnable's code, if the current runnable lists the operation as an {\em asynchronousServerCallPoint}).
+For the synchronous case, axiom (38) excludes {\bf nodata} values, which effectively blocks progress of the caller until a server result distinct from {\bf nodata} has been produced.
+The asynchronous case of axiom (39) has no value restrictions, so it will happily pass back the {\bf nodata} tag as well as an indication that a result is not yet available.
+
+The state keeping track of a call's status resides in the {\bf oper} term of a connection.
+Axiom (40) defines that {\bf nodata} is the result provided while a call is open, whereas axiom (41) returns the stored result once a call has completed.
+
+A server runnable instance is identified by a client parameter distinct from $\Tvoid$.
+Axiom (42) says that when such a runnable instance reaches the end of its code sequence, it must emit the produced value together with the sequence number and {\bf oper} address of its current invocation, before it is allowed to terminate (see Section~\ref{sect:SpawnTerm}).
+The addressed {\bf oper} term reacts according to axiom (43), by completing the call, storing the produced value and increasing its sequence counter in preparation for the next call.
+However, should the {\bf oper} timeout counter reach zero before a transition according to (43) can be taken, the client is obliged to unilaterally terminate the call and set the result to error code {\bf timeout} (axiom (44)).
+Axiom (45) enables the completion of a call---successfully or via a timeout---to trigger a subscribing runnable (can be distinct from the original client).
+
+Because of call timeouts, server runnables run the risk of producing results that the original client is not waiting for anymore.
+Detecting such mismatches is the job of sequence numbers, as witnessed by the requirement in axiom (43) that the sequence numbers returned by the server and expected by the client must be identical (variable $m$).
+However, there must also exist a means for server runnables to simply discard their results, otherwise they would not be able to terminate.
+Axiom (46) therefore defines an alternative transition for finished servers, applicable on the condition that the returned sequence number does {\em not} match what the client expects (axiom (47)).
+By means of axiom (48), runnables can ignore the corresponding transition.
+
+An example of a normal server result interaction can look as follows:
+$$
+\begin{array}{c}
+	{\Trinst b {\Tclient {i.p.o} {91} {v_0}} \Xs {\Treturn v}} \,\Opar  \\
+	{\Toper {i.p.o} {91} {\Tcalling t}} \vspace{0.6ex} \\
+	\red{\say{i.p.o}{\Tret {91} v}} \\
+	{\Trinst b \Tvoid \Xs {\Treturn \Tvoid}} \,\Opar \\
+	{\Toper {i.p.o} {92} {\Tdone v}}
+\end{array}
+$$
+As a contrast, this is the behavior observed when a call timeout occurs:
+$$
+\begin{array}{c}
+	{\Trinst b {\Tclient {i.p.o} {91} {v_0}} \Xs {\Treturn v}} \,\Opar  \\
+	{\Toper {i.p.o} {91} {\Tcalling 0}} \vspace{0.6ex} \\
+	\red{\say{i.p.o}{\Tret {91} \Ttimeout}} \\
+	{\Trinst b {\Tclient {i.p.o} {91} {v_0}} \Xs {\Treturn v}} \,\Opar  \\
+	{\Toper {i.p.o} {92} {\Tdone \Ttimeout}}
+\end{array}
+$$
+The processes are now in a state where runnable instance $b$ is prohibited from doing a normal return. Progress is only possible via axioms (46) and (47).
+$$
+\begin{array}{c}
+	{\Trinst b {\Tclient {i.p.o} {91} {v_0}} \Xs {\Treturn v}} \,\Opar  \\
+	{\Toper {i.p.o} {92} {\Tdone \Ttimeout}} \vspace{0.6ex} \\
+	\red{\say{i.p.o}{\Tskip {91}}} \\
+	{\Trinst b \Tvoid \Xs {\Treturn \Tvoid}} \,\Opar \\
+	{\Toper {i.p.o} {92} {\Tdone \Ttimeout}}
+\end{array}
+$$
 
@@
 
-Mixing OperationInvokedEvents with others for same runnable...
+Multiple runnables triggered by same OperationInvokedEvent?
 
 Multiple timingEvent for same runnable, possibly with same period?
+
+Timer initial offsets
 
@@
 Return value of qelem write with fanout > 1 and *some* LIMIT failures
-
-Timer initial offsets
 

### 2016-08-12 paper/paper.tex
More paper sections.

@@
 To this end, we have chosen to ignore all internal computations and just capture the observable effects of runnable execution as a pure sequence of RTE (VFB) calls.
-And because the result from an RTE call may in general affect subsequent runnable behavior, we use a continuation-passing style where every $\Code$ term except the final function return literally the sequence that follows it, as a continuation function $\Cont$.
+And because the result from an RTE call may in general affect subsequent runnable behavior, we use a continuation-passing style where every $\Code$ term except the final function return literally contain the sequence that follows it, as a continuation parameter $\Cont$.
 %
@@
 
-In contrast to $\Code$ sequences, process terms are completely unordered.
+In contrast to code sequences, process terms are completely unordered.
 This is expressed in the standard process calculus style using an associative operator $\Tpar{}{}$ that allows arbitrary sets of primitive processes to be composed in parallel.
@@
 Axiom (17) provides another means of intepreting a data element write.
-It expresses that a runnable set up to trigger on data reception on required data element $i.p.e$ should be marked as {\bf pending} whenever a write occurs on an element connected to $i.p.e$.
+It expresses that a runnable set up to trigger on data reception on required data element $a$ should be marked as {\bf pending} whenever a write occurs on an element connected to $a$.
 An example may be as follows (assuming $\Tconnect {i.p.e} a$ and $\TdataReceivedEvent b a$:
@@
 
-Buffered interprocess communication (axioms (22) - (30)) is essentially a minor variant of the unbuffered mechanism, where the single-value store of a {\bf delem} term is replaced by a {\bf qelem} process holding a sequence of values.
+Buffered interprocess communication (axioms (22) - (30)) is just a minor variant of the unbuffered mechanism, where the single-value store of a {\bf delem} term is replaced by a {\bf qelem} process holding a sequence of values.
 The buffer is consumed from the left (axiom (23)) and extended to the right (axiom (27)).
@@
 
+Runnables with an activation state distinct from $\Tidle$ or $\Teps$ are amenable to instantiation.
+Axiom (49) defines this crucial step for runnables triggered by all types of events except server invocation, which is instead handled by axiom (50).
+Both rules have many details in common:
+\begin{itemize}
+\item They apply only when the runnable instantiation timer (parameter two) has reached zero.
+\item They reset the timer to a model-defined {\em minimumStartInterval} to disable further instantiation immediately after a new instance is born.
+\item They are guarded by the condition that the current number of runnable instances (fourth parameter) is either zero, or the runnable has been defined to accept concurrent invocations.
+\item They make the runnable keep track of the new number of instances.
+\item They let the new instance inherit the address of its parent runnable.
+\item They create the new instance in parallel with its parent (runnable instances is actually the only example of dynamic process creation in this calculus).
+\item They initialize the new instance with code according to the model-defined implementation and an empty list of owned exclusive areas.
+\end{itemize}
+The differences all emanate from the server/non-server distinction:
+\begin{itemize}
+\item The non-server runnable instance gets a $\Tvoid$ client parameter whereas the server instance is assigned the client tuple from the head of the server's client queue.
+\item The non-server runnable toggles back to the $\Tidle$ activation state, while the server just chops of the head of its client queue.
+\item Only the server runnable instance is given a non-void parameter, which is taken from the same client queue head.
+\end{itemize}
+The transition label of axioms (49) and (59) is noteworthy because it is not matched by a corresponding hearing transition; all other process terms just ignore the information that a new runnable instance has been spawned.
+An example of process instantiation within a parallel context will thus merely be a trivial variant of one of the instantiation axioms, and is therefore left out.
+
+Complementing the notion of spawning is a mechanism for runnable instance termination.
+An instance with a $\Tvoid$ client parameter (either acquired natively, or via axioms (42) or (46)) transforms into the terminated process $\Tzero$ as defined in axiom (51).
+Since this term is the unit of parallel process composition, it is silently absorbed by any other process in parallel with it.
+The instance termination is also noted by the associated runnable parent, which reacts by decrementing its current instance counter (axiom (52)).
+An example of this interaction follows below.
+$$
+\begin{array}{c}
+	{\Trunnable b t \Act 3} \Opar
+	{\Trinst b \Tvoid \Teps {\Treturn v}} \vspace{0.6ex} \\
+	\red{\say{i.p.o}{\Tterm}} \\
+	{\Trunnable b t \Act 2} \Opar \Tzero \\
+	\equiv \\
+	{\Trunnable b t \Act 2}
+\end{array}
+$$
+
 \subsection{Passing time}
+
+In common with most real-time process calculi, the concept of passing time in our calculus is kept separate from the computational work expressed by the say/hear transitions.
+This has the advantage of freeing the semantics from possessing a particular computation speed, as any finite number of computation steps can be performed before time must advance.
+On the other hand, it also means that the  limitations of a particular platform cannot be directly studied unless the calculus is complemented with some form of constraint on the work-time relationship.
+Such an extension is beyond the scope of the current report, though.
+
+The passage of time is captured as a special transition relation $\red {\Tdelta t}$, whose only effect on the related process terms is to decrement any contained time variables by $t$.
+For some processes time cannot advance until a particular work transition has been taken; this corresponds to time events that must be noted (although not necessarily acted upon) at the exact time instance when they occur.
+Yet other processes can only take time transitions, which simply means that they represent an idle system that currently has no work to do.
+Mostly, though, work and time transitions are simltaneously available, indicating that the semantics considers the slightly slower and sligthly faster behavior alternatives to be equally correct.
+
+Most process terms actually ignore the passage of time, which is expressed by axioms (53) to (58).
+In axiom (59), an {\bf oper} term in the {\bf calling} state lets time pass by decrementing its timeout counter.
+The counter is not allowed to go negative, though, which forces the {\bf oper} term to handle the timeout once the counter reaches zero (c.f.\ axiom (44)).
+Axiom (60) defines a similar behavior for the start interval timer of a runnable, although this timer is also allowed to slip behind further time updates once it has reached zero (axiom (61)).
+The reason for this is that the interval timers do not enforce any particular action when they time out, they just enable new transition options that do not have to be taken immediately.
+
+A timer process, finally, decrements its counter according to axiom (62).
+When the counter has become zero, the timer is forced to take the tick transition of axiom (63), which also resets the counter to its static period value $t_p$.
+The net effect of such a transition is that the runnable to which the ticking timer belongs is put into the {\bf pending} state (axiom (64)).
+An example of a transition sequence involving a time-triggered runnable follows below.
+$$
+\begin{array}{c}
+	{\Trunnable a 0 \Tidle 0} \Opar
+	{\Ttimer a {2.5} {0.3}} \vspace{0.6ex} \\
+	\red{\Tdelta {0.3}} \\
+	{\Trunnable a 0 \Tidle 0} \Opar
+	{\Ttimer a {2.5} 0} \vspace{0.6ex} \\
+	\red{\say a \Ttick} \\
+	{\Trunnable a 0 \Tpending 0} \Opar
+	{\Ttimer a {2.5} {2.5}} \vspace{0.6ex} \\
+	\red{\say a \Tnew} \\
+	{\Trunnable a {0.8} \Tidle 1} \Opar
+	{\Ttimer a {2.5} {2.5}} \,\Opar \\
+	{\Trinst a \Tvoid \Teps {\Code_1}} \vspace{0.6ex} \\
+	\longrightarrow \cdots \longrightarrow \\
+	{\Trunnable a {0.8} \Tidle 1} \Opar
+	{\Ttimer a {2.5} {2.5}} \,\Opar \\
+	{\Trinst a \Tvoid \Teps {\Code_2}} \vspace{0.6ex} \\
+	\red{\Tdelta {0.8}} \\
+	{\Trunnable a 0 \Tidle 1} \Opar
+	{\Ttimer a {2.5} {1.7}} \,\Opar \\
+	{\Trinst a \Tvoid \Teps {\Code_2}} \vspace{0.6ex} \\
+	\longrightarrow \cdots \longrightarrow \\
+	{\Trunnable a 0 \Tidle 1} \Opar
+	{\Ttimer a {2.5} {2.5}} \,\Opar \\
+	{\Trinst a \Tvoid \Teps {\Treturn \Tvoid}} \vspace{0.6ex} \\
+	\red{\say a \Tterm} \\
+	{\Trunnable a 0 \Tidle 0} \Opar
+	{\Ttimer a {2.5} {1.7}} \vspace{0.6ex} \\
+	\red{\Tdelta {1.5}} \cdots \red{\Tdelta {0.2}} \\
+	{\Trunnable a 0 \Tidle 0} \Opar
+	{\Ttimer a {2.5} 0} \vspace{0.6ex} \\
+	\red{\say a \Ttick} \\
+	\cdots
+\end{array}
+$$
 

### 2016-08-14 paper/paper.tex
paper: small LaTeX fixes

@@
 
-The label $l$ of a transition can essentially be of two kinds, indicating that the affected process either "says" or "hears" something during the transition.
+The label $l$ of a transition can essentially be of two kinds, indicating that the affected process either ``says'' or ``hears'' something during the transition.
 Such a label always consists of an address $a$ paired with some additional detail $d$.
-There is also a special form of "hearing" that denotes letting time pass and which does not carry any address.
+There is also a special form of ``hearing'' that denotes letting time pass and which does not carry any address.
 The different labels are captured in the following grammar.
@@
 These transitions are just syntactic variants of shared variable reading and writing (axioms (5) and (7)).
-A data element ({\bf delem}) process term contains two pieces of state in addition to its address: a communicated data value and a boolean flag for keeping track of unread writes.
+A data element (\textbf{delem}) process term contains two pieces of state in addition to its address: a communicated data value and a boolean flag for keeping track of unread writes.
 Axioms (14) and (16) capture reads and writes to the data value and also set the update flag accordingly.
@@
 Axiom (17) provides another means of intepreting a data element write.
-It expresses that a runnable set up to trigger on data reception on required data element $a$ should be marked as {\bf pending} whenever a write occurs on an element connected to $a$.
+It expresses that a runnable set up to trigger on data reception on required data element $a$ should be marked as \textbf{pending} whenever a write occurs on an element connected to $a$.
 An example may be as follows (assuming $\Tconnect {i.p.e} a$ and $\TdataReceivedEvent b a$:
@@
 
-Buffered interprocess communication (axioms (22) - (30)) is just a minor variant of the unbuffered mechanism, where the single-value store of a {\bf delem} term is replaced by a {\bf qelem} process holding a sequence of values.
+Buffered interprocess communication (axioms (22) - (30)) is just a minor variant of the unbuffered mechanism, where the single-value store of a \textbf{delem} term is replaced by a \textbf{qelem} process holding a sequence of values.
 The buffer is consumed from the left (axiom (23)) and extended to the right (axiom (27)).
-When the buffer is empty, the special value (error code) {\bf nodata} is returned to receiving runnable instances (axiom (24)).
+When the buffer is empty, the special value (error code) \textbf{nodata} is returned to receiving runnable instances (axiom (24)).
 When the buffer is at its maximum capacity, the communicated value is simply dropped (axiom (28)).
@@
 However, the AUTOSAR specification requires the send command to inform its caller whether all connected buffers were able to successfully store the communicated value or whether some buffers had it dropped.
-Such information is naturally expressed as an additional parameter $\Tvar{as}$ in the {\em snd} transition, listing the addresses of connected {\bf qelem} processes that have no spare capacity.
+Such information is naturally expressed as an additional parameter $\Tvar{as}$ in the \emph{snd} transition, listing the addresses of connected \textbf{qelem} processes that have no spare capacity.
 The preconditions of axioms (27) and (28) ensure that $\Tvar{as}$ reflects the truth, and axioms (25) and (26) feed different error codes to the sender's continuation depending on whether $\Tvar{as}$ is empty or not.
-Precondition $\Tconnect {i.p.e} \Tvar{as}$ in axiom (26) should be read $\Tconnect {i.p.e} a$ for all $a \in \Tvar{as}$, and its purpose is to outrule irrelevant addresses in $\Tvar{as}$ causing bogus {\bf limit} results.
+Precondition $\Tconnect {i.p.e} \Tvar{as}$ in axiom (26) should be read $\Tconnect {i.p.e} a$ for all $a \in \Tvar{as}$, and its purpose is to outrule irrelevant addresses in $\Tvar{as}$ causing bogus \textbf{limit} results.
 
@@
 In the former case, the call command is treated as if it were immediately followed by a command for retrieving the server result (axiom (31)), whereas in the latter case, the call succeeds immediately and the result has to be retireved explicitly by the client's continuation code (axiom (32)).
-In both cases, however, an attempt to call the server before it is done processing a previous asynchronous call leads to immediate abortion with the error code {\bf limit} (axiom (35)).
+In both cases, however, an attempt to call the server before it is done processing a previous asynchronous call leads to immediate abortion with the error code \textbf{limit} (axiom (35)).
 
@@
 
-The {\em result} command can appear as an implicit effect of a previous call command, if the operation invoked is marked as an {\em synchronousServerCallPoint} (c.f.\ axiom (31)).
-It can also be referenced explicitly in a client runnable's code, if the current runnable lists the operation as an {\em asynchronousServerCallPoint}).
-For the synchronous case, axiom (38) excludes {\bf nodata} values, which effectively blocks progress of the caller until a server result distinct from {\bf nodata} has been produced.
-The asynchronous case of axiom (39) has no value restrictions, so it will happily pass back the {\bf nodata} tag as well as an indication that a result is not yet available.
+The \emph{result} command can appear as an implicit effect of a previous call command, if the operation invoked is marked as an \emph{synchronousServerCallPoint} (c.f.\ axiom (31)).
+It can also be referenced explicitly in a client runnable's code, if the current runnable lists the operation as an \emph{asynchronousServerCallPoint}).
+For the synchronous case, axiom (38) excludes \textbf{nodata} values, which effectively blocks progress of the caller until a server result distinct from \textbf{nodata} has been produced.
+The asynchronous case of axiom (39) has no value restrictions, so it will happily pass back the \textbf{nodata} tag as well as an indication that a result is not yet available.
 
-The state keeping track of a call's status resides in the {\bf oper} term of a connection.
-Axiom (40) defines that {\bf nodata} is the result provided while a call is open, whereas axiom (41) returns the stored result once a call has completed.
+The state keeping track of a call's status resides in the \textbf{oper} term of a connection.
+Axiom (40) defines that \textbf{nodata} is the result provided while a call is open, whereas axiom (41) returns the stored result once a call has completed.
 
 A server runnable instance is identified by a client parameter distinct from $\Tvoid$.
-Axiom (42) says that when such a runnable instance reaches the end of its code sequence, it must emit the produced value together with the sequence number and {\bf oper} address of its current invocation, before it is allowed to terminate (see Section~\ref{sect:SpawnTerm}).
-The addressed {\bf oper} term reacts according to axiom (43), by completing the call, storing the produced value and increasing its sequence counter in preparation for the next call.
-However, should the {\bf oper} timeout counter reach zero before a transition according to (43) can be taken, the client is obliged to unilaterally terminate the call and set the result to error code {\bf timeout} (axiom (44)).
+Axiom (42) says that when such a runnable instance reaches the end of its code sequence, it must emit the produced value together with the sequence number and \textbf{oper} address of its current invocation, before it is allowed to terminate (see Section~\ref{sect:SpawnTerm}).
+The addressed \textbf{oper} term reacts according to axiom (43), by completing the call, storing the produced value and increasing its sequence counter in preparation for the next call.
+However, should the \textbf{oper} timeout counter reach zero before a transition according to (43) can be taken, the client is obliged to unilaterally terminate the call and set the result to error code \textbf{timeout} (axiom (44)).
 Axiom (45) enables the completion of a call---successfully or via a timeout---to trigger a subscribing runnable (can be distinct from the original client).
@@
 However, there must also exist a means for server runnables to simply discard their results, otherwise they would not be able to terminate.
-Axiom (46) therefore defines an alternative transition for finished servers, applicable on the condition that the returned sequence number does {\em not} match what the client expects (axiom (47)).
+Axiom (46) therefore defines an alternative transition for finished servers, applicable on the condition that the returned sequence number does \emph{not} match what the client expects (axiom (47)).
 By means of axiom (48), runnables can ignore the corresponding transition.
@@
 \item They apply only when the runnable instantiation timer (parameter two) has reached zero.
-\item They reset the timer to a model-defined {\em minimumStartInterval} to disable further instantiation immediately after a new instance is born.
+\item They reset the timer to a model-defined \emph{minimumStartInterval} to disable further instantiation immediately after a new instance is born.
 \item They are guarded by the condition that the current number of runnable instances (fourth parameter) is either zero, or the runnable has been defined to accept concurrent invocations.
@@
 Most process terms actually ignore the passage of time, which is expressed by axioms (53) to (58).
-In axiom (59), an {\bf oper} term in the {\bf calling} state lets time pass by decrementing its timeout counter.
-The counter is not allowed to go negative, though, which forces the {\bf oper} term to handle the timeout once the counter reaches zero (c.f.\ axiom (44)).
+In axiom (59), an \textbf{oper} term in the \textbf{calling} state lets time pass by decrementing its timeout counter.
+The counter is not allowed to go negative, though, which forces the \textbf{oper} term to handle the timeout once the counter reaches zero (c.f.\ axiom (44)).
 Axiom (60) defines a similar behavior for the start interval timer of a runnable, although this timer is also allowed to slip behind further time updates once it has reached zero (axiom (61)).
@@
 When the counter has become zero, the timer is forced to take the tick transition of axiom (63), which also resets the counter to its static period value $t_p$.
-The net effect of such a transition is that the runnable to which the ticking timer belongs is put into the {\bf pending} state (axiom (64)).
+The net effect of such a transition is that the runnable to which the ticking timer belongs is put into the \textbf{pending} state (axiom (64)).
 An example of a transition sequence involving a time-triggered runnable follows below.
@@
 
-irv without initValue vs.\ delem without initValue -- "undefined" vs.\ NO\_DATA at initial read.
+irv without initValue vs.\ delem without initValue -- ``undefined'' vs.\ NO\_DATA at initial read.
 

### 2016-08-15 paper/paper.tex
More paper text.

--- 2016-08-12 paper/paper.tex
@@
 
+Although our chosen mechanism for coordinating transitions among process terms goes under the name {\em broadcast}, there is no implication that every process must react to everything being said --- the intent is rather that a process term not directly or indirectly addressed in a transition should be allowed to ignore it.
+One could try to express this as some form of "catch-up" rule $p \red{l} p$, that would apply only if none of the axioms (5) to (64) match $p$ and $l$.
+However, such a rule would be too liberal, as it would also enable transitions that are deliberately omitted from the previous axioms, like ${\Texcl {i.x} \Ttrue} \red{\say{i.x}{\Tent}} {\Texcl {i.x} \Ttrue}$ (granting entry to an exclusive area already taken).
+
+Instead we need a more restrictive rule, that applies to processes not addressed by earlier axioms but avoids those previously excluded by restrictive patterns and other side-conditions.
+To make this idea precise, however, we need to split the rule according to the different forms of $p$ so that the addressing notion of each process type can be individually captured.
+Axioms (65) to (72) contain the resulting definitions.
+
+Runnable instances and timers can ignore everything being said by others, as they define no hearing transitions proper that need to be excluded here (axioms (65) and (66)).
+Exclusive areas, inter-runnable variables and operation terms have hearing transitions previously defined, but only for label addresses matching their own; they can thus safely ignore transitions labelled with any other address.
+Axioms (70) to (72) constitute the non-trivial cases, because here there are multiple earlier definitions to exclude, which also link label and process addresses via static model references.
+Still, it can be verified that each of the axioms (70) to (72) is guarded by an exact negation of the addressing conditions of any previously defined hearing transition for the same process type.
+
+\subsection{TODO}
+
 TODO: some links from rules and choices made here to the AUTOSAR spec
@@
 
-\renewcommand\Pcomma{\nonumber \\&&\text{and}\;\;}
+\renewcommand\Pcomma{\nonumber \;\text{and}\;}
 
 \begin{eqnarray}      \CallServer              \notag \end{eqnarray}
+
+\renewcommand\Pcomma{\nonumber \\&&\text{and}\;\;}
 

### 2016-08-16 paper/paper.tex
Changed addresses in DataReceivedEvent, etc.

@@
 We write $\Timplementation{i.r}{k}$ to state the assumption that the static AUTOSAR model under study assigns implementation function $k$ to runnable $i.r$.
-Similar notations will be used to express other references to the underlying static model; for example $\TdataReceivedEvent {i.r} {p.e}$ to state that the model allows runnable $i.r$ to be triggered by data reception events on port $p.e$ (in component $i$).
+Similar notations will be used to express other references to the underlying static model; for example $\TdataReceivedEvent {i.r} {i.p.e}$ to state that the model allows runnable $i.r$ to be triggered by data reception events on port $i.p.e$.
 Although the chosen notation is sometimes significantly shorter than the UML/XML-based syntax used in the AUTOSAR specification, the intended meaning should nevertheless be clear.
@@
 
-\subsection{Buffered sending/receiving}
+\subsection{Buffered sending/receiving} \label{sect:BufSndRcv}
 
@@
 The preconditions of axioms (27) and (28) ensure that $\Tvar{as}$ reflects the truth, and axioms (25) and (26) feed different error codes to the sender's continuation depending on whether $\Tvar{as}$ is empty or not.
-Precondition $\Tconnect {i.p.e} \Tvar{as}$ in axiom (26) should be read $\Tconnect {i.p.e} a$ for all $a \in \Tvar{as}$, and its purpose is to outrule irrelevant addresses in $\Tvar{as}$ causing bogus \textbf{limit} results.
+By requiring connectivity for all $a$ in $\Tvar{as}$ in axiom (26), the possibility of letting irrelevant addresses in $\Tvar{as}$ cause bogus \textbf{limit} results is outruled.
 
-The following example illustrates a scenario where a sent value is only stored in a subset of the connected buffers. Assuming $\Tconnect {\Tep e p i} {a_1}$ and $\Tconnect {\Tep e p i} {a_2}$:
+The following example illustrates a scenario where a sent value is only stored in a subset of the connected buffers. Assuming $\Tconnect {\Tep e p i} a$ and $\Tconnect {\Tep e p i} b$:
 $$
@@
   \Trinst {i.r} c \Xs {\TrteSend {p.e} v k} \,\Opar  \\
-  {\Tqelem {a_1} 2 \Teps} \Opar {\Tqelem {a_2} 2 {\Tcons {v_1} {v_2}}} \vspace{0.6ex} \\
+  {\Tqelem a 2 \Teps} \Opar {\Tqelem b 2 {\Tcons {v_1} {v_2}}} \vspace{0.6ex} \\
   \red{\say{i.p.e}{\Tsnd v {a_2}}} \\
   \Trinst {i.r} c \Xs {\Tap k \Tlimit} \,\Opar \\
-  {\Tqelem {a_1} 2 v} \Opar {\Tqelem {a_2} 2 {\Tcons {v_1} {v_2}}}
+  {\Tqelem a 2 v} \Opar {\Tqelem b 2 {\Tcons {v_1} {v_2}}}
 \end{array}
@@
 
-Here is an example of how a client runnable instance, a client port operation, and a server runnable interact during a call transition (assuming $b = {i_2.r_2}$, $\ToperationInvokedEvent b {p_2.o}$, $\Tconnect {i_2.p_2.o} {i.p.o}$, and $\TserverCallPointTimeout{i.p.o} t $).
+Here is an example of how a client runnable instance, a client port operation, and a server runnable interact during a call transition (assuming $\ToperationInvokedEvent b a$, $\Tconnect a {i.p.o}$, and $\TserverCallPointTimeout{i.p.o} t $).
 $$
@@
 A timer process, finally, decrements its counter according to axiom (62).
-When the counter has become zero, the timer is forced to take the tick transition of axiom (63), which also resets the counter to its static period value $t_p$.
+When the counter has become zero, the timer is forced to take the $\payload{tick}$ transition of axiom (63), which also resets the counter to its static period value $t_p$.
 The net effect of such a transition is that the runnable to which the ticking timer belongs is put into the \textbf{pending} state (axiom (64)).
@@
 Still, it can be verified that each of the axioms (70) to (72) is guarded by an exact negation of the addressing conditions of any previously defined hearing transition for the same process type.
+
+As an example, consider the example of Section~\ref{sect:BufSndRcv} under the different assumption that only $\Tconnect{{i.p.e} a}$ holds, with no connection from ${i.p.e}$ to $b$.
+$$
+\begin{array}{c}
+  \Trinst {i.r} c \Xs {\TrteSend {p.e} v k} \,\Opar  \\
+  {\Tqelem a 2 \Teps} \Opar {\Tqelem b 2 {\Tcons {v_1} {v_2}}} \vspace{0.6ex} \\
+  \red{\say{i.p.e}{\Tsnd v \Teps}} \\
+  \Trinst {i.r} c \Xs {\Tap k \Tok} \,\Opar \\
+  {\Tqelem a 2 v} \Opar {\Tqelem b 2 {\Tcons {v_1} {v_2}}}
+\end{array}
+$$
 

### 2016-08-16 paper/paper.tex
Added two examples.

@@
   {\Tqelem a 2 \Teps} \Opar {\Tqelem b 2 {\Tcons {v_1} {v_2}}} \vspace{0.6ex} \\
-  \red{\say{i.p.e}{\Tsnd v {a_2}}} \\
+  \red{\say{i.p.e}{\Tsnd v b}} \\
   \Trinst {i.r} c \Xs {\Tap k \Tlimit} \,\Opar \\
@@
 
-As an example, consider the example of Section~\ref{sect:BufSndRcv} under the different assumption that only $\Tconnect{{i.p.e} a}$ holds, with no connection from ${i.p.e}$ to $b$.
+As an concrete illustration, consider the example of Section~\ref{sect:BufSndRcv}.
+The axioms that enable each individual process transition in that example are (26), (27) and (28), respectively (and rules (1) and (3) allow them to be combined).
+In particular, axiom (28) is applicable because of the assumption $\Tconnect {i.p.e} b$, which simultaneously blocks applicability of axiom (70).
+However, under the different assumption that there is no connection from ${i.p.e}$ to $b$, axioms (28) and (70) must trade places.
+This also forces the removal of $b$ from the second argument of the {\sc snd} payload (i.e., the list of connected buffers not able to store the value sent).
+As a consequence, axiom (26) is no longer applicable and must be replaced by axiom (25).
+The resulting parallel transition looks as follows, illustrating the correct generation of an {\bf ok} result despite the presence of a full (but unconnected) buffer.
 $$
@@
 
-\subsection{TODO}
-
 TODO: some links from rules and choices made here to the AUTOSAR spec
 
-\section{Examples}
-\label{sec:Examples}
+\section{Parallelism}
+\label{sec:Parallelism}
 
-\begin{itemize}
-\item More details on the simple examples + a bigger example.
-\item Some examples chosen to illustrate the value of a formal calculus
-\end{itemize}
+$$
+\begin{array}{c}
+  {\Trunnable a 0 \Tidle 0} \Opar 
+  {\Tqelem b 2 \Teps} \vspace{0.6ex} \\
+
+  \red{\hear{b}{\Tsnd {v_1} \Teps}} \\
+
+  {\Trunnable a 0 \Tpending 0} \Opar 
+  {\Tqelem b 2 {v_1}} \vspace{0.6ex} \\
+
+  \red{\say{a}{\Tnew}} \\
+
+  {\Trunnable a 0 \Tidle 1} \Opar 
+  {\Tqelem b 2 {v_1}} \,\Opar  \\
+  {\Trinst a \Tvoid \Teps {\Code}} \vspace{0.6ex} \\
+
+  \red{\hear{b}{\Tsnd {v_2} \Teps}} \\
+
+  {\Trunnable a 0 \Tpending 1} \Opar 
+  {\Tqelem b 2 {\Tcons{v_1}{v_2}}} \,\Opar  \\
+  {\Trinst a \Tvoid \Teps {\Code}} \vspace{0.6ex} \\
+
+  \red{\say{a}{\Tnew}} \\
+
+  {\Trunnable a 0 \Tidle 2} \Opar 
+  {\Tqelem b 2 {\Tcons{v_1}{v_2}}} \,\Opar  \\
+  {\Trinst a \Tvoid \Teps {\Code}} \Opar 
+  {\Trinst a \Tvoid \Teps {\Code}}
+\end{array}
+$$
+
+$$
+\begin{array}{c}
+  {\Trunnable a 0 \Tidle 0} \Opar 
+  {\Tqelem b 2 \Teps} \vspace{0.6ex} \\
+
+  \red{\hear{b}{\Tsnd {v_1} \Teps}} \\
+
+  {\Trunnable a 0 \Tpending 0} \Opar 
+  {\Tqelem b 2 {v_1}} \vspace{0.6ex} \\
+
+  \red{\hear{b}{\Tsnd {v_2} \Teps}} \\
+
+  {\Trunnable a 0 \Tpending 0} \Opar 
+  {\Tqelem b 2 {\Tcons{v_1}{v_2}}} \vspace{0.6ex} \\
+
+  \red{\say{a}{\Tnew}} \\
+
+  {\Trunnable a 0 \Tidle 2} \Opar 
+  {\Tqelem b 2 {\Tcons{v_1}{v_2}}} \,\Opar  \\
+  {\Trinst a \Tvoid \Teps {\Code}}
+\end{array}
+$$
 

### 2016-08-18 paper/paper.tex
Paper: spell-checking

@@
 It is also not very abstract, in that it makes frequent references to assumed implementation techniques for the purpose of defining its semantics.
-In practice, the AUTOSAR standard becomes blurred with the specific behavior of one's chosen platform and development tools.
+In practice, the AUTOSAR standard becomes blurred with the specific behaviour of one's chosen platform and development tools.
 And because the standard is open to interpretation, the interoperability of software components across tools and platforms is often seriously hampered.
-What is more, a single software component cannot easily be studied and understood in isolation, since its interactive behavior is only indirectly defined in terms of the concrete C-code and OS tasks that realize it and its interaction environment.
+What is more, a single software component cannot easily be studied and understood in isolation, since its interactive behaviour is only indirectly defined in terms of the concrete C-code and OS tasks that realize it and its interaction environment.
 
@@
 The formalization covers most of the Software Component Template and its accompanying Run-Time Environment (RTE) (Section~\ref{sec:Calc}), and is able to directly express every legal way a system of software components can evolve at run-time on an arbitrarily fast platform (Section~\ref{sec:Sem}).
-It can thus serve as a basis for both a concrete AUTOSAR implementation on a specific platform (whose supported behaviors must be a subset of those defined by the formal semantics), and a platform-independent AUTOSAR simulator (where behaviors from the legal set can be picked at random).
+It can thus serve as a basis for both a concrete AUTOSAR implementation on a specific platform (whose supported behaviours must be a subset of those defined by the formal semantics), and a platform-independent AUTOSAR simulator (where behaviours from the legal set can be picked at random).
 Examples of semantic derivations and their applications are found in Section~\ref{sec:Examples}.
@@
 A formal notation is nevertheless a good starting point for any discussion on the resolution of such issues.
-The semantics is furthermore written to be unambiguous, except where concurrent execution should allow for more than one observed behavior.
+The semantics is furthermore written to be unambiguous, except where concurrent execution should allow for more than one observed behaviour.
 At some points, the AUTOSAR documents have been found unclear; here a specific choice has been made but alternative interpretations will be discussed separately (Section~\ref{sec:DiscAmb}).
 At other points, the standard documents are clear but the resulting semantics is still dubious.
-These cases will also be emphasized and discussed, together with suggested ways forward (Section~\ref{sec:DiscImp}).
+These cases will also be emphasised and discussed, together with suggested ways forward (Section~\ref{sec:DiscImp}).
 
@@
 Some components merely act as containers of other components structures, these are called \textbf{composition} SWCs.
-In contrast, \textbf{atomic} SWCs define their own behavior in terms of \textbf{runnables} and \textbf{inter-runnable variables}, primarily.
-A runnable denotes a sequential and (normally) terminating piece of code, although its actual behavioral definition is typically relegated to external C or Matlab files rather than being part of the AUTOSAR model proper.
+In contrast, \textbf{atomic} SWCs define their own behaviour in terms of \textbf{runnables} and \textbf{inter-runnable variables}, primarily.
+A runnable denotes a sequential and (normally) terminating piece of code, although its actual behavioural definition is typically relegated to external C or Matlab files rather than being part of the AUTOSAR model proper.
 Instead, an AUTOSAR runnable  only provides constraints on what ports and inter-runnable variables the actual implementation can access; as well as \textbf{event} declarations that determine under what conditions it will get executed.
@@
 Each runnable indicates whether or not it may be instantiated concurrently with itself.
-Runnable code can also use \textbf{exclusive areas} (a form of mutex semaphores) to further control concurrent behavior.
+Runnable code can also use \textbf{exclusive areas} (a form of mutex semaphores) to further control concurrent behaviour.
 
@@
 We write $\Vs$ for a sequence of values \V{V}, $\Cs$ for a sequence of clients \V{C}, etc.
-Sequences may be empty, which is written $\epsilon$.
+Sequences may be empty, which is written~$\epsilon$.
 By $\Tlength{\Vs}$ we mean the length of sequence $\Vs$.
@@
 To this end, we have chosen to ignore all internal computations and just capture the observable effects of runnable execution as a pure sequence of RTE (VFB) calls.
-And because the result from an RTE call may in general affect subsequent runnable behavior, we use a continuation-passing style where every $\Code$ term except the final function return literally contain the sequence that follows it, as a continuation parameter $\Cont$.
+And because the result from an RTE call may in general affect subsequent runnable behaviour, we use a continuation-passing style where every $\Code$ term except the final function return literally contain the sequence that follows it, as a continuation parameter $\Cont$.
 %
@@
 The effect of these agreement rules closely resembles the idea of a \emph{broadcast}: what one process says may be heard by every other process in a large parallel composition.
-This behavior serves us well, because in general, what one AUTOSAR runnable does may have impact on many (if not all) parts of the executing system.
+This behaviour serves us well, because in general, what one AUTOSAR runnable does may have impact on many (if not all) parts of the executing system.
 It is however important not to confuse this broadcasting notion with any particular form of AUTOSAR communication.
@@
 
-We expect the implementation of a runnable to be available as a continuation function $\Cont$, even though the real AUTOSAR model will be refering to concrete functions in external C/Matlab files.
+We expect the implementation of a runnable to be available as a continuation function $\Cont$, even though the real AUTOSAR model will be referring to concrete functions in external C/Matlab files.
 We will further abstract away from actual numbers of input and output parameters to these functions by assuming that records (structs) are used as the single continuation argument and result whenever the arity so requires.
@@
 $$
-This means that any runnable instance in $p$ wishing to enter $x$ is effectively blocked from making progress until some other process in $p$ choses to exit $x$.
+This means that any runnable instance in $p$ wishing to enter $x$ is effectively blocked from making progress until some other process in $p$ chooses to exit $x$.
 In the same manner, a process wishing to exit an exclusive area not taken is also effectively blocked.
@@
 
-Axioms (13) and (15) define unbuffered interprocess communication from a runnable instance point of view.
+Axioms (13) and (15) define unbuffered inter-process communication from a runnable instance point of view.
 These transitions are just syntactic variants of shared variable reading and writing (axioms (5) and (7)).
@@
 Axioms (14) and (16) capture reads and writes to the data value and also set the update flag accordingly.
-The side-confition in axiom (16) makes the rule apply only if there is a static connection from the data element being written and the address of the matched term (recall that data element storage is associated with the receiving side of a connection).
+The side-condition in axiom (16) makes the rule apply only if there is a static connection from the data element being written and the address of the matched term (recall that data element storage is associated with the receiving side of a connection).
 Because AUTOSAR allows sender-receiver communication patterns such that $\Tconnect {i.p.e} a$ and $\Tconnect {i.p.e} b$ are simultaneously true, parallel data element updates like the following are possible:
@@
 
-Axiom (17) provides another means of intepreting a data element write.
+Axiom (17) provides another means of interpreting a data element write.
 It expresses that a runnable set up to trigger on data reception on required data element $a$ should be marked as \textbf{pending} whenever a write occurs on an element connected to $a$.
@@
 
-Buffered interprocess communication (axioms (22) - (30)) is just a minor variant of the unbuffered mechanism, where the single-value store of a \textbf{delem} term is replaced by a \textbf{qelem} process holding a sequence of values.
+Buffered inter-process communication (axioms (22) - (30)) is just a minor variant of the unbuffered mechanism, where the single-value store of a \textbf{delem} term is replaced by a \textbf{qelem} process holding a sequence of values.
 The buffer is consumed from the left (axiom (23)) and extended to the right (axiom (27)).
@@
 The preconditions of axioms (27) and (28) ensure that $\Tvar{as}$ reflects the truth, and axioms (25) and (26) feed different error codes to the sender's continuation depending on whether $\Tvar{as}$ is empty or not.
-By requiring connectivity for all $a$ in $\Tvar{as}$ in axiom (26), the possibility of letting irrelevant addresses in $\Tvar{as}$ cause bogus \textbf{limit} results is outruled.
+By requiring connectivity for all $a$ in $\Tvar{as}$ in axiom (26), the possibility of letting irrelevant addresses in $\Tvar{as}$ cause bogus \textbf{limit} results is ruled out.
 
@@
 
-The behavior of client-server communication in AUTOSAR depends on whether a client lists the required server port operation as a synchronous or an asynchronous callpoint (which are mutually exclusive model attributes).
-In the former case, the call command is treated as if it were immediately followed by a command for retrieving the server result (axiom (31)), whereas in the latter case, the call succeeds immediately and the result has to be retireved explicitly by the client's continuation code (axiom (32)).
+The behaviour of client-server communication in AUTOSAR depends on whether a client lists the required server port operation as a synchronous or an asynchronous callpoint (which are mutually exclusive model attributes).
+In the former case, the call command is treated as if it were immediately followed by a command for retrieving the server result (axiom (31)), whereas in the latter case, the call succeeds immediately and the result has to be retrieved explicitly by the client's continuation code (axiom (32)).
 In both cases, however, an attempt to call the server before it is done processing a previous asynchronous call leads to immediate abortion with the error code \textbf{limit} (axiom (35)).
 
-The AUTOSAR specification restricts client-server connections to the many-to-one pattern only, assigning a dedicated runnable resposible for implementing the offered service.
+The AUTOSAR specification restricts client-server connections to the many-to-one pattern only, assigning a dedicated runnable responsible for implementing the offered service.
 It furthermore forbids runnables to be triggered by any events other than operation invocations of compatible signatures, making it natural to associate the state needed to buffer up server invocations with the server runnables themselves.
@@
 $$
-As a contrast, this is the behavior observed when a call timeout occurs:
+As a contrast, this is the behaviour observed when a call timeout occurs:
 $$
@@
 \item They create the new instance in parallel with its parent (runnable instances is actually the only example of dynamic process creation in this calculus).
-\item They initialize the new instance with code according to the model-defined implementation and an empty list of owned exclusive areas.
+\item They initialise the new instance with code according to the model-defined implementation and an empty list of owned exclusive areas.
 \end{itemize}
@@
 Yet other processes can only take time transitions, which simply means that they represent an idle system that currently has no work to do.
-Mostly, though, work and time transitions are simltaneously available, indicating that the semantics considers the slightly slower and sligthly faster behavior alternatives to be equally correct.
+Mostly, though, work and time transitions are simultaneously available, indicating that the semantics considers the slightly slower and slightly faster behaviour alternatives to be equally correct.
 
@@
 The counter is not allowed to go negative, though, which forces the \textbf{oper} term to handle the timeout once the counter reaches zero (c.f.\ axiom (44)).
-Axiom (60) defines a similar behavior for the start interval timer of a runnable, although this timer is also allowed to slip behind further time updates once it has reached zero (axiom (61)).
+Axiom (60) defines a similar behaviour for the start interval timer of a runnable, although this timer is also allowed to slip behind further time updates once it has reached zero (axiom (61)).
 The reason for this is that the interval timers do not enforce any particular action when they time out, they just enable new transition options that do not have to be taken immediately.

### 2016-08-20 paper/paper.tex
Paper: started reading + fixing minor things.

@@
 \usepackage{mathpartir}
-\usepackage{ amssymb }
+\usepackage{amssymb}
 \usepackage[makeroom]{cancel}
@@
 \begin{abstract}
+TODO
 
@@
 
-TODO: Simple examples early (informally)
+TODO: Some simple (informal) example early to give the reader an early view of flavour of the results.
 
@@
 
-A runnable under execution is called a \textbf{runnable instance}, and these may execute concurrently even if they belong to the same SWC.
+A runnable under execution is called a \textbf{runnable instance}, and several instances may execute concurrently even if they belong to the same SWC.
 Each runnable indicates whether or not it may be instantiated concurrently with itself.
@@
 
-TODO: add fig:model
+TODO: add fig:model (from the slides)
 
@@
 \item $\Tdelem {\Tep e p i} u v$ \newline
-Non-queued sender-receiver data element $e$ in port $p$ of component $i$, currently holding value \V{V} with boolean update status flag $u$.
+Non-queued sender-receiver data element $e$ in port $p$ of component $i$, currently holding value $v$ with boolean update status flag $u$.
 \item $\Tqelem {\Tep e p i} n \Vs$ \newline
@@
 To this end, we have chosen to ignore all internal computations and just capture the observable effects of runnable execution as a pure sequence of RTE (VFB) calls.
-And because the result from an RTE call may in general affect subsequent runnable behaviour, we use a continuation-passing style where every $\Code$ term except the final function return literally contain the sequence that follows it, as a continuation parameter $\Cont$.
+And because the result from an RTE call may in general affect subsequent runnable behaviour, we use a continuation-passing style where every $\Code$ term except the final function return literally contains the sequence that follows it, as a continuation parameter $\Cont$.
 %
@@
 The way transition labels distribute over the parallel composition operator is merely a technical aspect of our process calculus, and will be used to express several different AUTOSAR communication semantics, among other things.
+
+%% TODO: PaJa: continue reading from here.
 

### 2016-08-23 NewARSim/NewARSim.hs
Fixed a bug where the trace shrinking scheduler could not terminate simulation.

@@ -1287,15 +1287,18 @@ checkEarlyPending :: Monad m
                   -> Scheduler m
 checkEarlyPending conn pm sched alts = do
   res <- sched alts
-  let (label, us) = case res of 
-          -- Invariant: A transition was made, result is never @None@ 
-          Warn w (trans, ps) -> (transLabel trans, ps)
-          Some   (trans, ps) -> (transLabel trans, ps)
-  case label of 
-    TICK a    -> warnIf a us res
-    WR a _    -> warnIf a us res 
-    SND a _ _ -> warnIf a us res 
-    _         -> return res
+  case res of 
+    None -> return None
+    _ -> do
+      let (label, us) = case res of 
+              -- Invariant: A transition was made, result is never @None@ 
+              Warn w (trans, ps) -> (transLabel trans, ps)
+              Some   (trans, ps) -> (transLabel trans, ps)
+      case label of 
+        TICK a    -> warnIf a us res
+        WR a _    -> warnIf a us res 
+        SND a _ _ -> warnIf a us res 
+        _         -> return res
   where
     warnIf a us res
       | null (pendTwice a us) = return res

### 2016-08-23 paper/paper.tex
Dump of current paper.

--- 2016-08-16 paper/paper.tex
@@
 
+\bibliographystyle{splncs03} % Latest bibtex style for llncs
+
 \setlength{\parindent}{0pt}
@@
 
-The AUTOSAR standard is an open software component architecture for the automotive industry.
+The AUTOSAR standard is an open software component architecture for the automotive industry \cite{AR:Homepage}.
 Its main purpose is to enable interoperability of software modules among different vendors and on heterogeneous platforms, via an extensive set of standardized interfaces and libraries, and a common software development methodology.
@@
 
-An AUTOSAR model is primarily a structure of interconnected \textbf{software components} (SWCs).
+An AUTOSAR model is primarily a structure of interconnected \textbf{software components} (SWCs) \cite{AR:SWC}.
 Links between SWCs are called \textbf{connectors}, which attach to \textbf{ports} exposed by the SWCs.
@@
 To interact with ports, inter-runnable variables and exclusive areas, runnables use what is abstractly known as the \textbf{virtual function bus} (VFB).
-Concretely, the VFB takes the shape of a C programming interface to the \textbf{run-time environment} (RTE), custom-generated for each runnable on the basis of its access constraints.
+Concretely, the VFB takes the shape of a C programming interface to the \textbf{run-time environment} (RTE), custom-generated for each runnable on the basis of its access constraints \cite{AR:RTE}.
 Behind the RTE, a collection of OS kernels, communication stacks and other basic software modules implement the VFB services for the platform at hand.
@@
 
-\subsection{Calling a server}
+\subsection{Calling a server} \label{sect:Call}
 
@@
 
-\subsection{Passing back a server result}
+\subsection{Passing back a server result} \label{sect:Res}
 
@@
 \item The non-server runnable toggles back to the $\Tidle$ activation state, while the server just chops of the head of its client queue.
-\item Only the server runnable instance is given a non-void parameter, which is taken from the same client queue head.
+\item Only the server runnable instance is given a non-void code parameter, which is taken from the same client queue head.
 \end{itemize}
@@
 
+The transition relation defined in Appendix~\ref{sec:Rules} is non-deterministic; i.e., it may very well be that for some process $p$,
 $$
 \begin{array}{c}
-  {\Trunnable a 0 \Tidle 0} \Opar 
-  {\Tqelem b 2 \Teps} \vspace{0.6ex} \\
-
-  \red{\hear{b}{\Tsnd {v_1} \Teps}} \\
-
-  {\Trunnable a 0 \Tpending 0} \Opar 
-  {\Tqelem b 2 {v_1}} \vspace{0.6ex} \\
-
-  \red{\say{a}{\Tnew}} \\
-
-  {\Trunnable a 0 \Tidle 1} \Opar 
-  {\Tqelem b 2 {v_1}} \,\Opar  \\
-  {\Trinst a \Tvoid \Teps {\Code}} \vspace{0.6ex} \\
-
-  \red{\hear{b}{\Tsnd {v_2} \Teps}} \\
-
-  {\Trunnable a 0 \Tpending 1} \Opar 
-  {\Tqelem b 2 {\Tcons{v_1}{v_2}}} \,\Opar  \\
-  {\Trinst a \Tvoid \Teps {\Code}} \vspace{0.6ex} \\
-
-  \red{\say{a}{\Tnew}} \\
-
-  {\Trunnable a 0 \Tidle 2} \Opar 
-  {\Tqelem b 2 {\Tcons{v_1}{v_2}}} \,\Opar  \\
-  {\Trinst a \Tvoid \Teps {\Code}} \Opar 
-  {\Trinst a \Tvoid \Teps {\Code}}
+	p \red{l_1} p_1 \hspace{10mm}\text{and}\hspace{10mm} p \red{l_2} p_2
 \end{array}
 $$
-
+are both possible. If it also holds that
 $$
 \begin{array}{c}
-  {\Trunnable a 0 \Tidle 0} \Opar 
-  {\Tqelem b 2 \Teps} \vspace{0.6ex} \\
-
-  \red{\hear{b}{\Tsnd {v_1} \Teps}} \\
-
-  {\Trunnable a 0 \Tpending 0} \Opar 
-  {\Tqelem b 2 {v_1}} \vspace{0.6ex} \\
-
-  \red{\hear{b}{\Tsnd {v_2} \Teps}} \\
-
-  {\Trunnable a 0 \Tpending 0} \Opar 
-  {\Tqelem b 2 {\Tcons{v_1}{v_2}}} \vspace{0.6ex} \\
-
-  \red{\say{a}{\Tnew}} \\
-
-  {\Trunnable a 0 \Tidle 2} \Opar 
-  {\Tqelem b 2 {\Tcons{v_1}{v_2}}} \,\Opar  \\
-  {\Trinst a \Tvoid \Teps {\Code}}
+	p_1 \red{l_2} q \hspace{10mm}\text{and}\hspace{10mm} p_2 \red{l_1} q
 \end{array}
 $$
+are derivable, we note that process $p$ may evolve into state $q$ either via an $l_1$ step followed by $l_2$, or by first taking the $l_2$ step and then $l_1$.
+Such freedom of choice is commonly taken as the definition of {\em parallelism} in the process calculus literature --- if the sequential ordering of steps $l_1$ and $l_2$ has no significance, they might just as well be considered to occur in parallel.
 
-\section{Discussion / results}
-\label{sec:Disc}
+However, it may also be that two possible step orderings do not lead to identical results. That happens if, instead of the two transitions leading to $q$ above, we have
+$$
+\begin{array}{c}
+	p_1 \red{l_2} q_1 \hspace{10mm}\text{and}\hspace{10mm} p_2 \red{l_1} q_2
+\end{array}
+$$
+where $q_1$ and $q_2$ are "similar" but not identical states.
+This non-determinism also has a concrete interpretation, either reflecting the non-deterministic nature of a system's environment (messages $l_1$ and $l_2$ just happen to arrive in some particular order) or its internals (processor speed or a scheduling choice just happens to pick $l_1$ before $l_2$).
+It is one of the primary merits of the process calculus approach to allow the scope of such inherent non-determinism to be precisely defined, because although a given system state may render many different behaviors possible, many more are not.
+A formal semantics makes it possible to clearly draw the line between these groups of behaviors.
 
-\subsection{Ambiguities and alternative interpretations}
-\label{sec:DiscAmb}
+The AUTOSAR specification contains a very subtle sensitivity to this kind of non-deterministism that is worth examplifying.
+Consider a queued data element $a$ together with a runnable $b$ set up to trigger on data reception events for $a$.
+A scenario where two values $v_1$ and $v_2$ are sent on a port connected to $a$ may lead to the following trace:
+$$
+\begin{array}{c}
+  {\Tqelem a 2 \Teps} \Opar 
+  {\Trunnable b 0 \Tidle 0} \vspace{0.6ex} \\
 
-\subsection{Clarification proposals and improvements}
-\label{sec:DiscImp}
+  \red{\hear{i.p.e}{\Tsnd {v_1} \Teps}} \\
 
-irv without initValue vs.\ delem without initValue -- ``undefined'' vs.\ NO\_DATA at initial read.
+  {\Tqelem a 2 {v_1}} \Opar 
+  {\Trunnable b 0 \Tpending 0} \vspace{0.6ex} \\
 
-Multiple runnables triggered by same OperationInvokedEvent?
+  \red{\say{b}{\Tnew}} \\
 
-Multiple timingEvent for same runnable, possibly with same period?
+  {\Tqelem a 2 {v_1}} \Opar 
+  {\Trunnable b 0 \Tidle 1} \,\Opar  \\
+  {\Trinst b \Tvoid \Teps {\Code}} \vspace{0.6ex} \\
 
-Timer initial offsets
+  \red{\hear{i.p.e}{\Tsnd {v_2} \Teps}} \\
+
+  {\Tqelem a 2 {\Tcons{v_1}{v_2}}} \Opar 
+  {\Trunnable b 0 \Tpending 1} \,\Opar  \\
+  {\Trinst b \Tvoid \Teps {\Code}} \vspace{0.6ex} \\
+
+  \red{\say{b}{\Tnew}} \\
+
+  {\Tqelem a 2 {\Tcons{v_1}{v_2}}} \Opar 
+  {\Trunnable b 0 \Tidle 2} \,\Opar  \\
+  {\Trinst b \Tvoid \Teps {\Code}} \Opar 
+  {\Trinst b \Tvoid \Teps {\Code}}
+\end{array}
+$$
+That is, a resulting state where $a$ holds the two received values together with two concurrent instances of $b$ ready to start reacting upon the two reception events.
+
+However, an alternative trace is equally valid for the same scenario:
+$$
+\begin{array}{c}
+  {\Tqelem a 2 \Teps} \Opar 
+  {\Trunnable b 0 \Tidle 0} \vspace{0.6ex} \\
+
+  \red{\hear{i.p.e}{\Tsnd {v_1} \Teps}} \\
+
+  {\Tqelem a 2 {v_1}} \Opar 
+  {\Trunnable b 0 \Tpending 0} \vspace{0.6ex} \\
+
+  \red{\hear{i.p.e}{\Tsnd {v_2} \Teps}} \\
+
+  {\Tqelem a 2 {\Tcons{v_1}{v_2}}} \Opar 
+  {\Trunnable b 0 \Tpending 0} \vspace{0.6ex} \\
+
+  \red{\say{b}{\Tnew}} \\
+
+  {\Tqelem a 2 {\Tcons{v_1}{v_2}}} \Opar 
+  {\Trunnable b 0 \Tidle 1} \,\Opar  \\
+  {\Trinst b \Tvoid \Teps {\Code}}
+\end{array}
+$$
+Here there is only one $b$ instance created and none is pending, even though $a$ correctly contains the two values received.
+The reason is that by swapping the order of the first instantiation and the second data reception steps, the system is put in a state where its inability to count pending instantiations becomes fatal.
+
+Unfortunately there is very little a programmer can do to influence which of these two traces is taken.
+Giving $b$ a positive {\em minimumStartInterval} just guarantees that any subsequent events are lost during that interval.
+Long delays between reception events will reduce the likelihood of losses, but the choice of when to take an instantiation step is still fully internal to the operating system scheduler.
+Moreover, when messages are produced by multiple parallel senders, maintaining control over their time distribution may be highly impractical.
+Non-concurrent runnables can work around the problem by iterating over all received data at each invocation, but that solution is not very attractive if concurrent processing of each received item is desired.
+
+One could argue that replacing the {\bf pending}/{\bf idle} flag of runnables with a small counter variable would render the event triggering mechanism in AUTOSAR much more robust.
+The standard explicitly excludes event counting for all events but operation invocation, though, so the possibility that events may be lost is undeniably intentional \cite[ch.~4.2.3]{AR:RTE}.
+However, should this design decision ever be revisited, the two trace alternatives above could serve as a succinct problem characterization.
+
+\section{Ambiguities and clarification proposals}
+\label{sec:Amb}
+
+The formal semantics presented in this report is aimed to be a faithful capture of the AUTOSAR specification, especially its Software Components template and RTE specification \cite{AR:SWC, AR:RTE}.
+Still, there are several areas where we have found the specification open to interpretation, or at least in need of some further clarification.
+We detail a few of these areas here, as an illustration to the kind of design decisions that may surface as the result of a formalization undertaking.
+
+Unqueued data elements that are read before first written should return the error code {\bf neverReceived} \cite[ch.~5.6.10]{AR:RTE}.
+Since no exceptions are mentioned in the specification, our intepretation is that this holds even when a data element possesses an {\em initValue} attribute  (third parameter of the initial {\bf delem} state defined in Section~\ref{sec:Sem}).
+However, as this renders any such assignments entirely redundant, an alternative interpretation creates a term
+$$
+  \Tdelem {i.p.e} \Tfalse v
+$$
+for each unqueued data element $i.p.e$, where $v$ is the {\em initValue} of $e$ if it exists, and $\TneverReceived$ otherwise.
+
+Inter-runnable variables can also be read before written, although here there is no special error code defined in the specification for the uninitialized case \cite[ch.~5.6.26]{AR:RTE}.
+Section~\ref{sec:Sem} sets each {\bf irv} contents to {\bf void} to indicate a truly non-existing value, although an entirely reasonable alternative would be
+$$
+  \Tirv {i.s} v
+$$
+where $v$ is the {\em initValue} of $s$ if it exists, and the error code {\bf nodata} otherwise.
+
+The specification restricts client-server connections such that a client may not be connected to multiple servers, for the purpose of avoiding an operation call to be handled by more than one server runnable \cite[ch.~4.2.3]{AR:RTE}.
+However, this restriction does not explicitly forbid multiple {\em OperationInvokedEvent}s refering to the same client-server operation to trigger separate runnables.
+Our formulation of client-server communication is based on the assumption that this is just an oversight, as allowing multiple events would have the same problem with ambiguous response values as a would one-to-many client-server connection pattern.
+%It is not easy to see what could constitute an alternative interpretation without introducing some programmer option for selecting among server results.
+
+Although the AUTOSAR specification only motivates timed runnable activation with the need to support plain periodic execution, there is really nothing that precludes such runnables from being triggered by additional events as well --- including other {\em TimingEvent}s, that might even carry identical periods.
+Our formulation naturally supports this generality, but it should be observed that the inability of AUTOSAR runnables to count pending activations (see Section~\ref{sec:Parallelism}) also renders the resolution of any coinciding timer events highly non-deterministic.
+Restricting
+
+It may further be noted that an alternative interpretation where conflicting timers are outruled just means a static model restriction, with no semantic implications on the timer configurations that are actually allowed.
+
+Related to multiple timers is the
 
@@
 Return to the limitation - what is the next step (which can be lifted more easily)
+
+\bibliography{refs}
 

### 2016-08-23 NewABS/Simulink.hs
Initial work on properties for NewABS

@@ -5,5 +5,5 @@ import NewABS2
 import System.Random
 
 main :: IO ()
-main = simulateUsingExternal absSystem
+main = simulateUsingExternal absSystem >> return ()
 

### 2016-08-23 paper/paper.tex

@@
 Some components merely act as containers of other components structures, these are called \textbf{composition} SWCs.
-In contrast, \textbf{atomic} SWCs define their own behavior in terms of \textbf{runnables} and \textbf{inter-runnable variables}, primarily.
-A runnable denotes a sequential and (normally) terminating piece of code, although its actual behavioral definition is typically relegated to external C or Matlab files rather than being part of the AUTOSAR model proper.
+In contrast, \textbf{atomic} SWCs define their own behaviour in terms of \textbf{runnables} and \textbf{inter-runnable variables}, primarily.
+A runnable denotes a sequential and (normally) terminating piece of code, although its actual behavioural definition is typically relegated to external C or Matlab files rather than being part of the AUTOSAR model proper.
 Instead, an AUTOSAR runnable only provides constraints on what ports and inter-runnable variables the actual implementation can access; as well as \textbf{event} declarations that determine under what conditions it will get executed.
 
-A runnable under execution is called a \textbf{runnable instance}, and these may execute concurrently even if they belong to the same SWC.
+A runnable under execution is called a \textbf{runnable instance}, and several instances may execute concurrently even if they belong to the same SWC.
 Each runnable indicates whether or not it may be instantiated concurrently with itself.
-Runnable code can also use \textbf{exclusive areas} (a form of mutex semaphores) to further control concurrent behavior.
+Runnable code can also use \textbf{exclusive areas} (a form of mutex semaphores) to further control concurrent behaviour.
 
@@
 To this end, we have chosen to ignore all internal computations and just capture the observable effects of runnable execution as a pure sequence of RTE (VFB) calls.
-And because the result from an RTE call may in general affect subsequent runnable behavior, we use a continuation-passing style where every $\Code$ term except the final function return literally contain the sequence that follows it, as a continuation parameter $\Cont$.
+And because the result from an RTE call may in general affect subsequent runnable behaviour, we use a continuation-passing style where every $\Code$ term except the final function return literally contains the sequence that follows it, as a continuation parameter $\Cont$.
 %
@@
 The effect of these agreement rules closely resembles the idea of a \emph{broadcast}: what one process says may be heard by every other process in a large parallel composition.
-This behavior serves us well, because in general, what one AUTOSAR runnable does may have impact on many (if not all) parts of the executing system.
+This behaviour serves us well, because in general, what one AUTOSAR runnable does may have impact on many (if not all) parts of the executing system.
 It is however important not to confuse this broadcasting notion with any particular form of AUTOSAR communication.
 The way transition labels distribute over the parallel composition operator is merely a technical aspect of our process calculus, and will be used to express several different AUTOSAR communication semantics, among other things.
+
+%% TODO: PaJa: continue reading from here.
 

### 2016-08-24 ACC/ACC.hs
Made a single-task assignment for the ACC example.

@@ -1,9 +1,6 @@
 -- | Adaptive Cruise Control. 
---
--- TODO: Fix panic braking.
 
 {-# LANGUAGE RecordWildCards   #-}
-{-# LANGUAGE DeriveDataTypeable #-}
 
 module ACC 
   ( -- * ACC system
@@ -17,7 +14,7 @@ import Control.Monad
 import Data.Maybe
 import Generic
 import Revlimit 
-import NewARSim      hiding (void)
+import NewARSim
 import PID
 
 -- * Types 
@@ -110,7 +107,8 @@ cruiseCtrl deltaT = atomic $
 
      -- Cruise control mode.
      -- ~~~~~~~~~~~~~~~~~~~~
-     runnable (MinInterval 0) [TimingEvent deltaT] $
+--      runnable (MinInterval 0) [TimingEvent deltaT] $
+     runnableT ["core1" :>> (3, 10)] (MinInterval 0) [TimingEvent deltaT] $ 
        do Ok c <- rteRead crVel
           Ok v <- rteRead vhVel
           Ok m <- rteRead target

### 2016-08-24 ACC/Gearbox.hs

@@ -40,7 +40,8 @@ gearController deltaT =
      gears <- interRunnableVariable initialGear
      timer <- interRunnableVariable (True, 0.8 * timerLimit)
 
-     runnable (MinInterval 0) [TimingEvent deltaT] $
+--      runnable (MinInterval 0) [TimingEvent deltaT] $
+     runnableT ["core1" :>> (6, 10)] (MinInterval 0) [TimingEvent deltaT] $ 
        do res <- rteRead engineRPM
           case res of 
             Ok rpm1 -> do

### 2016-08-24 ACC/Generic.hs

@@ -12,7 +12,7 @@ module Generic
   , sequencer
     -- * Feedthrough component
   , Feedthrough(..)
-  , feedthrough
+  , feedthrough, feedthroughT 
     -- * Signal routing
   , Switch(..)
   , switchRoute
@@ -100,10 +100,23 @@ data Feedthrough a b = Feedthrough
   , feedOut :: DataElem Unqueued b Provided
   }
 
+-- | Task assigned feedthrough component.
+feedthroughT :: (Data a, Data b)
+             => [Task]
+             -> (a -> b)
+             -> b
+             -> AUTOSAR (Feedthrough a b)
+feedthroughT tasks f init = atomic $ do
+  feedIn  <- requiredPort
+  feedOut <- providedPort
+  comSpec feedOut (InitValue init)
+
+  runnableT tasks (MinInterval 0) [DataReceivedEvent feedIn] $
+    do Ok t <- rteRead feedIn
+       rteWrite feedOut (f t)
+  return $ sealBy Feedthrough feedIn feedOut
+
 -- | Feedthrough component.
---
--- *** TODO ***
---  
 -- * Initial value not useful anymore.
 feedthrough :: (Data a, Data b)
             => (a -> b)
@@ -111,16 +124,7 @@ feedthrough :: (Data a, Data b)
             -> b 
             -- ^ Initial value for output
             -> AUTOSAR (Feedthrough a b)
-feedthrough f init = atomic $ do
-     feedIn  <- requiredPort
-     feedOut <- providedPort
-     comSpec feedOut (InitValue init)
-
-     -- Perform some sort of normalisation here
-     runnable (MinInterval 0) [DataReceivedEvent feedIn] $
-       do Ok t <- rteRead feedIn
-          rteWrite feedOut (f t)
-     return $ sealBy Feedthrough feedIn feedOut
+feedthrough = feedthroughT [] 
 
 -- * Signal routing
 -- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

### 2016-08-24 ACC/NewABS2.hs

@@ -84,7 +84,8 @@ bangBangCtrl = atomic $
  
      let width = 20 -- Maximum pulse width.
      
-     runnable (MinInterval 0) [TimingEvent 1e-3] $
+--      runnable (MinInterval 0) [TimingEvent 1e-3] $
+     runnableT ["core1" :>> (2, 1)] (MinInterval 0) [TimingEvent 1e-3] $ 
        do Ok c  <- rteIrvRead carrier
           Ok s0 <- rteRead slips
           Ok s  <- rteCall pidCall (0.2, s0)
@@ -210,7 +211,8 @@ mainLoop = atomic $
       
      mapM_ (`comSpec` InitValue 0.0) slipsOut
       
-     runnable (MinInterval 0) [DataReceivedEvent velo] $
+--      runnable (MinInterval 0) [DataReceivedEvent velo] $
+     runnableT ["core1" :>> (1, 10)] (MinInterval 0) [DataReceivedEvent velo] $ 
        do velos <- mapM (fmap fromOk . rteRead) velosIn
           Ok v1 <- rteRead velo
           forM (velos `zip` slipsOut) $ \(v, p) -> rteWrite p (slipRatio v1 v)

### 2016-08-25 paper/paper.tex
Completed the clarifications section.

@@
 $$
-This means that any runnable instance in $p$ wishing to enter $x$ is effectively blocked from making progress until some other process in $p$ chooses to exit $x$.
+This means that any runnable instance in $p$ wishing to enter $i.x$ is effectively blocked from making progress until some other process in $p$ chooses to exit $i.x$.
 In the same manner, a process wishing to exit an exclusive area not taken is also effectively blocked.
@@
 An exclusive area can thus only be exited by the process that entered it, and only in the reverse order of entering.
-Section~\ref{sect:DiscExcl} will bring up some alternatives to this semantics, and also discuss the interpretation of blocked processes in more detail.
+Section~\ref{sec:ExclNest} will bring up some alternatives to this semantics, and also discuss the interpretation of blocked processes in more detail.
 
@@
 
+\subsection{Data initialization}
+
 Unqueued data elements that are read before first written should return the error code {\bf neverReceived} \cite[ch.~5.6.10]{AR:RTE}.
-Since no exceptions are mentioned in the specification, our intepretation is that this holds even when a data element possesses an {\em initValue} attribute  (third parameter of the initial {\bf delem} state defined in Section~\ref{sec:Sem}).
-However, as this renders any such assignments entirely redundant, an alternative interpretation creates a term
+Since no exceptions to this rule are mentioned in the specification, our intepretation is that this holds even when a data element possesses an {\em initValue} attribute  (initial {\bf delem} state in Section~\ref{sec:Sem}).
+However, as this renders any such initial value assignments entirely redundant, an alternative interpretation would be to create a term
 $$
@@
 
+\subsection{Activation of server runnables}
+
 The specification restricts client-server connections such that a client may not be connected to multiple servers, for the purpose of avoiding an operation call to be handled by more than one server runnable \cite[ch.~4.2.3]{AR:RTE}.
 However, this restriction does not explicitly forbid multiple {\em OperationInvokedEvent}s refering to the same client-server operation to trigger separate runnables.
-Our formulation of client-server communication is based on the assumption that this is just an oversight, as allowing multiple events would have the same problem with ambiguous response values as a would one-to-many client-server connection pattern.
+Our formulation of client-server communication is based on the assumption that this is just an oversight, as allowing multiple events on a server operation would have the same problem with ambiguous response values as a would one-to-many client-server connection pattern.
 %It is not easy to see what could constitute an alternative interpretation without introducing some programmer option for selecting among server results.
 
-Although the AUTOSAR specification only motivates timed runnable activation with the need to support plain periodic execution, there is really nothing that precludes such runnables from being triggered by additional events as well --- including other {\em TimingEvent}s, that might even carry identical periods.
-Our formulation naturally supports this generality, but it should be observed that the inability of AUTOSAR runnables to count pending activations (see Section~\ref{sec:Parallelism}) also renders the resolution of any coinciding timer events highly non-deterministic.
-Restricting
+\subsection{Interleaving of timing events}
 
-It may further be noted that an alternative interpretation where conflicting timers are outruled just means a static model restriction, with no semantic implications on the timer configurations that are actually allowed.
+Although the AUTOSAR specification only motivates timed runnable activation with the need to support plain periodic execution, there is really nothing in the specification that precludes such runnables from being triggered by additional events as well --- including other {\em TimingEvent}s, perhaps even with identical periods \cite[table~7.1, ch.~7.2.3]{AR:SWC}.
+Our formulation naturally supports this generality, but it should be observed that the inability of AUTOSAR runnables to count pending activations (see Section~\ref{sec:Parallelism}) also renders the resolution of coinciding timing events highly non-deterministic.
+As an example, consider the event patterns of two timers with distinct periods that occasionally coincide:
+\begin{center}
+\rule{3mm}{1pt}%
+						\rule[-9pt]{1mm}{10pt}%
+\rule{2mm}{1pt}%
+													\rule{1mm}{10pt}%
+\rule{7mm}{1pt}%
+						\rule[-9pt]{1mm}{10pt}%
+\rule{1mm}{1pt}%
+													\rule{1mm}{10pt}%
+\rule{9mm}{1pt}%
+									\rule[-9pt]{1mm}{19pt}%
+\rule{9mm}{1pt}%
+													\rule{1mm}{10pt}%
+\rule{1mm}{1pt}%
+						\rule[-9pt]{1mm}{10pt}%
+\rule{4mm}{1pt}%
+\end{center}
+If the timers were activating two distinct runnables, the fragment above would be guaranteed to result in eight instantiations in total.
+But if directed towards a single runnable, the number of instances created could be seven as well as eight, with no means for the programmer to control the outcome.
 
-Related to multiple timers is the
+This anomaly may lead to the conclusion that the AUTOSAR specification does not really intend more than one timing event to be defined per runnable.
+Luckily, our formalization is also compatible with this interpretation, as the restriction only concerns what set of static models to expect as starting states, not their dynamic behavior.
 
-DataReceivedEvent at Invalidate?
+Closely related to the interleaving of timing events, and of interest even if only distinct runnables are being triggered, is the notion of initial timer offsets.
+That is, at what time will each timer tick for the first time --- immediately at startup or at some other (later) time?
+The specification documents are totally silent on the issue, and the simplest conclusion is arguably to let all timers fire simultaneously at $0$ seconds from startup.
+Our {\bf timer} terms are thus initialized to the dynamic value $0$ (Section~\ref{sec:Sem}), although it is not hard to see that other values could be equally reasonable, like
+$$
+	\Ttimer {i.r} {t_p} {t_p}
+$$ 
+which lets each timer wait one full cycle before triggering for the first time.
+However, real systems tend to exhibit an arbitrary delay before each timer gets started, and if that behavior is meant to count as legal, the initial timer state must rather be
+$$
+	\Ttimer {i.r} {t_p} {t_r}
+$$ 
+for some arbitrarily chosen non-negative $t_r$.
+One might even want to add the condition $t_r < t_p$, but it must be emphasized that all these initilization options are mere semantic speculations --- the AUTOSAR specification must be clarified before any particular choice can be put on firm ground.
 
-DataReceivedEvent at limit response?
+\subsection{Interpretation of data reception events}
 
-Exclusive area nesting
+Activation of runnables due to {\em DataReceivedEvent}s is faithfully captured in the formal semantics for unqueued as well as queued data elements (c.f.~axioms (17) and (29) with \cite[ch.~7.5.1.6]{AR:SWC}).
+But for unqueued elements AUTOSAR also provides an invalidation operation, whose net effect is very similar to that of a data write (see axiom (21)).
+Whether data invalidation should count as a reception event is not clear from the specification; we have chosen not to include it as a conservative conjecture.
+Revising this choice would not be hard, though: all that is necessary is to add an axiom that mimics (17) but matches on {\sc inv} rather than {\sc wr} labels.
 
-Return value of qelem write with fanout > 1 and *some* LIMIT failures
+A similar specification ambiguity concerns {\em DataReceivedEvent}s in the presence of a full receiver buffer.
+We have chosen to block these events from activating connected runnables (axiom (30)), although the opposite choice could also be justified on grounds that activation of a reciever must be even more pressing when its buffer is overflowing.
+Expressing this formally would amount to simply deleting axiom (30) together with the last precondition of axiom (29).
 
-Rule and examples of provided delegation port compatibility don't match
+\subsection{Exclusive area nesting} \label{sec:ExclNest}
 
-Multiple inconsistent CompSpecs
+The AUTOSAR specification is apparently ambiguous when it comes to exclusive areas, stating that "The RTE is not required to support nested invocations of [enter/exit] for the same exclusive area" \cite[ch.~5.6.28-29]{AR:RTE}.
+Whether entering a previously acquired area should succeed or not is thus left entirely open, something we have taken as a justification not to support it formally (axioms (9) to (12)).
+Still, it is interesting to ponder what would be required to actually allow this optional behavior.
+One aspect is that the referenced exclusive areas themselves need not really be involved in the resulting transitions; they must already be in the occupied state they should retain.
+Another detail is that if a runnable is already holding its referenced exclusive area, both the enter and exit commands essentially turn into {\em no-operations}.
+Thus, the necessary additions to the axiom set would look something like this:
+$$
+\begin{array}{lcll}
+		{\Trinst {i.r} c \Xs {\TrteEnter x k}}
+		\;\; \red{\Tsay {i.r} \Tnop} \\
+		{\Trinst {i.r} c {\Tcons x \Xs} {\Tap k \Tvoid}}
+		& \Pif\;x \in \Xs
+\\ \\
+		{\Trinst {i.r} c {\Tcons x \Xs} {\TrteExit x k}}
+		\;\; \red{\Tsay {i.r} \Tnop} \\
+		{\Trinst {i.r} c \Xs {\Tap k \Tvoid}}
+		& \Pif\;x \in \Xs
+\end{array}
+$$
 
-Crashed runnables
+Another concern regarding exclusive areas is that although the specification clearly demands that exiting is done in the reverse order of entering, it does not provide any means for the exit command to report detected violations of that condition \cite[ch.~5.6.29]{AR:RTE}.
+This means that if a runnable instance tries to exit exclusive area $x$ and $x$ is not the latest exclusive area acquired by that instance, the operation is neither allowed to succeed nor able to flag an error due to a {\em void} return type.
+This leaves our transition semantics with no other option than to indefinitely block progress for such a runnable, which is the consequence of providing no transition alternative besides axiom (11) to a runnable instance about to execute the exit command.
+Had the specification allowed for an error result, however, the following axioms could have been added to resolve the order violation cases:
+$$
+\begin{array}{lcll}
+		{\Trinst {i.r} c {\Tcons y \Xs} {\TrteExit x k}}
+		\;\; \red{\Tsay {i.r} \Tnop} \\
+		{\Trinst {i.r} c {\Tcons y \Xs} {\Tap k {\kw{orderErr}}}}
+		& \Pif\;x \neq y
+\\ \\
+		{\Trinst {i.r} c \Teps {\TrteExit x k}}
+		\;\; \red{\Tsay {i.r} \Tnop} \\
+		{\Trinst {i.r} c \Teps {\Tap k {\kw{orderErr}}}}
+
+\end{array}
+$$
+
+\subsection{Failing runnables}
+
+A transition semantics like the one defined in this report provides rules for deriving the correct behaviors, while leaving the incorrect behaviors without derivation alternatives.
+This has the effect that a process term which attempts to do something illegal is assigned no meaningful remaining trace; it appears indefinitely stuck in its current state as if it were deadlocked and can participate in subsequent transitions only by ignoring what is being said.
+
+An alternative way of handling program failures in general is of course to throw exceptions or trigger some other non-local error-handling mechanism.
+AUTOSAR actually specifies a set of such generic error-handling approaches, but it does so in terms of implementation artefacts like tasks and memory partitions only, rather than the software component concepts under study in this report \cite{AR:ERR}.
+Still, the presence of exceptional errors raises a string of questions when combined with the concurrency and communication features of software components, and a clarification of this aspect could go a long way towards making AUTOSAR a self-contained programming model in its own right.
+We outline one possible way of formalizing the semantics of failing runnables below.
+
+First the code syntax needs to be extended so that failing computations can be represented:
+$$
+\begin{array}{lll}
+  \Code & ::= & \ldots               \\
+        & |   & \Return v            \\
+        & |   & \Throw
+\end{array}
+$$
+A "throw" command has no continuation and is thus similar to the exception generating commands of many modern programming languages.
+It could be given an informative value parameter as well, but this will not be needed in our simplistic approach.
+Throw commands can appear explicitly in the code to indicate software-detected errors, or they can be the result of special failure transitions, like the following alternative handling of an exit order violation:
+$$
+\begin{array}{lcll}
+		{\Trinst {i.r} c {\Tcons y \Xs} {\TrteExit x k}}
+		\;\; \red {\Tsay {i.r} \Tfail} \\
+		{\Trinst {i.r} c {\Tcons y \Xs} \Throw}
+		& \Pif\;x \neq y
+\\ \\
+		{\Trinst {i.r} c \Teps {\TrteExit x k}}
+		\;\; \red {\Tsay {i.r} \Tfail}  \\
+		{\Trinst {i.r} c \Teps \Throw}
+\end{array}
+$$
+Because a failed runnable instance now has an explicit representation, we can specify details of how such a process should interact with its enviroment.
+We may for example wish to define that a special error code is returned if our runnable is running on behalf of a client, and that any exclusive areas held are automatically exited in the right order (c.f.~axioms (42) and (11)):
+$$
+\begin{array}{lcll}
+		{\Trinst a {\Tclient b m \_} \Xs \Throw}
+		\;\; \red{\Tsay b {\Tret m {\kw{serverErr}}}} \\
+		{\Trinst a \Tvoid \Xs \Throw}
+\\ \\
+		{\Trinst {i.r} \Tvoid {\Tcons x \Xs} \Throw}
+		\;\; \red{\Tsay {i.x} \Tex} \\
+		{\Trinst {i.r} \Tvoid \Xs \Throw}
+\end{array}
+$$
+Finally, we can ensure that the instance counter of a runnable is properly decremented even if an instance terminates due to a failure (in interaction with axiom (52)):
+$$
+\begin{array}{rcll}
+		{\Trinst a \Tvoid \Teps \Throw}
+		& \red{\Tsay a \Tterm}
+		\Tzero
+\end{array}
+$$
+Other design choices are of course perfectly possible, and more elaborate rules, for example defining an explicit exception-catching mechanism, could be considered as well.
+We must emphasize, though, that the rules in this section are just hypothetical examples; they are not based on the current AUTOSAR specification and are only presented to illustrate the succinctness and level of detail that can be obtained by chosing a formal semantics aproach.
 

### 2016-08-25 paper/paper.tex
Some more minor fixes to the paper

--- 2016-08-23 paper/paper.tex
@@
 
-\bibliographystyle{splncs03} % Latest bibtex style for llncs
-
 \setlength{\parindent}{0pt}
@@
 
-The AUTOSAR standard is an open software component architecture for the automotive industry \cite{AR:Homepage}.
+The AUTOSAR standard is an open software component architecture for the automotive industry.
 Its main purpose is to enable interoperability of software modules among different vendors and on heterogeneous platforms, via an extensive set of standardized interfaces and libraries, and a common software development methodology.
@@
 
-An AUTOSAR model is primarily a structure of interconnected \textbf{software components} (SWCs) \cite{AR:SWC}.
+An AUTOSAR model is primarily a structure of interconnected \textbf{software components} (SWCs).
 Links between SWCs are called \textbf{connectors}, which attach to \textbf{ports} exposed by the SWCs.
@@
 To interact with ports, inter-runnable variables and exclusive areas, runnables use what is abstractly known as the \textbf{virtual function bus} (VFB).
-Concretely, the VFB takes the shape of a C programming interface to the \textbf{run-time environment} (RTE), custom-generated for each runnable on the basis of its access constraints \cite{AR:RTE}.
+Concretely, the VFB takes the shape of a C programming interface to the \textbf{run-time environment} (RTE), custom-generated for each runnable on the basis of its access constraints.
 Behind the RTE, a collection of OS kernels, communication stacks and other basic software modules implement the VFB services for the platform at hand.
@@
 This behaviour serves us well, because in general, what one AUTOSAR runnable does may have impact on many (if not all) parts of the executing system.
-It is however important not to confuse this broadcasting notion with any particular form of AUTOSAR communication.
+It is, however, important not to confuse this broadcasting notion with any particular form of AUTOSAR communication.
 The way transition labels distribute over the parallel composition operator is merely a technical aspect of our process calculus, and will be used to express several different AUTOSAR communication semantics, among other things.
-
-%% TODO: PaJa: continue reading from here.
 
@@
 
-The AUTOSAR naming scheme guarantees that all processes terms in such an initial composition will carry unique addresses, with exception of timer terms which share addresses with the runnables they belong to.
+The AUTOSAR naming scheme guarantees that all processes terms in such an initial composition will carry unique addresses, with the exception of timer terms which share addresses with the runnables they belong to.
 We will later see how runnable instances reuse the address of their runnables in the same manner.
@@
 Some further static model information will also be referenced by the transition rules.
-We refer to the connectors of a model using a binary relation $\Rightarrow$ on port addresses, where the arrow points in the \emph{provided-to-required} direction.
-We also assume that $\Rightarrow$ is transitively closed (i.e., it expresses end-to-end connectivity across \emph{Delegation-} as well as \emph{AssemblySwConnector}s), and that every connection $a \Rightarrow b$ is lifted to the \emph{DataElement}s and \emph{ClientServerOperation}s of the connected ports such that
+We refer to the connectors of a model using a binary relation $\Tconnect{}{}$ on port addresses, where the arrow points in the \emph{provided-to-required} direction.
+We also assume that $\Tconnect{}{}$ is transitively closed (i.e., it expresses end-to-end connectivity across \emph{Delegation-} as well as \emph{AssemblySwConnector}s), and that every connection $\Tconnect{a}{b}$ is lifted to the \emph{DataElement}s and \emph{ClientServerOperation}s of the connected ports such that
 \[
@@
 How the runnable instances themselves come into existence will be detailed in Section~\ref{sect:SpawnTerm}.
+
+%% TODO: PaJa: continue reading from here.
 
@@
 
-\subsection{Calling a server} \label{sect:Call}
+\subsection{Calling a server}
 
@@
 
-\subsection{Passing back a server result} \label{sect:Res}
+\subsection{Passing back a server result}
 
@@
 \item The non-server runnable toggles back to the $\Tidle$ activation state, while the server just chops of the head of its client queue.
-\item Only the server runnable instance is given a non-void code parameter, which is taken from the same client queue head.
+\item Only the server runnable instance is given a non-void parameter, which is taken from the same client queue head.
 \end{itemize}
@@
 
-The transition relation defined in Appendix~\ref{sec:Rules} is non-deterministic; i.e., it may very well be that for some process $p$,
 $$
 \begin{array}{c}
-	p \red{l_1} p_1 \hspace{10mm}\text{and}\hspace{10mm} p \red{l_2} p_2
+  {\Trunnable a 0 \Tidle 0} \Opar
+  {\Tqelem b 2 \Teps} \vspace{0.6ex} \\
+
+  \red{\hear{b}{\Tsnd {v_1} \Teps}} \\
+
+  {\Trunnable a 0 \Tpending 0} \Opar
+  {\Tqelem b 2 {v_1}} \vspace{0.6ex} \\
+
+  \red{\say{a}{\Tnew}} \\
+
+  {\Trunnable a 0 \Tidle 1} \Opar
+  {\Tqelem b 2 {v_1}} \,\Opar  \\
+  {\Trinst a \Tvoid \Teps {\Code}} \vspace{0.6ex} \\
+
+  \red{\hear{b}{\Tsnd {v_2} \Teps}} \\
+
+  {\Trunnable a 0 \Tpending 1} \Opar
+  {\Tqelem b 2 {\Tcons{v_1}{v_2}}} \,\Opar  \\
+  {\Trinst a \Tvoid \Teps {\Code}} \vspace{0.6ex} \\
+
+  \red{\say{a}{\Tnew}} \\
+
+  {\Trunnable a 0 \Tidle 2} \Opar
+  {\Tqelem b 2 {\Tcons{v_1}{v_2}}} \,\Opar  \\
+  {\Trinst a \Tvoid \Teps {\Code}} \Opar
+  {\Trinst a \Tvoid \Teps {\Code}}
 \end{array}
 $$
-are both possible. If it also holds that
+
 $$
 \begin{array}{c}
-	p_1 \red{l_2} q \hspace{10mm}\text{and}\hspace{10mm} p_2 \red{l_1} q
+  {\Trunnable a 0 \Tidle 0} \Opar
+  {\Tqelem b 2 \Teps} \vspace{0.6ex} \\
+
+  \red{\hear{b}{\Tsnd {v_1} \Teps}} \\
+
+  {\Trunnable a 0 \Tpending 0} \Opar
+  {\Tqelem b 2 {v_1}} \vspace{0.6ex} \\
+
+  \red{\hear{b}{\Tsnd {v_2} \Teps}} \\
+
+  {\Trunnable a 0 \Tpending 0} \Opar
+  {\Tqelem b 2 {\Tcons{v_1}{v_2}}} \vspace{0.6ex} \\
+
+  \red{\say{a}{\Tnew}} \\
+
+  {\Trunnable a 0 \Tidle 2} \Opar
+  {\Tqelem b 2 {\Tcons{v_1}{v_2}}} \,\Opar  \\
+  {\Trinst a \Tvoid \Teps {\Code}}
 \end{array}
 $$
-are derivable, we note that process $p$ may evolve into state $q$ either via an $l_1$ step followed by $l_2$, or by first taking the $l_2$ step and then $l_1$.
-Such freedom of choice is commonly taken as the definition of {\em parallelism} in the process calculus literature --- if the sequential ordering of steps $l_1$ and $l_2$ has no significance, they might just as well be considered to occur in parallel.
 
-However, it may also be that two possible step orderings do not lead to identical results. That happens if, instead of the two transitions leading to $q$ above, we have
-$$
-\begin{array}{c}
-	p_1 \red{l_2} q_1 \hspace{10mm}\text{and}\hspace{10mm} p_2 \red{l_1} q_2
-\end{array}
-$$
-where $q_1$ and $q_2$ are "similar" but not identical states.
-This non-determinism also has a concrete interpretation, either reflecting the non-deterministic nature of a system's environment (messages $l_1$ and $l_2$ just happen to arrive in some particular order) or its internals (processor speed or a scheduling choice just happens to pick $l_1$ before $l_2$).
-It is one of the primary merits of the process calculus approach to allow the scope of such inherent non-determinism to be precisely defined, because although a given system state may render many different behaviors possible, many more are not.
-A formal semantics makes it possible to clearly draw the line between these groups of behaviors.
+\section{Discussion / results}
+\label{sec:Disc}
 
-The AUTOSAR specification contains a very subtle sensitivity to this kind of non-deterministism that is worth examplifying.
-Consider a queued data element $a$ together with a runnable $b$ set up to trigger on data reception events for $a$.
-A scenario where two values $v_1$ and $v_2$ are sent on a port connected to $a$ may lead to the following trace:
-$$
-\begin{array}{c}
-  {\Tqelem a 2 \Teps} \Opar 
-  {\Trunnable b 0 \Tidle 0} \vspace{0.6ex} \\
+\subsection{Ambiguities and alternative interpretations}
+\label{sec:DiscAmb}
 
-  \red{\hear{i.p.e}{\Tsnd {v_1} \Teps}} \\
+\subsection{Clarification proposals and improvements}
+\label{sec:DiscImp}
 
-  {\Tqelem a 2 {v_1}} \Opar 
-  {\Trunnable b 0 \Tpending 0} \vspace{0.6ex} \\
+irv without initValue vs.\ delem without initValue -- ``undefined'' vs.\ NO\_DATA at initial read.
 
-  \red{\say{b}{\Tnew}} \\
+Multiple runnables triggered by same OperationInvokedEvent?
 
-  {\Tqelem a 2 {v_1}} \Opar 
-  {\Trunnable b 0 \Tidle 1} \,\Opar  \\
-  {\Trinst b \Tvoid \Teps {\Code}} \vspace{0.6ex} \\
+Multiple timingEvent for same runnable, possibly with same period?
 
-  \red{\hear{i.p.e}{\Tsnd {v_2} \Teps}} \\
-
-  {\Tqelem a 2 {\Tcons{v_1}{v_2}}} \Opar 
-  {\Trunnable b 0 \Tpending 1} \,\Opar  \\
-  {\Trinst b \Tvoid \Teps {\Code}} \vspace{0.6ex} \\
-
-  \red{\say{b}{\Tnew}} \\
-
-  {\Tqelem a 2 {\Tcons{v_1}{v_2}}} \Opar 
-  {\Trunnable b 0 \Tidle 2} \,\Opar  \\
-  {\Trinst b \Tvoid \Teps {\Code}} \Opar 
-  {\Trinst b \Tvoid \Teps {\Code}}
-\end{array}
-$$
-That is, a resulting state where $a$ holds the two received values together with two concurrent instances of $b$ ready to start reacting upon the two reception events.
-
-However, an alternative trace is equally valid for the same scenario:
-$$
-\begin{array}{c}
-  {\Tqelem a 2 \Teps} \Opar 
-  {\Trunnable b 0 \Tidle 0} \vspace{0.6ex} \\
-
-  \red{\hear{i.p.e}{\Tsnd {v_1} \Teps}} \\
-
-  {\Tqelem a 2 {v_1}} \Opar 
-  {\Trunnable b 0 \Tpending 0} \vspace{0.6ex} \\
-
-  \red{\hear{i.p.e}{\Tsnd {v_2} \Teps}} \\
-
-  {\Tqelem a 2 {\Tcons{v_1}{v_2}}} \Opar 
-  {\Trunnable b 0 \Tpending 0} \vspace{0.6ex} \\
-
-  \red{\say{b}{\Tnew}} \\
-
-  {\Tqelem a 2 {\Tcons{v_1}{v_2}}} \Opar 
-  {\Trunnable b 0 \Tidle 1} \,\Opar  \\
-  {\Trinst b \Tvoid \Teps {\Code}}
-\end{array}
-$$
-Here there is only one $b$ instance created and none is pending, even though $a$ correctly contains the two values received.
-The reason is that by swapping the order of the first instantiation and the second data reception steps, the system is put in a state where its inability to count pending instantiations becomes fatal.
-
-Unfortunately there is very little a programmer can do to influence which of these two traces is taken.
-Giving $b$ a positive {\em minimumStartInterval} just guarantees that any subsequent events are lost during that interval.
-Long delays between reception events will reduce the likelihood of losses, but the choice of when to take an instantiation step is still fully internal to the operating system scheduler.
-Moreover, when messages are produced by multiple parallel senders, maintaining control over their time distribution may be highly impractical.
-Non-concurrent runnables can work around the problem by iterating over all received data at each invocation, but that solution is not very attractive if concurrent processing of each received item is desired.
-
-One could argue that replacing the {\bf pending}/{\bf idle} flag of runnables with a small counter variable would render the event triggering mechanism in AUTOSAR much more robust.
-The standard explicitly excludes event counting for all events but operation invocation, though, so the possibility that events may be lost is undeniably intentional \cite[ch.~4.2.3]{AR:RTE}.
-However, should this design decision ever be revisited, the two trace alternatives above could serve as a succinct problem characterization.
-
-\section{Ambiguities and clarification proposals}
-\label{sec:Amb}
-
-The formal semantics presented in this report is aimed to be a faithful capture of the AUTOSAR specification, especially its Software Components template and RTE specification \cite{AR:SWC, AR:RTE}.
-Still, there are several areas where we have found the specification open to interpretation, or at least in need of some further clarification.
-We detail a few of these areas here, as an illustration to the kind of design decisions that may surface as the result of a formalization undertaking.
-
-Unqueued data elements that are read before first written should return the error code {\bf neverReceived} \cite[ch.~5.6.10]{AR:RTE}.
-Since no exceptions are mentioned in the specification, our intepretation is that this holds even when a data element possesses an {\em initValue} attribute  (third parameter of the initial {\bf delem} state defined in Section~\ref{sec:Sem}).
-However, as this renders any such assignments entirely redundant, an alternative interpretation creates a term
-$$
-  \Tdelem {i.p.e} \Tfalse v
-$$
-for each unqueued data element $i.p.e$, where $v$ is the {\em initValue} of $e$ if it exists, and $\TneverReceived$ otherwise.
-
-Inter-runnable variables can also be read before written, although here there is no special error code defined in the specification for the uninitialized case \cite[ch.~5.6.26]{AR:RTE}.
-Section~\ref{sec:Sem} sets each {\bf irv} contents to {\bf void} to indicate a truly non-existing value, although an entirely reasonable alternative would be
-$$
-  \Tirv {i.s} v
-$$
-where $v$ is the {\em initValue} of $s$ if it exists, and the error code {\bf nodata} otherwise.
-
-The specification restricts client-server connections such that a client may not be connected to multiple servers, for the purpose of avoiding an operation call to be handled by more than one server runnable \cite[ch.~4.2.3]{AR:RTE}.
-However, this restriction does not explicitly forbid multiple {\em OperationInvokedEvent}s refering to the same client-server operation to trigger separate runnables.
-Our formulation of client-server communication is based on the assumption that this is just an oversight, as allowing multiple events would have the same problem with ambiguous response values as a would one-to-many client-server connection pattern.
-%It is not easy to see what could constitute an alternative interpretation without introducing some programmer option for selecting among server results.
-
-Although the AUTOSAR specification only motivates timed runnable activation with the need to support plain periodic execution, there is really nothing that precludes such runnables from being triggered by additional events as well --- including other {\em TimingEvent}s, that might even carry identical periods.
-Our formulation naturally supports this generality, but it should be observed that the inability of AUTOSAR runnables to count pending activations (see Section~\ref{sec:Parallelism}) also renders the resolution of any coinciding timer events highly non-deterministic.
-Restricting
-
-It may further be noted that an alternative interpretation where conflicting timers are outruled just means a static model restriction, with no semantic implications on the timer configurations that are actually allowed.
-
-Related to multiple timers is the
+Timer initial offsets
 
@@
 Return to the limitation - what is the next step (which can be lifted more easily)
-
-\bibliography{refs}
 

### 2016-08-25 paper/paper.tex
Paper updates

@@
 
-%% TODO: PaJa: continue reading from here.
+\subsection{Inter-runnable variables}
 
-\subsection{Inter-runnable variables}
+TODO: Perhaps match up the order of introducing the terms of the initial process term \V{P_0} with these subsections
 
@@
 
+TODO: Dangling pointer to Section~\ref{sect:DiscExcl}
+
 \subsection{Unbuffered sending/receiving}
@@
 The side-condition in axiom (16) makes the rule apply only if there is a static connection from the data element being written and the address of the matched term (recall that data element storage is associated with the receiving side of a connection).
-Because AUTOSAR allows sender-receiver communication patterns such that $\Tconnect {i.p.e} a$ and $\Tconnect {i.p.e} b$ are simultaneously true, parallel data element updates like the following are possible:
+Because AUTOSAR allows sender-receiver communication patterns to be one-to-many (such as $\Tconnect {i.p.e} a$ and $\Tconnect {i.p.e} b$), parallel data element updates like the following are possible:
 $$
@@
 As an aside, note that the associativity of the parallel composition operator is made entirely irrelevant by the symmetric shape of transition rules (1) and (2).
+
+TODO: ? I don't get that aside ? (associativity is about equality of terms: (a|b)|c == a|(b|c) but the rules deal with transitions, not equality.)
 
@@
 By requiring connectivity for all $a$ in $\Tvar{as}$ in axiom (26), the possibility of letting irrelevant addresses in $\Tvar{as}$ cause bogus \textbf{limit} results is ruled out.
+
+%% TODO: PaJa: continue reading from here.
 

### 2016-08-26 paper/paper.tex
More on limitations.

--- 2016-08-25 paper/paper.tex
@@
 
-This paper takes an important step towards a remedy to these problems, by contributing a formal specification of a substantial core of the AUTOSAR standard.
+This paper takes a first step towards a remedy to these problems, by contributing a formal specification of a substantial core of the AUTOSAR standard.
 The formalization covers most of the Software Component Template and its accompanying Run-Time Environment (RTE) (Section~\ref{sec:Calc}), and is able to directly express every legal way a system of software components can evolve at run-time on an arbitrarily fast platform (Section~\ref{sec:Sem}).
 It can thus serve as a basis for both a concrete AUTOSAR implementation on a specific platform (whose supported behaviours must be a subset of those defined by the formal semantics), and a platform-independent AUTOSAR simulator (where behaviours from the legal set can be picked at random).
-Examples of semantic derivations and their applications are found in Section~\ref{sec:Examples}.
+Examples of semantic derivations and their applications can be found throughout the presentation.
 
@@
 The semantics is furthermore written to be unambiguous, except where concurrent execution should allow for more than one observed behaviour.
-At some points, the AUTOSAR documents have been found unclear; here a specific choice has been made but alternative interpretations will be discussed separately (Section~\ref{sec:DiscAmb}).
-At other points, the standard documents are clear but the resulting semantics is still dubious.
-These cases will also be emphasised and discussed, together with suggested ways forward (Section~\ref{sec:DiscImp}).
+At some points, the standard documents are clear but the resulting semantics is dubious.
+These cases will be emphasised and discussed, together with suggested ways forward (Section~\ref{sec:NonDet}).
+At other points, the AUTOSAR documents have been found unclear; here a specific choice has been made but alternative interpretations will be discussed separately (Section~\ref{sec:Amb}).
 
@@
 
-TODO: Some simple (informal) example early to give the reader an early view of flavour of the results.
+%TODO: Some simple (informal) example early to give the reader an early view of flavour of the results.
 
@@
 Some components merely act as containers of other components structures, these are called \textbf{composition} SWCs.
-In contrast, \textbf{atomic} SWCs define their own behaviour in terms of \textbf{runnables} and \textbf{inter-runnable variables}, primarily.
+In contrast, \textbf{atomic} SWCs define their own behaviour, primarily in terms of \textbf{runnables} and \textbf{inter-runnable variables}.
 A runnable denotes a sequential and (normally) terminating piece of code, although its actual behavioural definition is typically relegated to external C or Matlab files rather than being part of the AUTOSAR model proper.
@@
 To this end, we have chosen to ignore all internal computations and just capture the observable effects of runnable execution as a pure sequence of RTE (VFB) calls.
-And because the result from an RTE call may in general affect subsequent runnable behaviour, we use a continuation-passing style where every $\Code$ term except the final function return literally contains the sequence that follows it, as a continuation parameter $\Cont$.
+And because the result from an RTE call may in general affect subsequent runnable behaviour, we use a continuation-passing style where every $\Code$ term except the final return command literally contains the sequence that follows it, as a continuation parameter $\Cont$.
 %
@@
 
-TODO: details on limitations: what's not covered by the calculus
-
 \section{Semantics}
@@
 \renewcommand{\Tstep}[3]{#1 \red{#2} #3}
-\renewcommand\Pcomma{\;\text{and}\;}
+\renewcommand\Pcomma{\;\wedge\;}
 
@@
 This behaviour serves us well, because in general, what one AUTOSAR runnable does may have impact on many (if not all) parts of the executing system.
-It is, however, important not to confuse this broadcasting notion with any particular form of AUTOSAR communication.
+It is however important not to confuse this broadcasting notion with any particular form of AUTOSAR communication.
 The way transition labels distribute over the parallel composition operator is merely a technical aspect of our process calculus, and will be used to express several different AUTOSAR communication semantics, among other things.
+
+%% TODO: PaJa: continue reading from here.
 
@@
 where $\V{T_p}$ is the period of the event.
-%TODO: This means that all timed runnables want to act _immediately_ (at the same time - with no init delay).
 
@@
 
-The AUTOSAR naming scheme guarantees that all processes terms in such an initial composition will carry unique addresses, with the exception of timer terms which share addresses with the runnables they belong to.
+The AUTOSAR naming scheme guarantees that all processes terms in such an initial composition will carry unique addresses, with exception of timer terms which share addresses with the runnables they belong to.
 We will later see how runnable instances reuse the address of their runnables in the same manner.
@@
 Some further static model information will also be referenced by the transition rules.
-We refer to the connectors of a model using a binary relation $\Tconnect{}{}$ on port addresses, where the arrow points in the \emph{provided-to-required} direction.
-We also assume that $\Tconnect{}{}$ is transitively closed (i.e., it expresses end-to-end connectivity across \emph{Delegation-} as well as \emph{AssemblySwConnector}s), and that every connection $\Tconnect{a}{b}$ is lifted to the \emph{DataElement}s and \emph{ClientServerOperation}s of the connected ports such that
+We refer to the connectors of a model using a binary relation $\Rightarrow$ on port addresses, where the arrow points in the \emph{provided-to-required} direction.
+We also assume that $\Rightarrow$ is transitively closed (i.e., it expresses end-to-end connectivity across \emph{Delegation-} as well as \emph{AssemblySwConnector}s), and that every connection $a \Rightarrow b$ is lifted to the \emph{DataElement}s and \emph{ClientServerOperation}s of the connected ports such that
 \[
@@
 How the runnable instances themselves come into existence will be detailed in Section~\ref{sect:SpawnTerm}.
-
-%% TODO: PaJa: continue reading from here.
 
@@
 
-\section{Parallelism}
-\label{sec:Parallelism}
+\section{Non-determinism}
+\label{sec:NonDet}
 
@@
 It is one of the primary merits of the process calculus approach to allow the scope of such inherent non-determinism to be precisely defined, because although a given system state may render many different behaviors possible, many more are not.
-A formal semantics makes it possible to clearly draw the line between these groups of behaviors.
+A formal semantics enables us to clearly draw the line between these groups of behaviors.
 
@@
 
+\subsection{Activation of server runnables}
+
 The specification restricts client-server connections such that a client may not be connected to multiple servers, for the purpose of avoiding an operation call to be handled by more than one server runnable \cite[ch.~4.2.3]{AR:RTE}.
 However, this restriction does not explicitly forbid multiple {\em OperationInvokedEvent}s refering to the same client-server operation to trigger separate runnables.
-Our formulation of client-server communication is based on the assumption that this is just an oversight, as allowing multiple events would have the same problem with ambiguous response values as a would one-to-many client-server connection pattern.
+Our formulation of client-server communication is based on the assumption that this is just an oversight, as allowing multiple events on a server operation would have the same problem with ambiguous response values as a would one-to-many client-server connection pattern.
 %It is not easy to see what could constitute an alternative interpretation without introducing some programmer option for selecting among server results.
 
-Although the AUTOSAR specification only motivates timed runnable activation with the need to support plain periodic execution, there is really nothing that precludes such runnables from being triggered by additional events as well --- including other {\em TimingEvent}s, that might even carry identical periods.
-Our formulation naturally supports this generality, but it should be observed that the inability of AUTOSAR runnables to count pending activations (see Section~\ref{sec:Parallelism}) also renders the resolution of any coinciding timer events highly non-deterministic.
-Restricting
+\subsection{Interleaving of timing events}
 
-It may further be noted that an alternative interpretation where conflicting timers are outruled just means a static model restriction, with no semantic implications on the timer configurations that are actually allowed.
+Although the AUTOSAR specification only motivates timed runnable activation with the need to support plain periodic execution, there is really nothing in the specification that precludes such runnables from being triggered by additional events as well --- including other {\em TimingEvent}s, perhaps even with identical periods \cite[table~7.1, ch.~7.2.3]{AR:SWC}.
+Our formulation naturally supports this generality, but it should be observed that the inability of AUTOSAR runnables to count pending activations (see Section~\ref{sec:NonDet}) also renders the resolution of coinciding timing events highly non-deterministic.
+As an example, consider the event patterns of two timers with distinct periods that occasionally coincide:
+\begin{center}
+\rule{3mm}{1pt}%
+						\rule[-9pt]{1mm}{10pt}%
+\rule{2mm}{1pt}%
+													\rule{1mm}{10pt}%
+\rule{7mm}{1pt}%
+						\rule[-9pt]{1mm}{10pt}%
+\rule{1mm}{1pt}%
+													\rule{1mm}{10pt}%
+\rule{9mm}{1pt}%
+									\rule[-9pt]{1mm}{19pt}%
+\rule{9mm}{1pt}%
+													\rule{1mm}{10pt}%
+\rule{1mm}{1pt}%
+						\rule[-9pt]{1mm}{10pt}%
+\rule{4mm}{1pt}%
+\end{center}
+If the timers were activating two distinct runnables, the fragment above would be guaranteed to result in eight instantiations in total.
+But if directed towards a single runnable, the number of instances created could be seven as well as eight, with no means for the programmer to control the outcome.
 
-Related to multiple timers is the
+This anomaly may lead to the conclusion that the AUTOSAR specification does not really intend more than one timing event to be defined per runnable.
+Luckily, our formalization is also compatible with this interpretation, as the restriction only concerns what set of static models to expect as starting states, not their dynamic behavior.
 
-DataReceivedEvent at Invalidate?
+Closely related to the interleaving of timing events, and of interest even if only distinct runnables are being triggered, is the notion of initial timer offsets.
+That is, at what time will each timer tick for the first time --- immediately at startup or at some other (later) time?
+The AUTOSAR specification only introduces the notion of timing offset in relation to the operating system tasks that are expected to implement runnables; the Software Component Template itself is totally silent on the issue \cite{AR:SWC}.
+Lacking an abstract definition of runnable offsets, we have chosen to simply fire all timers simultaneously at $0$ seconds from startup (hence the value $0$ in the initial {\bf timer} term definition in Section~\ref{sec:Sem}).
+It is not hard to see that other values may be equally reasonable, though, like
+$$
+	\Ttimer {i.r} {t_p} {t_p}
+$$ 
+which lets each timer wait one full cycle before triggering for the first time.
+However, should we really want to take the assumed implementation offsets into account, the initial timer state would rather be
+$$
+	\Ttimer {i.r} {t_p} {t_o}
+$$ 
+for some arbitrarily chosen non-negative $t_o$.
+One might even want to add the condition $t_o < t_p$, but it must be emphasized that all these initilization options are speculative --- the AUTOSAR Software Component Template must be clarified before a sufficiently abstract model of timed behavior can be obtained.
 
-DataReceivedEvent at limit response?
+\subsection{Interpretation of data reception events}
 
-Exclusive area nesting
+Activation of runnables due to {\em DataReceivedEvent}s is faithfully captured in the formal semantics for unqueued as well as queued data elements (c.f.~axioms (17) and (29) with \cite[ch.~7.5.1.6]{AR:SWC}).
+But for unqueued elements AUTOSAR also provides an invalidation operation, whose net effect is very similar to that of a data write (see axiom (21)).
+Whether data invalidation should count as a reception event is not clear from the specification; we have chosen not to include it as a conservative conjecture.
+Revising this choice would not be hard, though: all that is necessary is to add an axiom that mimics (17) but matches on {\sc inv} rather than {\sc wr} labels.
 
-Return value of qelem write with fanout > 1 and *some* LIMIT failures
+A similar specification ambiguity concerns {\em DataReceivedEvent}s in the presence of a full receiver buffer.
+We have chosen to block these events from activating connected runnables (axiom (30)), although the opposite choice could also be justified on grounds that activation of a reciever must be even more pressing when its buffer is overflowing.
+Expressing this formally would amount to simply deleting axiom (30) together with the last precondition of axiom (29).
 
-Rule and examples of provided delegation port compatibility don't match
+\subsection{Exclusive area nesting} \label{sec:ExclNest}
 
-Multiple inconsistent CompSpecs
+The AUTOSAR specification is apparently ambiguous when it comes to exclusive areas, stating that "The RTE is not required to support nested invocations of [enter/exit] for the same exclusive area" \cite[ch.~5.6.28-29]{AR:RTE}.
+Whether entering a previously acquired area should succeed or not is thus left entirely open, something we have taken as a justification not to support it formally (axioms (9) to (12)).
+Still, it is interesting to ponder what would be required to actually allow this optional behavior.
+One aspect is that the referenced exclusive areas themselves need not really be involved in the resulting transitions; they must already be in the occupied state they should retain.
+Another detail is that if a runnable is already holding its referenced exclusive area, both the enter and exit commands essentially turn into {\em no-operations}.
+Thus, the necessary additions to the axiom set would look something like this:
+$$
+\begin{array}{lcll}
+		{\Trinst {i.r} c \Xs {\TrteEnter x k}}
+		\;\; \red{\Tsay {i.r} \Tnop} \\
+		{\Trinst {i.r} c {\Tcons x \Xs} {\Tap k \Tvoid}}
+		& \Pif\;x \in \Xs
+\\ \\
+		{\Trinst {i.r} c {\Tcons x \Xs} {\TrteExit x k}}
+		\;\; \red{\Tsay {i.r} \Tnop} \\
+		{\Trinst {i.r} c \Xs {\Tap k \Tvoid}}
+		& \Pif\;x \in \Xs
+\end{array}
+$$
 
-Crashed runnables
+Another concern regarding exclusive areas is that although the specification clearly demands that exiting is done in the reverse order of entering, it does not provide any means for the exit command to report detected violations of that condition \cite[ch.~5.6.29]{AR:RTE}.
+This means that if a runnable instance tries to exit exclusive area $x$, and $x$ is not the latest exclusive area acquired by that instance, the operation is neither allowed to succeed nor able to flag an error due to a {\em void} return type.
+This leaves our transition semantics with no other option than to indefinitely block progress for such a runnable, which is the consequence of providing no transition alternative besides axiom (11) to a runnable instance about to execute the exit command.
+Had the specification allowed for an error result, however, the following axioms could have been added to resolve the order violation cases:
+$$
+\begin{array}{lcll}
+		{\Trinst {i.r} c {\Tcons y \Xs} {\TrteExit x k}}
+		\;\; \red{\Tsay {i.r} \Tnop} \\
+		{\Trinst {i.r} c {\Tcons y \Xs} {\Tap k {\kw{orderErr}}}}
+		& \Pif\;x \neq y
+\\ \\
+		{\Trinst {i.r} c \Teps {\TrteExit x k}}
+		\;\; \red{\Tsay {i.r} \Tnop} \\
+		{\Trinst {i.r} c \Teps {\Tap k {\kw{orderErr}}}}
+
+\end{array}
+$$
+
+\subsection{Failing runnables}
+
+A transition semantics like the one defined in this report provides rules for deriving the correct behaviors, while leaving the incorrect behaviors without derivation alternatives.
+This has the effect that a process term which attempts to do something illegal is assigned no meaningful remaining trace; it appears indefinitely stuck in its current state as if it were deadlocked and can participate in subsequent transitions only by ignoring what is being said.
+
+An alternative way of handling program failures in general is of course to throw exceptions or trigger some other non-local error-handling mechanism.
+AUTOSAR actually specifies a set of such generic error-handling approaches, but it does so in terms of implementation artefacts like tasks and memory partitions only, rather than the software component concepts under study in this report \cite{AR:ERR}.
+Still, the presence of exceptional errors raises a string of questions when combined with the concurrency and communication features of software components, and a clarification of this aspect could go a long way towards making AUTOSAR a self-contained programming model in its own right.
+We outline one possible way of formalizing the semantics of failing runnables below.
+
+First the code syntax needs to be extended so that failing computations can be represented:
+$$
+\begin{array}{lll}
+  \Code & ::= & \ldots               \\
+        & |   & \Return v            \\
+        & |   & \Throw
+\end{array}
+$$
+A "throw" command has no continuation and is thus similar to the exception generating commands of many modern programming languages.
+It could be given an informative value parameter as well, but this will not be needed in our simplistic approach.
+Throw commands can appear explicitly in the code to indicate software-detected errors, or they can be the result of special failure transitions, like the following alternative handling of an exit order violation:
+$$
+\begin{array}{lcll}
+		{\Trinst {i.r} c {\Tcons y \Xs} {\TrteExit x k}}
+		\;\; \red {\Tsay {i.r} \Tfail} \\
+		{\Trinst {i.r} c {\Tcons y \Xs} \Throw}
+		& \Pif\;x \neq y
+\\ \\
+		{\Trinst {i.r} c \Teps {\TrteExit x k}}
+		\;\; \red {\Tsay {i.r} \Tfail}  \\
+		{\Trinst {i.r} c \Teps \Throw}
+\end{array}
+$$
+Because a failed runnable instance now has an explicit representation, we can specify details of how such a process should interact with its enviroment.
+We may for example wish to define that a special error code is returned if our runnable is running on behalf of a client, and that any exclusive areas held are automatically exited in the right order (c.f.~axioms (42) and (11)):
+$$
+\begin{array}{lcll}
+		{\Trinst a {\Tclient b m \_} \Xs \Throw}
+		\;\; \red{\Tsay b {\Tret m {\kw{serverErr}}}} \\
+		{\Trinst a \Tvoid \Xs \Throw}
+\\ \\
+		{\Trinst {i.r} \Tvoid {\Tcons x \Xs} \Throw}
+		\;\; \red{\Tsay {i.x} \Tex} \\
+		{\Trinst {i.r} \Tvoid \Xs \Throw}
+\end{array}
+$$
+Finally, we can ensure that the instance counter of a runnable is properly decremented even if an instance terminates due to a failure (in interaction with axiom (52)):
+$$
+\begin{array}{rcll}
+		{\Trinst a \Tvoid \Teps \Throw}
+		& \red{\Tsay a \Tterm}
+		\Tzero
+\end{array}
+$$
+Other design choices are of course perfectly possible, and more elaborate rules, for example defining an explicit exception-catching mechanism, could be considered as well.
+We must emphasize, though, that the rules in this section are just hypothetical examples; they are not based on the current AUTOSAR specification and are only presented to illustrate the succinctness and level of detail that can be obtained by chosing a formal semantics aproach.
+
+\section{Limitations}
+
+We claim that the formalized AUTOSAR semantics of this report covers a substantial core of the Software Component Template and Specification of RTE \cite{AR:SWC, AR:RTE}.
+But even these two documents alone add up to almost 2000 pages of specification text in total, covering a wide range of features on many levels of abstraction.
+We have therefore found it necessary to further limit our scope; both by leaving out parts that have very little to do with the concurrency and communication aspects that are our primary interest, and by ignoring parts that to some extent duplicate the features that we are already covering.
+We discuss the chosen omissions in more detail below.
+
+\subsection{Out of scope}
+
+All aspects of the AUTOSAR specification that concern the representation and selection of data values are beyond the scope of the present study. This includes
+\begin{itemize}
+\item Data types on any level of abstraction, and their associated mappings.
+\item Port interface mapping and data scaling.
+\item Measurements and calibration parameters.
+\item Variant handling.
+\end{itemize}
+These modelling elements only impose restrictions and demands on the values $v$ and continuations $k$ we reference in our semantics, and which we in turn do not constrain in any way.
+
+\subsection{Minor variants}
+
+The only atomic software component type we include on our formalization is the {\em ApplicationSwComponentType}.
+Other atomic component types describe dynamic behaviors that are either just minor variants of the formalized component type ({\em SensorActuatorSwComponentType}, {\em ComplexDeviceDriverSwComponentType}, {\em ServiceSwComponentType}, {\em ServiceProxySwComponentType}), or trivial in comparison ({\em NvBlockSwComponentType}, {\em EcuAbstractionSwComponentType}, {\em ParameterSwComponentType}).
+
+A similar argument justifies our decision to omit port types beyond sender-receiver and client-server, as well as per instance memories.
+A port typed by a {\em TriggerInterface} is essentially a sender-receiver port with no data contents.
+Mode management is arguably a much more complex concept in principle, but since the related AUTOSAR mechanism is limited to communication of integer-valued mode requests and acknowledgements, it can be seen as an instance of sender-receiver communication as well.
+Per instance memories essentially duplicate the behavior of inter-runnable variables.
+
+Basic Software modules implement behaviors that cannot be captured as ordinary software components due to their dependency on hardware and other platform artefacts.
+However, our abstract treatment of computation dependencies in general confines such differences to one's choice of $k$ continuations, which elimiates the need to formalize Basic Software components through any special means.
+
+\subsection{Left out for brevity}
+
+A few areas of the AUTOSAR SWC and RTE specifications have been omitted for the purpose of keeping the size of the resulting formalism at a reasonable level.
+These are areas where a worked out formalization could prove interesting, and possibly even discover new design ambiguities.
+However, we also believe that the techniques required for undertaking such an effort are already demonstrated by our current work, and that making the resulting semantics accessible is an even more important goal than striving for completeness.
+The major omissions of this kind are:
+\begin{itemize}
+\item Implicit communication, where designated port elements are automatically read/written at the start/end the of runnables that access them.
+\item Category 2 runnables, which declare certain read or receive operations as {\em WaitPoints} that block progress until the associated events occur.
+While this form of event handling is fundamentally different from the standard mechansim of activating runnables, it may be noted that the implicit result command that follows from calling a {\em synchronousServerCallPoint} already defines the essence of such a blocking {\em WaitPoint}.
+\item Most details found in the {\em ComSpec} annotations that can be attached to ports (currently only {\em initValues} are acknowledged).
+\item COM layer
+\item DataWriteCompletedEvent, DataSendCompletedEvent, DataReceiveErrorEvent, the feedback command
+\item RTE lifecycle and callback APIs
+\end{itemize}
 
@@
 
-\setcounter{equation}{4}
+\setcounter{equation}{0}
+
+\subsection{Parallel composition}
+{
+\renewcommand{\Prule}[2]{#1 \quad \Pif\; #2\\}
+\renewcommand{\Tstep}[3]{#1 \red{#2} #3}
+\renewcommand\Pcomma{\;\text{and}\;}
+
+\begin{eqnarray}  \CombRed  \notag \end{eqnarray}
+}
 

### 2016-08-26 paper/paper.tex

@@
 \subsection{Inter-runnable variables}
+
+TODO: Perhaps match up the order of introducing the terms of the initial process term \V{P_0} with these subsections
 
@@
 
+TODO: ? I don't get that aside ? (associativity is about equality of terms: (a|b)|c == a|(b|c) but the rules deal with transitions, not equality.)
+
 Axiom (17) provides another means of interpreting a data element write.

### 2016-08-27 paper/paper.tex
More on limitations.

@@
 
-A similar argument justifies our decision to omit port types beyond sender-receiver and client-server, as well as per instance memories.
+A similar argument justifies our decision to omit port types beyond the sender-receiver and client-server kinds, as well as per instance memories.
 A port typed by a {\em TriggerInterface} is essentially a sender-receiver port with no data contents.
@@
 \item Category 2 runnables, which declare certain read or receive operations as {\em WaitPoints} that block progress until the associated events occur.
-While this form of event handling is fundamentally different from the standard mechansim of activating runnables, it may be noted that the implicit result command that follows from calling a {\em synchronousServerCallPoint} already defines the essence of such a blocking {\em WaitPoint}.
+While this form of event handling is fundamentally different from the standard mechansim of activating runnables, it may be noted that the result command that follows from calling a {\em synchronousServerCallPoint} already defines the essence of such a blocking {\em WaitPoint}.
+\item The AUTOSAR COM layer and related events and commands ({\em DataWriteCompletedEvent}, {\em DataSendCompletedEvent}, {\em DataReceiveErrorEvent}, the feedback command) \cite{AR:COM}.
+COM is the communication substrate between nodes in a distributed AUTOSAR system and its limited storage capacity introduces an element of resource competition among communicating components.
+It also leads to the separation of sending and reception into distinct events, as well as a new class of communication errors.
 \item Most details found in the {\em ComSpec} annotations that can be attached to ports (currently only {\em initValues} are acknowledged). 
-\item COM layer
-\item DataWriteCompletedEvent, DataSendCompletedEvent, DataReceiveErrorEvent, the feedback command
-\item RTE lifecycle and callback APIs
+\item The meta-programming oriented RTE Lifecycle and Callback APIs.
 \end{itemize}

### 2016-08-27 paper/paper.tex
Added figure and refs.
Paper: just a few fixes + added forgotten file.

@@
 \usepackage[makeroom]{cancel}
+\usepackage{graphicx}
+
 % \usepackage{minted}
@@
 
+\begin{figure*}
+\includegraphics[page=1,width=\textwidth]{Fig}
+%\includegraphics[page=1,scale=0.73]{Fig}
+\caption{\label{fig:model}{Example of a graphical AUTOSAR software components model.}}
+\end{figure*}
+
 Some components merely act as containers of other components structures, these are called \textbf{composition} SWCs.
@@
 A graphical model example is shown in Figure~\ref{fig:model}.
-
-TODO: add fig:model (from the slides)
 
@@
 
-Although our chosen mechanism for coordinating transitions among process terms goes under the name {\em broadcast}, there is no implication that every process must react to everything being said --- the intent is rather that a process term not directly or indirectly addressed in a transition should be allowed to ignore it.
+Although our chosen mechanism for coordinating transitions among process terms goes under the name \emph{broadcast}, there is no implication that every process must react to everything being said --- the intent is rather that a process term not directly or indirectly addressed in a transition should be allowed to ignore it.
 One could try to express this as some form of "catch-up" rule $p \red{l} p$, that would apply only if none of the axioms (5) to (64) match $p$ and $l$.
@@
 However, under the different assumption that there is no connection from ${i.p.e}$ to $b$, axioms (28) and (70) must trade places.
-This also forces the removal of $b$ from the second argument of the {\sc snd} payload (i.e., the list of connected buffers not able to store the value sent).
+This also forces the removal of $b$ from the second argument of the \payload{snd} payload (i.e., the list of connected buffers not able to store the value sent).
 As a consequence, axiom (26) is no longer applicable and must be replaced by axiom (25).
-The resulting parallel transition looks as follows, illustrating the correct generation of an {\bf ok} result despite the presence of a full (but unconnected) buffer.
+The resulting parallel transition looks as follows, illustrating the correct generation of an \Tatom{ok} result despite the presence of a full (but unconnected) buffer.
 $$
@@
 are derivable, we note that process $p$ may evolve into state $q$ either via an $l_1$ step followed by $l_2$, or by first taking the $l_2$ step and then $l_1$.
-Such freedom of choice is commonly taken as the definition of {\em parallelism} in the process calculus literature --- if the sequential ordering of steps $l_1$ and $l_2$ has no significance, they might just as well be considered to occur in parallel.
+Such freedom of choice is commonly taken as the definition of \emph{parallelism} in the process calculus literature --- if the sequential ordering of steps $l_1$ and $l_2$ has no significance, they might just as well be considered to occur in parallel.
 
@@
 Unfortunately there is very little a programmer can do to influence which of these two traces is taken.
-Giving $b$ a positive {\em minimumStartInterval} just guarantees that any subsequent events are lost during that interval.
+Giving $b$ a positive \emph{minimumStartInterval} just guarantees that any subsequent events are lost during that interval.
 Long delays between reception events will reduce the likelihood of losses, but the choice of when to take an instantiation step is still fully internal to the operating system scheduler.
@@
 Unqueued data elements that are read before first written should return the error code {\bf neverReceived} \cite[ch.~5.6.10]{AR:RTE}.
-Since no exceptions to this rule are mentioned in the specification, our intepretation is that this holds even when a data element possesses an {\em initValue} attribute  (initial {\bf delem} state in Section~\ref{sec:Sem}).
+Since no exceptions to this rule are mentioned in the specification, our intepretation is that this holds even when a data element possesses an \emph{initValue} attribute  (initial {\bf delem} state in Section~\ref{sec:Sem}).
 However, as this renders any such initial value assignments entirely redundant, an alternative interpretation would be to create a term
@@
 $$
-for each unqueued data element $i.p.e$, where $v$ is the {\em initValue} of $e$ if it exists, and $\TneverReceived$ otherwise.
+for each unqueued data element $i.p.e$, where $v$ is the \emph{initValue} of $e$ if it exists, and $\TneverReceived$ otherwise.
 
@@
 $$
-where $v$ is the {\em initValue} of $s$ if it exists, and the error code {\bf nodata} otherwise.
+where $v$ is the \emph{initValue} of $s$ if it exists, and the error code {\bf nodata} otherwise.
 
@@
 The specification restricts client-server connections such that a client may not be connected to multiple servers, for the purpose of avoiding an operation call to be handled by more than one server runnable \cite[ch.~4.2.3]{AR:RTE}.
-However, this restriction does not explicitly forbid multiple {\em OperationInvokedEvent}s refering to the same client-server operation to trigger separate runnables.
+However, this restriction does not explicitly forbid multiple \emph{OperationInvokedEvent}s refering to the same client-server operation to trigger separate runnables.
 Our formulation of client-server communication is based on the assumption that this is just an oversight, as allowing multiple events on a server operation would have the same problem with ambiguous response values as a would one-to-many client-server connection pattern.
@@
 
-Although the AUTOSAR specification only motivates timed runnable activation with the need to support plain periodic execution, there is really nothing in the specification that precludes such runnables from being triggered by additional events as well --- including other {\em TimingEvent}s, perhaps even with identical periods \cite[table~7.1, ch.~7.2.3]{AR:SWC}.
+Although the AUTOSAR specification only motivates timed runnable activation with the need to support plain periodic execution, there is really nothing in the specification that precludes such runnables from being triggered by additional events as well --- including other \emph{TimingEvent}s, perhaps even with identical periods \cite[table~7.1, ch.~7.2.3]{AR:SWC}.
 Our formulation naturally supports this generality, but it should be observed that the inability of AUTOSAR runnables to count pending activations (see Section~\ref{sec:NonDet}) also renders the resolution of coinciding timing events highly non-deterministic.
@@
 
-Activation of runnables due to {\em DataReceivedEvent}s is faithfully captured in the formal semantics for unqueued as well as queued data elements (c.f.~axioms (17) and (29) with \cite[ch.~7.5.1.6]{AR:SWC}).
+Activation of runnables due to \emph{DataReceivedEvent}s is faithfully captured in the formal semantics for unqueued as well as queued data elements (c.f.~axioms (17) and (29) with \cite[ch.~7.5.1.6]{AR:SWC}).
 But for unqueued elements AUTOSAR also provides an invalidation operation, whose net effect is very similar to that of a data write (see axiom (21)).
@@
 
-A similar specification ambiguity concerns {\em DataReceivedEvent}s in the presence of a full receiver buffer.
+A similar specification ambiguity concerns \emph{DataReceivedEvent}s in the presence of a full receiver buffer.
 We have chosen to block these events from activating connected runnables (axiom (30)), although the opposite choice could also be justified on grounds that activation of a reciever must be even more pressing when its buffer is overflowing.
@@
 One aspect is that the referenced exclusive areas themselves need not really be involved in the resulting transitions; they must already be in the occupied state they should retain.
-Another detail is that if a runnable is already holding its referenced exclusive area, both the enter and exit commands essentially turn into {\em no-operations}.
+Another detail is that if a runnable is already holding its referenced exclusive area, both the enter and exit commands essentially turn into \emph{no-operations}.
 Thus, the necessary additions to the axiom set would look something like this:
@@
 Another concern regarding exclusive areas is that although the specification clearly demands that exiting is done in the reverse order of entering, it does not provide any means for the exit command to report detected violations of that condition \cite[ch.~5.6.29]{AR:RTE}.
-This means that if a runnable instance tries to exit exclusive area $x$, and $x$ is not the latest exclusive area acquired by that instance, the operation is neither allowed to succeed nor able to flag an error due to a {\em void} return type.
+This means that if a runnable instance tries to exit exclusive area $x$, and $x$ is not the latest exclusive area acquired by that instance, the operation is neither allowed to succeed nor able to flag an error due to a \emph{void} return type.
 This leaves our transition semantics with no other option than to indefinitely block progress for such a runnable, which is the consequence of providing no transition alternative besides axiom (11) to a runnable instance about to execute the exit command.
@@
 
-The only atomic software component type we include on our formalization is the {\em ApplicationSwComponentType}.
-Other atomic component types describe dynamic behaviors that are either just minor variants of the formalized component type ({\em SensorActuatorSwComponentType}, {\em ComplexDeviceDriverSwComponentType}, {\em ServiceSwComponentType}, {\em ServiceProxySwComponentType}), or trivial in comparison ({\em NvBlockSwComponentType}, {\em EcuAbstractionSwComponentType}, {\em ParameterSwComponentType}).
+The only atomic software component type we include on our formalization is the \emph{ApplicationSwComponentType}.
+Other atomic component types describe dynamic behaviors that are either just minor variants of the formalized component type (\emph{SensorActuatorSwComponentType}, \emph{ComplexDeviceDriverSwComponentType}, \emph{ServiceSwComponentType}, \emph{ServiceProxySwComponentType}), or trivial in comparison (\emph{NvBlockSwComponentType}, \emph{EcuAbstractionSwComponentType}, \emph{ParameterSwComponentType}).
 
 A similar argument justifies our decision to omit port types beyond the sender-receiver and client-server kinds, as well as per instance memories.
-A port typed by a {\em TriggerInterface} is essentially a sender-receiver port with no data contents.
+A port typed by a \emph{TriggerInterface} is essentially a sender-receiver port with no data contents.
 Mode management is arguably a much more complex concept in principle, but since the related AUTOSAR mechanism is limited to communication of integer-valued mode requests and acknowledgements, it can be seen as an instance of sender-receiver communication as well.
@@
 \item Implicit communication, where designated port elements are automatically read/written at the start/end the of runnables that access them.
-\item Category 2 runnables, which declare certain read or receive operations as {\em WaitPoints} that block progress until the associated events occur.
-While this form of event handling is fundamentally different from the standard mechansim of activating runnables, it may be noted that the result command that follows from calling a {\em synchronousServerCallPoint} already defines the essence of such a blocking {\em WaitPoint}.
-\item The AUTOSAR COM layer and related events and commands ({\em DataWriteCompletedEvent}, {\em DataSendCompletedEvent}, {\em DataReceiveErrorEvent}, the feedback command) \cite{AR:COM}.
+\item Category 2 runnables, which declare certain read or receive operations as \emph{WaitPoints} that block progress until the associated events occur.
+While this form of event handling is fundamentally different from the standard mechansim of activating runnables, it may be noted that the result command that follows from calling a \emph{synchronousServerCallPoint} already defines the essence of such a blocking \emph{WaitPoint}.
+\item The AUTOSAR COM layer and related events and commands (\emph{DataWriteCompletedEvent}, \emph{DataSendCompletedEvent}, \emph{DataReceiveErrorEvent}, the feedback command) \cite{AR:COM}.
 COM is the communication substrate between nodes in a distributed AUTOSAR system and its limited storage capacity introduces an element of resource competition among communicating components.
 It also leads to the separation of sending and reception into distinct events, as well as a new class of communication errors.
-\item Most details found in the {\em ComSpec} annotations that can be attached to ports (currently only {\em initValues} are acknowledged). 
+\item Most details found in the \emph{ComSpec} annotations that can be attached to ports (currently only \emph{initValues} are acknowledged).
 \item The meta-programming oriented RTE Lifecycle and Callback APIs.

### 2016-08-28 paper/paper.tex
Paper: read and made minor changes to the last part

@@
 
-The transition relation defined in Appendix~\ref{sec:Rules} is non-deterministic; i.e., it may very well be that for some process $p$,
+The transition relation defined in Appendix~\ref{sec:Rules} is non-deterministic; i.e., it may very well be that for some process~$p$,
 $$
@@
 Consider a queued data element $a$ together with a runnable $b$ set up to trigger on data reception events for $a$.
+%**TODO: The explanation is rather long so it may be good to indicate in a few words what it will show already at start: "... lead to the runnable triggering once or twice"
 A scenario where two values $v_1$ and $v_2$ are sent on a port connected to $a$ may lead to the following trace:
+%
+TODO: The i.p.e in the transition label is not explained.
 $$
@@
 The reason is that by swapping the order of the first instantiation and the second data reception steps, the system is put in a state where its inability to count pending instantiations becomes fatal.
+%*TODO: Perhaps abbreviate the terms to make the difference more clear: let Q(vs) = \Tqelem a 2 {vs}; R(n) = \Trunnable b 0 \Tidle n; RI = \Trinst b \Tvoid \Teps {\Code}
+
+%   Q(\Teps) \Opar R(0)  ->*  Q(\Tcons{v_1}{v_2}}) \Opar R(1) \Opar RI
+% and
+%   Q(\Teps) \Opar R(0)  ->*  Q(\Tcons{v_1}{v_2}}) \Opar R(2) \Opar RI \Opar RI
 
@@
 
+%**TODO: The end of the previous section also seems to be a clarification proposal
+
 The formal semantics presented in this report is aimed to be a faithful capture of the AUTOSAR specification, especially its Software Components template and RTE specification \cite{AR:SWC, AR:RTE}.
@@
 for each unqueued data element $i.p.e$, where $v$ is the \emph{initValue} of $e$ if it exists, and $\TneverReceived$ otherwise.
+%*TODO: perhaps replace i.p.e by a?
 
@@
 The specification restricts client-server connections such that a client may not be connected to multiple servers, for the purpose of avoiding an operation call to be handled by more than one server runnable \cite[ch.~4.2.3]{AR:RTE}.
-However, this restriction does not explicitly forbid multiple \emph{OperationInvokedEvent}s refering to the same client-server operation to trigger separate runnables.
+However, this restriction does not explicitly forbid multiple \emph{Operation\-InvokedEvent}s refering to the same client-server operation to trigger separate runnables.
 Our formulation of client-server communication is based on the assumption that this is just an oversight, as allowing multiple events on a server operation would have the same problem with ambiguous response values as a would one-to-many client-server connection pattern.
@@
 
-Activation of runnables due to \emph{DataReceivedEvent}s is faithfully captured in the formal semantics for unqueued as well as queued data elements (c.f.~axioms (17) and (29) with \cite[ch.~7.5.1.6]{AR:SWC}).
+Activation of runnables due to \emph{DataReceived\-Event}s is faithfully captured in the formal semantics for unqueued as well as queued data elements (c.f.~axioms (17) and (29) with \cite[ch.~7.5.1.6]{AR:SWC}).
 But for unqueued elements AUTOSAR also provides an invalidation operation, whose net effect is very similar to that of a data write (see axiom (21)).
@@
 
-A similar specification ambiguity concerns \emph{DataReceivedEvent}s in the presence of a full receiver buffer.
+A similar specification ambiguity concerns \emph{DataReceived\-Event}s in the presence of a full receiver buffer.
 We have chosen to block these events from activating connected runnables (axiom (30)), although the opposite choice could also be justified on grounds that activation of a reciever must be even more pressing when its buffer is overflowing.
@@
 We claim that the formalized AUTOSAR semantics of this report covers a substantial core of the Software Component Template and Specification of RTE \cite{AR:SWC, AR:RTE}.
-But even these two documents alone add up to almost 2000 pages of specification text in total, covering a wide range of features on many levels of abstraction.
+But these two documents alone add up to almost 2000 pages of specification text in total, covering a wide range of features on many levels of abstraction.
 We have therefore found it necessary to further limit our scope; both by leaving out parts that have very little to do with the concurrency and communication aspects that are our primary interest, and by ignoring parts that to some extent duplicate the features that we are already covering.
@@
 
+%*TODO: Perhaps avoid repeating the common suffix SwComponentType (with explanation).
 The only atomic software component type we include on our formalization is the \emph{ApplicationSwComponentType}.

### 2016-08-29 paper/paper.tex
Addressed Patrik's comments.

@@
 An instance of runnable $r$ of component $i$, currently executing $\Code$ and owning the exclusive areas $\Xs$, possibly running on behalf of client $c$.
+\item $\Tirv {\Ts s i} v$ \newline
+Inter-runnable variable $s$ of component $i$, currently holding value $v$.
 \item $\Texcl {\Tx x i} u$ \newline
 Exclusive area $x$ of component $i$, with boolean occupation state $u$.
-\item $\Tirv {\Ts s i} v$ \newline
-Inter-runnable variable $s$ of component $i$, currently holding value $v$.
 \item $\Tdelem {\Tep e p i} u v$ \newline
@@
         & |   & \Trinst a c {\Xs} {\Code}    \\
+        & |   & \Tirv a v                    \\
         & |   & \Texcl a u                   \\
-        & |   & \Tirv a v                    \\
         & |   & \Tdelem a u v                \\
@@
 
+\item For each inter-runnable variable $s$ of each atomic component $i$, a term
+\[
+\begin{array}{c}
+  \Tirv {\Ts s i} v
+\end{array}
+\]
+where $v$ is the \emph{initValue} attributed to $s$ if it exists, otherwise $v$ is $\Tvoid$.
+
 \item For each exclusive area $x$ of each atomic component $i$, a term
@@
 \]
-
-\item For each inter-runnable variable $s$ of each atomic component $i$, a term
-\[
-\begin{array}{c}
-  \Tirv {\Ts s i} v
-\end{array}
-\]
-where $v$ is the \emph{initValue} attributed to $s$ if it exists, otherwise $v$ is $\Tvoid$.
 
@@
 \subsection{Inter-runnable variables}
-
-TODO: Perhaps match up the order of introducing the terms of the initial process term \V{P_0} with these subsections
 
@@
 
-TODO: Dangling pointer to Section~\ref{sect:DiscExcl}
-
 \subsection{Unbuffered sending/receiving}
@@
 $$
-As an aside, note that the associativity of the parallel composition operator is made entirely irrelevant by the symmetric shape of transition rules (1) and (2).
-
-TODO: ? I don't get that aside ? (associativity is about equality of terms: (a|b)|c == a|(b|c) but the rules deal with transitions, not equality.)
+As an aside, note that the ordering of terms in a parallel composition is made entirely irrelevant by the symmetric formulation of transition rules (1) and (2).
 
@@
 By requiring connectivity for all $a$ in $\Tvar{as}$ in axiom (26), the possibility of letting irrelevant addresses in $\Tvar{as}$ cause bogus \textbf{limit} results is ruled out.
-
-%% TODO: PaJa: continue reading from here.
 
@@
 Consider a queued data element $a$ together with a runnable $b$ set up to trigger on data reception events for $a$.
-%**TODO: The explanation is rather long so it may be good to indicate in a few words what it will show already at start: "... lead to the runnable triggering once or twice"
-A scenario where two values $v_1$ and $v_2$ are sent on a port connected to $a$ may lead to the following trace:
-%
-TODO: The i.p.e in the transition label is not explained.
+A scenario where two values $v_1$ and $v_2$ are sent on a port element $i.p.e$ connected to $a$ may lead to the following trace:
 $$
@@
 The reason is that by swapping the order of the first instantiation and the second data reception steps, the system is put in a state where its inability to count pending instantiations becomes fatal.
-%*TODO: Perhaps abbreviate the terms to make the difference more clear: let Q(vs) = \Tqelem a 2 {vs}; R(n) = \Trunnable b 0 \Tidle n; RI = \Trinst b \Tvoid \Teps {\Code}
-
-%   Q(\Teps) \Opar R(0)  ->*  Q(\Tcons{v_1}{v_2}}) \Opar R(1) \Opar RI
-% and
-%   Q(\Teps) \Opar R(0)  ->*  Q(\Tcons{v_1}{v_2}}) \Opar R(2) \Opar RI \Opar RI
+
+So, in an abbreviated form, both the following derivations are possible:
+$$
+\begin{array}{l}
+  q \Opar p \rightarrow \cdots \rightarrow q_1 \Opar p_1 \Opar p_{\text{inst}} \\
+  q \Opar p \rightarrow \cdots \rightarrow q_1 \Opar p_2 \Opar p_{\text{inst}} \Opar p_{\text{inst}}
+\end{array}
+$$
+That is, two data reception events in sequence can lead to the creation of either one or two runnable instances, and both behaviors are equally correct according to the AUTOSAR semantics.
 
@@
 Giving $b$ a positive \emph{minimumStartInterval} just guarantees that any subsequent events are lost during that interval.
-Long delays between reception events will reduce the likelihood of losses, but the choice of when to take an instantiation step is still fully internal to the operating system scheduler.
+Long delays between reception events reduce the likelihood of losses, but the choice of when to take an instantiation step is still fully internal to the operating system scheduler.
 Moreover, when messages are produced by multiple parallel senders, maintaining control over their time distribution may be highly impractical.
@@
 
-One could argue that replacing the {\bf pending}/{\bf idle} flag of runnables with a small counter variable would render the event triggering mechanism in AUTOSAR much more robust.
-The standard explicitly excludes event counting for all events but operation invocation, though, so the possibility that events may be lost is undeniably intentional \cite[ch.~4.2.3]{AR:RTE}.
+One could argue that replacing the {\bf pending}/{\bf idle} flag of runnables with a small counter variable would render the AUTOSAR event triggering mechanism much more robust.
+The standard explicitly excludes counting for all events but operation invocation, though, so the possibility of event losses is undeniably intentional \cite[ch.~4.2.3]{AR:RTE}.
 However, should this design decision ever be revisited, the two trace alternatives above could serve as a succinct problem characterization.
@@
 for each unqueued data element $i.p.e$, where $v$ is the \emph{initValue} of $e$ if it exists, and $\TneverReceived$ otherwise.
-%*TODO: perhaps replace i.p.e by a?
 
@@
 $$
-for some arbitrarily chosen non-negative $t_o$.
+for some chosen non-negative $t_o$.
 One might even want to add the condition $t_o < t_p$, but it must be emphasized that all these initilization options are speculative --- the AUTOSAR Software Component Template must be clarified before a sufficiently abstract model of timed behavior can be obtained.
@@
 A "throw" command has no continuation and is thus similar to the exception generating commands of many modern programming languages.
-It could be given an informative value parameter as well, but this will not be needed in our simplistic approach.
-Throw commands can appear explicitly in the code to indicate software-detected errors, or they can be the result of special failure transitions, like the following alternative handling of an exit order violation:
+It could be given an informative value parameter as well, but this will not be needed in this simplistic proposal.
+Throw commands can appear explicitly in the code to indicate software-detected errors, or they can be the result of special failure transitions, like the following alternative resolution of an exit order violation:
 $$
@@
 $$
-Other design choices are of course perfectly possible, and more elaborate rules, for example defining an explicit exception-catching mechanism, could be considered as well.
-We must emphasize, though, that the rules in this section are just hypothetical examples; they are not based on the current AUTOSAR specification and are only presented to illustrate the succinctness and level of detail that can be obtained by chosing a formal semantics aproach.
+Other design choices are of course perfectly possible, and more elaborate rules, for example involving an explicit exception-catching mechanism, could be considered as well.
+We must emphasize, though, that the rules in this section are just hypothetical examples; they are not based on the current AUTOSAR specification and are only presented to illustrate the succinctness and level of detail that can be obtained by a formal semantics aproach.
 
@@
 
-All aspects of the AUTOSAR specification that concern the representation and selection of data values are beyond the scope of the present study. This includes
+All aspects of the AUTOSAR specification that concern representation and selection of data values are beyond the scope of the present study. This includes
 \begin{itemize}
@@
 
-%*TODO: Perhaps avoid repeating the common suffix SwComponentType (with explanation).
-The only atomic software component type we include on our formalization is the \emph{ApplicationSwComponentType}.
-Other atomic component types describe dynamic behaviors that are either just minor variants of the formalized component type (\emph{SensorActuatorSwComponentType}, \emph{ComplexDeviceDriverSwComponentType}, \emph{ServiceSwComponentType}, \emph{ServiceProxySwComponentType}), or trivial in comparison (\emph{NvBlockSwComponentType}, \emph{EcuAbstractionSwComponentType}, \emph{ParameterSwComponentType}).
+The only atomic software component type we include on our formalization is the \emph{Application} software component type.
+Other atomic component types describe dynamic behaviors that are either just minor variants of the formalized component type (\emph{SensorActuator}, \emph{ComplexDeviceDriver}, \emph{Service}, \emph{ServiceProxy}), or trivial in comparison (\emph{NvBlock}, \emph{EcuAbstraction}, \emph{Parameter}).
 
@@
 While this form of event handling is fundamentally different from the standard mechansim of activating runnables, it may be noted that the result command that follows from calling a \emph{synchronousServerCallPoint} already defines the essence of such a blocking \emph{WaitPoint}.
-\item The AUTOSAR COM layer and related events and commands (\emph{DataWriteCompletedEvent}, \emph{DataSendCompletedEvent}, \emph{DataReceiveErrorEvent}, the feedback command) \cite{AR:COM}.
+\item The AUTOSAR COM layer and related events and commands (\emph{DataWriteCompleted}, \emph{DataSendCompleted}, \emph{DataReceiveError}, the {\em Feedback} command) \cite{AR:COM}.
 COM is the communication substrate between nodes in a distributed AUTOSAR system and its limited storage capacity introduces an element of resource competition among communicating components.
@@
 \item Most details found in the \emph{ComSpec} annotations that can be attached to ports (currently only \emph{initValues} are acknowledged).
-\item The meta-programming oriented RTE Lifecycle and Callback APIs.
+\item The meta-programming oriented RTE {\em Lifecycle} and {\em Callback} APIs.
 \end{itemize}
@@
 
-%TODO: RuleLong?
 \subsection{Calling a server}

### 2016-08-29 paper/paper.tex
Interim update.

@@
 
-%TODO: Some simple (informal) example early to give the reader an early view of flavour of the results.
-
 \section{AUTOSAR}
@@
 
-The formal approach we have chosen is that of a typical process calculus \cite{TODO}, where a system of concurrent processes evolves in a sequence of atomic steps, as determined by a global set of transition rules.
+The formal approach we have chosen is that of a typical process calculus \cite{milner:ComMob}, where a system of concurrent processes evolves in a sequence of atomic steps, as determined by a global set of transition rules.
 Our process terms correspond to the individual state-carrying parts of an AUTOSAR system, like the runnables, runnable instances, inter-runnable variables and port elements.
@@
 Since AUTOSAR components can be arbitrarily nested, the component names \V{I} can also be considered to have a similarly nested internal structure (component \V{I_1} within component \V{I_2} within component \V{I_3}, etc).
-We can fully ignore this detail here, though, as the hierarchical layout of components in an AUTOSAR system has no run-time implications \cite{TODO2}.
+We can fully ignore this detail here, though, as the hierarchical layout of components in an AUTOSAR system has no run-time implications \cite[ch.~3.3.1]{AR:SWC}.
 
@@
 
-The effect of these agreement rules closely resembles the idea of a \emph{broadcast}: what one process says may be heard by every other process in a large parallel composition.
+The effect of these agreement rules closely resembles the idea of a \emph{broadcast}: what one process says may be heard by every other process in a large parallel composition \cite{prasad:cbs}.
 This behaviour serves us well, because in general, what one AUTOSAR runnable does may have impact on many (if not all) parts of the executing system.
@@
 
-%**TODO: The end of the previous section also seems to be a clarification proposal
-
 The formal semantics presented in this report is aimed to be a faithful capture of the AUTOSAR specification, especially its Software Components template and RTE specification \cite{AR:SWC, AR:RTE}.

### 2016-08-29 paper/paper.tex
Paper: merge
Nuked the port meta-variables!

@@
 Components and ports do not carry any dynamic state besides the state of their constituent parts, so they need no explicit representation as process terms.
-However, in order to still be able to talk about model elements using their locally scoped names, we employ a hierarchical addressing scheme, such that address $\Tr{\V{R}}{\V{I}}$ means runnable \V{R} of component \V{I} and $\Tep{\V{E}}{\V{P}}{\V{I}}$ means element \V{E} of port \V{P} within component \V{I}.
+However, in order to still be able to talk about model elements using their locally scoped names, we employ a hierarchical addressing scheme, such that address $i.r$ means runnable $r$ of component $i$ and $i.e$ means port element $e$ within component $i$.
 We use the following meta-variable conventions:
@@
 \begin{array}{rcll}
-  \V{I} & \in & \text{Component names} \\
-  \V{R} & \in & \text{Runnable names} \\
-  \V{P} & \in & \text{Port names} \\
-  \V{E} & \in & \text{Sender-receiver element names} \\
-  \V{O} & \in & \text{Client-server operation names} \\
-  \V{S} & \in & \text{Inter-runnable variable names} \\
-  \V{X} & \in & \text{Exclusive area names} \\
+  i & \in & \text{Component names} \\
+  r & \in & \text{Runnable names} \\
+  e & \in & \text{Sender-receiver port element names} \\
+  o & \in & \text{Client-server port operation names} \\
+  s & \in & \text{Inter-runnable variable names} \\
+  x & \in & \text{Exclusive area names} \\
 \\
-  \V{A},\V{B} & \in & \text{Addresses} \\
-  \V{A},\V{B} & ::= & \Tr{\V{R}}{\V{I}}   \sep   i.p  \sep  \Tep{\V{E}}{\V{P}}{\V{I}}   \sep \Top{\V{O}}{\V{P}}{\V{I}}   \sep   \Ts{\V{S}}{\V{I}}   \sep   \Tx{\V{X}}{\V{I}}   \\
+  a,b & \in & \text{Addresses} \\
+  a,b & ::= & i.r  \sep  i.e   \sep i.o   \sep   i.s   \sep   i.x   \\
 \end{array}
@@
 %
-Since AUTOSAR components can be arbitrarily nested, the component names \V{I} can also be considered to have a similarly nested internal structure (component \V{I_1} within component \V{I_2} within component \V{I_3}, etc).
+AUTOSAR elements and operations are actually only unique within the {\em ports} they belong to, so technically we ought to address them by reference to an intended {\em port name} as well, as in $i.p.e$ or $i.p.o$ for some port $p$.
+However, port names play no role of their own in our study, so we choose to leave them be implied by the element and operation names we reference.
+Moreover, component names \V{I} can also be considered to have a nested internal structure (component \V{I_1} within component \V{I_2} within component \V{I_3}, etc).
 We can fully ignore this detail here, though, as the hierarchical layout of components in an AUTOSAR system has no run-time implications \cite[ch.~3.3.1]{AR:SWC}.
@@
 \begin{itemize}
-\item $\Trunnable {\Tr r i} t \Act n$  \newline
+\item $\Trunnable {i.r} t \Act n$  \newline
 Dynamic state for runnable $r$ of component $i$, indicating $n$ current instances, $t$ seconds left until next possible instantiation, and activation state $\Act$.
-\item $\Trinst {\Tr r i} c \Xs \Code$ \newline
+\item $\Trinst {i.r} c \Xs \Code$ \newline
 An instance of runnable $r$ of component $i$, currently executing $\Code$ and owning the exclusive areas $\Xs$, possibly running on behalf of client $c$.
-\item $\Tirv {\Ts s i} v$ \newline
+\item $\Tirv {i.s} v$ \newline
 Inter-runnable variable $s$ of component $i$, currently holding value $v$.
-\item $\Texcl {\Tx x i} u$ \newline
+\item $\Texcl {i.x} u$ \newline
 Exclusive area $x$ of component $i$, with boolean occupation state $u$.
-\item $\Tdelem {\Tep e p i} u v$ \newline
-Non-queued sender-receiver data element $e$ in port $p$ of component $i$, currently holding value $v$ with boolean update status flag $u$.
-\item $\Tqelem {\Tep e p i} n \Vs$ \newline
-Queued sender-receiver data element $e$ in port $p$ of component $i$ with maximum capacity $n$, currently holding value sequence $\Vs$.
-\item $\Toper {\Top o p i} m \Srv$ \newline
-Client-server operation $o$ in port $p$ of component $i$, currently holding sequence number $m$ and call state $\Srv$.
-\item $\Ttimer {\Tr r i} {t_p} t$ \newline
+\item $\Tdelem {i.e} u v$ \newline
+Non-queued sender-receiver port element $e$ of component $i$, currently holding value $v$ with boolean update status flag $u$.
+\item $\Tqelem {i.e} n \Vs$ \newline
+Queued sender-receiver port element $e$ of component $i$ with maximum capacity $n$, currently holding value sequence $\Vs$.
+\item $\Toper {i.o} m \Srv$ \newline
+Client-server port operation $o$ of component $i$, currently holding sequence number $m$ and call state $\Srv$.
+\item $\Ttimer {i.r} {t_p} t$ \newline
 Periodic timer for runnable $r$ of component $i$, with a period of $t_p$ and $t$ seconds left of its current cycle.
@@
   \Srv     & \in & \text{Call states} \\ \\
-  c        & ::= & \Tclient {\Top o p i} m v \sep \Tvoid \\
+  c        & ::= & \Tclient {i.o} m v \sep \Tvoid \\
   \Act     & ::= & \Tidle \sep \Tpending \sep \Cs \\
@@
         & |   & \IrvRead s \Cont             \\
-        & |   & \Send {\Te e p} v \Cont      \\
-        & |   & \Receive {\Te e p} \Cont     \\
-        & |   & \Write {\Te e p} v \Cont     \\
-        & |   & \Read {\Te e p} \Cont        \\
-        & |   & \Invalidate {\Te e p} \Cont  \\
-        & |   & \IsUpdated {\Te e p} \Cont   \\
-        & |   & \Call {\To o p} v \Cont      \\
-        & |   & \Result {\To o p} \Cont      \\
+        & |   & \Send e v \Cont      \\
+        & |   & \Receive e \Cont     \\
+        & |   & \Write e v \Cont     \\
+        & |   & \Read e \Cont        \\
+        & |   & \Invalidate e \Cont  \\
+        & |   & \IsUpdated e \Cont   \\
+        & |   & \Call o v \Cont      \\
+        & |   & \Result o \Cont      \\
         & |   & \Return v                    \\
@@
 \]
-The possible forms of the extra label information \V{D} will be introduced as the different transition rules are defined.
+The possible forms of the extra label detail \V{D} will be introduced as the different transition rules are defined.
 
@@
 
-\item For each data element $e$ of each required port $p$ of each atomic component $i$, a term
-\[
-\begin{array}{c}
-  \Tqelem {\Tep e p i} n \epsilon
+\item For each {\em required} port element $e$ of each atomic component $i$, a term
+\[
+\begin{array}{c}
+  \Tqelem {i.e} n \epsilon
 \end{array}
@@
 \begin{array}{c}
-  \Tdelem {\Tep e p i} \Tfalse \TneverReceived
-\end{array}
-\]
-
-\item For each operation $o$ of each required port $p$ of each atomic component $i$, a term
-\[
-\begin{array}{c}
-  \Toper{\Top o p i} 0 {\Tdone \Tvoid}
+  \Tdelem {i.e} \Tfalse \TneverReceived
+\end{array}
+\]
+
+\item For each {\em required} port operation $o$ of each atomic component $i$, a term
+\[
+\begin{array}{c}
+  \Toper{i.o} 0 {\Tdone \Tvoid}
 \end{array}
@@
 \begin{array}{c}
-  \Ttimer {\Tr r i } {t_p} 0
+  \Ttimer {i.r} {t_p} 0
 \end{array}
@@
 Some further static model information will also be referenced by the transition rules.
-We refer to the connectors of a model using a binary relation $\Tconnect{}{}$ on port addresses, where the arrow points in the \emph{provided-to-required} direction.
-We also assume that $\Tconnect{}{}$ is transitively closed (i.e., it expresses end-to-end connectivity across \emph{Delegation-} as well as \emph{AssemblySwConnector}s), and that every connection $\Tconnect{a}{b}$ is lifted to the \emph{DataElement}s and \emph{ClientServerOperation}s of the connected ports such that
-\[
-\begin{array}{lll}
-  \Tconnect{a.e}{b.e} & \text{for all elements $e$ common to $a$ and $b$} \\
-  \Tconnect{a.o}{b.o} & \text{for all operations $o$ common to $a$ and $b$} \\
-\end{array}
-\]
+We refer to the connectors of a model using a binary relation $\Tconnect{}{}$ on port elements and operations, where the arrow points in the \emph{provided-to-required} direction.
+Although connectors are tecnically defined for ports, we assume $\Tconnect{}{}$ is lifted to the common elements and operations of connected ports, and that it is transitively closed (i.e., expresses end-to-end connectivity across \emph{Delegation-} as well as \emph{AssemblySwConnector}s).
 
@@
 We write $\Timplementation{i.r}{k}$ to state the assumption that the static AUTOSAR model under study assigns implementation function $k$ to runnable $i.r$.
-Similar notations will be used to express other references to the underlying static model; for example $\TdataReceivedEvent {i.r} {i.p.e}$ to state that the model allows runnable $i.r$ to be triggered by data reception events on port $i.p.e$.
+Similar notations will be used to express other references to the underlying static model; for example $\TdataReceivedEvent {i.r} {i.e}$ to state that the model allows runnable $i.r$ to be triggered by data reception events on port element $i.e$.
 Although the chosen notation is sometimes significantly shorter than the UML/XML-based syntax used in the AUTOSAR specification, the intended meaning should nevertheless be clear.
@@
 The side-condition in axiom (16) makes the rule apply only if there is a static connection from the data element being written and the address of the matched term (recall that data element storage is associated with the receiving side of a connection).
-Because AUTOSAR allows sender-receiver communication patterns to be one-to-many (such as $\Tconnect {i.p.e} a$ and $\Tconnect {i.p.e} b$), parallel data element updates like the following are possible:
+Because AUTOSAR allows sender-receiver communication patterns to be one-to-many (such as $\Tconnect {i.e} a$ and $\Tconnect {i.e} b$), parallel data element updates like the following are possible:
 $$
@@
   {\Tdelem a \_ \_} \Opar {\Tdelem b \_ \_} \vspace{0.6ex} \\
-  \red{\say{i.p.e}{\Twr v}} \\
+  \red{\say{i.e}{\Twr v}} \\
   \Trinst {i.r} c \Xs {\Tap k \Tvoid} \,\Opar \\
@@
 It expresses that a runnable set up to trigger on data reception on required data element $a$ should be marked as \textbf{pending} whenever a write occurs on an element connected to $a$.
-An example may be as follows (assuming $\Tconnect {i.p.e} a$ and $\TdataReceivedEvent b a$:
+An example may be as follows (assuming $\Tconnect {i.e} a$ and $\TdataReceivedEvent b a$:
 $$
@@
   {\Tdelem a \_ \_} \Opar {\Trunnable b t \_ n} \vspace{0.6ex} \\
-  \red{\say{i.p.e}{\Twr v}} \\
+  \red{\say{i.e}{\Twr v}} \\
   \Trinst {i.r} c \Xs {\Tap k \Tvoid} \,\Opar \\
@@
 
-The following example illustrates a scenario where a sent value is only stored in a subset of the connected buffers. Assuming $\Tconnect {\Tep e p i} a$ and $\Tconnect {\Tep e p i} b$:
+The following example illustrates a scenario where a sent value is only stored in a subset of the connected buffers. Assuming $\Tconnect {i.e} a$ and $\Tconnect {i.e} b$:
 $$
@@
   {\Tqelem a 2 \Teps} \Opar {\Tqelem b 2 {\Tcons {v_1} {v_2}}} \vspace{0.6ex} \\
-  \red{\say{i.p.e}{\Tsnd v b}} \\
+  \red{\say{i.e}{\Tsnd v b}} \\
   \Trinst {i.r} c \Xs {\Tap k \Tlimit} \,\Opar \\
@@
 
-Here is an example of how a client runnable instance, a client port operation, and a server runnable interact during a call transition (assuming $\ToperationInvokedEvent b a$, $\Tconnect a {i.p.o}$, and $\TserverCallPointTimeout{i.p.o} t $).
+Here is an example of how a client runnable instance, a client port operation, and a server runnable interact during a call transition (assuming $\ToperationInvokedEvent b a$, $\Tconnect a {i.o}$, and $\TserverCallPointTimeout{i.o} t $).
 $$
@@
   {\Trinst {i.r} c \Xs {\TrteCall {p.o} v k}} \,\Opar  \\
-  {\Toper {i.p.o} m {\Tdone \Tnodata}} \Opar \\
+  {\Toper {i.o} m {\Tdone \Tnodata}} \Opar \\
   {\Trunnable b 0 \Cs 0} \vspace{0.6ex} \\
-  \red{\say{i.p.o}{\Tcall m v}} \\
+  \red{\say{i.o}{\Tcall m v}} \\
   {\Trinst {i.r} c \Xs {\Tap k \Tok}} \,\Opar \\
-  {\Toper {i.p.o} m {\Tcalling t}} \Opar \\
-  {\Trunnable b 0 {\Tcons \Cs {\Tclient {i.p.o} m v}}  0}
+  {\Toper {i.o} m {\Tcalling t}} \Opar \\
+  {\Trunnable b 0 {\Tcons \Cs {\Tclient {i.o} m v}}  0}
 \end{array}
@@
 \begin{array}{c}
-  {\Trinst b {\Tclient {i.p.o} {91} {v_0}} \Xs {\Treturn v}} \,\Opar  \\
-  {\Toper {i.p.o} {91} {\Tcalling t}} \vspace{0.6ex} \\
-  \red{\say{i.p.o}{\Tret {91} v}} \\
+  {\Trinst b {\Tclient {i.o} {91} {v_0}} \Xs {\Treturn v}} \,\Opar  \\
+  {\Toper {i.o} {91} {\Tcalling t}} \vspace{0.6ex} \\
+  \red{\say{i.o}{\Tret {91} v}} \\
   {\Trinst b \Tvoid \Xs {\Treturn \Tvoid}} \,\Opar \\
-  {\Toper {i.p.o} {92} {\Tdone v}}
+  {\Toper {i.o} {92} {\Tdone v}}
 \end{array}
@@
 \begin{array}{c}
-  {\Trinst b {\Tclient {i.p.o} {91} {v_0}} \Xs {\Treturn v}} \,\Opar  \\
-  {\Toper {i.p.o} {91} {\Tcalling 0}} \vspace{0.6ex} \\
-  \red{\say{i.p.o}{\Tret {91} \Ttimeout}} \\
-  {\Trinst b {\Tclient {i.p.o} {91} {v_0}} \Xs {\Treturn v}} \,\Opar  \\
-  {\Toper {i.p.o} {92} {\Tdone \Ttimeout}}
-\end{array}
-$$
-The processes are now in a state where runnable instance $b$ is prohibited from doing a normal return. Progress is only possible via axioms (46) and (47).
-$$
-\begin{array}{c}
-  {\Trinst b {\Tclient {i.p.o} {91} {v_0}} \Xs {\Treturn v}} \,\Opar  \\
-  {\Toper {i.p.o} {92} {\Tdone \Ttimeout}} \vspace{0.6ex} \\
-  \red{\say{i.p.o}{\Tskip {91}}} \\
+  {\Trinst b {\Tclient {i.o} {91} {v_0}} \Xs \Code} \,\Opar  \\
+  {\Toper {i.o} {91} {\Tcalling 0}} \vspace{0.6ex} \\
+  \red{\say{i.o}{\Tret {91} \Ttimeout}} \\
+  {\Trinst b {\Tclient {i.o} {91} {v_0}} \Xs \Code} \,\Opar  \\
+  {\Toper {i.o} {92} {\Tdone \Ttimeout}}
+\end{array}
+$$
+The processes are now in a state where runnable instance $b$ is prohibited from doing a normal return. Termination of $b$ is only possible via axioms (46) and (47).
+$$
+\begin{array}{c}
+  {\Trinst b {\Tclient {i.o} {91} {v_0}} \Xs {\Treturn v}} \,\Opar  \\
+  {\Toper {i.o} {92} {\Tdone \Ttimeout}} \vspace{0.6ex} \\
+  \red{\say{i.o}{\Tskip {91}}} \\
   {\Trinst b \Tvoid \Xs {\Treturn \Tvoid}} \,\Opar \\
-  {\Toper {i.p.o} {92} {\Tdone \Ttimeout}}
+  {\Toper {i.o} {92} {\Tdone \Ttimeout}}
 \end{array}
@@
   {\Trinst b \Tvoid \Teps {\Treturn v}} \vspace{0.6ex} \\
-  \red{\say{i.p.o}{\Tterm}} \\
+  \red{\say{i.o}{\Tterm}} \\
   {\Trunnable b t \Act 2} \Opar \Tzero \\
@@
 The axioms that enable each individual process transition in that example are (26), (27) and (28), respectively (and rules (1) and (3) allow them to be combined).
-In particular, axiom (28) is applicable because of the assumption $\Tconnect {i.p.e} b$, which simultaneously blocks applicability of axiom (70).
-However, under the different assumption that there is no connection from ${i.p.e}$ to $b$, axioms (28) and (70) must trade places.
+In particular, axiom (28) is applicable because of the assumption $\Tconnect {i.e} b$, which simultaneously blocks applicability of axiom (70).
+However, under the different assumption that there is no connection from ${i.e}$ to $b$, axioms (28) and (70) must trade places.
 This also forces the removal of $b$ from the second argument of the \payload{snd} payload (i.e., the list of connected buffers not able to store the value sent).
@@
   {\Tqelem a 2 \Teps} \Opar {\Tqelem b 2 {\Tcons {v_1} {v_2}}} \vspace{0.6ex} \\
-  \red{\say{i.p.e}{\Tsnd v \Teps}} \\
+  \red{\say{i.e}{\Tsnd v \Teps}} \\
   \Trinst {i.r} c \Xs {\Tap k \Tok} \,\Opar \\
@@
 Consider a queued data element $a$ together with a runnable $b$ set up to trigger on data reception events for $a$.
-A scenario where two values $v_1$ and $v_2$ are sent on a port element $i.p.e$ connected to $a$ may lead to the following trace:
+A scenario where two values $v_1$ and $v_2$ are sent on a port element $i.e$ connected to $a$ may lead to the following trace:
 $$
@@
 
-  \red{\hear{i.p.e}{\Tsnd {v_1} \Teps}} \\
+  \red{\hear{i.e}{\Tsnd {v_1} \Teps}} \\
 
@@
 
-  \red{\hear{i.p.e}{\Tsnd {v_2} \Teps}} \\
+  \red{\hear{i.e}{\Tsnd {v_2} \Teps}} \\
 
@@
 $$
-  \Tdelem {i.p.e} \Tfalse v
-$$
-for each unqueued data element $i.p.e$, where $v$ is the \emph{initValue} of $e$ if it exists, and $\TneverReceived$ otherwise.
+  \Tdelem {i.e} \Tfalse v
+$$
+for each unqueued data element $i.e$, where $v$ is the \emph{initValue} of $e$ if it exists, and $\TneverReceived$ otherwise.
 

### 2016-08-29 arsim-examples/AUTOSAR.hs
Added suite of examples under package "arsim-examples" (prev. NewABS and ACC)

@@ -1,0 +1,26 @@
+-- Placeholder, not sure to export here.
+--
+-- Some good & bad examples maybe?
+module AUTOSAR
+  ( -- * NewABS, NewABS2
+    standaloneNewABS
+  , simulinkNewABS2
+    -- * ACC example
+  , simulinkACC
+  ) where
+
+import qualified AUTOSAR.NewABS.Standalone
+import qualified AUTOSAR.NewABS2.Simulink
+import qualified AUTOSAR.ACC.Simulink
+
+-- | Standalone ABS example.
+standaloneNewABS :: IO ()
+standaloneNewABS = AUTOSAR.NewABS.Standalone.main
+
+-- | Simulink ABS example.
+simulinkNewABS2 :: IO ()
+simulinkNewABS2 = AUTOSAR.NewABS2.Simulink.main
+
+-- | Simulink ACC example.
+simulinkACC :: IO ()
+simulinkACC = AUTOSAR.ACC.Simulink.main

### 2016-08-29 arsim-examples/AUTOSAR/ACC/ACC.hs

@@ -1,0 +1,129 @@
+-- | Adaptive Cruise Control. 
+
+{-# LANGUAGE RecordWildCards   #-}
+
+module AUTOSAR.ACC.ACC
+  ( -- * ACC system
+    ACCSystem(..)
+  , accSystem
+    -- * Types
+  , Velo, Throttle
+  ) where
+
+import AUTOSAR.ACC.Revlimit 
+import AUTOSAR.Shared.Generic
+import AUTOSAR.Shared.PID
+import Control.Monad
+import Data.Maybe
+import NewARSim
+
+-- * Types 
+-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+
+type Dist     = Double
+type Velo     = Double
+type Throttle = Double
+
+-- * ACC composition
+-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+
+data ACCSystem = ACCSystem
+  { -- Inputs
+    accCruise   :: DataElem Unqueued Velo Required
+  , accVeloIn   :: DataElem Unqueued Velo Required
+  , accTarget   :: DataElem Unqueued (Maybe (Velo, Dist)) Required
+    -- Outputs
+  , accThrottle :: DataElem Unqueued Throttle Provided
+  , accBrake    :: DataElem Unqueued Throttle Provided
+  }
+
+-- | Adaptive Cruise Control system.
+accSystem :: Time -> AUTOSAR ACCSystem
+accSystem timeStep = composition $
+  do -- Sample time resolution of the entire system. Inherited 
+     -- by the target vehicle sensor and the ACC ECU.
+    
+     accECU <- cruiseCtrl timeStep
+
+     -- PD controller
+     -- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+     -- Create a PD controller for the ACC. Parameters are 
+     -- found by trial-and-error.
+     pidCtrl <- pdController (timeStep / 10) 5e-3 5e-1
+
+     -- Connections
+     connect (pidOp pidCtrl)    (ctrl accECU)
+     
+     accCruise <- requiredDelegate [crVel accECU]
[…]

### 2016-08-29 arsim-examples/AUTOSAR/ACC/Gearbox.hs

@@ -1,0 +1,94 @@
+-- | Automatic transmission for Simulink 4-speed gearbox.
+--
+-- The transmission component 'gearController' runs on a set sample time
+-- (preferably but not necessarily the same as the rest of the system) and
+-- periodically reads engine RPM and attempts to shift up/down based on this.
+--
+-- The code is not intended as a faithful model of an automatic transmission,
+-- but is rather there to give us larger freedom with speed settings in
+-- Simulink, since the vehicle's engine model is somewhat sensitive to RPM.
+module AUTOSAR.ACC.Gearbox 
+  ( GearCtrl(..)
+  , gearController
+  ) where
+
+import AUTOSAR.Shared.Generic
+import Control.Monad
+import NewARSim
+
+-- | The gear logic requires engine rpm as input, and provides a gear signal
+-- in the range @[-1, 4]@ as output.
+data GearCtrl = GearCtrl
+  { engineRPM  :: DataElem Unqueued Double  Required
+  , gearSignal :: DataElem Unqueued Integer Provided
+  }
+
+-- | @'gearController' dt@ creates a gear controller reading RPM inputs
+-- periodically with an interval length of @dt@. 
+gearController :: Time -> AUTOSAR GearCtrl
+gearController deltaT = 
+  let initialGear = -1   -- Initial gear setting
+      timerLimit  = 0.5  -- Time delay after a gear change has been made.
+  in atomic $ 
+  do engineRPM  <- requiredPort
+     gearSignal <- providedPort
+     comSpec gearSignal (InitValue initialGear)
+   
+     -- Initial setup: Gearbox in neutral and timer enabled to prevent
+     -- gear lock-in before Simulink vehicle model stabilizes.
+     revs  <- interRunnableVariable 0.0
+     gears <- interRunnableVariable initialGear
+     timer <- interRunnableVariable (True, 0.8 * timerLimit)
+
+--      runnable (MinInterval 0) [TimingEvent deltaT] $
+     runnableT ["core1" :>> (6, 10)] (MinInterval 0) [TimingEvent deltaT] $ 
+       do res <- rteRead engineRPM
+          case res of 
+            Ok rpm1 -> do
+              Ok rpm0  <- rteIrvRead revs
+              Ok gear0 <- rteIrvRead gears
+
+              -- Check timer, we want delays between shifts to let the engine
+              -- adjust revs.
+              Ok (counting, t) <- rteIrvRead timer
[…]

### 2016-08-29 arsim-examples/AUTOSAR/ACC/Simulink.hs

@@ -1,0 +1,37 @@
+-- | External simulation of the ACC component using Simulink (or similar). 
+module AUTOSAR.ACC.Simulink where
+
+import AUTOSAR.ACC.Vehicle
+import Control.Monad
+import NewARSim 
+
+-- * External simulation
+-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+
+-- Current task assignment:
+--
+--  TASK "core1", 1 ms
+--  ORDER    Runnable       Execution policy
+--  ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ 
+--  0        velocityCtrl   Every 10 cycles, 10 ms 
+--  1        mainLoop       (Data reception)
+--  2        bangBang       Every cycle, 1 ms
+--  3        radarCtrl      Every 10 cycles, 10 ms
+--  4        cruiseCtrl     Every 10 cycles, 10 ms
+--  5        brakeCtrl      (Data reception)
+--  6        throttleCtrl   (Data reception)
+--  7        gearController Every 10 cycles, 10 ms
+--
+
+-- TODO: Perhaps we should be able to set task assignments from here,
+-- and we could keep some copies of this file instead?
+
+system :: AUTOSAR IOModule
+system = do
+  declareTask "core1" (TimingEvent 1e-3)
+  vehicleIO 
+
+main :: IO ()
+main = do 
+  simulateUsingExternal True system 
+  return ()

### 2016-08-29 arsim-examples/AUTOSAR/ACC/Vehicle.hs

@@ -1,0 +1,178 @@
+-- | This module contains all vehicle-related things, such as IO/control of
+-- vehicle subsystems, velocity and slip estimation, et cetera. 
+
+{-# LANGUAGE RecordWildCards #-}
+
+module AUTOSAR.ACC.Vehicle 
+  ( IOModule(..)
+  , vehicleIO
+  ) where
+
+import AUTOSAR.ACC.ACC
+import AUTOSAR.ACC.Gearbox
+import AUTOSAR.ACC.Revlimit
+import AUTOSAR.NewABS2.NewABS2
+import AUTOSAR.Shared.Generic
+import AUTOSAR.Shared.Velocity
+
+import Control.Monad
+import NewARSim 
+
+type Dist = Double
+
+-- * Radar control
+-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+
+data RadarCtrl = RadarCtrl
+  { radarDistIn :: DataElem Unqueued Double                   Required
+  , radarRelIn  :: DataElem Unqueued Double                   Required
+  , radarOutput :: DataElem Unqueued (Maybe (Double, Double)) Provided
+  }
+
+radarCtrl :: Time -> Double -> AUTOSAR RadarCtrl
+radarCtrl deltaT thresh = atomic $ do
+  distIn <- requiredPort
+  relIn  <- requiredPort
+  out    <- providedPort
+
+  comSpec out    (InitValue Nothing)
+  comSpec distIn (InitValue 100)
+  comSpec relIn  (InitValue 0.0)
+
+  runnable (MinInterval 0) [TimingEvent deltaT] $ do
+    Ok d <- rteRead distIn
+    Ok r <- rteRead relIn
+    let output
+          | d < thresh = Just (r, d)
+          | otherwise  = Nothing
+    rteWrite out output
+
+  return $ sealBy RadarCtrl distIn relIn out
+
+-- * Throttle control
+-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+
+data ThrottleCtrl = ThrottleCtrl
+  { throttleCtrlIn  :: DataElem Unqueued Double Required
+  , throttleCtrlOut :: DataElem Unqueued Double Provided
[…]

### 2016-08-29 arsim-examples/AUTOSAR/NewABS2/NewABS2.hs

@@ -1,0 +1,277 @@
+{-
+Copyright (c) 2014-2016, Johan Nordlander, Jonas Duregård, Michał Pałka,
+                         Patrik Jansson and Josef Svenningsson
+All rights reserved.
+
+Redistribution and use in source and binary forms, with or without
+modification, are permitted provided that the following conditions are met:
+
+   * Redistributions of source code must retain the above copyright notice,
+     this list of conditions and the following disclaimer.
+   * Redistributions in binary form must reproduce the above copyright
+     notice, this list of conditions and the following disclaimer in the
+     documentation and/or other materials provided with the distribution.
+   * Neither the name of the Chalmers University of Technology nor the names of its
+     contributors may be used to endorse or promote products derived from this
+     software without specific prior written permission.
+
+THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE
+DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE
+FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
+DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
+SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER
+CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY,
+OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
+OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
+-}
+
+{-# LANGUAGE TypeFamilies       #-}
+{-# LANGUAGE RecordWildCards    #-}
+{-# LANGUAGE FlexibleInstances  #-}
+
+module AUTOSAR.NewABS2.NewABS2
+  ( -- * ABS system
+    ABS(..)
+  , absSystem
+    -- * Valve ports
[…]

### 2016-08-29 arsim-examples/AUTOSAR/NewABS2/Simulink.hs

@@ -1,0 +1,46 @@
+module AUTOSAR.NewABS2.Simulink where
+
+import AUTOSAR.NewABS2.NewABS2
+import AUTOSAR.Shared.Velocity
+import Control.Monad
+import NewARSim
+import System.Random
+
+enumLabel :: [(Address, String)] -> [(Address, String)]
+enumLabel is = map (\((a, l), i) -> (a, l ++ show i)) $ is `zip` [1..]
+
+data Vehicle = Vehicle
+  { wheelsIn  :: [DataElem Unqueued Double Required]
+  , accelIn   :: DataElem Unqueued Double Required 
+  , valvesOut :: [ValveP Provided]
+  }
+
+instance External Vehicle where
+  fromExternal vh = relabel "ACCEL" (fromExternal (accelIn vh)) ++
+                    enumLabel (relabel "WHEEL" (fromExternal (wheelsIn vh)))
+  toExternal = toExternal . valvesOut 
+
+-- | Export wheel ports and accelerometer connection from the ABS system.
+-- TODO: Export tasks + time delta
+vehicleIO :: AUTOSAR Vehicle
+vehicleIO = do
+  declareTask "core1" (TimingEvent 1e-3)
+  absCtrl  <- absSystem
+  veloCtrl <- velocityCtrl 1e-2
+
+  connect (velocity veloCtrl) (absVeloIn absCtrl)
+  
+  let ws = wheels veloCtrl `zip` map veloIn (wheelPorts absCtrl)
+  wheelsIn  <- forM ws $ \(vc, wp) -> requiredDelegate [vc, wp]
+  valvesOut <- forM (map valveOut (wheelPorts absCtrl)) $ \vp ->
+    providedDelegate [vp]
+  
+  accelIn <- requiredDelegate [accel veloCtrl]
+
+  return $ sealBy Vehicle wheelsIn accelIn valvesOut 
+
+main :: IO ()
+main = do
+  simulateUsingExternal True vehicleIO
+  return ()
+

### 2016-08-29 arsim-examples/AUTOSAR/Shared/Generic.hs

@@ -1,0 +1,214 @@
+{-# LANGUAGE DeriveDataTypeable #-}
+
+-- | Generic components / skeletons. 
+-- 
+-- Provides some basic generic components/skeletons, such as Johan's sequencer
+-- and the generic feedthrough component. 
+-- 
+-- NOTE: Most things in here are largely unused at this point.
+module AUTOSAR.Shared.Generic
+  ( -- * Sequencer skeleton
+    SeqState(..), Ticks, Limit, Index
+  , sequencer
+    -- * Feedthrough component
+  , Feedthrough(..)
+  , feedthrough, feedthroughT 
+    -- * Signal routing
+  , Switch(..)
+  , switchRoute
+    -- * Signal conversion
+  , Trigger(..)
+  , trigger
+  ) where
+
+import Control.Monad
+import NewARSim      hiding (void)
+
+-- * Sequencer skeleton
+-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+-- Johan's generic sequencer.
+
+-- | `SeqState` current tick (ms).
+type Ticks = Int
+
+-- | `SeqState` tick limit (ms).
+type Limit = Int
+
+-- | `SeqState` index.
+type Index = Int
+
+-- | A sequencer state is either @Stopped@ or @Running ticks limit index@, where 
+-- @limit@ is the step size in milliseconds, @ticks@ is the time in milliseconds
+-- since the last step, and @index@ is the new step index. A sequencer can be 
+-- started and stopped by means of an exported boolean service operation.
+data SeqState 
+  = Stopped 
+  | Running Ticks Limit Index 
+  deriving (Typeable, Data)
+
+-- | A generic skeleton for creating sequencer components; i.e.,components that 
+-- produce sequences of observable effects at programmable points in time. The 
+-- skeleton is parametric in the actual step function (which takes takes a step 
+-- index as a parameter, performs any observable effects, and returns a new 
+-- new sequencer state). 
+sequencer :: Data a 
+          => RTE c SeqState 
+          -- ^ Initial state/sequencer setup program
+          -> (Index -> RTE c SeqState) 
+          -- ^ Step function
+          -> (a -> RTE c SeqState)
+          -- ^ Controller
+          -> Atomic c (ClientServerOperation a () Provided c)
[…]

### 2016-08-29 arsim-examples/AUTOSAR/Shared/Velocity.hs

@@ -1,0 +1,160 @@
+{-# LANGUAGE DeriveDataTypeable #-}
+
+module AUTOSAR.Shared.Velocity 
+  ( VelocityCtrl(..)
+  , velocityCtrl
+  ) where
+
+import Control.Monad
+import NewARSim 
+
+-- * Velocity estimation
+-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+
+-- | The velocity estimation algorithm requires wheel speeds and (filtered)
+-- accelerometer inputs. It is assumed that the first two wheel inputs belong to
+-- the non-driving wheels (if any).
+data VelocityCtrl = VelocityCtrl
+  { wheels   :: [DataElem Unqueued Double Required]
+    -- ^ Wheel velocities
+  , accel    :: DataElem Unqueued Double Required
+    -- ^ Accelerometer
+  , velocity :: DataElem Unqueued Double Provided
+    -- ^ Velocity approximation
+  }
+
+-- | Performs vehicle velocity estimation based on vehicle longtitudal
+-- acceleration and wheel speeds. Requires a longtitudal accelerometer to be
+-- mounted on the vehicle. Moreover, it is assumed that the data from this
+-- accelerometer has been filtered.
+-- 
+-- The algorithm used assumes that the first two wheel velocity inputs come from
+-- the non-driving wheels of the vehicle, and that a longtitudal accelerometer
+-- is mounted on the unit. The slip calculation algorithm is borrowed from 
+-- Active Braking Control Systems: Design for Vehicles by Savaresi and Tanelli,
+-- chapter 5.
+velocityCtrl :: Time 
+             -- ^ Sample time 
+             -> AUTOSAR VelocityCtrl
+velocityCtrl deltaT = atomic $ 
+  do wheels   <- replicateM 4 requiredPort
+     accel    <- requiredPort 
+     velocity <- providedPort
+      
+     mapM_ (`comSpec` InitValue 0.0) wheels 
+     comSpec accel    (InitValue 0.0)
+     comSpec velocity (InitValue 0.0)
+      
+     let memory = 100
+    
+     -- Keep acceleration memory for backwards integration phase.
+     accMem  <- interRunnableVariable (replicate memory 0.0)
+     avgMem  <- interRunnableVariable (replicate memory 0.0)
[…]

### 2016-08-29 arsim-examples/Examples/ABS2.hs

@@ -1,0 +1,6 @@
+module Main where
+
+import AUTOSAR
+
+main :: IO ()
+main = simulinkNewABS2 

### 2016-08-29 paper/paper.tex
Language polishing.

@@
-\documentclass[twocolumn]{article}
+\documentclass[10pt,conference]{IEEEtran}
+%\documentclass[twocolumn]{article}
 \usepackage{verbatim}
@@
 
+%TODO: Some simple (informal) example early to give the reader an early view of flavour of the results.
+%  Suggestion: use something like the "two values sent to a queue with a trigger runnable can lead to the runnable triggering once or twice" example
+
 \section{AUTOSAR}
@@
 
-Components and ports do not carry any dynamic state besides the state of their constituent parts, so they need no explicit representation as process terms.
+Components do not carry any dynamic state besides the state of their constituent parts, so they need no explicit representation as process terms.
 However, in order to still be able to talk about model elements using their locally scoped names, we employ a hierarchical addressing scheme, such that address $i.r$ means runnable $r$ of component $i$ and $i.e$ means port element $e$ within component $i$.
@@
 It furthermore forbids runnables to be triggered by any events other than operation invocations of compatible signatures, making it natural to associate the state needed to buffer up server invocations with the server runnables themselves.
-On the other hand, the specification also introduces timeouts and sequence counters that are private to each connection, which is why our calculus needs process terms that correspond to each required (i.e., client-side) operation as well.
+On the other hand, the specification also introduces timeouts and sequence counters that are private to each connection, which is why our calculus also needs process terms that correspond to each required (i.e., client-side) operation.
 Axiom (33) shows how the current sequence counter $m$ blazes a call transition if the client-side operation is not already busy, while the busy case is detected in axiom (36).
@@
 
-Here is an example of how a client runnable instance, a client port operation, and a server runnable interact during a call transition (assuming $\ToperationInvokedEvent b a$, $\Tconnect a {i.o}$, and $\TserverCallPointTimeout{i.o} t $).
+Here is an example of how a client runnable instance, a client port operation and a server runnable interact during a call transition (assuming $\ToperationInvokedEvent b a$, $\Tconnect a {i.o}$, and $\TserverCallPointTimeout{i.o} t $).
 $$
@@
 A server runnable instance is identified by a client parameter distinct from $\Tvoid$.
-Axiom (42) says that when such a runnable instance reaches the end of its code sequence, it must emit the produced value together with the sequence number and \textbf{oper} address of its current invocation, before it is allowed to terminate (see Section~\ref{sect:SpawnTerm}).
+Axiom (42) says that when such a runnable instance reaches the end of its code sequence, it must announce the produced value, together with the sequence number and \textbf{oper} address of its current invocation, before it is allowed to terminate (to be defined in Section~\ref{sect:SpawnTerm}).
 The addressed \textbf{oper} term reacts according to axiom (43), by completing the call, storing the produced value and increasing its sequence counter in preparation for the next call.
@@
 Because of call timeouts, server runnables run the risk of producing results that the original client is not waiting for anymore.
-Detecting such mismatches is the job of sequence numbers, as witnessed by the requirement in axiom (43) that the sequence numbers returned by the server and expected by the client must be identical (variable $m$).
+Detecting such mismatches is the job of the sequence numbers, as witnessed by the requirement in axiom (43): the sequence number returned by the server and the one expected by the client must be identical (variable $m$).
 However, there must also exist a means for server runnables to simply discard their results, otherwise they would not be able to terminate.
 Axiom (46) therefore defines an alternative transition for finished servers, applicable on the condition that the returned sequence number does \emph{not} match what the client expects (axiom (47)).
-By means of axiom (48), runnables can ignore the corresponding transition.
+By means of axiom (48), subscribing runnables are allowed to ignore the corresponding transition.
 
@@
 
-So, in an abbreviated form, both the following derivations are possible:
+So, in an abbreviated form, both the following derivations are possible for the same reception event sequence:
 $$
@@
 $$
-That is, two data reception events in sequence can lead to the creation of either one or two runnable instances, and both behaviors are equally correct according to the AUTOSAR semantics.
+That is, the system can respond by creating either one or two runnable instances, and both behaviors are equally correct according to the AUTOSAR semantics.
 
@@
 
-This anomaly may lead to the conclusion that the AUTOSAR specification does not really intend more than one timing event to be defined per runnable.
+This anomaly may lead to the conclusion that the AUTOSAR specification does not really intend to allow more than one timing event per runnable.
 Luckily, our formalization is also compatible with this interpretation, as the restriction only concerns what set of static models to expect as starting states, not their dynamic behavior.
@@
 That is, at what time will each timer tick for the first time --- immediately at startup or at some other (later) time?
-The AUTOSAR specification only introduces the notion of timing offset in relation to the operating system tasks that are expected to implement runnables; the Software Component Template itself is totally silent on the issue \cite{AR:SWC}.
+The AUTOSAR specification only introduces a concept of timing offsets in relation to the operating system tasks that are expected to implement runnables; the Software Component Template itself is totally silent on the issue \cite{AR:SWC}.
 Lacking an abstract definition of runnable offsets, we have chosen to simply fire all timers simultaneously at $0$ seconds from startup (hence the value $0$ in the initial {\bf timer} term definition in Section~\ref{sec:Sem}).
@@
 $$
-which lets each timer wait one full cycle before triggering for the first time.
-However, should we really want to take the assumed implementation offsets into account, the initial timer state would rather be
+which lets each timer wait one full cycle before triggering for the first time. However, should we really want to take the assumed task offsets into account, the initial timer state would rather be
 $$
@@
 $$
-for some chosen non-negative $t_o$.
+for some non-negative $t_o$.
 One might even want to add the condition $t_o < t_p$, but it must be emphasized that all these initilization options are speculative --- the AUTOSAR Software Component Template must be clarified before a sufficiently abstract model of timed behavior can be obtained.
@@
 Another concern regarding exclusive areas is that although the specification clearly demands that exiting is done in the reverse order of entering, it does not provide any means for the exit command to report detected violations of that condition \cite[ch.~5.6.29]{AR:RTE}.
-This means that if a runnable instance tries to exit exclusive area $x$, and $x$ is not the latest exclusive area acquired by that instance, the operation is neither allowed to succeed nor able to flag an error due to a \emph{void} return type.
-This leaves our transition semantics with no other option than to indefinitely block progress for such a runnable, which is the consequence of providing no transition alternative besides axiom (11) to a runnable instance about to execute the exit command.
+This means that if a runnable instance tries to exit exclusive area $x$, and $x$ is not the latest exclusive area acquired by that instance, the operation is neither allowed to succeed, nor able to flag an error (due to a \emph{void} return type).
+This leaves our transition semantics with no other option than to indefinitely block progress for such a runnable, which is the consequence of providing no transition alternative besides axiom (11) to runnable instances about to execute the exit command.
 Had the specification allowed for an error result, however, the following axioms could have been added to resolve the order violation cases:
@@
 A transition semantics like the one defined in this report provides rules for deriving the correct behaviors, while leaving the incorrect behaviors without derivation alternatives.
-This has the effect that a process term which attempts to do something illegal is assigned no meaningful remaining trace; it appears indefinitely stuck in its current state as if it were deadlocked and can participate in subsequent transitions only by ignoring what is being said.
+This has the effect that a process term which attempts to do something illegal is assigned no meaningful remaining trace; it appears indefinitely stuck in its current state as if it were deadlocked, and can participate in subsequent transitions only by ignoring what is being said.
 
@@
 A "throw" command has no continuation and is thus similar to the exception generating commands of many modern programming languages.
-It could be given an informative value parameter as well, but this will not be needed in this simplistic proposal.
+It could be given an informative value parameter as well, but this will not be needed in our simplistic proposal.
 Throw commands can appear explicitly in the code to indicate software-detected errors, or they can be the result of special failure transitions, like the following alternative resolution of an exit order violation:

### 2016-08-29 arsim-examples/AUTOSAR/ACC/Gearbox.hs
Fixed ABS code: Transitioned into a a standstill state at too high velocities.

@@ -25,8 +25,12 @@ data GearCtrl = GearCtrl
 
 -- | @'gearController' dt@ creates a gear controller reading RPM inputs
 -- periodically with an interval length of @dt@. 
-gearController :: Time -> AUTOSAR GearCtrl
-gearController deltaT = 
+gearController :: Time 
+               -- ^ Sample time
+               -> Task
+               -- ^ Task assignment
+               -> AUTOSAR GearCtrl
+gearController deltaT task = 
   let initialGear = -1   -- Initial gear setting
       timerLimit  = 0.5  -- Time delay after a gear change has been made.
   in atomic $ 
@@ -40,8 +44,7 @@ gearController deltaT =
      gears <- interRunnableVariable initialGear
      timer <- interRunnableVariable (True, 0.8 * timerLimit)
 
---      runnable (MinInterval 0) [TimingEvent deltaT] $
-     runnableT ["core1" :>> (6, 10)] (MinInterval 0) [TimingEvent deltaT] $ 
+     runnableT [task] (MinInterval 0) [TimingEvent deltaT] $ 
        do res <- rteRead engineRPM
           case res of 
             Ok rpm1 -> do

### 2016-08-29 arsim-examples/AUTOSAR/NewABS2/NewABS2.hs

@@ -73,8 +73,8 @@ data BangBang = BangBang
 -- of the system uses, since the actuator pulses need to be of a fairly high
 -- resolution. This also means that this is the bottleneck of the ACC example
 -- (since the rest of the system runs on a 20 ms period).
-bangBangCtrl :: AUTOSAR BangBang
-bangBangCtrl = atomic $ 
+bangBangCtrl :: Time -> Task -> AUTOSAR BangBang
+bangBangCtrl deltaT task = atomic $ 
   do valvePort <- providedPort
      slips     <- requiredPort 
      pidCall   <- requiredPort
@@ -82,14 +82,14 @@ bangBangCtrl = atomic $
 
      carrier <- interRunnableVariable (0 :: Int)
  
-     let width = 20 -- Maximum pulse width.
+     let width = 20 -- Maximum pulse width. Whatever resolution we have in the
+                    -- timing event in the runnable.
      
---      runnable (MinInterval 0) [TimingEvent 1e-3] $
-     runnableT ["core1" :>> (2, 1)] (MinInterval 0) [TimingEvent 1e-3] $ 
+     runnableT [task] (MinInterval 0) [TimingEvent deltaT] $ 
        do Ok c  <- rteIrvRead carrier
           Ok s0 <- rteRead slips
           Ok s  <- rteCall pidCall (0.2, s0)
-
+          
           -- Control signal modulation
           -- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
           -- Cut-off to @[-1, 1]@. If width * |u0| exceeds carrier 
@@ -168,10 +168,10 @@ data WheelCtrl = WheelCtrl
 -- | Wheel control module. Connects the pressure valves to the ABS controller
 -- and the carrier wave unit. Better results are achieved when the PID uses a
 -- time step below the rest of the system. 
-wheelCtrl :: Index -> AUTOSAR WheelCtrl
-wheelCtrl i = composition $ 
-  do bang <- bangBangCtrl
-     pid  <- pdController 1e-4 1e-3 6
+wheelCtrl :: Time -> Task -> Index -> AUTOSAR WheelCtrl
+wheelCtrl deltaT task i = composition $ 
+  do bang <- bangBangCtrl deltaT task
+     pid  <- pdController (deltaT / 10) 1e-2 6
 
      connect (pidOp pid) (pidCall bang)
      
[…]

### 2016-08-29 arsim-examples/AUTOSAR/NewABS2/Simulink.hs

@@ -21,12 +21,16 @@ instance External Vehicle where
   toExternal = toExternal . valvesOut 
 
 -- | Export wheel ports and accelerometer connection from the ABS system.
--- TODO: Export tasks + time delta
 vehicleIO :: AUTOSAR Vehicle
 vehicleIO = do
-  declareTask "core1" (TimingEvent 1e-3)
-  absCtrl  <- absSystem
-  veloCtrl <- velocityCtrl 1e-2
+  let delta = 1e-3
+      veloT = "core1" :-> 0
+      absT  = "core1" :-> 1
+      bangT = "core1" :-> 2
+  declareTask "core1" (TimingEvent delta)
+
+  absCtrl  <- absSystem    delta absT  bangT
+  veloCtrl <- velocityCtrl delta veloT
 
   connect (velocity veloCtrl) (absVeloIn absCtrl)
   

### 2016-08-29 paper/paper.tex
Most refs done.

@@
 AUTOSAR elements and operations are actually only unique within the {\em ports} they belong to, so technically we ought to address them by reference to an intended {\em port name} as well, as in $i.p.e$ or $i.p.o$ for some port $p$.
-However, port names play no role of their own in our study, so we choose to leave them be implied by the element and operation names we reference.
+However, port names play no role of their own in our study, so we choose to leave them implied by the element and operation names we reference.
 Moreover, component names \V{I} can also be considered to have a nested internal structure (component \V{I_1} within component \V{I_2} within component \V{I_3}, etc).
@@
 
-In the untyped setting of this report, a value $v$ can be any data item computed and communicated by an AUTOSAR system, including the unit value $\Tvoid$ and the error codes that might be returned from RTE calls.
+In the untyped setting of this report, a value $v$ can be any data item computed and communicated by an AUTOSAR system, including the unit value $\Tvoid$ and the error codes that might be returned from RTE calls \cite[ch.~5.5.1]{AR:RTE}.
 Meta-variables $n$ and $m$ stand for natural numbers, while $u$ and $t$ range over boolean values and (floating point) time values, respectively.
@@
 By $\Tlength{\Vs}$ we mean the length of sequence $\Vs$.
-Left (right) concatenation of an element with a sequence is written $\Tcons{\V{V}}{\V{Vs}}$ ($\Tsnoc{\V{Vs}}{\V{V}}$).
+Left (right) concatenation of an element with a sequence is written $\Tcons v \Vs$ ($\Tsnoc \Vs v$).
+We use meta-variable conventions to resolve whether concatenation implies a fixed number of elements (as in $\Tcons {v_1} {v_2}$) or is open-ended (as in $\Tcons v \Vs$ or $\Tsnoc \Vs v$).
 
@@
 
-To this end, we have chosen to ignore all internal computations and just capture the observable effects of runnable execution as a pure sequence of RTE (VFB) calls.
-And because the result from an RTE call may in general affect subsequent runnable behaviour, we use a continuation-passing style where every $\Code$ term except the final return command literally contains the sequence that follows it, as a continuation parameter $\Cont$.
+To this end, we have chosen to ignore all internal computations and just capture the observable effects of runnable execution as a pure string of RTE (VFB) calls \cite[ch.~5.6]{AR:RTE}.
+And because the result from an RTE call may in general affect subsequent behaviour, we use a continuation-passing style \cite{Reynolds93} where every $\Code$ term except the final return command literally contains the commands that may follow, as a continuation parameter $\Cont$.
 %
@@
 
-TODO: links from syntax and concepts introduced here to the AUTOSAR spec
-
 \section{Semantics}
@@
 This behaviour serves us well, because in general, what one AUTOSAR runnable does may have impact on many (if not all) parts of the executing system.
-It is, however, important not to confuse this broadcasting notion with any particular form of AUTOSAR communication.
+It is, however, important not to confuse this broadcasting notion with any particular form of AUTOSAR communication \cite[ch.~4.2]{AR:SWC}.
 The way transition labels distribute over the parallel composition operator is merely a technical aspect of our process calculus, and will be used to express several different AUTOSAR communication semantics, among other things.
@@
 
-The AUTOSAR naming scheme guarantees that all processes terms in such an initial composition will carry unique addresses, with the exception of timer terms which share addresses with the runnables they belong to.
+The AUTOSAR naming scheme \cite[ch.~3.2.5]{AR:SWC} guarantees that all processes terms in such an initial composition will carry unique addresses, with the exception of timer terms which share addresses with the runnables they belong to.
 We will later see how runnable instances reuse the address of their runnables in the same manner.
@@
 Some further static model information will also be referenced by the transition rules.
-We refer to the connectors of a model using a binary relation $\Tconnect{}{}$ on port elements and operations, where the arrow points in the \emph{provided-to-required} direction.
+We refer to the connectors of a model using a binary relation $\Tconnect{}{}$ on port elements and operations, where the arrow points in the \emph{provided-to-required} direction \cite[ch.~3.3.3]{AR:SWC}.
 Although connectors are tecnically defined for ports, we assume $\Tconnect{}{}$ is lifted to the common elements and operations of connected ports, and that it is transitively closed (i.e., expresses end-to-end connectivity across \emph{Delegation-} as well as \emph{AssemblySwConnector}s).
 
-We expect the implementation of a runnable to be available as a continuation function $\Cont$, even though the real AUTOSAR model will be referring to concrete functions in external C/Matlab files.
+We expect the implementation of a runnable to be available as a continuation function $\Cont$, even though the real AUTOSAR model will be referring to concrete functions in external C files \cite[ch.~5.1.1]{AR:RTE}.
 We will further abstract away from actual numbers of input and output parameters to these functions by assuming that records (structs) are used as the single continuation argument and result whenever the arity so requires.
@@
 
-The semantic transition axioms are written in full in Appendix~\ref{sec:Rules}.
-We will discuss each group of axioms in turn, starting with the relatively simple behaviour of inter-runnable variables and exclusive areas.
+The semantic transition rulesand axioms  are written in full in Appendix~\ref{sec:Rules}.
+We will discuss each definion group in turn, starting with the relatively simple behaviour of inter-runnable variables and exclusive areas.
 In most cases, the driving force behind a transition is some runnable instance wishing to say something which other processes either react to or ignore.
@@
 
-An inter-runnable variable (irv) represents state that is local to a particular component and accessible by all runnables of that component.
+An inter-runnable variable (irv) represents state that is local to a particular component and accessible by all runnables of that component \cite[ch.~4.3.3.1]{AR:RTE}.
 Axioms (5) and (7) show the \emph{say} transitions taken by a runnable instance wishing to read and write an irv $s$.
@@
 
-An exclusive area is the AUTOSAR equivalent of a binary semaphore, that can be acquired and released in an atomic fashion by competing runnable instances.
+An exclusive area is the AUTOSAR equivalent of a binary semaphore, that can be acquired and released in an atomic fashion by competing runnable instances \cite[ch.~4.2.2.5]{AR:RTE}.
 Axiom (9) states that a runnable instance wishing to enter exclusive area $x$ may successfully proceed to its continuation $k$ by broadcasting $\say {i.x}{\Tent}$.
@@
 The conditions on exiting are actually stronger than those that govern entering, since axiom (11) is only applicable to a runnable instance for which the exited area $x$ is at the top of its stack of acquired exclusive areas (as expressed by the sequence pattern $\Tcons x \Xs$).
-An exclusive area can thus only be exited by the process that entered it, and only in the reverse order of entering.
+An exclusive area can thus only be exited by the process that entered it, and only in the reverse order of entering \cite[ch.~5.6.28]{AR:RTE}.
 Section~\ref{sec:ExclNest} will bring up some alternatives to this semantics, and also discuss the interpretation of blocked processes in more detail.
@@
 These transitions are just syntactic variants of shared variable reading and writing (axioms (5) and (7)).
-A data element (\textbf{delem}) process term contains two pieces of state in addition to its address: a communicated data value and a boolean flag for keeping track of unread writes.
+A data element (\textbf{delem}) process term contains two pieces of state in addition to its address: a communicated data value and a boolean flag for keeping track of unread writes \cite[ch.~4.3.1.10.1]{AR:RTE}.
 Axioms (14) and (16) capture reads and writes to the data value and also set the update flag accordingly.
 The side-condition in axiom (16) makes the rule apply only if there is a static connection from the data element being written and the address of the matched term (recall that data element storage is associated with the receiving side of a connection).
-Because AUTOSAR allows sender-receiver communication patterns to be one-to-many (such as $\Tconnect {i.e} a$ and $\Tconnect {i.e} b$), parallel data element updates like the following are possible:
+Because AUTOSAR allows sender-receiver communication patterns to be one-to-many (such as $\Tconnect {i.e} a$ and $\Tconnect {i.e} b$), parallel data element updates like the following are possible \cite[ch.~4.3.1.4]{AR:RTE}:
 $$
@@
 It expresses that a runnable set up to trigger on data reception on required data element $a$ should be marked as \textbf{pending} whenever a write occurs on an element connected to $a$.
-An example may be as follows (assuming $\Tconnect {i.e} a$ and $\TdataReceivedEvent b a$:
+An example may be as follows (assuming $\Tconnect {i.e} a$ and $\TdataReceivedEvent b a$):
 $$
@@
 
-A data element can also be queried for its update status flag, and be marked as carrying no valid value. Axioms (18)-(21) define the corresponding transitions.
+A data element can also be queried for its update status flag \cite[ch.~5.6.35]{AR:RTE}, and be marked as carrying no valid value \cite[ch.~5.6.7]{AR:RTE}.
+Axioms (18)-(21) define the corresponding transitions.
 
@@
 
-Buffered inter-process communication (axioms (22) - (30)) is just a minor variant of the unbuffered mechanism, where the single-value store of a \textbf{delem} term is replaced by a \textbf{qelem} process holding a sequence of values.
+Buffered inter-process communication (axioms (22) - (30)) is just a minor variant of the unbuffered mechanism, where the single-value store of a \textbf{delem} term is replaced by a \textbf{qelem} process holding a sequence of values \cite[ch.~4.3.1.10.2]{AR:RTE}.
 The buffer is consumed from the left (axiom (23)) and extended to the right (axiom (27)).
@@
 
-However, the AUTOSAR specification requires the send command to inform its caller whether all connected buffers were able to successfully store the communicated value or whether some buffers had it dropped.
-Such information is naturally expressed as an additional parameter $\Tvar{as}$ in the \emph{snd} transition, listing the addresses of connected \textbf{qelem} processes that have no spare capacity.
+However, the AUTOSAR specification requires the send command to inform its caller whether all connected buffers were able to successfully store the communicated value or whether some buffers had it dropped \cite[ch.~5.6.5]{AR:RTE}.
+Such non-local information can be expressed as an additional parameter $\Tvar{as}$ to the {\sc snd} transition label, listing the addresses of connected \textbf{qelem} processes that have no spare capacity.
 The preconditions of axioms (27) and (28) ensure that $\Tvar{as}$ reflects the truth, and axioms (25) and (26) feed different error codes to the sender's continuation depending on whether $\Tvar{as}$ is empty or not.
@@
 
-The following example illustrates a scenario where a sent value is only stored in a subset of the connected buffers. Assuming $\Tconnect {i.e} a$ and $\Tconnect {i.e} b$:
+The following example illustrates a scenario where a sent value is only stored in a subset of the connected buffers.
+Assuming $\Tconnect {i.e} a$ and $\Tconnect {i.e} b$ (and let $\Tvar{as}$ be the singleton sequence $b$):
 $$
@@
 
-The behaviour of client-server communication in AUTOSAR depends on whether a client lists the required server port operation as a synchronous or an asynchronous callpoint (which are mutually exclusive model attributes).
+The behaviour of client-server communication in AUTOSAR \cite[ch.~4.3.2]{AR:RTE} depends on whether a client lists the required server port operation as a synchronous or an asynchronous callpoint (which are mutually exclusive model attributes).
 In the former case, the call command is treated as if it were immediately followed by a command for retrieving the server result (axiom (31)), whereas in the latter case, the call succeeds immediately and the result has to be retrieved explicitly by the client's continuation code (axiom (32)).
@@
 
-The AUTOSAR specification restricts client-server connections to the many-to-one pattern only, assigning a dedicated runnable responsible for implementing the offered service.
+The AUTOSAR specification restricts client-server connections to the many-to-one pattern only, assigning a dedicated runnable responsible for implementing the offered service \cite[ch.~4.3.2]{AR:SWC}.
 It furthermore forbids runnables to be triggered by any events other than operation invocations of compatible signatures, making it natural to associate the state needed to buffer up server invocations with the server runnables themselves.
-On the other hand, the specification also introduces timeouts and sequence counters that are private to each connection, which is why our calculus also needs process terms that correspond to each required (i.e., client-side) operation.
+On the other hand, the specification also introduces timeouts \cite[ch.~4.3.2.3]{AR:RTE} and sequence counters \cite[ch.~4.3.2.6.1]{AR:RTE} that are private to each connection, which is why our calculus also needs process terms that correspond to each required (i.e., client-side) operation.
 Axiom (33) shows how the current sequence counter $m$ blazes a call transition if the client-side operation is not already busy, while the busy case is detected in axiom (36).
@@
 The \emph{result} command can appear as an implicit effect of a previous call command, if the operation invoked is marked as an \emph{synchronousServerCallPoint} (c.f.\ axiom (31)).
-It can also be referenced explicitly in a client runnable's code, if the current runnable lists the operation as an \emph{asynchronousServerCallPoint}).
+It can also be referenced explicitly in a client runnable's code, if the current runnable lists the operation as an \emph{asynchronousServerCallPoint} \cite[ch.~5.6.14]{AR:RTE}.
 For the synchronous case, axiom (38) excludes \textbf{nodata} values, which effectively blocks progress of the caller until a server result distinct from \textbf{nodata} has been produced.
@@
 
-Runnables with an activation state distinct from $\Tidle$ or $\Teps$ are amenable to instantiation.
+Runnables with an activation state distinct from $\Tidle$ or $\Teps$ are amenable to instantiation \cite[ch.~4.2.3]{AR:RTE}.
 Axiom (49) defines this crucial step for runnables triggered by all types of events except server invocation, which is instead handled by axiom (50).
@@
 
-In common with most real-time process calculi, the concept of passing time in our calculus is kept separate from the computational work expressed by the say/hear transitions.
+In common with most real-time process calculi, the concept of passing time in our calculus is kept separate from the computational work expressed by the say/hear transitions \cite{Baeten00processalgebra}.
 This has the advantage of freeing the semantics from possessing a particular computation speed, as any finite number of computation steps can be performed before time must advance.
@@
 The passage of time is captured as a special transition relation $\red {\Tdelta t}$, whose only effect on the related process terms is to decrement any contained time variables by $t$.
-For some processes time cannot advance until a particular work transition has been taken; this corresponds to time events that must be noted (although not necessarily acted upon) at the exact time instance when they occur.
+For some processes, time cannot advance until a particular work transition has been taken; this corresponds to time events that must be noted (although not necessarily acted upon) at the exact time instance when they occur.
 Yet other processes can only take time transitions, which simply means that they represent an idle system that currently has no work to do.
@@
 Axiom (60) defines a similar behaviour for the start interval timer of a runnable, although this timer is also allowed to slip behind further time updates once it has reached zero (axiom (61)).
-The reason for this is that the interval timers do not enforce any particular action when they time out, they just enable new transition options that do not have to be taken immediately.
+The reason for this is that the interval timers do not enforce any particular action when they time out, they just enable new transition options that do not have to be taken immediately \cite[ch.~4.2.3]{AR:RTE}.
 
@@
 The net effect of such a transition is that the runnable to which the ticking timer belongs is put into the \textbf{pending} state (axiom (64)).
-An example of a transition sequence involving a time-triggered runnable follows below.
+An example of a transition sequence involving a time-activated runnable follows below \cite[ch.~4.2.2.8]{AR:RTE}.
 $$
@@
   \Trinst {i.r} c \Xs {\Tap k \Tok} \,\Opar \\
-  {\Tqelem a 2 v} \Opar {\Tqelem b 2 {\Tcons {v_1} {v_2}}}
-\end{array}
-$$
-
-TODO: some links from rules and choices made here to the AUTOSAR spec
+  {\Tqelem a 2 v} \Opar {\Tqelem b 2 {\Tcons {v_1} {v_2}}} \\
+\end{array}
+$$
 
@@
 are derivable, we note that process $p$ may evolve into state $q$ either via an $l_1$ step followed by $l_2$, or by first taking the $l_2$ step and then $l_1$.
-Such freedom of choice is commonly taken as the definition of \emph{parallelism} in the process calculus literature --- if the sequential ordering of steps $l_1$ and $l_2$ has no significance, they might just as well be considered to occur in parallel.
+Such freedom of choice is commonly taken as the definition of \emph{parallelism} in the process calculus literature --- if the sequential ordering of steps $l_1$ and $l_2$ has no significance, they might just as well be considered to occur in parallel \cite{milner:ComMob}.
 
@@
 Here there is only one $b$ instance created and none is pending, even though $a$ correctly contains the two values received.
-The reason is that by swapping the order of the first instantiation and the second data reception steps, the system is put in a state where its inability to count pending instantiations becomes fatal.
+The reason is that by swapping the order of the first instantiation and the second data reception steps, the system is put in a state where its inability to count pending instantiations \cite[ch.~4.2.3]{AR:RTE} becomes fatal.
 
@@
 One could argue that replacing the {\bf pending}/{\bf idle} flag of runnables with a small counter variable would render the AUTOSAR event triggering mechanism much more robust.
-The standard explicitly excludes counting for all events but operation invocation, though, so the possibility of event losses is undeniably intentional \cite[ch.~4.2.3]{AR:RTE}.
+The standard explicitly excludes counting for all events but operation invocation, though, so the possibility of event losses is undeniably intentional \cite[p.~128]{AR:RTE}.
 However, should this design decision ever be revisited, the two trace alternatives above could serve as a succinct problem characterization.
@@
 \begin{itemize}
-\item Data types on any level of abstraction, and their associated mappings.
-\item Port interface mapping and data scaling.
-\item Measurements and calibration parameters.
-\item Variant handling.
+\item Data types on any level of abstraction, and their associated mappings \cite[ch.~5]{AR:SWC}.
+\item Port interface mapping and data scaling \cite[ch.~4.3]{AR:SWC}.
+\item Measurements and calibration parameters \cite[ch.~2.2]{AR:SWC}.
+\item Variant handling \cite[ch.~2.4]{AR:SWC}.
 \end{itemize}
@@
 The only atomic software component type we include on our formalization is the \emph{Application} software component type.
-Other atomic component types describe dynamic behaviors that are either just minor variants of the formalized component type (\emph{SensorActuator}, \emph{ComplexDeviceDriver}, \emph{Service}, \emph{ServiceProxy}), or trivial in comparison (\emph{NvBlock}, \emph{EcuAbstraction}, \emph{Parameter}).
-
-A similar argument justifies our decision to omit port types beyond the sender-receiver and client-server kinds, as well as per instance memories.
-A port typed by a \emph{TriggerInterface} is essentially a sender-receiver port with no data contents.
-Mode management is arguably a much more complex concept in principle, but since the related AUTOSAR mechanism is limited to communication of integer-valued mode requests and acknowledgements, it can be seen as an instance of sender-receiver communication as well.
+Other atomic component types describe dynamic behaviors that are either just minor variants of the formalized component type (\emph{SensorActuator}, \emph{ComplexDeviceDriver}, \emph{Service}, \emph{ServiceProxy}), or trivial in comparison (\emph{NvBlock}, \emph{EcuAbstraction}, \emph{Parameter}) \cite[ch.~3.2.3]{AR:SWC}.
+
+A similar argument justifies our decision to omit port types beyond the sender-receiver and client-server kinds \cite[ch.~4.2]{AR:SWC}, as well as per instance memories \cite[ch.~7.7]{AR:SWC}.
+A port typed by a \emph{TriggerInterface} \cite[ch.~4.4.7]{AR:SWC} is essentially a sender-receiver port with no data contents.
+Mode management is arguably a much more complex concept in principle, but since the related AUTOSAR mechanism is limited to communication of integer-valued mode requests and acknowledgements, it can be seen as an instance of sender-receiver communication as well \cite[ch.~4.4.6]{AR:SWC}.
 Per instance memories essentially duplicate the behavior of inter-runnable variables.

### 2016-08-30 paper/paper.tex
More references and intro tweak.

@@
 \usepackage{graphicx}
+\usepackage{authblk}
 
@@
 \title{A semantics of core AUTOSAR}
-\author{Johan Nordlander \and Patrik Jansson}
+\author{Johan Nordlander \quad\quad Patrik Jansson}
+
+\affil{Chalmers University of Technology}
+
 %%%%%%%
@@
 
-This paper takes a first step towards a remedy to these problems, by contributing a formal specification of a substantial core of the AUTOSAR standard.
+In this paper we take a first step towards a remedy to these problems, by contributing a formal specification of a substantial core of the AUTOSAR standard.
+AUTOSAR itself is summarized in Section~\ref{sec:autosar}.
 The formalization covers most of the Software Component Template and its accompanying Run-Time Environment (RTE) (Section~\ref{sec:Calc}), and is able to directly express every legal way a system of software components can evolve at run-time on an arbitrarily fast platform (Section~\ref{sec:Sem}).
 It can thus serve as a basis for both a concrete AUTOSAR implementation on a specific platform (whose supported behaviours must be a subset of those defined by the formal semantics), and a platform-independent AUTOSAR simulator (where behaviours from the legal set can be picked at random).
-Examples of semantic derivations and their applications can be found throughout the presentation.
+Non-determinism and concurrency is emphasized in Section~\ref{sec:NonDet}, together with an example of an unfortunate consequence of the specified standard.
+Section~\ref{sec:Amb} identifies some areas where the AUTOSAR documents has been found unclear; here our chosen readings are put in relation to alternative interpretations.
+Section~\ref{sec:Lim}, finally, gives a detailed account on the scope limitations we have imposed on our work.
 
@@
 The semantics is furthermore written to be unambiguous, except where concurrent execution should allow for more than one observed behaviour.
-At some points, the standard documents are clear but the resulting semantics is dubious.
-These cases will be emphasised and discussed, together with suggested ways forward (Section~\ref{sec:NonDet}).
-At other points, the AUTOSAR documents have been found unclear; here a specific choice has been made but alternative interpretations will be discussed separately (Section~\ref{sec:Amb}).
-
-TODO: Limitations
-
-%TODO: Some simple (informal) example early to give the reader an early view of flavour of the results.
-%  Suggestion: use something like the "two values sent to a queue with a trigger runnable can lead to the runnable triggering once or twice" example
+Examples of questions that can be firmly resolved on basis of our formalized semantics include:
+\begin{itemize}
+\item {\em How many runnable instances will a sequence of buffered data receptions give rise to?}
+\item {\em What values may ever be returned when an inter-runnable variable is first read?}
+\item {\em At what times will two timing events coincide, if at all?}
+\end{itemize}
 
@@
 However, this restriction does not explicitly forbid multiple \emph{Operation\-InvokedEvent}s refering to the same client-server operation to trigger separate runnables.
-Our formulation of client-server communication is based on the assumption that this is just an oversight, as allowing multiple events on a server operation would have the same problem with ambiguous response values as a would one-to-many client-server connection pattern.
-%It is not easy to see what could constitute an alternative interpretation without introducing some programmer option for selecting among server results.
+Our formulation of client-server communication is based on the assumption that this is just an oversight, as allowing multiple events on a server operation would have the same problem with conflicting response values as a would one-to-many client-server connection pattern.
+It is actually not easy to see what could constitute an alternative interpretation here, without introducing a need for some new programming mechanism for selecting among server results.
 
@@
 
-\section{Limitations}
+\section{Limitations} \label{sec:Lim}
 
@@
 
-Basic Software modules implement behaviors that cannot be captured as ordinary software components due to their dependency on hardware and other platform artefacts.
+Basic Software modules implement behaviors that cannot be captured as ordinary software components due to their dependency on hardware and other platform artefacts \cite{AR:BSW} \cite[ch.~4.2]{AR:RTE}.
 However, our abstract treatment of computation dependencies in general confines such differences to one's choice of $k$ continuations, which elimiates the need to formalize Basic Software components through any special means.
@@
 \begin{itemize}
-\item Implicit communication, where designated port elements are automatically read/written at the start/end the of runnables that access them.
-\item Category 2 runnables, which declare certain read or receive operations as \emph{WaitPoints} that block progress until the associated events occur.
+\item Implicit communication, where designated port elements are automatically read/written at the start/end the of runnables that access them \cite[ch.~4.3.1.5]{AR:RTE}.
+\item Category 2 runnables, which declare certain read or receive operations as \emph{WaitPoints} that block progress until the associated events occur \cite[ch.~7.2.4.4]{AR:SWC}.
 While this form of event handling is fundamentally different from the standard mechansim of activating runnables, it may be noted that the result command that follows from calling a \emph{synchronousServerCallPoint} already defines the essence of such a blocking \emph{WaitPoint}.
-\item The AUTOSAR COM layer and related events and commands (\emph{DataWriteCompleted}, \emph{DataSendCompleted}, \emph{DataReceiveError}, the {\em Feedback} command) \cite{AR:COM}.
+\item The AUTOSAR COM layer \cite{AR:COM} and related events and commands (\emph{DataWriteCompleted}, \emph{DataSendCompleted}, \emph{DataReceiveError}, the {\em Feedback} command \cite[ch.~5.7.5, 5.6.8]{AR:RTE}).
 COM is the communication substrate between nodes in a distributed AUTOSAR system and its limited storage capacity introduces an element of resource competition among communicating components.
-It also leads to the separation of sending and reception into distinct events, as well as a new class of communication errors.
-\item Most details found in the \emph{ComSpec} annotations that can be attached to ports (currently only \emph{initValues} are acknowledged).
-\item The meta-programming oriented RTE {\em Lifecycle} and {\em Callback} APIs.
+It also leads to the separation of sending and reception into distinct events, as well as additional communication errors.
+\item Most details found in the \emph{ComSpec} annotations that can be attached to ports (currently only \emph{initValues} are acknowledged) \cite[ch.~4.5]{AR:SWC}.
+\item The meta-programming oriented RTE {\em Lifecycle} and {\em Callback} APIs \cite[ch.~5.8-9]{AR:RTE}.
 \end{itemize}

### 2016-08-30 paper/paper.tex
Conclusion and abstract.

@@
 
-\affil{Chalmers University of Technology}
+\affil{Department of Computer Science and Engineering, Chalmers University of Technology}
 
@@
 \begin{abstract}
-TODO
-
+The AUTOSAR system architecture for the automotive industry is rich in features and large in size, but lacks a comprehensive definition that allows its Software Components and Runtime Environment to be treated as an abstract programming model in its own right.
+As a result, AUTOSAR components are hard to analyze, test and evaluate without first committing to a particular execution platform and component context.
+This paper addresses this shortcoming by means of a formalization of the core AUTOSAR specification as a labelled transition system.
+The resulting semantic rules can be used to clarify the specification, emphasize design tradeoffs and serve as a foundation for simulation tools that fulfill the AUTOSAR promise of platform- and vendor-independent automotive software development.
 \end{abstract}
@@
 
-Return to the limitation - what is the next step (which can be lifted more easily)
+In this paper we have provided a formal specification of a substantial core of the AUTOSAR specification.
+Transition rules and axioms have been defined and linked to the AUTOSAR specification documents, and issues related to choices or ambiguities in the specification have been discussed.
+
+The next step forward for this work is to seek involvement of the AUTOSAR community, in order to gain feedback on our formal interpretation, stimulate clarifying discussions on open design choices, and concievably also arrive at a consensus regarding the semantics of AUTOSAR as a programming model of its own.
+
+Another line of work is obviously to lift more of the scope limitations listed in Section~\ref{sec:Lim}.
+A formalization of omitted component, port and event types is likely to present very few problems besides an inevitable size increase, which could be mitigated by a compartementalization of the semantics into independent fragments.
+More serious work may be required to incorporate the COM layer and the added set of behaviors that follow from component distribution, but the effort would also lead to a significantly better formal coverage of the AUTOSAR specification.
+
+The formal semantics may also lay the foundation of tools that analyze or transform AUTOSAR models.
+One such example is our AUTOSAR simulator \cite{ARSim}, which interprets AUTOSAR models entirely on basis of the transition system, utilizing truly random scheduling and a maximum progress assumption to select among available transitions at each step.
+The benefit of this simulator is its ability to dynamically execute AUTOSAR Software Component models as they are, without prior selection of a particular platform or translation into a low level C representation.
+A further avenue of exploration concerns the potential to prove properties of AUTOSAR system on basis of their semantics.
+To prepare for this opportunity, we have defined the transition system used in this paper as a Horn clause file readable by common theorem provers \cite{AR.eprover}, and in fact produced Appendix~\ref{sec:Rules} through a simple formatting translation of this file.
 

### 2016-08-30 paper/paper.tex
Spell check (back to U.S. English).

@@
 It is also not very abstract, in that it makes frequent references to assumed implementation techniques for the purpose of defining its semantics.
-In practice, the AUTOSAR standard becomes blurred with the specific behaviour of one's chosen platform and development tools.
+In practice, the AUTOSAR standard becomes blurred with the specific behavior of one's chosen platform and development tools.
 And because the standard is open to interpretation, the interoperability of software components across tools and platforms is often seriously hampered.
-What is more, a single software component cannot easily be studied and understood in isolation, since its interactive behaviour is only indirectly defined in terms of the concrete C-code and OS tasks that realize it and its interaction environment.
+What is more, a single software component cannot easily be studied and understood in isolation, since its interactive behavior is only indirectly defined in terms of the concrete C-code and OS tasks that realize it and its interaction environment.
 
@@
 The formalization covers most of the Software Component Template and its accompanying Run-Time Environment (RTE) (Section~\ref{sec:Calc}), and is able to directly express every legal way a system of software components can evolve at run-time on an arbitrarily fast platform (Section~\ref{sec:Sem}).
-It can thus serve as a basis for both a concrete AUTOSAR implementation on a specific platform (whose supported behaviours must be a subset of those defined by the formal semantics), and a platform-independent AUTOSAR simulator (where behaviours from the legal set can be picked at random).
+It can thus serve as a basis for both a concrete AUTOSAR implementation on a specific platform (whose supported behaviors must be a subset of those defined by the formal semantics), and a platform-independent AUTOSAR simulator (where behaviors from the legal set can be picked at random).
 Non-determinism and concurrency is emphasized in Section~\ref{sec:NonDet}, together with an example of an unfortunate consequence of the specified standard.
@@
 A formal notation is nevertheless a good starting point for any discussion on the resolution of such issues.
-The semantics is furthermore written to be unambiguous, except where concurrent execution should allow for more than one observed behaviour.
+The semantics is furthermore written to be unambiguous, except where concurrent execution should allow for more than one observed behavior.
 Examples of questions that can be firmly resolved on basis of our formalized semantics include:
@@
 Some components merely act as containers of other components structures, these are called \textbf{composition} SWCs.
-In contrast, \textbf{atomic} SWCs define their own behaviour, primarily in terms of \textbf{runnables} and \textbf{inter-runnable variables}.
-A runnable denotes a sequential and (normally) terminating piece of code, although its actual behavioural definition is typically relegated to external C or Matlab files rather than being part of the AUTOSAR model proper.
+In contrast, \textbf{atomic} SWCs define their own behavior, primarily in terms of \textbf{runnables} and \textbf{inter-runnable variables}.
+A runnable denotes a sequential and (normally) terminating piece of code, although its actual behavioral definition is typically relegated to external C or Matlab files rather than being part of the AUTOSAR model proper.
 Instead, an AUTOSAR runnable only provides constraints on what ports and inter-runnable variables the actual implementation can access; as well as \textbf{event} declarations that determine under what conditions it will get executed.
@@
 Each runnable indicates whether or not it may be instantiated concurrently with itself.
-Runnable code can also use \textbf{exclusive areas} (a form of mutex semaphores) to further control concurrent behaviour.
+Runnable code can also use \textbf{exclusive areas} (a form of mutex semaphores) to further control concurrent behavior.
 
@@
 To this end, we have chosen to ignore all internal computations and just capture the observable effects of runnable execution as a pure string of RTE (VFB) calls \cite[ch.~5.6]{AR:RTE}.
-And because the result from an RTE call may in general affect subsequent behaviour, we use a continuation-passing style \cite{Reynolds93} where every $\Code$ term except the final return command literally contains the commands that may follow, as a continuation parameter $\Cont$.
+And because the result from an RTE call may in general affect subsequent behavior, we use a continuation-passing style \cite{Reynolds93} where every $\Code$ term except the final return command literally contains the commands that may follow, as a continuation parameter $\Cont$.
 %
@@
 The effect of these agreement rules closely resembles the idea of a \emph{broadcast}: what one process says may be heard by every other process in a large parallel composition \cite{prasad:cbs}.
-This behaviour serves us well, because in general, what one AUTOSAR runnable does may have impact on many (if not all) parts of the executing system.
+This behavior serves us well, because in general, what one AUTOSAR runnable does may have impact on many (if not all) parts of the executing system.
 It is, however, important not to confuse this broadcasting notion with any particular form of AUTOSAR communication \cite[ch.~4.2]{AR:SWC}.
@@
 We refer to the connectors of a model using a binary relation $\Tconnect{}{}$ on port elements and operations, where the arrow points in the \emph{provided-to-required} direction \cite[ch.~3.3.3]{AR:SWC}.
-Although connectors are tecnically defined for ports, we assume $\Tconnect{}{}$ is lifted to the common elements and operations of connected ports, and that it is transitively closed (i.e., expresses end-to-end connectivity across \emph{Delegation-} as well as \emph{AssemblySwConnector}s).
+Although connectors are technically defined for ports, we assume $\Tconnect{}{}$ is lifted to the common elements and operations of connected ports, and that it is transitively closed (i.e., expresses end-to-end connectivity across \emph{Delegation-} as well as \emph{AssemblySwConnector}s).
 
@@
 
-The semantic transition rulesand axioms  are written in full in Appendix~\ref{sec:Rules}.
-We will discuss each definion group in turn, starting with the relatively simple behaviour of inter-runnable variables and exclusive areas.
+The semantic transition rules and axioms  are written in full in Appendix~\ref{sec:Rules}.
+We will discuss each definition group in turn, starting with the relatively simple behavior of inter-runnable variables and exclusive areas.
 In most cases, the driving force behind a transition is some runnable instance wishing to say something which other processes either react to or ignore.
@@
 
-The behaviour of client-server communication in AUTOSAR \cite[ch.~4.3.2]{AR:RTE} depends on whether a client lists the required server port operation as a synchronous or an asynchronous callpoint (which are mutually exclusive model attributes).
+The behavior of client-server communication in AUTOSAR \cite[ch.~4.3.2]{AR:RTE} depends on whether a client lists the required server port operation as a synchronous or an asynchronous callpoint (which are mutually exclusive model attributes).
 In the former case, the call command is treated as if it were immediately followed by a command for retrieving the server result (axiom (31)), whereas in the latter case, the call succeeds immediately and the result has to be retrieved explicitly by the client's continuation code (axiom (32)).
@@
 $$
-As a contrast, this is the behaviour observed when a call timeout occurs:
+As a contrast, this is the behavior observed when a call timeout occurs:
 $$
@@
 \item They create the new instance in parallel with its parent (runnable instances is actually the only example of dynamic process creation in this calculus).
-\item They initialise the new instance with code according to the model-defined implementation and an empty list of owned exclusive areas.
+\item They initialize the new instance with code according to the model-defined implementation and an empty list of owned exclusive areas.
 \end{itemize}
@@
 Yet other processes can only take time transitions, which simply means that they represent an idle system that currently has no work to do.
-Mostly, though, work and time transitions are simultaneously available, indicating that the semantics considers the slightly slower and slightly faster behaviour alternatives to be equally correct.
+Mostly, though, work and time transitions are simultaneously available, indicating that the semantics considers the slightly slower and slightly faster behavior alternatives to be equally correct.
 
@@
 The counter is not allowed to go negative, though, which forces the \textbf{oper} term to handle the timeout once the counter reaches zero (c.f.\ axiom (44)).
-Axiom (60) defines a similar behaviour for the start interval timer of a runnable, although this timer is also allowed to slip behind further time updates once it has reached zero (axiom (61)).
+Axiom (60) defines a similar behavior for the start interval timer of a runnable, although this timer is also allowed to slip behind further time updates once it has reached zero (axiom (61)).
 The reason for this is that the interval timers do not enforce any particular action when they time out, they just enable new transition options that do not have to be taken immediately \cite[ch.~4.2.3]{AR:RTE}.
@@
 
-The AUTOSAR specification contains a very subtle sensitivity to this kind of non-deterministism that is worth examplifying.
+The AUTOSAR specification contains a very subtle sensitivity to this kind of non-determinism that is worth exemplifying.
 Consider a queued data element $a$ together with a runnable $b$ set up to trigger on data reception events for $a$.
@@
 Unqueued data elements that are read before first written should return the error code {\bf neverReceived} \cite[ch.~5.6.10]{AR:RTE}.
-Since no exceptions to this rule are mentioned in the specification, our intepretation is that this holds even when a data element possesses an \emph{initValue} attribute  (initial {\bf delem} state in Section~\ref{sec:Sem}).
+Since no exceptions to this rule are mentioned in the specification, our interpretation is that this holds even when a data element possesses an \emph{initValue} attribute  (initial {\bf delem} state in Section~\ref{sec:Sem}).
 However, as this renders any such initial value assignments entirely redundant, an alternative interpretation would be to create a term
@@
 The specification restricts client-server connections such that a client may not be connected to multiple servers, for the purpose of avoiding an operation call to be handled by more than one server runnable \cite[ch.~4.2.3]{AR:RTE}.
-However, this restriction does not explicitly forbid multiple \emph{Operation\-InvokedEvent}s refering to the same client-server operation to trigger separate runnables.
+However, this restriction does not explicitly forbid multiple \emph{Operation\-InvokedEvent}s referring to the same client-server operation to trigger separate runnables.
 Our formulation of client-server communication is based on the assumption that this is just an oversight, as allowing multiple events on a server operation would have the same problem with conflicting response values as a would one-to-many client-server connection pattern.
@@
 for some non-negative $t_o$.
-One might even want to add the condition $t_o < t_p$, but it must be emphasized that all these initilization options are speculative --- the AUTOSAR Software Component Template must be clarified before a sufficiently abstract model of timed behavior can be obtained.
+One might even want to add the condition $t_o < t_p$, but it must be emphasized that all these initialization options are speculative --- the AUTOSAR Software Component Template must be clarified before a sufficiently abstract model of timed behavior can be obtained.
 
@@
 A similar specification ambiguity concerns \emph{DataReceived\-Event}s in the presence of a full receiver buffer.
-We have chosen to block these events from activating connected runnables (axiom (30)), although the opposite choice could also be justified on grounds that activation of a reciever must be even more pressing when its buffer is overflowing.
+We have chosen to block these events from activating connected runnables (axiom (30)), although the opposite choice could also be justified on grounds that activation of a receiver must be even more pressing when its buffer is overflowing.
 Expressing this formally would amount to simply deleting axiom (30) together with the last precondition of axiom (29).
@@
 An alternative way of handling program failures in general is of course to throw exceptions or trigger some other non-local error-handling mechanism.
-AUTOSAR actually specifies a set of such generic error-handling approaches, but it does so in terms of implementation artefacts like tasks and memory partitions only, rather than the software component concepts under study in this report \cite{AR:ERR}.
+AUTOSAR actually specifies a set of such generic error-handling approaches, but it does so in terms of implementation artifacts like tasks and memory partitions only, rather than the software component concepts under study in this report \cite{AR:ERR}.
 Still, the presence of exceptional errors raises a string of questions when combined with the concurrency and communication features of software components, and a clarification of this aspect could go a long way towards making AUTOSAR a self-contained programming model in its own right.
@@
 $$
-Because a failed runnable instance now has an explicit representation, we can specify details of how such a process should interact with its enviroment.
+Because a failed runnable instance now has an explicit representation, we can specify details of how such a process should interact with its environment.
 We may for example wish to define that a special error code is returned if our runnable is running on behalf of a client, and that any exclusive areas held are automatically exited in the right order (c.f.~axioms (42) and (11)):
@@
 Other design choices are of course perfectly possible, and more elaborate rules, for example involving an explicit exception-catching mechanism, could be considered as well.
-We must emphasize, though, that the rules in this section are just hypothetical examples; they are not based on the current AUTOSAR specification and are only presented to illustrate the succinctness and level of detail that can be obtained by a formal semantics aproach.
+We must emphasize, though, that the rules in this section are just hypothetical examples; they are not based on the current AUTOSAR specification and are only presented to illustrate the succinctness and level of detail that can be obtained by a formal semantics approach.
 
@@
 \end{itemize}
-These modelling elements only impose restrictions and demands on the values $v$ and continuations $k$ we reference in our semantics, and which we in turn do not constrain in any way.
+These modeling elements only impose restrictions and demands on the values $v$ and continuations $k$ we reference in our semantics, and which we in turn do not constrain in any way.
 
@@
 
-Basic Software modules implement behaviors that cannot be captured as ordinary software components due to their dependency on hardware and other platform artefacts \cite{AR:BSW} \cite[ch.~4.2]{AR:RTE}.
-However, our abstract treatment of computation dependencies in general confines such differences to one's choice of $k$ continuations, which elimiates the need to formalize Basic Software components through any special means.
+Basic Software modules implement behaviors that cannot be captured as ordinary software components due to their dependency on hardware and other platform artifacts \cite{AR:BSW} \cite[ch.~4.2]{AR:RTE}.
+However, our abstract treatment of computation dependencies in general confines such differences to one's choice of $k$ continuations, which eliminates the need to formalize Basic Software components through any special means.
 
@@
 \item Category 2 runnables, which declare certain read or receive operations as \emph{WaitPoints} that block progress until the associated events occur \cite[ch.~7.2.4.4]{AR:SWC}.
-While this form of event handling is fundamentally different from the standard mechansim of activating runnables, it may be noted that the result command that follows from calling a \emph{synchronousServerCallPoint} already defines the essence of such a blocking \emph{WaitPoint}.
+While this form of event handling is fundamentally different from the standard mechanism of activating runnables, it may be noted that the result command that follows from calling a \emph{synchronousServerCallPoint} already defines the essence of such a blocking \emph{WaitPoint}.
 \item The AUTOSAR COM layer \cite{AR:COM} and related events and commands (\emph{DataWriteCompleted}, \emph{DataSendCompleted}, \emph{DataReceiveError}, the {\em Feedback} command \cite[ch.~5.7.5, 5.6.8]{AR:RTE}).
@@
 
-The next step forward for this work is to seek involvement of the AUTOSAR community, in order to gain feedback on our formal interpretation, stimulate clarifying discussions on open design choices, and concievably also arrive at a consensus regarding the semantics of AUTOSAR as a programming model of its own.
+The next step forward for this work is to seek involvement of the AUTOSAR community, in order to gain feedback on our formal interpretation, stimulate clarifying discussions on open design choices, and conceivably also arrive at a consensus regarding the semantics of AUTOSAR as a programming model of its own.
 
 Another line of work is obviously to lift more of the scope limitations listed in Section~\ref{sec:Lim}.
-A formalization of omitted component, port and event types is likely to present very few problems besides an inevitable size increase, which could be mitigated by a compartementalization of the semantics into independent fragments.
+A formalization of omitted component, port and event types is likely to present very few problems besides an inevitable size increase, which could be mitigated by a compartmentalization of the semantics into independent fragments.
 More serious work may be required to incorporate the COM layer and the added set of behaviors that follow from component distribution, but the effort would also lead to a significantly better formal coverage of the AUTOSAR specification.

### 2016-08-30 paper/paper.tex
Micro tweak.

@@
 This paper addresses this shortcoming by means of a formalization of the core AUTOSAR specification as a labelled transition system.
-The resulting semantic rules can be used to clarify the specification, emphasize design tradeoffs and serve as a foundation for simulation tools that fulfill the AUTOSAR promise of platform- and vendor-independent automotive software development.
+The resulting semantic rules can be used to clarify the specification, emphasize design tradeoffs and serve as a foundation for tools that fulfill the AUTOSAR promise of platform- and vendor-independent automotive software development.
 \end{abstract}
@@
 \begin{itemize}
-\item {\em How many runnable instances will a sequence of buffered data receptions give rise to?}
+\item {\em How many runnable instances will a sequence of received values give rise to?}
 \item {\em What values may ever be returned when an inter-runnable variable is first read?}

### 2016-08-30 paper/paper.tex
Tweaks.

@@
 
-The formal semantics may also lay the foundation of tools that analyze or transform AUTOSAR models.
+The formal semantics may also provide a foundation for tools that analyze or transform AUTOSAR models.
 One such example is our AUTOSAR simulator \cite{ARSim}, which interprets AUTOSAR models entirely on basis of the transition system, utilizing truly random scheduling and a maximum progress assumption to select among available transitions at each step.
-The benefit of this simulator is its ability to dynamically execute AUTOSAR Software Component models as they are, without prior selection of a particular platform or translation into a low level C representation.
-A further avenue of exploration concerns the potential to prove properties of AUTOSAR system on basis of their semantics.
-To prepare for this opportunity, we have defined the transition system used in this paper as a Horn clause file readable by common theorem provers \cite{AR.eprover}, and in fact produced Appendix~\ref{sec:Rules} through a simple formatting translation of this file.
+The benefit of this simulator is its ability to dynamically execute AUTOSAR Software Component models as they are, without prior selection of a particular platform, or translation into a low level representation.
+A further avenue of exploration concerns the potential of proving properties of AUTOSAR system on basis of their semantics.
+To prepare for this opportunity, we have defined the transition system used in this paper as a Horn clause file readable by common theorem provers, and in fact produced Appendix~\ref{sec:Rules} through a simple formatting translation of this file \cite{AR.eprover}.
 

### 2016-08-30 paper/paper.tex
Paper: allowed pagebreaks in equation groups in the appendix
Paper: minor fix + formatting of urls in bibliography

@@
 \documentclass[10pt,conference]{IEEEtran}
+\usepackage{hyperref}
 %\documentclass[twocolumn]{article}
@@
 
-Another line of work is obviously to lift more of the scope limitations listed in Section~\ref{sec:Lim}.
+Another line of work is to lift more of the scope limitations listed in Section~\ref{sec:Lim}.
 A formalization of omitted component, port and event types is likely to present very few problems besides an inevitable size increase, which could be mitigated by a compartmentalization of the semantics into independent fragments.
@@
 The benefit of this simulator is its ability to dynamically execute AUTOSAR Software Component models as they are, without prior selection of a particular platform, or translation into a low level representation.
-A further avenue of exploration concerns the potential of proving properties of AUTOSAR system on basis of their semantics.
-To prepare for this opportunity, we have defined the transition system used in this paper as a Horn clause file readable by common theorem provers, and in fact produced Appendix~\ref{sec:Rules} through a simple formatting translation of this file \cite{AR.eprover}.
+A further avenue of exploration concerns the potential of proving properties of AUTOSAR systems on the basis of their semantics.
+To prepare for this opportunity, we have defined the transition system used in this paper as a Horn clause file readable by common theorem provers, and, in fact, Appendix~\ref{sec:Rules} is automatically produced through a simple formatting translation of this file \cite{AR.eprover}.
 
@@
 {
-\renewcommand{\Prule}[2]{#1 \quad \Pif\; #2\\}
+\renewcommand{\Prule}[2]{#1 \quad &\Pif\; #2\\}
 \renewcommand{\Tstep}[3]{#1 \red{#2} #3}
@@
 
-\begin{eqnarray}  \CombRed  \notag \end{eqnarray}
+\begin{align}  \CombRed  \notag \end{align}
 }
@@
 \subsection{Inter-runnable variables}
-\begin{eqnarray}      \InterrunnableVariables  \notag \end{eqnarray}
+\begin{align}      \InterrunnableVariables  \notag \end{align}
 
 \subsection{Exclusive areas}
-\begin{eqnarray}      \ExclusiveAreas          \notag \end{eqnarray}
+\begin{align}      \ExclusiveAreas          \notag \end{align}
 
 \subsection{Unbuffered sending/receiving}
-\begin{eqnarray}      \ReadingWriting          \notag \end{eqnarray}
+\begin{align}      \ReadingWriting          \notag \end{align}
 
 \subsection{Buffered sending/receiving}
-\begin{eqnarray}      \SendingReceiving        \notag \end{eqnarray}
+\begin{align}      \SendingReceiving        \notag \end{align}
 
@@
 
-\begin{eqnarray}      \CallServer              \notag \end{eqnarray}
-
-\renewcommand\Pcomma{\nonumber \\&&\text{and}\;\;}
+\begin{align}      \CallServer              \notag \end{align}
+
+\renewcommand\Pcomma{\nonumber \\&\text{and}\;\;}
 
 \subsection{Passing back a server result}
-\begin{eqnarray}      \ServerResult            \notag \end{eqnarray}
+\begin{align}      \ServerResult            \notag \end{align}
 
 \subsection{Spawning and terminating}
-\begin{eqnarray}      \SpawnTerminate          \notag \end{eqnarray}
+\begin{align}      \SpawnTerminate          \notag \end{align}
 
 \subsection{Passing time}
-\begin{eqnarray}      \PassingTime             \notag \end{eqnarray}
+\begin{align}      \PassingTime             \notag \end{align}
 
 %\renewcommand{\Prule}[2]{#1 & \Pif \;\; #2\\}
+%**TODO: check page-break
+\newpage
 \subsection{Ignoring broadcasts}
-\begin{eqnarray}      \IgnoreBroadcast         \notag \end{eqnarray}
+\begin{align}      \IgnoreBroadcast         \notag \end{align}
 

### 2016-08-31 paper/paper.tex
Added RAWFP ack. + hidden links
Polishing.

@@
 \documentclass[10pt,conference]{IEEEtran}
-\usepackage{hyperref}
+\usepackage[hidelinks]{hyperref}
 %\documentclass[twocolumn]{article}
@@
 A sender-receiver interface is an aggregation of one or more \textbf{data-elements}, with each element characterized by the type of data it carries and optional further communication specification details.
-A client-server interface likewise consists of one or more \textbf{operations}, that each give the signature of a procedure call and the possible errors that may result.
+A client-server interface likewise consists of one or more \textbf{operations}, that each gives the signature of a procedure call and the possible errors that may result.
 
@@
 Moreover, component names \V{I} can also be considered to have a nested internal structure (component \V{I_1} within component \V{I_2} within component \V{I_3}, etc).
-We can fully ignore this detail here, though, as the hierarchical layout of components in an AUTOSAR system has no run-time implications \cite[ch.~3.3.1]{AR:SWC}.
+We can ignore this detail too, though, as the hierarchical layout of components in an AUTOSAR system has no run-time implications \cite[ch.~3.3.1]{AR:SWC}.
 
@@
 In both cases the embedded continuation $k$ is applied to the command result.
-The corresponding \emph{hear} transitions of an irv appear as rules (6) and (8).
+The corresponding \emph{hear} transitions of an irv appear as axioms (6) and (8).
 Thanks to inference rule (1), a runnable instance and an irv can now engage in a parallel transition, for example expressing the writing of shared data.
@@
 
+\section{Acknowledgements}
+
+This work was mainly carried out in the project ``Resource Aware
+Functional Programming (RAWFP)'' funded by the Swedish Foundation for
+Strategic Research.
+
 \bibliography{refs}

### 2016-08-31 ACC/ACC.hs
Major updates to directory and package structure.
* arsim-examples now import the arsim package module AUTOSAR.ARSim.
* Added ARSim/ directory and renamed NewARSim to AUTOSAR.ARSim under package "arsim"
* Removed old NewARSim, NewABS, ACC directories (now under ARSim/). Moved MATLAB to ARSim/.

@@ -1,129 +1,0 @@
--- | Adaptive Cruise Control. 
-
-{-# LANGUAGE RecordWildCards   #-}
-
-module ACC 
-  ( -- * ACC system
-    ACCSystem(..)
-  , accSystem
-    -- * Types
-  , Velo, Throttle
-  ) where
-
-import Control.Monad
-import Data.Maybe
-import Generic
-import Revlimit 
-import NewARSim
-import PID
-
--- * Types 
--- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-
-type Dist     = Double
-type Velo     = Double
-type Throttle = Double
-
--- * ACC composition
--- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-
-data ACCSystem = ACCSystem
-  { -- Inputs
-    accCruise   :: DataElem Unqueued Velo Required
-  , accVeloIn   :: DataElem Unqueued Velo Required
-  , accTarget   :: DataElem Unqueued (Maybe (Velo, Dist)) Required
-    -- Outputs
-  , accThrottle :: DataElem Unqueued Throttle Provided
-  , accBrake    :: DataElem Unqueued Throttle Provided
-  }
-
--- | Adaptive Cruise Control system.
-accSystem :: Time -> AUTOSAR ACCSystem
-accSystem timeStep = composition $
-  do -- Sample time resolution of the entire system. Inherited 
-     -- by the target vehicle sensor and the ACC ECU.
-    
-     accECU <- cruiseCtrl timeStep
-
-     -- PD controller
-     -- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-     -- Create a PD controller for the ACC. Parameters are 
-     -- found by trial-and-error.
-     pidCtrl <- pdController (timeStep / 10) 5e-3 5e-1
-
-     -- Connections
-     connect (pidOp pidCtrl)    (ctrl accECU)
-     
-     accCruise <- requiredDelegate [crVel accECU]
-     accVeloIn <- requiredDelegate [vhVel accECU]
-     accTarget <- requiredDelegate [target accECU]
-
[…]

### 2016-08-31 ACC/Gearbox.hs

@@ -1,94 +1,0 @@
--- | Automatic transmission for Simulink 4-speed gearbox.
---
--- The transmission component 'gearController' runs on a set sample time
--- (preferably but not necessarily the same as the rest of the system) and
--- periodically reads engine RPM and attempts to shift up/down based on this.
---
--- The code is not intended as a faithful model of an automatic transmission,
--- but is rather there to give us larger freedom with speed settings in
--- Simulink, since the vehicle's engine model is somewhat sensitive to RPM.
-module Gearbox 
-  ( GearCtrl(..)
-  , gearController
-  ) where
-
-import Control.Monad
-import Generic
-import NewARSim
-
--- | The gear logic requires engine rpm as input, and provides a gear signal
--- in the range @[-1, 4]@ as output.
-data GearCtrl = GearCtrl
-  { engineRPM  :: DataElem Unqueued Double  Required
-  , gearSignal :: DataElem Unqueued Integer Provided
-  }
-
--- | @'gearController' dt@ creates a gear controller reading RPM inputs
--- periodically with an interval length of @dt@. 
-gearController :: Time -> AUTOSAR GearCtrl
-gearController deltaT = 
-  let initialGear = -1   -- Initial gear setting
-      timerLimit  = 0.5  -- Time delay after a gear change has been made.
-  in atomic $ 
-  do engineRPM  <- requiredPort
-     gearSignal <- providedPort
-     comSpec gearSignal (InitValue initialGear)
-   
-     -- Initial setup: Gearbox in neutral and timer enabled to prevent
-     -- gear lock-in before Simulink vehicle model stabilizes.
-     revs  <- interRunnableVariable 0.0
-     gears <- interRunnableVariable initialGear
-     timer <- interRunnableVariable (True, 0.8 * timerLimit)
-
---      runnable (MinInterval 0) [TimingEvent deltaT] $
-     runnableT ["core1" :>> (6, 10)] (MinInterval 0) [TimingEvent deltaT] $ 
-       do res <- rteRead engineRPM
-          case res of 
-            Ok rpm1 -> do
-              Ok rpm0  <- rteIrvRead revs
-              Ok gear0 <- rteIrvRead gears
-
-              -- Check timer, we want delays between shifts to let the engine
-              -- adjust revs.
-              Ok (counting, t) <- rteIrvRead timer
-              case (counting, t >= timerLimit) of
-                (True, True)  -> 
[…]

### 2016-08-31 ACC/Vehicle.hs

@@ -1,177 +1,0 @@
--- | This module contains all vehicle-related things, such as IO/control of
--- vehicle subsystems, velocity and slip estimation, et cetera. 
-
-{-# LANGUAGE RecordWildCards #-}
-
-module Vehicle 
-  ( IOModule(..)
-  , vehicleIO
-  ) where
-
-import ACC
-import Control.Monad
-import Gearbox
-import Generic
-import NewARSim 
-import NewABS2
-import Revlimit
-import Velocity
-
-type Dist = Double
-
--- * Radar control
--- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-
-data RadarCtrl = RadarCtrl
-  { radarDistIn :: DataElem Unqueued Double                   Required
-  , radarRelIn  :: DataElem Unqueued Double                   Required
-  , radarOutput :: DataElem Unqueued (Maybe (Double, Double)) Provided
-  }
-
-radarCtrl :: Time -> Double -> AUTOSAR RadarCtrl
-radarCtrl deltaT thresh = atomic $ do
-  distIn <- requiredPort
-  relIn  <- requiredPort
-  out    <- providedPort
-
-  comSpec out    (InitValue Nothing)
-  comSpec distIn (InitValue 100)
-  comSpec relIn  (InitValue 0.0)
-
-  runnable (MinInterval 0) [TimingEvent deltaT] $ do
-    Ok d <- rteRead distIn
-    Ok r <- rteRead relIn
-    let output
-          | d < thresh = Just (r, d)
-          | otherwise  = Nothing
-    rteWrite out output
-
-  return $ sealBy RadarCtrl distIn relIn out
-
--- * Throttle control
--- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-
-data ThrottleCtrl = ThrottleCtrl
-  { throttleCtrlIn  :: DataElem Unqueued Double Required
-  , throttleCtrlOut :: DataElem Unqueued Double Provided
-  , throttleCtrlRpm :: DataElem Unqueued Double Required
-  , throttleCtrlRev :: ClientServerOp (Double, Double) Double Required
-  }
-
--- Vehicle throttle controller. Limits throughput signal to the @[-1, 1]@
--- range. Makes use of the rev-limiter.
[…]

### 2016-08-31 ARSim/arsim-examples/AUTOSAR.hs

@@ -1,0 +1,39 @@
+module AUTOSAR
+  ( -- * NewABS, NewABS2
+    standaloneNewABS
+  , simulinkNewABS2
+    -- * ACC example
+  , simulinkACCgood1
+  , simulinkACCbad1
+    -- * Re-export simulation functions
+  , simulateUsingExternal, simulateStandalone
+  ) where
+
+import qualified AUTOSAR.NewABS.Standalone
+import qualified AUTOSAR.NewABS2.Simulink
+import qualified AUTOSAR.ACC.Simulink
+import           AUTOSAR.ACC.Vehicle
+import           AUTOSAR.ARSim
+
+-- * ABS examples
+-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+
+-- | Standalone ABS example. No task mapping.
+standaloneNewABS :: IO ()
+standaloneNewABS = AUTOSAR.NewABS.Standalone.main
+
+-- | Simulink ABS example, one task, good task mapping.
+simulinkNewABS2 :: IO ()
+simulinkNewABS2 = AUTOSAR.NewABS2.Simulink.main
+
+-- * ACC examples
+-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+
+-- | Simulink ACC example, one task, good task mapping.
+simulinkACCgood1 :: AUTOSAR IOModule
+simulinkACCgood1 = AUTOSAR.ACC.Simulink.singleTaskGood
+
+-- | Simulink ACC example, one task, bad task mapping.
+simulinkACCbad1 :: AUTOSAR IOModule
+simulinkACCbad1 = AUTOSAR.ACC.Simulink.singleTaskBad
+

### 2016-08-31 ARSim/arsim-examples/AUTOSAR/ACC/ACC.hs

@@ -1,0 +1,136 @@
+-- | Adaptive Cruise Control. 
+
+{-# LANGUAGE RecordWildCards   #-}
+
+module AUTOSAR.ACC.ACC
+  ( -- * ACC system
+    ACCSystem(..)
+  , accSystem
+    -- * Types
+  , Velo, Throttle
+  ) where
+
+import AUTOSAR.ACC.Revlimit 
+import AUTOSAR.ARSim 
+import AUTOSAR.Shared.Generic
+import AUTOSAR.Shared.PID
+import Control.Arrow
+import Control.Monad
+import Data.Maybe
+
+-- * Types 
+-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+
+type Dist     = Double
+type Velo     = Double
+type Throttle = Double
+
+-- * ACC composition
+-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+
+data ACCSystem = ACCSystem
+  { -- Inputs
+    accCruise   :: DataElem Unqueued Velo Required
+  , accVeloIn   :: DataElem Unqueued Velo Required
+  , accTarget   :: DataElem Unqueued (Maybe (Velo, Dist)) Required
+    -- Outputs
+  , accThrottle :: DataElem Unqueued Throttle Provided
+  , accBrake    :: DataElem Unqueued Throttle Provided
+  }
+
+-- | Adaptive Cruise Control system.
+accSystem :: Time 
+          -- ^ Sample time
+          -> Task 
+          -- ^ Task assignment for the cruise control
+          -> AUTOSAR ACCSystem
+accSystem deltaT task = composition $
+  do -- Sample time resolution of the entire system. Inherited 
+     -- by the target vehicle sensor and the ACC ECU.
+    
+     accECU <- cruiseCtrl deltaT task
+
+     -- PD controller
+     -- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+     -- Create a PD controller for the ACC. Parameters are 
+     -- found by trial-and-error.
+     pidCtrl <- pdController (deltaT / 10) 5e-3 5e-1
+
+     -- Connections
[…]

### 2016-08-31 ARSim/arsim-examples/AUTOSAR/ACC/Revlimit.hs

@@ -1,0 +1,23 @@
+-- | Limit engine revs by cutting off throttle.
+module AUTOSAR.ACC.Revlimit
+  ( RevLimit(..)
+  , revLimit
+  ) where
+
+import AUTOSAR.ARSim
+
+-- | The rev-limiter exports an operation which takes a @(throttle, rpm)@ tuple
+-- and returns a @throttle@ value.
+newtype RevLimit = RevLimit
+  { revOp :: ClientServerOp (Double, Double) Double Provided }
+
+-- | Limits the engine speed to a fixed number of revolutions per minute. Simply
+-- cuts off throttle until engine revs have dropped below threshold. 
+revLimit :: Double -> AUTOSAR RevLimit
+revLimit revLimit = atomic $
+  do revOp <- providedPort
+     serverRunnable (MinInterval 0) [OperationInvokedEvent revOp] $ 
+       \(throttle, rpm) -> 
+         return $ if rpm >= revLimit then -1 else throttle
+     return $ sealBy RevLimit revOp
+

### 2016-08-31 ARSim/arsim-examples/AUTOSAR/ACC/Simulink.hs

@@ -1,0 +1,79 @@
+-- | External simulation of the ACC component using Simulink (or similar). 
+module AUTOSAR.ACC.Simulink 
+  ( module AUTOSAR.ACC.Vehicle
+  , singleTaskGood
+  , singleTaskBad
+  ) where
+
+import AUTOSAR.ACC.Vehicle
+import AUTOSAR.ARSim 
+import Control.Monad
+
+-- | Entire system assigned to a single task:
+--
+--      ORDER    Runnable       Execution policy
+--      ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ 
+--      0        velocityCtrl   Every cycle, 1 ms 
+--      1        mainLoop       (Data reception)
+--      2        bangBang       Every cycle, 1 ms
+--      3        radarCtrl      Every 10 cycles, 10 ms
+--      4        cruiseCtrl     Every 10 cycles, 10 ms
+--      5        brakeCtrl      (Data reception)
+--      6        throttleCtrl   (Data reception)
+--      7        gearController Every 10 cycles, 10 ms
+--
+--  Sample time set to 1 ms.
+tasks1 :: [Task]
+tasks1 =
+  [ "core1" :>> (0, 1)
+  , "core1" :>> (1, 1)
+  , "core1" :>> (2, 1)
+  , "core1" :>> (3, 10)
+  , "core1" :>> (4, 10)
+  , "core1" :>> (5, 10)
+  , "core1" :>> (6, 10)
+  , "core1" :>> (7, 10)
+  ]
+
+-- | Entire system assigned to a single task with some bad choice. Velocity
+-- approximation (or anything else that acts as a source in the data
+-- dependencies) appears too late in the task ordering.
+--
+--      ORDER    Runnable       Execution policy
+--      ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ 
+--      4        velocityCtrl   Every cycle, 1 ms 
+--      0        mainLoop       (Data reception)
+--      1        bangBang       Every cycle, 1 ms
+--      2        radarCtrl      Every 10 cycles, 10 ms
+--      3        cruiseCtrl     Every 10 cycles, 10 ms
+--      5        brakeCtrl      (Data reception)
+--      6        throttleCtrl   (Data reception)
[…]

### 2016-08-31 ARSim/arsim-examples/AUTOSAR/ACC/Vehicle.hs

@@ -1,0 +1,187 @@
+-- | This module contains all vehicle-related things, such as IO/control of
+-- vehicle subsystems, velocity and slip estimation, et cetera. 
+
+{-# LANGUAGE RecordWildCards #-}
+
+module AUTOSAR.ACC.Vehicle 
+  ( IOModule(..)
+  , vehicleIO
+  ) where
+
+import AUTOSAR.ACC.ACC
+import AUTOSAR.ACC.Gearbox
+import AUTOSAR.ACC.Revlimit
+import AUTOSAR.ARSim 
+import AUTOSAR.NewABS2.NewABS2
+import AUTOSAR.Shared.Generic
+import AUTOSAR.Shared.Velocity
+
+import Control.Monad
+
+type Dist = Double
+
+-- * Radar control
+-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+
+data RadarCtrl = RadarCtrl
+  { radarDistIn :: DataElem Unqueued Double                   Required
+  , radarRelIn  :: DataElem Unqueued Double                   Required
+  , radarOutput :: DataElem Unqueued (Maybe (Double, Double)) Provided
+  }
+
+radarCtrl :: Time -> Task -> Double -> AUTOSAR RadarCtrl
+radarCtrl deltaT task thresh = atomic $ do
+  distIn <- requiredPort
+  relIn  <- requiredPort
+  out    <- providedPort
+
+  comSpec out    (InitValue Nothing)
+  comSpec distIn (InitValue 100)
+  comSpec relIn  (InitValue 0.0)
+
+  runnableT [task] (MinInterval 0) [TimingEvent deltaT] $ do
+    Ok d <- rteRead distIn
+    Ok r <- rteRead relIn
+    let output
+          | d < thresh = Just (r, d)
+          | otherwise  = Nothing
+    rteWrite out output
+
+  return $ sealBy RadarCtrl distIn relIn out
+
+-- * Throttle control
+-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+
+data ThrottleCtrl = ThrottleCtrl
+  { throttleCtrlIn  :: DataElem Unqueued Double Required
[…]

### 2016-08-31 ARSim/arsim-examples/AUTOSAR/NewABS/NewABS.hs

@@ -1,0 +1,389 @@
+{-
+Copyright (c) 2014-2016, Johan Nordlander, Jonas Duregård, Michał Pałka,
+                         Patrik Jansson and Josef Svenningsson
+All rights reserved.
+
+Redistribution and use in source and binary forms, with or without
+modification, are permitted provided that the following conditions are met:
+
+   * Redistributions of source code must retain the above copyright notice,
+     this list of conditions and the following disclaimer.
+   * Redistributions in binary form must reproduce the above copyright
+     notice, this list of conditions and the following disclaimer in the
+     documentation and/or other materials provided with the distribution.
+   * Neither the name of the Chalmers University of Technology nor the names of its
+     contributors may be used to endorse or promote products derived from this
+     software without specific prior written permission.
+
+THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE
+DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE
+FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
+DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
+SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER
+CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY,
+OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
+OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
+-}
+
+{-# LANGUAGE DeriveDataTypeable #-}
+{-# LANGUAGE TypeFamilies       #-}
+{-# LANGUAGE RecordWildCards    #-}
+{-# LANGUAGE FlexibleInstances  #-}
+
[…]

### 2016-08-31 ARSim/arsim-examples/AUTOSAR/NewABS/Standalone.hs

@@ -1,0 +1,125 @@
+module AUTOSAR.NewABS.Standalone where
+
+import AUTOSAR.ARSim
+import AUTOSAR.NewABS.NewABS
+import Control.Monad
+import Graphics.EasyPlot
+
+-------------------------------------------------------------------
+--  A simulated physical car
+-------------------------------------------------------------------
+wheel_f :: Index -> Time -> Bool -> Bool -> Velo -> (Accel, Velo)
+wheel_f i time pressure relief velo
+        | time < 1.0                = veloStep 0          velo
+        | i /= 2                    = veloStep (-4.5)     velo
+        -- let wheel 2 skid...
+        -- We "postulate" a reasonable approximation of the wheel
+        -- speed (ignoring the actual valves). Ideally the whole car
+        -- dynamics whould be in another module (in Simulink).
+        | time < 1.6                = veloStep (-10)      velo
+        | time < 2                  = veloStep (-4)       velo
+        | time < 2.5                = veloStep (-3)       velo
+        | time < 3                  = veloStep 0.0        velo
+        | time < 3.4                = veloStep (-4.5)     velo
+        | time < 4                  = veloStep (-5)       velo
+        | time < 4.3                = veloStep (-8.4)     velo
+        | time < 4.7                = veloStep (-4)       velo
+        | otherwise                 = veloStep 0          velo
+-- The "pressure logic" is something like this, but a proper
+-- treatment need integration over time and knowledge of vehicle speed
+-- and other physical parameters.
+--        | pressure && not relief    = veloStep (-10)      velo
+--        | relief && not pressure    = veloStep (-1)       velo
+--        | otherwise                 = veloStep (-3)       velo
+
+timeStep :: Time
+timeStep = 0.01
+
+veloStep :: Accel -> Velo -> (Accel, Velo)
+veloStep a v = (a, v + a*timeStep)
+
+data Car = Car {
+            actuators  :: [ValveP Required],
+            v_sensors  :: [DataElem Unqueued Velo  Provided],
+            a_sensors  :: [DataElem Unqueued Accel Provided] }
+
+simulated_car :: AUTOSAR Car
+simulated_car = atomic $ do
+        wheels <- forM [1..4] $ \i -> do
+            actuator <- requiredPort
+            v_sensor <- providedPort
+            a_sensor <- providedPort
+            comSpec actuator (InitValue (False, True))
[…]

### 2016-08-31 ARSim/arsim-examples/AUTOSAR/NewABS2/NewABS2.hs

@@ -1,0 +1,279 @@
+{-
+Copyright (c) 2014-2016, Johan Nordlander, Jonas Duregård, Michał Pałka,
+                         Patrik Jansson and Josef Svenningsson
+All rights reserved.
+
+Redistribution and use in source and binary forms, with or without
+modification, are permitted provided that the following conditions are met:
+
+   * Redistributions of source code must retain the above copyright notice,
+     this list of conditions and the following disclaimer.
+   * Redistributions in binary form must reproduce the above copyright
+     notice, this list of conditions and the following disclaimer in the
+     documentation and/or other materials provided with the distribution.
+   * Neither the name of the Chalmers University of Technology nor the names of its
+     contributors may be used to endorse or promote products derived from this
+     software without specific prior written permission.
+
+THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE
+DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE
+FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
+DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
+SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER
+CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY,
+OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
+OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
+-}
+
+{-# LANGUAGE TypeFamilies       #-}
+{-# LANGUAGE RecordWildCards    #-}
+{-# LANGUAGE FlexibleInstances  #-}
+
+module AUTOSAR.NewABS2.NewABS2
+  ( -- * ABS system
[…]

### 2016-08-31 ARSim/arsim-examples/AUTOSAR/NewABS2/Simulink.hs

@@ -1,0 +1,50 @@
+module AUTOSAR.NewABS2.Simulink where
+
+import AUTOSAR.ARSim
+import AUTOSAR.NewABS2.NewABS2
+import AUTOSAR.Shared.Velocity
+import Control.Monad
+import System.Random
+
+enumLabel :: [(Address, String)] -> [(Address, String)]
+enumLabel is = map (\((a, l), i) -> (a, l ++ show i)) $ is `zip` [1..]
+
+data Vehicle = Vehicle
+  { wheelsIn  :: [DataElem Unqueued Double Required]
+  , accelIn   :: DataElem Unqueued Double Required 
+  , valvesOut :: [ValveP Provided]
+  }
+
+instance External Vehicle where
+  fromExternal vh = relabel "ACCEL" (fromExternal (accelIn vh)) ++
+                    enumLabel (relabel "WHEEL" (fromExternal (wheelsIn vh)))
+  toExternal = toExternal . valvesOut 
+
+-- | Export wheel ports and accelerometer connection from the ABS system.
+vehicleIO :: AUTOSAR Vehicle
+vehicleIO = do
+  let delta = 1e-3
+      veloT = "core1" :-> 0
+      absT  = "core1" :-> 1
+      bangT = "core1" :-> 2
+  declareTask "core1" (TimingEvent delta)
+
+  absCtrl  <- absSystem    delta absT  bangT
+  veloCtrl <- velocityCtrl delta veloT
+
+  connect (velocity veloCtrl) (absVeloIn absCtrl)
+  
+  let ws = wheels veloCtrl `zip` map veloIn (wheelPorts absCtrl)
+  wheelsIn  <- forM ws $ \(vc, wp) -> requiredDelegate [vc, wp]
+  valvesOut <- forM (map valveOut (wheelPorts absCtrl)) $ \vp ->
+    providedDelegate [vp]
+  
+  accelIn <- requiredDelegate [accel veloCtrl]
+
+  return $ sealBy Vehicle wheelsIn accelIn valvesOut 
+
+main :: IO ()
+main = do
+  simulateUsingExternal True vehicleIO
+  return ()
+

### 2016-08-31 ARSim/arsim-examples/AUTOSAR/NewABS2/SimulinkDrive.hs

@@ -1,0 +1,16 @@
+module Main where
+
+import AUTOSAR.ARSim
+import AUTOSAR.NewABS2.NewABS2
+import System.Random
+import Test.QuickCheck
+import Test.QuickCheck.Monadic
+
+main :: IO ()
+--main = simulateDriveExternal "./CarModel" absSystem >> return ()
+main = quickCheckWith stdArgs { maxSuccess = 5 } prop_example
+
+prop_example :: () -> Property
+prop_example () = monadicIO $ run $ do
+  simulateDriveExternal "./CarModel" absSystem
+  return True

### 2016-08-31 ARSim/arsim-examples/AUTOSAR/Shared/PID.hs

@@ -1,0 +1,74 @@
+-- | AUTOSAR PID controller.
+--
+-- A set of P/PI/PD/PID controllers. Note that these are somewhat sensitive to
+-- parameters, sample times and noise in the input. It's a good idea to filter
+-- noisy inputs for this reason.
+module AUTOSAR.Shared.PID 
+  ( -- * Full PID controller
+    PIDCtrl(..)
+  , pidController
+    -- * PD controller
+  , pdController
+    -- * PI controller
+  , piController
+  ) where
+
+import AUTOSAR.ARSim
+
+-- * PID controller
+-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+-- The PID controller produces a linear combination of Propotional-, 
+-- Integrating and Differentiating gain. A few useful special cases (PI-, PD-)
+-- are exported as well.
+
+-- | The PID controller requires a @(cruise, vehicle)@ velocity tuple and
+-- provides a throttle control.
+newtype PIDCtrl = PIDCtrl
+  { pidOp :: ClientServerOp (Double, Double) Double Provided }
+
+-- | Create a PID controller.
+pidController :: Time 
+              -- ^ Sample time
+              -> Time
+              -- ^ Derivative time 
+              -> Time
+              -- ^ Integral time
+              -> Double
+              -- ^ Proportional scale 
+              -> AUTOSAR PIDCtrl
+              -- ^ Throttle control (output)
+pidController dt td ti k = atomic $ do
+     state <- interRunnableVariable (0.0, 0.0)
+     pidOp <- providedPort
+     serverRunnable (MinInterval 0) [OperationInvokedEvent pidOp] $
+       \(ctrl, feedback) -> do
+          Ok (prevErr, prevInt) <- rteIrvRead state
+         
+          let err        = ctrl - feedback
+              derivative = (err - prevErr) / dt
+              integral   = prevInt + dt / ti * err
+              output     = k * (err + integral + td * derivative) 
+          rteIrvWrite state (err, integral) 
+          return output
+     return $ sealBy PIDCtrl pidOp
+
+-- | Create a PD controller (proportional and derivative control).
+pdController :: Time 
+             -- ^ Sample time
+             -> Time
+             -- ^ Derivative time
+             -> Double 
+             -- ^ Proportional scale
+             -> AUTOSAR PIDCtrl
[…]

### 2016-08-31 ARSim/arsim-examples/AUTOSAR/Shared/Velocity.hs

@@ -1,0 +1,162 @@
+{-# LANGUAGE DeriveDataTypeable #-}
+
+module AUTOSAR.Shared.Velocity 
+  ( VelocityCtrl(..)
+  , velocityCtrl
+  ) where
+
+import AUTOSAR.ARSim 
+import Control.Monad
+
+-- * Velocity estimation
+-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+
+-- | The velocity estimation algorithm requires wheel speeds and (filtered)
+-- accelerometer inputs. It is assumed that the first two wheel inputs belong to
+-- the non-driving wheels (if any).
+data VelocityCtrl = VelocityCtrl
+  { wheels   :: [DataElem Unqueued Double Required]
+    -- ^ Wheel velocities
+  , accel    :: DataElem Unqueued Double Required
+    -- ^ Accelerometer
+  , velocity :: DataElem Unqueued Double Provided
+    -- ^ Velocity approximation
+  }
+
+-- | Performs vehicle velocity estimation based on vehicle longtitudal
+-- acceleration and wheel speeds. Requires a longtitudal accelerometer to be
+-- mounted on the vehicle. Moreover, it is assumed that the data from this
+-- accelerometer has been filtered.
+-- 
+-- The algorithm used assumes that the first two wheel velocity inputs come from
+-- the non-driving wheels of the vehicle, and that a longtitudal accelerometer
+-- is mounted on the unit. The slip calculation algorithm is borrowed from 
+-- Active Braking Control Systems: Design for Vehicles by Savaresi and Tanelli,
+-- chapter 5.
+velocityCtrl :: Time 
+             -- ^ Sample time 
+             -> Task
+             -- ^ Task assignment
+             -> AUTOSAR VelocityCtrl
+velocityCtrl deltaT task = atomic $ 
+  do wheels   <- replicateM 4 requiredPort
+     accel    <- requiredPort 
+     velocity <- providedPort
+      
+     mapM_ (`comSpec` InitValue 0.0) wheels 
+     comSpec accel    (InitValue 0.0)
+     comSpec velocity (InitValue 0.0)
+      
+     let memory = 200
+    
+     -- Keep acceleration memory for backwards integration phase.
+     accMem  <- interRunnableVariable (replicate memory 0.0)
[…]

### 2016-08-31 ARSim/arsim-examples/Examples/ACC.hs

@@ -1,0 +1,9 @@
+module Main where
+
+import AUTOSAR
+
+-- | Set this flag to @False@ to disable all task assignments.
+main :: IO ()
+main = do
+  simulateUsingExternal True simulinkACCgood1
+  return ()

### 2016-08-31 ARSim/arsim/AUTOSAR/ARSim.hs

@@ -1,0 +1,2194 @@
+{-
+Copyright (c) 2014-2016, Johan Nordlander, Jonas Duregård, Michał Pałka,
+                         Patrik Jansson and Josef Svenningsson
+All rights reserved.
+
+Redistribution and use in source and binary forms, with or without
+modification, are permitted provided that the following conditions are met:
+
+   * Redistributions of source code must retain the above copyright notice,
+     this list of conditions and the following disclaimer.
+   * Redistributions in binary form must reproduce the above copyright
+     notice, this list of conditions and the following disclaimer in the
+     documentation and/or other materials provided with the distribution.
+   * Neither the name of the Chalmers University of Technology nor the names of its
+     contributors may be used to endorse or promote products derived from this
+     software without specific prior written permission.
+
+THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE
+DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE
+FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
+DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
+SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER
+CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY,
+OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
+OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
+-}
+
+{-# LANGUAGE DeriveAnyClass             #-}
+{-# LANGUAGE DeriveFunctor              #-}
+{-# LANGUAGE ExistentialQuantification  #-}
[…]

### 2016-08-31 NewABS/Simulink.hs

@@ -1,11 +1,0 @@
-module Main where
-
-import NewARSim
-import NewABS2
-import System.Random
-
-main :: IO ()
-main = do
-  simulateUsingExternal True absSystem
-  return ()
-

### 2016-08-31 arsim-examples/AUTOSAR/ACC/ACC.hs

@@ -1,136 +1,0 @@
--- | Adaptive Cruise Control. 
-
-{-# LANGUAGE RecordWildCards   #-}
-
-module AUTOSAR.ACC.ACC
-  ( -- * ACC system
-    ACCSystem(..)
-  , accSystem
-    -- * Types
-  , Velo, Throttle
-  ) where
-
-import AUTOSAR.ACC.Revlimit 
-import AUTOSAR.Shared.Generic
-import AUTOSAR.Shared.PID
-import Control.Arrow
-import Control.Monad
-import Data.Maybe
-import NewARSim
-
--- * Types 
--- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-
-type Dist     = Double
-type Velo     = Double
-type Throttle = Double
-
--- * ACC composition
--- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-
-data ACCSystem = ACCSystem
-  { -- Inputs
-    accCruise   :: DataElem Unqueued Velo Required
-  , accVeloIn   :: DataElem Unqueued Velo Required
-  , accTarget   :: DataElem Unqueued (Maybe (Velo, Dist)) Required
-    -- Outputs
-  , accThrottle :: DataElem Unqueued Throttle Provided
-  , accBrake    :: DataElem Unqueued Throttle Provided
-  }
-
--- | Adaptive Cruise Control system.
-accSystem :: Time 
-          -- ^ Sample time
-          -> Task 
-          -- ^ Task assignment for the cruise control
-          -> AUTOSAR ACCSystem
-accSystem deltaT task = composition $
-  do -- Sample time resolution of the entire system. Inherited 
-     -- by the target vehicle sensor and the ACC ECU.
-    
-     accECU <- cruiseCtrl deltaT task
-
-     -- PD controller
-     -- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-     -- Create a PD controller for the ACC. Parameters are 
-     -- found by trial-and-error.
-     pidCtrl <- pdController (deltaT / 10) 5e-3 5e-1
-
-     -- Connections
[…]

### 2016-08-31 arsim-examples/AUTOSAR/NewABS/Standalone.hs

@@ -1,125 +1,0 @@
-module AUTOSAR.NewABS.Standalone where
-
-import AUTOSAR.NewABS.NewABS
-import Control.Monad
-import Graphics.EasyPlot
-import NewARSim
-
--------------------------------------------------------------------
---  A simulated physical car
--------------------------------------------------------------------
-wheel_f :: Index -> Time -> Bool -> Bool -> Velo -> (Accel, Velo)
-wheel_f i time pressure relief velo
-        | time < 1.0                = veloStep 0          velo
-        | i /= 2                    = veloStep (-4.5)     velo
-        -- let wheel 2 skid...
-        -- We "postulate" a reasonable approximation of the wheel
-        -- speed (ignoring the actual valves). Ideally the whole car
-        -- dynamics whould be in another module (in Simulink).
-        | time < 1.6                = veloStep (-10)      velo
-        | time < 2                  = veloStep (-4)       velo
-        | time < 2.5                = veloStep (-3)       velo
-        | time < 3                  = veloStep 0.0        velo
-        | time < 3.4                = veloStep (-4.5)     velo
-        | time < 4                  = veloStep (-5)       velo
-        | time < 4.3                = veloStep (-8.4)     velo
-        | time < 4.7                = veloStep (-4)       velo
-        | otherwise                 = veloStep 0          velo
--- The "pressure logic" is something like this, but a proper
--- treatment need integration over time and knowledge of vehicle speed
--- and other physical parameters.
---        | pressure && not relief    = veloStep (-10)      velo
---        | relief && not pressure    = veloStep (-1)       velo
---        | otherwise                 = veloStep (-3)       velo
-
-timeStep :: Time
-timeStep = 0.01
-
-veloStep :: Accel -> Velo -> (Accel, Velo)
-veloStep a v = (a, v + a*timeStep)
-
-data Car = Car {
-            actuators  :: [ValveP Required],
-            v_sensors  :: [DataElem Unqueued Velo  Provided],
-            a_sensors  :: [DataElem Unqueued Accel Provided] }
-
-simulated_car :: AUTOSAR Car
-simulated_car = atomic $ do
-        wheels <- forM [1..4] $ \i -> do
-            actuator <- requiredPort
-            v_sensor <- providedPort
-            a_sensor <- providedPort
-            comSpec actuator (InitValue (False, True))
-            comSpec v_sensor (InitValue init_v)
[…]

### 2016-08-31 arsim-examples/AUTOSAR/NewABS2/SimulinkDrive.hs

@@ -1,16 +1,0 @@
-module Main where
-
-import NewARSim
-import NewABS2
-import System.Random
-import Test.QuickCheck
-import Test.QuickCheck.Monadic
-
-main :: IO ()
---main = simulateDriveExternal "./CarModel" absSystem >> return ()
-main = quickCheckWith stdArgs { maxSuccess = 5 } prop_example
-
-prop_example :: () -> Property
-prop_example () = monadicIO $ run $ do
-  simulateDriveExternal "./CarModel" absSystem
-  return True

### 2016-08-31 arsim-examples/AUTOSAR/Shared/PID.hs

@@ -1,74 +1,0 @@
--- | AUTOSAR PID controller.
---
--- A set of P/PI/PD/PID controllers. Note that these are somewhat sensitive to
--- parameters, sample times and noise in the input. It's a good idea to filter
--- noisy inputs for this reason.
-module AUTOSAR.Shared.PID 
-  ( -- * Full PID controller
-    PIDCtrl(..)
-  , pidController
-    -- * PD controller
-  , pdController
-    -- * PI controller
-  , piController
-  ) where
-
-import NewARSim
-
--- * PID controller
--- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
--- The PID controller produces a linear combination of Propotional-, 
--- Integrating and Differentiating gain. A few useful special cases (PI-, PD-)
--- are exported as well.
-
--- | The PID controller requires a @(cruise, vehicle)@ velocity tuple and
--- provides a throttle control.
-newtype PIDCtrl = PIDCtrl
-  { pidOp :: ClientServerOp (Double, Double) Double Provided }
-
--- | Create a PID controller.
-pidController :: Time 
-              -- ^ Sample time
-              -> Time
-              -- ^ Derivative time 
-              -> Time
-              -- ^ Integral time
-              -> Double
-              -- ^ Proportional scale 
-              -> AUTOSAR PIDCtrl
-              -- ^ Throttle control (output)
-pidController dt td ti k = atomic $ do
-     state <- interRunnableVariable (0.0, 0.0)
-     pidOp <- providedPort
-     serverRunnable (MinInterval 0) [OperationInvokedEvent pidOp] $
-       \(ctrl, feedback) -> do
-          Ok (prevErr, prevInt) <- rteIrvRead state
-         
-          let err        = ctrl - feedback
-              derivative = (err - prevErr) / dt
-              integral   = prevInt + dt / ti * err
-              output     = k * (err + integral + td * derivative) 
-          rteIrvWrite state (err, integral) 
-          return output
-     return $ sealBy PIDCtrl pidOp
-
--- | Create a PD controller (proportional and derivative control).
-pdController :: Time 
-             -- ^ Sample time
-             -> Time
-             -- ^ Derivative time
-             -> Double 
-             -- ^ Proportional scale
-             -> AUTOSAR PIDCtrl
-pdController deltaT td = pidController deltaT td (1/0)
-
[…]

### 2016-08-31 arsim-examples/Examples/ACC.hs

@@ -1,6 +1,0 @@
-module Main where
-
-import AUTOSAR
-
-main :: IO ()
-main = simulinkACC

### 2016-08-31 ARSim/MATLAB/makeMEX.m
Renamed NewABS2 to ABS. Some removals and comment updates.

@@ -1,11 +1,0 @@
-function makeMEX(~)
-%MAKEMEX Build the MEX file including the protocol code.
-
-if nargin < 1
-  mex swrapper.c protocol.c
-else
-  mex -g swrapper.c protocol.c
-end
-
-end
-

### 2016-08-31 ARSim/MATLAB/protocol.c

@@ -2,7 +2,7 @@
 #include <stdio.h>
 #include <unistd.h>
 
-#include "protocol.h"
+#include "protocol.h" // what
 #include "debugm.h"
 
 char PROTOCOL_MAGIC_NUMBERS[2] = {0, 1};

### 2016-08-31 ARSim/MATLAB/setMaskLabel.m

@@ -6,7 +6,7 @@ function setMaskLabel(block_path, label, num, mode, init)
 %   This is called from swrapper.c to set port labels for the mask
 %   surrounding the S-function. For this to work properly the S-function
 %   will need to be connected to MUX and DEMUX blocks. Either of the masks
-%   used in CarModel.slx or ACCModel.slx can be used as templates.
+%   used in ABS_demo.slx or ACC_demo.slx could be used as templates. 
 %
 %   This function overwrites all previous mask initialization commands when
 %   called from swrapper.c.
@@ -19,7 +19,7 @@ function setMaskLabel(block_path, label, num, mode, init)
   % set_param(mask, 'Mask', 'on');
   if strcmp(bdroot, mask)
     fprintf(['[ERROR] setMaskLabel: S-function block %s not properly ', ...
-             'masked.\nPlease refer to CarModel.slx or ACCModel.slx ',  ...
+             'masked.\nPlease refer to ABS_demo.slx or ACC_demo.slx ',  ...
              'for usage examples.\n'], block_path); 
     return
   end

### 2016-08-31 ARSim/MATLAB/swrapper.c

@@ -2,7 +2,7 @@
  *
  * Oskar Abrahamsson, <aboskar@chalmers.se> 
  *
- * S-function wrapper for ARSIM. Connects to the pipe-based interface of ARSIM
+ * S-function wrapper for ARSim. Connects to the pipe-based interface of ARSim
  * allowing for Simulink to replace external components in the simulated AUTOSAR 
  * system.
  *
@@ -20,6 +20,10 @@
  * ARSim with empty inputs one time before proceeding with regular
  * simulation. This is done in init_simulator() which is called from
  * mdlInitializeSizes.
+ *
+ * NOTE: Most of the transmissions use read/write directly on file descriptors.
+ * The same goes Haskell-side, meaning that we pipes could probably be replaced
+ * by sockets fairly easily.
  *
  */
 #include <fcntl.h>

### 2016-08-31 ARSim/arsim-examples/AUTOSAR.hs

@@ -1,7 +1,6 @@
 module AUTOSAR
-  ( -- * NewABS, NewABS2
-    standaloneNewABS
-  , simulinkNewABS2
+  ( -- * ABS example
+    simulinkABS
     -- * ACC example
   , simulinkACCgood1
   , simulinkACCbad1
@@ -9,22 +8,21 @@ module AUTOSAR
   , simulateUsingExternal, simulateStandalone
   ) where
 
-import qualified AUTOSAR.NewABS.Standalone
-import qualified AUTOSAR.NewABS2.Simulink
+import qualified AUTOSAR.ABS.Simulink
 import qualified AUTOSAR.ACC.Simulink
 import           AUTOSAR.ACC.Vehicle
 import           AUTOSAR.ARSim
 
+
+-- TODO: Good/bad for ABS
+
 -- * ABS examples
 -- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 
--- | Standalone ABS example. No task mapping.
-standaloneNewABS :: IO ()
-standaloneNewABS = AUTOSAR.NewABS.Standalone.main
-
+-- TODO: Do this like ACC
 -- | Simulink ABS example, one task, good task mapping.
-simulinkNewABS2 :: IO ()
-simulinkNewABS2 = AUTOSAR.NewABS2.Simulink.main
+simulinkABS :: IO ()
+simulinkABS = AUTOSAR.ABS.Simulink.main
 
 -- * ACC examples
 -- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

### 2016-08-31 ARSim/arsim-examples/AUTOSAR/ABS/ABS.hs

@@ -27,11 +27,11 @@ OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
 OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
 -}
 
-{-# LANGUAGE TypeFamilies       #-}
-{-# LANGUAGE RecordWildCards    #-}
-{-# LANGUAGE FlexibleInstances  #-}
+{-# LANGUAGE TypeFamilies      #-}
+{-# LANGUAGE RecordWildCards   #-}
+{-# LANGUAGE FlexibleInstances #-}
 
-module AUTOSAR.NewABS2.NewABS2
+module AUTOSAR.ABS.ABS
   ( -- * ABS system
     ABS(..)
   , absSystem

### 2016-08-31 ARSim/arsim-examples/AUTOSAR/ABS/Simulink.hs

@@ -1,7 +1,7 @@
-module AUTOSAR.NewABS2.Simulink where
+module AUTOSAR.ABS.Simulink where
 
 import AUTOSAR.ARSim
-import AUTOSAR.NewABS2.NewABS2
+import AUTOSAR.ABS.ABS
 import AUTOSAR.Shared.Velocity
 import Control.Monad
 import System.Random

### 2016-08-31 ARSim/arsim-examples/AUTOSAR/ABS/SimulinkDrive.hs

@@ -1,7 +1,7 @@
-module Main where
+module AUTOSAR.ABS.SimulinkDrive where
 
 import AUTOSAR.ARSim
-import AUTOSAR.NewABS2.NewABS2
+import AUTOSAR.ABS.ABS
 import System.Random
 import Test.QuickCheck
 import Test.QuickCheck.Monadic

### 2016-08-31 ARSim/arsim-examples/AUTOSAR/ACC/Simulink.hs

@@ -35,17 +35,20 @@ tasks1 =
   , "core1" :>> (7, 10)
   ]
 
+-- TODO: Add a bad task assignment which actually /does/ something.
+-- 
 -- | Entire system assigned to a single task with some bad choice. Velocity
 -- approximation (or anything else that acts as a source in the data
--- dependencies) appears too late in the task ordering.
+-- dependencies) appears too late in the task ordering. This effectively
+-- prevents the system from doing /anything/ useful.
 --
 --      ORDER    Runnable       Execution policy
 --      ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ 
---      4        velocityCtrl   Every cycle, 1 ms 
+--      3        velocityCtrl   Every cycle, 1 ms 
 --      0        mainLoop       (Data reception)
 --      1        bangBang       Every cycle, 1 ms
 --      2        radarCtrl      Every 10 cycles, 10 ms
---      3        cruiseCtrl     Every 10 cycles, 10 ms
+--      4        cruiseCtrl     Every 10 cycles, 10 ms
 --      5        brakeCtrl      (Data reception)
 --      6        throttleCtrl   (Data reception)
 --      7        gearController Every 10 cycles, 10 ms
@@ -53,11 +56,11 @@ tasks1 =
 --  Sample time set to 1 ms.
 tasks1bad :: [Task]
 tasks1bad =
-  [ "core1" :>> (4, 1)
+  [ "core1" :>> (1, 1)
   , "core1" :>> (0, 1)
-  , "core1" :>> (1, 1)
-  , "core1" :>> (2, 10)
+  , "core1" :>> (2, 1)
   , "core1" :>> (3, 10)
+  , "core1" :>> (4, 10)
   , "core1" :>> (5, 10)
   , "core1" :>> (6, 10)
   , "core1" :>> (7, 10)

### 2016-08-31 ARSim/arsim-examples/AUTOSAR/ACC/Vehicle.hs

@@ -8,11 +8,11 @@ module AUTOSAR.ACC.Vehicle
   , vehicleIO
   ) where
 
+import AUTOSAR.ABS.ABS
 import AUTOSAR.ACC.ACC
 import AUTOSAR.ACC.Gearbox
 import AUTOSAR.ACC.Revlimit
 import AUTOSAR.ARSim 
-import AUTOSAR.NewABS2.NewABS2
 import AUTOSAR.Shared.Generic
 import AUTOSAR.Shared.Velocity
 

### 2016-08-31 ARSim/arsim-examples/AUTOSAR/NewABS/NewABS.hs

@@ -1,389 +1,0 @@
-{-
-Copyright (c) 2014-2016, Johan Nordlander, Jonas Duregård, Michał Pałka,
-                         Patrik Jansson and Josef Svenningsson
-All rights reserved.
-
-Redistribution and use in source and binary forms, with or without
-modification, are permitted provided that the following conditions are met:
-
-   * Redistributions of source code must retain the above copyright notice,
-     this list of conditions and the following disclaimer.
-   * Redistributions in binary form must reproduce the above copyright
-     notice, this list of conditions and the following disclaimer in the
-     documentation and/or other materials provided with the distribution.
-   * Neither the name of the Chalmers University of Technology nor the names of its
-     contributors may be used to endorse or promote products derived from this
-     software without specific prior written permission.
-
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
-AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
-IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE
-FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
-DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
-SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER
-CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY,
-OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
-OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
--}
-
-{-# LANGUAGE DeriveDataTypeable #-}
-{-# LANGUAGE TypeFamilies       #-}
-{-# LANGUAGE RecordWildCards    #-}
-{-# LANGUAGE FlexibleInstances  #-}
-
[…]

### 2016-08-31 ARSim/arsim-examples/AUTOSAR/NewABS/Standalone.hs

@@ -1,125 +1,0 @@
-module AUTOSAR.NewABS.Standalone where
-
-import AUTOSAR.ARSim
-import AUTOSAR.NewABS.NewABS
-import Control.Monad
-import Graphics.EasyPlot
-
--------------------------------------------------------------------
---  A simulated physical car
--------------------------------------------------------------------
-wheel_f :: Index -> Time -> Bool -> Bool -> Velo -> (Accel, Velo)
-wheel_f i time pressure relief velo
-        | time < 1.0                = veloStep 0          velo
-        | i /= 2                    = veloStep (-4.5)     velo
-        -- let wheel 2 skid...
-        -- We "postulate" a reasonable approximation of the wheel
-        -- speed (ignoring the actual valves). Ideally the whole car
-        -- dynamics whould be in another module (in Simulink).
-        | time < 1.6                = veloStep (-10)      velo
-        | time < 2                  = veloStep (-4)       velo
-        | time < 2.5                = veloStep (-3)       velo
-        | time < 3                  = veloStep 0.0        velo
-        | time < 3.4                = veloStep (-4.5)     velo
-        | time < 4                  = veloStep (-5)       velo
-        | time < 4.3                = veloStep (-8.4)     velo
-        | time < 4.7                = veloStep (-4)       velo
-        | otherwise                 = veloStep 0          velo
--- The "pressure logic" is something like this, but a proper
--- treatment need integration over time and knowledge of vehicle speed
--- and other physical parameters.
---        | pressure && not relief    = veloStep (-10)      velo
---        | relief && not pressure    = veloStep (-1)       velo
---        | otherwise                 = veloStep (-3)       velo
-
-timeStep :: Time
-timeStep = 0.01
-
-veloStep :: Accel -> Velo -> (Accel, Velo)
-veloStep a v = (a, v + a*timeStep)
-
-data Car = Car {
-            actuators  :: [ValveP Required],
-            v_sensors  :: [DataElem Unqueued Velo  Provided],
-            a_sensors  :: [DataElem Unqueued Accel Provided] }
-
-simulated_car :: AUTOSAR Car
-simulated_car = atomic $ do
-        wheels <- forM [1..4] $ \i -> do
-            actuator <- requiredPort
-            v_sensor <- providedPort
-            a_sensor <- providedPort
-            comSpec actuator (InitValue (False, True))
-            comSpec v_sensor (InitValue init_v)
[…]

### 2016-08-31 ARSim/arsim-examples/Examples/ABS.hs

@@ -3,4 +3,4 @@ module Main where
 import AUTOSAR
 
 main :: IO ()
-main = standaloneNewABS
+main = simulinkABS

### 2016-08-31 ARSim/arsim-examples/Examples/ACC.hs

@@ -5,5 +5,6 @@ import AUTOSAR
 -- | Set this flag to @False@ to disable all task assignments.
 main :: IO ()
 main = do
-  simulateUsingExternal True simulinkACCgood1
+--   simulateUsingExternal True simulinkACCgood1
+  simulateUsingExternal True simulinkACCbad1
   return ()

### 2016-08-31 ARSim/arsim-examples/AUTOSAR/ABS/Simulink.hs
Bad task assignment for ABS available.
Defaulting to good task assignments to minimize confusion.

@@ -21,13 +21,9 @@ instance External Vehicle where
   toExternal = toExternal . valvesOut 
 
 -- | Export wheel ports and accelerometer connection from the ABS system.
-vehicleIO :: AUTOSAR Vehicle
-vehicleIO = do
-  let delta = 1e-3
-      veloT = "core1" :-> 0
-      absT  = "core1" :-> 1
-      bangT = "core1" :-> 2
-  declareTask "core1" (TimingEvent delta)
+vehicleIO :: Time -> [Task] -> AUTOSAR Vehicle
+vehicleIO delta tasks = composition $ do
+  let [veloT, absT, bangT] = tasks
 
   absCtrl  <- absSystem    delta absT  bangT
   veloCtrl <- velocityCtrl delta veloT
@@ -43,8 +39,35 @@ vehicleIO = do
 
   return $ sealBy Vehicle wheelsIn accelIn valvesOut 
 
-main :: IO ()
-main = do
-  simulateUsingExternal True vehicleIO
-  return ()
+
+-- | Entire system assigned to a single task:
+--
+--      ORDER    Runnable       Execution policy
+--      ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ 
+--      0        velocityCtrl   Every cycle, 1 ms 
+--      1        mainLoop       (Data reception)
+--      2        bangBang       Every cycle, 1 ms
+--
+--  Sample time set to 1 ms.
+tasks1 :: [Task]
+tasks1 =
+  [ "core1" :-> 0
+  , "core1" :-> 1 
+  , "core1" :-> 2
+  ]
+
+-- | System using the task assignment in 'tasks1'.
+singleTaskGood :: AUTOSAR Vehicle
+singleTaskGood = do
+  let delta = 1e-3
+  declareTask "core1" (TimingEvent delta)
+  vehicleIO delta tasks1
+
+-- | System using the task assignment in 'tasks1', but with an (obviously) bad
+-- triggering event for the task.
+singleTaskBad :: AUTOSAR Vehicle
+singleTaskBad = do
+  let delta = 1e-3
+  declareTask "core1" (TimingEvent (1.5 * delta))
+  vehicleIO delta tasks1
 

### 2016-08-31 ARSim/arsim-examples/Examples/ACC.hs

@@ -5,6 +5,6 @@ import AUTOSAR
 -- | Set this flag to @False@ to disable all task assignments.
 main :: IO ()
 main = do
---   simulateUsingExternal True simulinkACCgood1
-  simulateUsingExternal True simulinkACCbad1
+  simulateUsingExternal True simulinkACCgood1
+--   simulateUsingExternal True simulinkACCbad1
   return ()

### 2016-08-31 paper/paper.tex
Moved back to eqnarray (but kept displaybreak!)

@@
 One such example is our AUTOSAR simulator \cite{ARSim}, which interprets AUTOSAR models entirely on basis of the transition system, utilizing truly random scheduling and a maximum progress assumption to select among available transitions at each step.
-The benefit of this simulator is its ability to dynamically execute AUTOSAR Software Component models as they are, without prior selection of a particular platform, or translation into a low level representation.
+The benefit of this simulator is its ability to dynamically execute AUTOSAR Software Component models as they are, without prior commitment to a particular platform, or translation into a low level representation.
 A further avenue of exploration concerns the potential of proving properties of AUTOSAR systems on the basis of their semantics.
@@
 {
-\renewcommand{\Prule}[2]{#1 \quad &\Pif\; #2\\}
+\renewcommand{\Prule}[2]{#1 \quad \Pif\; #2\\}
 \renewcommand{\Tstep}[3]{#1 \red{#2} #3}
@@
 
-\begin{align}  \CombRed  \notag \end{align}
+\begin{eqnarray}  \CombRed  \notag \end{eqnarray}
 }
@@
 \subsection{Inter-runnable variables}
-\begin{align}      \InterrunnableVariables  \notag \end{align}
+\begin{eqnarray}      \InterrunnableVariables  \notag \end{eqnarray}
 
 \subsection{Exclusive areas}
-\begin{align}      \ExclusiveAreas          \notag \end{align}
+\begin{eqnarray}      \ExclusiveAreas          \notag \end{eqnarray}
 
 \subsection{Unbuffered sending/receiving}
-\begin{align}      \ReadingWriting          \notag \end{align}
+\begin{eqnarray}      \ReadingWriting          \notag \end{eqnarray}
 
 \subsection{Buffered sending/receiving}
-\begin{align}      \SendingReceiving        \notag \end{align}
+\begin{eqnarray}      \SendingReceiving        \notag \end{eqnarray}
 
@@
 
-\begin{align}      \CallServer              \notag \end{align}
-
-\renewcommand\Pcomma{\nonumber \\&\text{and}\;\;}
+\begin{eqnarray}      \CallServer              \notag \end{eqnarray}
+
+\renewcommand\Pcomma{\nonumber \\&&\text{and}\;\;}
 
 \subsection{Passing back a server result}
-\begin{align}      \ServerResult            \notag \end{align}
+\begin{eqnarray}      \ServerResult            \notag \end{eqnarray}
 
 \subsection{Spawning and terminating}
-\begin{align}      \SpawnTerminate          \notag \end{align}
+\begin{eqnarray}      \SpawnTerminate          \notag \end{eqnarray}
 
 \subsection{Passing time}
-\begin{align}      \PassingTime             \notag \end{align}
+\begin{eqnarray}      \PassingTime             \notag \end{eqnarray}
 
@@
 \subsection{Ignoring broadcasts}
-\begin{align}      \IgnoreBroadcast         \notag \end{align}
+\begin{eqnarray}      \IgnoreBroadcast         \notag \end{eqnarray}
 

### 2016-09-01 paper/paper.tex
Tiny fixes.

@@
 for some non-negative $t_o$.
-One might even want to add the condition $t_o < t_p$, but it must be emphasized that all these initialization options are speculative --- the AUTOSAR Software Component Template must be clarified before a sufficiently abstract model of timed behavior can be obtained.
+One might even want to add the condition $t_o \leq t_p$, but it must be emphasized that all these initialization options are speculative --- the AUTOSAR Software Component Template must be clarified before a sufficiently abstract model of timed behavior can be obtained.
 
@@
 
-The formal semantics may also provide a foundation for tools that analyze or transform AUTOSAR models.
+The formal semantics may furthermore provide a foundation for tools that analyze or transform AUTOSAR models.
 One such example is our AUTOSAR simulator \cite{ARSim}, which interprets AUTOSAR models entirely on basis of the transition system, utilizing truly random scheduling and a maximum progress assumption to select among available transitions at each step.
-The benefit of this simulator is its ability to dynamically execute AUTOSAR Software Component models as they are, without prior commitment to a particular platform, or translation into a low level representation.
+The benefit of this simulator is its ability to dynamically execute AUTOSAR Software Component models as they are, without prior commitment to a particular platform or translation into a low level representation.
 A further avenue of exploration concerns the potential of proving properties of AUTOSAR systems on the basis of their semantics.

### 2016-09-02 paper/paper.tex
Minor fixups.

@@
 This has the advantage of freeing the semantics from possessing a particular computation speed, as any finite number of computation steps can be performed before time must advance.
-On the other hand, it also means that the  limitations of a particular platform cannot be directly studied unless the calculus is complemented with some form of constraint on the work-time relationship.
+On the other hand, it also means that the  limitations of a particular platform cannot be directly studied unless the calculus is complemented with some form of constraint on the work/time relationship.
 Such an extension is beyond the scope of the current report, though.
@@
 However, this restriction does not explicitly forbid multiple \emph{Operation\-InvokedEvent}s referring to the same client-server operation to trigger separate runnables.
-Our formulation of client-server communication is based on the assumption that this is just an oversight, as allowing multiple events on a server operation would have the same problem with conflicting response values as a would one-to-many client-server connection pattern.
+Our formulation of client-server communication is based on the assumption that this is just an oversight, as allowing multiple events on a server operation would have the same problem with conflicting response values as would a one-to-many client-server connection pattern.
 It is actually not easy to see what could constitute an alternative interpretation here, without introducing a need for some new programming mechanism for selecting among server results.
@@
 A similar specification ambiguity concerns \emph{DataReceived\-Event}s in the presence of a full receiver buffer.
-We have chosen to block these events from activating connected runnables (axiom (30)), although the opposite choice could also be justified on grounds that activation of a receiver must be even more pressing when its buffer is overflowing.
+We have chosen to block these events from activating connected runnables (axiom (30)), although the opposite choice could also be justified on grounds that activation of a receiver might be even more pressing when its buffer is overflowing.
 Expressing this formally would amount to simply deleting axiom (30) together with the last precondition of axiom (29).
@@
 Because a failed runnable instance now has an explicit representation, we can specify details of how such a process should interact with its environment.
-We may for example wish to define that a special error code is returned if our runnable is running on behalf of a client, and that any exclusive areas held are automatically exited in the right order (c.f.~axioms (42) and (11)):
+We may for example wish to define that a special error code is returned if our runnable is a server running on behalf of a client, and that any held exclusive areas are automatically exited in the right order (c.f.~axioms (42) and (11)):
 $$
@@
 \end{itemize}
-These modeling elements only impose restrictions and demands on the values $v$ and continuations $k$ we reference in our semantics, and which we in turn do not constrain in any way.
+These modeling elements only impose restrictions and demands on the values $v$ and continuations $k$ we reference in our semantics, and which we, in our turn, do not constrain or interpret in any way.
 
@@
 
-The only atomic software component type we include on our formalization is the \emph{Application} software component type.
+The only atomic software component type we include in our formalization is the \emph{Application} software component type.
 Other atomic component types describe dynamic behaviors that are either just minor variants of the formalized component type (\emph{SensorActuator}, \emph{ComplexDeviceDriver}, \emph{Service}, \emph{ServiceProxy}), or trivial in comparison (\emph{NvBlock}, \emph{EcuAbstraction}, \emph{Parameter}) \cite[ch.~3.2.3]{AR:SWC}.
